### 2020-01-03 yfcc100m/paper/vldb2020/design.tex
Add paper files

\section{Design and Implementation}
\label{arch}

VDMS implements a client-server design.
The VDMS Server handles client requests concurrently and coordinates
request execution across the metadata and data components in order to return a
unified response. The metadata component is the Persistent Memory Graph
Database (PMGD). The data component is our custom library, the
Visual Compute Library (VCL). The VCL enables machine friendly enhancements to
visual data.
VDMS and its components are fully available as open source tools
\footnote{https://github.com/IntelLabs/\{vdms, pmgd, vcl\}}.

\textbf{Persistent Memory Graph Database:}
Recent developments in persistent memory technologies
like 3D XPoint~\cite{IntelXPoint15}
promise storage elements  providing  nearly  the  speed  of  DRAM  and  the
durability of block-oriented storage. To provide an efficient storage
solution addressing the increasing popularity of connected data and
applications that benefit from graph like processing, we have designed
and implemented an in-persistent-memory graph database, PMGD, optimized
to run on a platform equipped with persistent memory.
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability properties expected from
databases. The graph model makes it very suitable for the data model and
access patterns shown by visual metadata.
With its natural ability to extend the schema very
easily (due to the use of a property graph model),
we can support new developments in machine learning that can lead to
enhancements to existing metadata over time.
The specific details on how the data structures are optimized for persistent
memory, and the performance comparison with other graph databases is on-going
work that will be publish separately.

\textbf{Visual Compute Library:} The VCL was designed to provide an interface
through which users can interact with visual data. For traditional formats,
the interface is an abstraction layer over OpenCV. However, it also provides a
way to use novel formats that are better suited for visual analytics: a novel,
array-based lossless image format. This format is built on the array data
manager TileDB~\cite{TileDB} and is well suited for images that are used in
visual analytics.
The VCL currently provides limited support for videos but we are enhancing
its capabilities as part of our future work.
Feature vector support is provided through an implementation based
on high-dimensional sparse arrays, also using TileDB, which contains both
storage and search functionality over feature vectors.
In addition, the VCL provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.

\begin{comment}

\begin{figure*}[t!]
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.92\linewidth]{figures/query}
  \caption{Simple Metadata Query.}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.52\textwidth}
  \centering
  \includegraphics[width=.94\linewidth]{figures/query2}
  \caption{Query involving visual transformations.}
  \label{fig:sfig2}
\end{subfigure}
\begin{subfigure}{\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results}
  \caption{Original image (left) and the result of query in (b)}
  \label{fig:sfig3}
\end{subfigure}
\caption{Snapshot of a Python notebook with sample queries}
\label{fig:query}
\end{figure*}

\end{comment}

\textbf{Request Server:}
Developers and users of machine learning frameworks and data science
applications favor simpler interfaces to access and process data and cannot
be expected to deal with two different ways of interacting with information (
metadata and visual data) instead of focusing on the algorithmic parts of their
pipelines.
VDMS takes care of coordinating client requests across the metadata and the
data as well as efficiently manages multiple clients through its Request
Server component, by implementing a JSON-based API.
It decomposes the command into
metadata and data requests, invokes the relevant calls behind the scene,
and returns a coherent response to the user after applying any additional
operations (explained in the next section).

% \textbf{Client Library:}
% A user application can use the VDMS API by defining metadata conforming to the
% query protocol we have defined and passing along blobs. The client side VDMS
% library provides a simple query function that accepts a JSON string with
% commands and an array or vector of blobs.
% Internally, the library wraps the query string and blob using Google Protobufs
% and sends it to VDMS. It also receives a similarly wrapped response from VDMS
% and returns it to the client. The responses require JSON parsing on the client
% side for the metadata string that indicates how to interpret blobs.
% VDMS provides Python and C++ client libraries with mainly
% one function: a query that passes metadata as a JSON string and,
% optionally, data as a vector of blobs. The VDMS client library manages the
% rest.

\section{VDMS API}
\label{arch}
One of our goals with VDMS was to define an API that is easy to use.
Our API explicitly predefines certain
visual primitives associated with images, videos, and feature vectors. In
addition, while we use a graph database to store our metadata, the API is not
graph specific.
We have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based
API\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description},
which is very popular across various application domains.
We understand that by defining a new JSON-based API we had to trade-off
expressiveness (compared to SPARQL or Gremlim) for the ability to design
native support for visual data, but we believe that, as development of VDMS
evolves, we will be able to achieve similar levels of expressiveness compared
to more mature query languages.
We have developed Python and C++ client
library to provide a simple query function that accepts a JSON string with
commands and an array or vector of blobs.

\begin{comment} 
% All of the folliwng is commented out.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.9\columnwidth]{figures/features}
\caption{Feature Vectors natively supported in VDMS}
\label{fig:features}
\end{figure}

We use a medical imaging dataset as a driving example, and use it
here to showcase our API. More details about this use case are discussed in
Section~\ref{demo}.
Figure~\ref{fig:query} shows a snapshot from a Jupyter Notebook with two
simple queries.
Figure~\ref{fig:sfig1} shows a simple metadata query
where information about properties of patients matching some specific
characteristics (admitted with 85 years of age or more)
is used to constrain the search.
The returned JSON structure shows 2 patients found in the database.
Note that in this examples the queries are expressed as plain strings to
help the reader understand its structure. In real applications, the queries
are structured using the preferred method or library
for handling JSON structures (e.g. Python dictionaries or jsoncpp in C++).

Figure~\ref{fig:sfig2} shows a query involving visual transformations.
In this case, the query is looking for an image with a certain \textit{id}
value, and expects  VDMS to return the image twice: once with a thresholding
transformation (e.g., zero out all pixels less than a threshold), and the
second time after applying a threshold and then resizing it.
\ref{fig:sfig3} shows the image originally inserted in VDMS (left),
along with the returned images corresponding to the query in \ref{fig:sfig2}.
Adding new operations to VDMS is easy, as our software architecture
encapsulates all operations in VCL, and any new operation can use
or wrap around OpenCV, which is used internally.

\end{comment} 
% End of Comment

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities or visual objects.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
These vectors can be labeled and classified to build search indexes,
and there are many in-memory libraries that are designed for
this task~\cite{flann, faiss}.
Using the VDMS API, users can manage feature vector indexes,
query previously inserted elements (images),
run a k-nearest neighbor search (knn) and express relationships
between existing images or feature vectors and
the newly inserted feature vectors.
By natively supporting feature vectors and knn,
VDMS allows out-of-the-box classification
functionalities for many applications.
Figure~\ref{fig:features} shows an example of how this functionality
can be integrated in a medical imaging analytics pipeline:
(1) A feature vector is extracted from an bounding
box in a brain image and labeled;
(2) the feature vector is inserted together
with all the associated metadata (type of cancer, for example)
and a link to the original image.
(3) A new feature vector is extracted from a new image,
(4) a query to VDMS is issued to classify that feature vector based on the
indexed features in VDMS, and
(5) VDMS can respond to the query with the label
associated to that feature vector.
Note that, even if the example is based on the medical images used for the
performance evaluation, this methodology is applicable in many other contexts
and use cases, such as face detection and matching.
More details about the JSON API for this functionality can be found on
our Github wiki page.

### 2020-01-03 yfcc100m/paper/vldb2020/intro.tex
Add paper files

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning techniques to
extract information out of the visual data which can subsequently be used
for informed decision making.
The insights this information can provide depend on the
application: a retail vendor might be interested in the amount of time
shoppers spend in front of a specific product, while a medical expert might
want to see the effect of a specific treatment on the size of a tumor.

Despite this rich and varied usage environment, there has been very little
research on the management of visual data.
Most of the current storage solutions are
an ad-hoc collection of tools combined with custom scripts to tie them
together, unique not only to a specific discipline but often to individual
researchers. For example, consider an ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. This requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database. If the queries require
search over some patient information, then patients have to be associated
with their brain scans. Finally, if the ML pipeline needs images that
are of a size different than the stored ones, there is additional compute
diverted towards preprocessing after the potentially larger images are
fetched. All these steps require investigation of different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to evolve the schema on the fly.
Not only do these ad hoc solutions make replicating experiments
difficult, they do not scale well to real-world applications.
Addressing the storage and retrieval of visual data necessitates a complete
overhaul of the storage architecture,
preferably using emerging breakthroughs in
heterogeneous memory and storage for efficiency.

We present the Visual Data Management System (VDMS)\cite{darkside},
an Open Source project designed to enable efficient access of visual data.
Since visual data often contains
rich metadata (such as objects, locations, and time), VDMS stores this
information in a high performance graph database. Using this metadata, VDMS
can quickly identify which data is relevant to a given query.
Additionally, VDMS uses a custom library to store and retrieve visual data,
which provides an interface for machine friendly formats as well as
traditional formats. These new formats are designed to support applications
that are often interested in specific areas of images or videos,
particularly when the individual object is large.

While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
that can be used to store metadata~\cite{memsql, vertica}, and systems that
manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
be distinguished from them on the following aspects:

\begin{itemize}
\item {\em Design for analytics and machine learning}: by targeting
visual data for use cases that require manipulation
of visual information and associated metadata,
\item {\em Ease-of-use}: By defining a common API that allows applications to
combine their complex metadata searches with operations on resulting visual
data, and together with full support for feature vectors, VDMS goes beyond the
traditional SQL or OpenCV level interfaces that do one or the other. Given our
focus on enabling machine learning, we also provide a client API in Python.
\item {\em Performance}: We show how a unified system such as VDMS can
outperform an ad-hoc system constructed with well-known discrete components.
Because of the capabilities we have built into VDMS, it handles complex
queries significantly better than the ad-hoc system without compromising the
performance of simple queries.
\end{itemize}

### 2020-01-03 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Add paper files

% THIS IS AN EXAMPLE DOCUMENT FOR VLDB 2012
% based on ACM SIGPROC-SP.TEX VERSION 2.7
% Modified by  Gerald Weber <gerald@cs.auckland.ac.nz>
% Removed the requirement to include *bbl file in here. (AhmetSacan, Sep2012)
% Fixed the equation on page 3 to prevent line overflow. (AhmetSacan, Sep2012)

\documentclass{vldb}
\usepackage{graphicx}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{hyperref}       % hyperlinks
\usepackage{graphicx}
\usepackage{comment}
\usepackage{multicol}
\usepackage{framed}
\usepackage{subcaption}
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}

% Include information below and uncomment for camera ready
\vldbTitle{A Sample Proceedings of the VLDB Endowment Paper in LaTeX Format}
\vldbAuthors{Ben Trovato, G. K. M. Tobin, Lars Th{\sf{\o}}rv{$\ddot{\mbox{a}}$}ld, Lawrence P. Leipuner, Sean Fogarty, Charles Palmer, John Smith, Julius P.~Kumquat, and Ahmet Sacan}
\vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
\vldbVolume{12}
\vldbNumber{xxx}
\vldbYear{2019}

\begin{document}

% ****************** TITLE ****************************************

\title{Using VDMS to Search 100M Images: A Piece of Cake}

% possible, but not really needed or used for PVLDB:
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available 
% as\textit{Author's Guide to Preparing ACM SIG Proceedings Using 
% \LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}

% ****************** AUTHORS **************************************

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% Authors
\alignauthor
Luis Remis\titlenote{Most of the work was done while at Intel Labs.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{ApertureData Inc.}\\
       \email{luis@aperturedata.io}
\alignauthor
Chaunte Lacewell\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
       \affaddr{Intel Labs}\\
       \email{chaunte.w.lacewell@intel.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv\"{a}ld Group, {\texttt{jsmith@affiliation.org}}), Julius P.~Kumquat
%(The \raggedright{Kumquat} Consortium, {\small \texttt{jpkumquat@consortium.net}}), and Ahmet Sacan (Drexel University, {\small \texttt{ahmetdevel@gmail.com}})}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle

\begin{abstract}
We introduce the Visual Data Management System (VDMS),
which enables faster access to big-{\em visual}-data
and adds support to visual analytics.
This is achieved by searching for relevant visual data via metadata stored as
a graph, and enabling faster access to visual data through new
machine-friendly storage formats. VDMS differs from existing large scale photo
serving, video streaming, and textual big-data management systems due to
its primary focus on supporting machine learning and
data analytics pipelines that use visual data
(images, videos, and feature vectors), treating these as first class entities.
We describe how to use VDMS via its user friendly interface and how
it enables rich and efficient vision analytics through a machine learning
pipeline for processing medical images. We show the improved
performance of 2x in complex queries over a comparable set-up.
\end{abstract}

\input{intro.tex}
\input{design.tex}
%\input{demo.tex}
%\input{conclusion.tex}

% \balance

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{vdms-vldb2020}
% vldb_sample.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

\input{legal.tex}

\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references).

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
%pagebreak

\begin{appendix}
You can use an appendix for optional proofs or details of your evaluation which are not absolutely necessary to the core understanding of your paper.

\section{Final Thoughts on Good Layout}
Please use readable font sizes in the figures and graphs. Avoid tempering with the correct border values, and the spacing (and format) of both text and captions of the PVLDB format (e.g. captions are bold).

At the end, please check for an overall pleasant layout, e.g. by ensuring a readable and logical positioning of any floating figures and tables. Please also check for any line overflows, which are only allowed in extraordinary circumstances (such as wide formulas or URLs where a line wrap would be counterintuitive).

Use the \texttt{balance} package together with a \texttt{\char'134 balance} command at the end of your document to ensure that the last page has balanced (i.e. same length) columns.

\end{appendix}

\end{document}

### 2020-01-15 yfcc100m/paper/vldb2020/conclusion.tex
Add more figures and other minor additions

\section{Conclusion}

We introduced the Visual Data Management System,
designed to enable efficient access of visual data.
We presented our rich, JSON-based API, designed to simplify visual data access
for data scientists and analytics pipelines.
We compared the performance of our system to a
baseline of a combination of widely available and used systems.
Our findings showed that VDMS efficiently deals with complex queries,
providing a performance improvement
of up to 2x in the examined medical data search use-case.
Furthermore, VDMS requires significantly fewer
lines of code to execute complex queries in complex visual pipelines.
We intend to continue with the evaluation of
VDMS performance different use cases, and to identify more
opportunities to optimize and simplify the access of visual data
for machine learning and analytics applications.

### 2020-01-15 yfcc100m/paper/vldb2020/design.tex
Add more figures and other minor additions

@@
 \footnote{https://github.com/IntelLabs/\{vdms, pmgd, vcl\}}.
+
+\begin{figure}[]
+\centering
+\includegraphics[width=1\columnwidth]{figures/vdms_arch.pdf}
+\caption{VDMS Architecture}
+\label{fig:arch}
+\end{figure}
 
@@
 
-% \textbf{Client Library:}
-% A user application can use the VDMS API by defining metadata conforming to the
-% query protocol we have defined and passing along blobs. The client side VDMS
-% library provides a simple query function that accepts a JSON string with
-% commands and an array or vector of blobs.
-% Internally, the library wraps the query string and blob using Google Protobufs
-% and sends it to VDMS. It also receives a similarly wrapped response from VDMS
-% and returns it to the client. The responses require JSON parsing on the client
-% side for the metadata string that indicates how to interpret blobs.
-% VDMS provides Python and C++ client libraries with mainly
-% one function: a query that passes metadata as a JSON string and,
-% optionally, data as a vector of blobs. The VDMS client library manages the
-% rest.
+\textbf{Client Library:}
+A user application can use the VDMS API by defining metadata conforming to the
+query protocol we have defined and passing along blobs. The client side VDMS
+library provides a simple query function that accepts a JSON string with
+commands and an array or vector of blobs.
+Internally, the library wraps the query string and blob using Google Protobufs
+and sends it to VDMS. It also receives a similarly wrapped response from VDMS
+and returns it to the client. The responses require JSON parsing on the client
+side for the metadata string that indicates how to interpret blobs.
+VDMS provides Python and C++ client libraries with mainly
+one function: a query that passes metadata as a JSON string and,
+optionally, data as a vector of blobs. The VDMS client library manages the
+rest.
 

### 2020-01-15 yfcc100m/paper/vldb2020/eval.tex
Add more figures and other minor additions

\section{Evaluation}
\label{eval}

\subsection{YFCC100M Dataset}
\label{dataset}

Here goes an explanation of the dataset and all the source
(images, metadata, videos, features).

\subsection{Experimental Setup}

Here goes explanation of the comparison systems.
Need versions of MYSQL, OS, amount of memory and CPU info.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/comparison_system}
\caption{Comparison Systems}
\label{fig:systems}
\end{figure*}

\subsection{Images + Metadata}

Explanation of the evaluation and description of some of the queries.

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/q1_latency}
\caption{Query 1 - Latency}
\label{fig:q1_latency}
\end{figure}

\begin{figure}[]
\centering
\includegraphics[width=\columnwidth]{figures/q1_throughput}
\caption{Query 1 - Throughput}
\label{fig:q1_throughput}
\end{figure}

\subsection{Videos}

Explanation of the video queries and comparison.
Explain concurrency.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/video_overhead}
\caption{Concurrency and Overhead}
\label{fig:video}
\end{figure*}

\subsection{Feature Vectors}

Explain different engines supported, and how the features vectors
were ingested and handled.

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_alternatives}
\caption{Feature Vector Evaluation}
\label{fig:features_eval}
\end{figure*}

\begin{figure*}[]
\centering
\includegraphics[width=\textwidth]{figures/features_disksize}
\caption{Feature Collection Size in Disk}
\label{fig:features_size_does_matter}
\end{figure*}

### 2020-01-22 yfcc100m/paper/vldb2020/eval.tex
Added build time table, vdms/mysql edge/row tables, and initial build notes

@@
 \end{figure*}
+
+\begin{table}[h]
+\caption{Edges in VDMS database}
+\centering
+\begin{tabular}{c c c c}
+\hline\hline
+Database Size & No. Connections & MetaData & Tags\\
+\hline
+100k & 848,432 & 100,000 & 1,570\\
+500k & 4,249,500 & 500,000 & 1,570\\
+1M & 8,503,045 & 1,000,000 & 1,570\\
+5M & 42,505,478 & 5,000,000 & 1,570\\
+10M & 85,040,404 & 10,000,000 & 1,570\\
+50M & 425,162,070 & 50,000,000 & 1,570\\
+100M & 895,572,430 & 99,205,984 & 1,570\\
+\hline
+\end{tabular}
+\label{table:vdmsedges}
+\end{table}
+
+The MySQL setup was implemented in a similar manner.
+For each database, three tables were created: taglist, metadata, and autotags.
+The taglist table contains the unique YFCC tags where each entry is given an integer index named tagid.
+The metadata table contains the YFCC metadata where the id column is used as the primary index.
+The autotags table contains the probabilities for the tags associated with the metadata entries of the metadata table using id and tagid as indices.
+Each metadata entry has approximately nine tags; therefore, the number of rows in autotags is linearly related to the size of the MySQL database as shown in Table~\ref{table:mysqltables}.
+
+\begin{table}[h]
+\caption{Rows in MySQL database}
+\centering
+\begin{tabular}{c c c c}
+\hline\hline
+ & \multicolumn{3}{c}{Table}\\
+\cline{2-4}
+Database Size & Autotags & MetaData & Tag List\\
+\hline
+100k & 848,912 & 100,000 & 1,570\\
+500k & 4,241,200 & 498,707 & 1,570\\
+1M & 8,508,380 & 1,000,000 & 1,570\\
+5M & 42,425,905 & 4,987,379 & 1,570\\
+10M & 85,095,265 & 10,000,000 & 1,570\\
+50M & 425,446,208 & 50,000,000 & 1,570\\
+100M & 896,002,496 & 99,206,564 & 1,570\\
+\hline
+\end{tabular}
+\label{table:mysqltables}
+\end{table}
+
+Table~\ref{table:metadata} compares the time to build and the size of the YFCC datasets as VDMS and MySQL databases.
+On average, it takes MySQL 3.7x hours longer than VDMS to build the databases even though [insert relationship VDMS edges to mysql rows].
+As a graph database, VDMS requires approximately 1.4x more storage than MySQL to store each database.
+
+\begin{table*}[h]
+\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
+\centering
+\begin{tabular}{c c c c c}
+\hline\hline
+ & \multicolumn{2}{c}{Building Metadata Time} & \multicolumn{2}{c}{Metadata Size}\\
+\cline{2-3} \cline{4-5}%\hline
+Database Size & MySQL & VDMD & MySQL & VDMS\\
+\hline
+100k & 0.2 & 0.1 & 0.173 & 0.226\\
+500k & 1.1 & 0.4 & 0.805 & 1.1\\
+1M & 2.2 & 0.6 & 1.6 & 2.2\\
+5M & 11.7 & 3.0 & 7.8 & 11\\
+10M & 22.1 & 6.0 & 16 & 22\\
+50M & 147.1 & 24.2 & 78 & 109\\
+100M & 263.0 & 72.5 & 163 & 228\\
+\hline
+\end{tabular}
+\label{table:metadata}
+\end{table*}
 

### 2020-01-24 yfcc100m/paper/vldb2020/eval.tex
Update latests figures
Changed build time from table to fig, added info about yfcc database

--- 2020-01-15 yfcc100m/paper/vldb2020/eval.tex
@@
 
-Here goes an explanation of the dataset and all the source
-(images, metadata, videos, features).
+\textbf{\textit{Here goes an explanation of the dataset and all the source
+(images, metadata, videos, features).}}
+
+We use the Yahoo!
+Flickr Creative Commons 100m (YFCC100M) dataset ~\cite{Thomee_2016} for a large collection of 100 million public Flickr media objects.
+This dataset contains approximately 99.2 million images and 0.8 million videos along with respective metadata characterized by 25 fields such as the identifier (metaid), the user who created the media (image/video), the date the media was taken and uploaded, the longitude/latitude it was taken, and the device the media was captured from.
+The metadata also provides information including the URL to download the media, a title, a description of the media, user and machine tags, and license information.
+To understand the content of the media, the YFCC100M also provides autotags which is a set of comma-separated tags and confidence scores generated from 1,570 trained Caffe binary SVM classifiers.
 
@@
 
-Here goes explanation of the comparison systems.
-Need versions of MYSQL, OS, amount of memory and CPU info.
+\textbf{\textit{Here goes explanation of the comparison systems.
+Need versions of MySQL, OS, amount of memory and CPU info.}}
 
@@
 
+As a baseline, we implemented a similar database system, depicted in Figure~\ref{fig:systems}, with popular components: MySQL Server 5.7, Apache Web Server, and OpenCV.
+Each database system used the following configuration: Intel\textsuperscript{\textregistered} Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake) server running Ubuntu 16.04.
+
+For the experiments, we built VDMS and MySQL databases for the YFCC100M dataset for dataset sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+The VDMS Python client library was used to build queries to insert the necessary elements to build the database.
+In the database, we insert the edges for the 1570 autotags and the YFCC media objects with the appropriate metadata.
+To add the autotag confidence scores, we parse the autotag data and add a connection between the media object and tag edges with metadata denoting its appropriate confidence score.
+Table~\ref{table:vdmsedges} shows the number of elements in each database.
+
+\begin{table}[h]
+\caption{Edges in VDMS database}
+\centering
+\begin{tabular}{c c c c}
+\hline\hline
+Database Size & No. Connections & Metadata (Images) & Autotags\\
+\hline
+100k & 848,432 & 100,000 & 1,570\\
+500k & 4,249,500 & 500,000 & 1,570\\
+1M & 8,503,045 & 1,000,000 & 1,570\\
+5M & 42,505,478 & 5,000,000 & 1,570\\
+10M & 85,040,404 & 10,000,000 & 1,570\\
+50M & 425,162,070 & 50,000,000 & 1,570\\
+100M & 895,572,430 & 99,205,984 & 1,570\\
+\hline
+\end{tabular}
+\label{table:vdmsedges}
+\end{table}
+
+Alternatively, three tables were created for the MySQL system: taglist, metadata, and autotags.
+The number of rows in the MySQL databases are shown in Table~\ref{table:mysqltables}.
+The taglist table contains the 1570 YFCC100M autotags where each entry is given an integer key named tagid.
+The metadata table contains image metadata where the metadata identifier is the primary key.
+The autotags table contains the confidence scores for the generated autotags for metadata entries of the metadata table, the tagid associated with the autotag, and the metaid associated with the metadata identifier.
+
+\begin{table}[h]
+\caption{Rows in MySQL database}
+\centering
+\begin{tabular}{c c c c}
+\hline\hline
+ & \multicolumn{3}{c}{Table}\\
+\cline{2-4}
+Database Size & Autotags & MetaData & Tag List\\
+\hline
+100k & 848,912 & 100,000 & 1,570\\
+500k & 4,241,200 & 498,707 & 1,570\\
+1M & 8,508,380 & 1,000,000 & 1,570\\
+5M & 42,425,905 & 4,987,379 & 1,570\\
+10M & 85,095,265 & 10,000,000 & 1,570\\
+50M & 425,446,208 & 50,000,000 & 1,570\\
+100M & 896,002,496 & 99,206,564 & 1,570\\
+\hline
+\end{tabular}
+\label{table:mysqltables}
+\end{table}
+
+The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsedges} and ~\ref{table:mysqltables} but on average it takes MySQL 3.7x hours longer to build than VDMS.
+Figure~\ref{fig:db_time_size} compares the time to build and the size of the VDMS and MySQL databases.
+VDMS requires approximately 1.4x more storage than MySQL each database.
+This is due to the need to store relationship information for the graph database.
+
+\begin{figure}[]
+\centering
+\includegraphics[width=\columnwidth]{figures/db_time_size}
+\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
+\label{fig:db_time_size}
+\end{figure}
+
 \subsection{Images + Metadata}
 
-Explanation of the evaluation and description of some of the queries.
+\textbf{\textit{Explanation of the evaluation and description of some of the queries.}}
+
+\begin{figure}[]
+\centering
+\includegraphics[width=\columnwidth]{figures/concurrency_vdms}
+\caption{Concurrency Analisys - VDMS}
+\label{fig:concurrency_vdms}
+\end{figure}
 
@@
 
-\begin{figure}[]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=\columnwidth]{figures/q1_throughput}
-\caption{Query 1 - Throughput}
-\label{fig:q1_throughput}
+\includegraphics[width=\columnwidth]{figures/queries_throughput_32}
+\caption{Queries - Throughput - 32 Clients}
+\label{fig:q_throughput_32}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/queries_throughput_56}
+\caption{Queries - Throughput - 56 Clients}
+\label{fig:q_throughput_56}
 \end{figure}
@@
 
-Explanation of the video queries and comparison.
-Explain concurrency.
+\textbf{\textit{Explanation of the video queries and comparison.
+Explain concurrency.}}
 
@@
 
-Explain different engines supported, and how the features vectors
-were ingested and handled.
+\textbf{\textit{Explain different engines supported, and how the features vectors
+were ingested and handled.}}
 

### 2020-01-24 yfcc100m/paper/vldb2020/eval.tex
update db_time_size.pdf, edits to eval.tex

@@
 
-\textbf{\textit{Here goes an explanation of the dataset and all the source
-(images, metadata, videos, features).}}
-
 We use the Yahoo!
 Flickr Creative Commons 100m (YFCC100M) dataset ~\cite{Thomee_2016} for a large collection of 100 million public Flickr media objects.
-This dataset contains approximately 99.2 million images and 0.8 million videos along with respective metadata characterized by 25 fields such as the identifier (metaid), the user who created the media (image/video), the date the media was taken and uploaded, the longitude/latitude it was taken, and the device the media was captured from.
+This dataset contains approximately 99.2 million images and 0.8 million videos along with respective metadata characterized by 25 fields such as the identifier (metaid), the user who created the media (image/video), the date the media was taken and uploaded, the location in longitude/latitude coordinates, and the device the media was captured.
 The metadata also provides information including the URL to download the media, a title, a description of the media, user and machine tags, and license information.
@@
 \subsection{Experimental Setup}
-
-\textbf{\textit{Here goes explanation of the comparison systems.
-Need versions of MySQL, OS, amount of memory and CPU info.}}
 
@@
 
-As a baseline, we implemented a similar database system, depicted in Figure~\ref{fig:systems}, with popular components: MySQL Server 5.7, Apache Web Server, and OpenCV.
+As a baseline, we implemented a similar database system, depicted in Figure~\ref{fig:systems}, comprised of a combination of widely available and out-of-the-box components: MySQL Server 5.7, Apache Web Server, and OpenCV.
 Each database system used the following configuration: Intel\textsuperscript{\textregistered} Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake) server running Ubuntu 16.04.
 
-For the experiments, we built VDMS and MySQL databases for the YFCC100M dataset for dataset sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
-The VDMS Python client library was used to build queries to insert the necessary elements to build the database.
-In the database, we insert the edges for the 1570 autotags and the YFCC media objects with the appropriate metadata.
-To add the autotag confidence scores, we parse the autotag data and add a connection between the media object and tag edges with metadata denoting its appropriate confidence score.
-Table~\ref{table:vdmsedges} shows the number of elements in each database.
+For the experiments, we built VDMS and MySQL databases for the YFCC100M dataset sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+The VDMS Python client library was used to add the necessary elements to the database.
+In the database, we insert the nodes for the 1,570 autotags and the YFCC media objects with the appropriate metadata.
+To add the autotag confidence scores, we parse the autotag data and add a connection between the media object and tag nodes with metadata denoting its appropriate confidence score.
+Table~\ref{table:vdmsnodes} shows the number of elements in each database.
+
+Alternatively, three tables were created for the MySQL system: taglist, metadata, and autotags.
+The number of rows in the MySQL databases are shown in Table~\ref{table:mysqltables}.
+The taglist table contains the 1,570 autotags where each entry is given an integer key named tagid.
+The metadata table contains image metadata where the metadata identifier is the primary key.
+The autotags table contains the confidence scores for the generated autotags for metadata entries of the metadata table.
+This table also relates the other two tables and includes the tagid associated with the autotag and the metaid associated with the metadata identifier which are both keys to allow faster retrieval of data.
 
 \begin{table}[h]
-\caption{Edges in VDMS database}
+\caption{Nodes in VDMS database}
 \centering
@@
 \hline\hline
-Database Size & No. Connections & Metadata (Images) & Autotags\\
+Database Size & Connections & Images & Autotags\\
 \hline
@@
 \end{tabular}
-\label{table:vdmsedges}
+\label{table:vdmsnodes}
 \end{table}
-
-Alternatively, three tables were created for the MySQL system: taglist, metadata, and autotags.
-The number of rows in the MySQL databases are shown in Table~\ref{table:mysqltables}.
-The taglist table contains the 1570 YFCC100M autotags where each entry is given an integer key named tagid.
-The metadata table contains image metadata where the metadata identifier is the primary key.
-The autotags table contains the confidence scores for the generated autotags for metadata entries of the metadata table, the tagid associated with the autotag, and the metaid associated with the metadata identifier.
 
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsedges} and ~\ref{table:mysqltables} but on average it takes MySQL 3.7x hours longer to build than VDMS.
-Figure~\ref{fig:db_time_size} compares the time to build and the size of the VDMS and MySQL databases.
-VDMS requires approximately 1.4x more storage than MySQL each database.
-This is due to the need to store relationship information for the graph database.
+The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
+Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
+On average, it took MySQL 3.72x hours longer to build than VDMS.
+For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
+This is a difference of over a week of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
+Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store relationship information about each node/connection.
+This may become a factor if storage is a limitation but for the YFCC databases, VDMS required 30-41\% more storage than MySQL.
 
@@
 \includegraphics[width=\columnwidth]{figures/concurrency_vdms}
-\caption{Concurrency Analisys - VDMS}
+\caption{Concurrency Analysis - VDMS}
 \label{fig:concurrency_vdms}
@@
 \end{figure}
+
+Figure~\ref{fig:q1_latency} illustrates the performance of a simple query (find metadata/images for an alligator and resize the image to 224x224) over 15 iterations.
+In this case, the performance of VDMS is always better than MySQL and increases as the size of the database grows.
+For the 100M database, the VDMS metadata and image query times are almost \textbf{6x?} and \textbf{3x?} faster than MySQL, respectively.
 

### 2020-01-27 yfcc100m/paper/vldb2020/eval.tex
Update eval section

@@
 
-We use the Yahoo!
-Flickr Creative Commons 100m (YFCC100M) dataset ~\cite{Thomee_2016} for a large collection of 100 million public Flickr media objects.
-This dataset contains approximately 99.2 million images and 0.8 million videos along with respective metadata characterized by 25 fields such as the identifier (metaid), the user who created the media (image/video), the date the media was taken and uploaded, the location in longitude/latitude coordinates, and the device the media was captured.
-The metadata also provides information including the URL to download the media, a title, a description of the media, user and machine tags, and license information.
-To understand the content of the media, the YFCC100M also provides autotags which is a set of comma-separated tags and confidence scores generated from 1,570 trained Caffe binary SVM classifiers.
+The Yahoo!
+Flickr Creative Commons 100m (YFCC100M) dataset is a large collection of 100 million public Flickr media objects.
+It contains approximately 99.2 million images and 0.8 million videos with metadata characterized by 25 fields such as the identifier, the user, the date the media was taken/uploaded, the location in longitude/latitude coordinates, and the device the media was captured.
+The metadata also provides information including the URL to download the media object, a title, a description, user and auto-generated tags, and Creative Commons license information.
+The YFCC100M dataset also used a deep learning approach to generate autotags data which is a set of comma-separated concepts such as people, scenery, objects, and animals and confidence scores generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
 
@@
 
-For the experiments, we built VDMS and MySQL databases for the YFCC100M dataset sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
-The VDMS Python client library was used to add the necessary elements to the database.
-In the database, we insert the nodes for the 1,570 autotags and the YFCC media objects with the appropriate metadata.
-To add the autotag confidence scores, we parse the autotag data and add a connection between the media object and tag nodes with metadata denoting its appropriate confidence score.
-Table~\ref{table:vdmsnodes} shows the number of elements in each database.
-
-Alternatively, three tables were created for the MySQL system: taglist, metadata, and autotags.
-The number of rows in the MySQL databases are shown in Table~\ref{table:mysqltables}.
-The taglist table contains the 1,570 autotags where each entry is given an integer key named tagid.
-The metadata table contains image metadata where the metadata identifier is the primary key.
-The autotags table contains the confidence scores for the generated autotags for metadata entries of the metadata table.
-This table also relates the other two tables and includes the tagid associated with the autotag and the metaid associated with the metadata identifier which are both keys to allow faster retrieval of data.
+We built VDMS and MySQL databases using the YFCC100M dataset with incremental sample sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+For each dataset size, we created a PMGD graph from the YFCC100m metadata, the YFCC100m autotags associated with each metadata identifier, and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
+With simple queries, we insert 1,570 nodes for the autotags and insert nodes for the YFCC media objects with the appropriate metadata.
+Adding the autotag confidence scores is a recursive process and requires more complex queries than adding the previous nodes.
+For each metadata identifier in autotags, the query must find the associated metadata node and a tag node for an autotag listed for the identifier.
+A connection between the two nodes is created with the confidence score as a property.
+The number of metadata nodes are dependent on the database size and the connections are responsible for 90\% of the elements in each database as shown in Table~\ref{table:vdmsnodes}.
+It is important to note that the metadata identifier, autotags, and longitude/latitude coordinates are set as indexes of the database to allow faster retrieval of the metadata.
 
@@
 \end{table}
+
+Alternatively, each MySQL database is created from the same data using three tables: taglist, metadata, and autotags.
+By default, MySQL accounts for thread concurrency when processing requests but we increased the buffer pool size to handle the large amounts of data to avoid locking ~\cite{mysql}.
+Using a Python client and simple queries, the taglist table is read from the list of tags with an auto-incremented tagid as an index and the metadata table is read from the YFCC100m metadata using the identifier as an index.
+The autotags table contains the generated autotags and confidence scores for metadata entries of the metadata table.
+To generate the table, we split the autotags data for each database by the metadata identifier and autotag into new files.
+The new files are read into the respective autotags table with the metadata identifier and tagid as indexes.
 
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
-On average, it took MySQL 3.72x hours longer to build than VDMS.
-For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
-This is a difference of over a week of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
-Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store relationship information about each node/connection.
-This may become a factor if storage is a limitation but for the YFCC databases, VDMS required 30-41\% more storage than MySQL.
-
 \begin{figure}[]
@@
 
+The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
+Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
+A key difference in ingestion times is the low-level implementation of how MySQL reads data from the files and saves them to the tables versus VDMS ability to add components concurrently.
+In comparison, on average, it took MySQL 3.72x hours longer to build each database than VDMS.
+For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
+This is a difference of over 7 days of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
+Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store relationship information about each node/connection.
+This may become a factor if storage is a limitation but for the YFCC databases, VDMS required 30-41\% more storage than MySQL.
+
 \subsection{Images + Metadata}
@@
 \textbf{\textit{Explanation of the evaluation and description of some of the queries.}}
+
+To evaluate the access to metadata and images, we use the following four queries:
+\begin{enumerate}
+\item {\bf {\em 1tag}}: Find metadata/images of alligator.
+\item {\bf {\em 1tag\_resize}}: Find metadata/images of alligator and resize to 224x224.
+\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images of alligator, resize to 224x224, and in a particular geolocation (20 degrees radius).
+\item {\bf {\em 2tag\_resize}}: Find metadata/images of BOTH alligator and lake and resize to 224x224.
+\end{enumerate}
 
@@
 \end{figure}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/q1_latency}
-\caption{Query 1 - Latency}
-\label{fig:q1_latency}
-\end{figure}
-
-Figure~\ref{fig:q1_latency} illustrates the performance of a simple query (find metadata/images for an alligator and resize the image to 224x224) over 15 iterations.
-In this case, the performance of VDMS is always better than MySQL and increases as the size of the database grows.
-For the 100M database, the VDMS metadata and image query times are almost \textbf{6x?} and \textbf{3x?} faster than MySQL, respectively.
 
@@
 \label{fig:q_throughput_56}
+\end{figure}
+
+An image and metadata search on the YFCC100M databases are very extensive and would benefit from multithreading to process queries simultaneously.
+Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for the \textit{1tag} query using VDMS to investigate the number of concurrent clients to use for the per-query performance analysis.
+This figure shows the metadata performance plateaus at 32 clients and the image performance at 56 clients for a simple query.
+To further investigate, Figure~\ref{fig:q_throughput_32} and Figure~\ref{fig:q_throughput_56} illustrate the throughput performance for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients, respectively.
+On the 56-client experiment, a 10x difference occurs between VDMS and MySQL in metadata throughput while 32 clients is similar but not as high.
+The overall performance for both MySQL and VDMS is better with 56 clients; therefore, we used this value for our per-query analysis.
+
+\begin{figure}[]
+\centering
+\includegraphics[width=\columnwidth]{figures/q1_latency}
+\caption{Query 1 - Latency}
+\label{fig:q1_latency}
 \end{figure}

### 2020-02-01 yfcc100m/paper/vldb2020/eval.tex
Update eval and comparison system figure

@@
 
-The Yahoo!
-Flickr Creative Commons 100m (YFCC100M) dataset is a large collection of 100 million public Flickr media objects.
-It contains approximately 99.2 million images and 0.8 million videos with metadata characterized by 25 fields such as the identifier, the user, the date the media was taken/uploaded, the location in longitude/latitude coordinates, and the device the media was captured.
-The metadata also provides information including the URL to download the media object, a title, a description, user and auto-generated tags, and Creative Commons license information.
-The YFCC100M dataset also used a deep learning approach to generate autotags data which is a set of comma-separated concepts such as people, scenery, objects, and animals and confidence scores generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
+The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a
+large collection of 100 million public Flickr media objects.
+It contains approximately 99.2 million images and 0.8 million videos
+with metadata characterized by 25 fields such as the identifier,
+the user, the date the media was taken/uploaded,
+the location in longitude/latitude coordinates,
+and the device the media was captured.
+The metadata also provides information including the URL to
+download the media object, a title, a description, user and
+auto-generated tags, and Creative Commons license information.
+The YFCC100M dataset also used a deep learning approach to generate
+autotags data which is a set of comma-separated concepts such as people,
+scenery, objects, and animals and confidence scores
+generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
 
 \subsection{Experimental Setup}
+
+For all our experiments, we use 2 servers, one hosting VDMS server and
+another hosting MySQL server. Both servers have a dual-socket INTEL XEON XXXXX,
+each cpu with 28 physical cores with multithreading enabeled,
+for a total of 112 logical cores per server.
+The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
+has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
+We use MySQL version XXXXX.
+
+Figure \ref{figure:systems} shows a logical view of the difference between the
+interaction of the client application (which wants to retrieve metadata and
+imaged) with VDMS (left) and MySQL + Apache HTTP Server (right).
+It is worth noting that the images are stored in a shared repository
+(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
+and VDMS have direct access. In the case of MySQL, metadata is stored in an
+attached SSD disk. Even if VDMS has native support for Optane Persistent Memory,
+we do not use it in this experiment because of fairness of comparison w.r.t
+MySQL. The benefits of Persistent Memory on metadata operations is left
+for another paper, and outside the scope of this evaluation.
+For this experiment then, we simply use a similar attached SSD disk to store
+metadata (even if VDMS Graph store is designed for PM, it can deliver good
+performance when using SSDs directly).
 
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
-A key difference in ingestion times is the low-level implementation of how MySQL reads data from the files and saves them to the tables versus VDMS ability to add components concurrently.
-In comparison, on average, it took MySQL 3.72x hours longer to build each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
-This is a difference of over 7 days of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
-Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store relationship information about each node/connection.
-This may become a factor if storage is a limitation but for the YFCC databases, VDMS required 30-41\% more storage than MySQL.
+The VDMS and MySQL databases have comparable number of elements as shown
+in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but
+VDMS outperforms MySQL in build speed.
+Figure~\ref{fig:db_time_size} illustrates how VDMS can build
+databases faster than MySQL as the database size grows.
+A key difference in ingestion times is the low-level implementation of
+how MySQL reads data from the files and saves them to the tables
+versus VDMS ability to add components concurrently.
+In comparison, on average, it took MySQL 3.72x hours longer to build
+each database than VDMS.
+For example, to build the 100M database, MySQL took 263 hours
+while VDMS only needed 72.5 hours.
+This is a difference of over 7 days of processing data with
+VDMS missing less than 0.001\% images and 0.048\% connections
+from the entire dataset.
+Alternatively, VDMS requires more storage, shown in
+Figure~\ref{fig:db_time_size}, to store relationship information
+about each node/connection.
+This may become a factor if storage is a limitation but for the YFCC databases,
+VDMS required 30-41\% more storage than MySQL.
+
+% \begin{tabular}{c c c c c}
+% \hline\hline
+%  & \multicolumn{2}{c}{Building Metadata Time} & \multicolumn{2}{c}{Metadata Size}\\
+% \cline{2-3} \cline{4-5}%\hline
+% Database Size & MySQL & VDMD & MySQL & VDMS\\
+% \hline
+% 100k & 0.2   & 0.1  & 0.173 & 0.226\\
+% 500k & 1.1   & 0.4  & 0.805 & 1.1\\
+% 1M   & 2.2   & 0.6  & 1.6   & 2.2\\
+% 5M   & 11.7  & 3.0  & 7.8   & 11\\
+% 10M  & 22.1  & 6.0  & 16    & 22\\
+% 50M  & 147.1 & 24.2 & 78    & 109\\
+% 100M & 263.0 & 72.5 & 163   & 228\\
+% \hline
+% \end{tabular}
+% \label{table:metadata}
+% \end{table*}
 

### 2020-02-01 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Add more figures and other minor additions
Added build time table, vdms/mysql edge/row tables, and initial build notes
Update eval and comparison system figure

@@
 \alignauthor
-Chaunte Lacewell\\
+Chaunt\'e W. Lacewell\\
 %       \affaddr{Institute for Clarity in Documentation}\\
@@
 \begin{abstract}
-We introduce the Visual Data Management System (VDMS),
-which enables faster access to big-{\em visual}-data
-and adds support to visual analytics.
-This is achieved by searching for relevant visual data via metadata stored as
-a graph, and enabling faster access to visual data through new
-machine-friendly storage formats. VDMS differs from existing large scale photo
+Data scientists spend most of their time dealing with data preparation,
+rather than doing what they know best:
+build machine learning models and algorithms to solve problems.
+The problem grows as the scale of the data that is used for analysis or training
+grows. In this paper, we demonstrate how VDMS (Visual Data Management System)
+can be used to simplify the data preparation process and we show how we gain
+in efficient just because we are using a tool that was design for the job.
+To demonstrate this, we use the largest available public dataset (YFCC100M),
+with 99+ million images, 800+ thousand videos plus add-ons that include
+machine-generated tags and 4K dimensional feature vectors,
+a total of ± 13TB of data.
+VDMS differs from existing large scale photo
 serving, video streaming, and textual big-data management systems due to
@@
 data analytics pipelines that use visual data
-(images, videos, and feature vectors), treating these as first class entities.
-We describe how to use VDMS via its user friendly interface and how
-it enables rich and efficient vision analytics through a machine learning
-pipeline for processing medical images. We show the improved
-performance of 2x in complex queries over a comparable set-up.
+(images, videos, and feature vectors), treating these as first class citizens.
+We demonstrate how VDMS outperforms well-know and widely used tools for
+data access by a 10x factor in certain queries, while simplifying the process
+of data preparation and data access, and providing functionalities non-existent
+in alternative options.
 \end{abstract}
@@
 \input{design.tex}
+\input{eval.tex}
 %\input{demo.tex}
-%\input{conclusion.tex}
+\input{conclusion.tex}
 
 % \balance
+
+\clearpage
 

### 2020-02-01 yfcc100m/paper/vldb2020/eval.tex
Minor changes and reformating

@@
 scenery, objects, and animals and confidence scores
-generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
+generated from 1,570 trained Caffe classifiers \cite{Thomee_2016}.
+We have also used feature vectors generated for every image and first frame
+of every video \cite{features} to implement similarity search.
 
@@
 For all our experiments, we use 2 servers, one hosting VDMS server and
-another hosting MySQL server. Both servers have a dual-socket INTEL XEON XXXXX,
-each cpu with 28 physical cores with multithreading enabeled,
+another hosting MySQL server. Both servers have a dual-socket
+Intel\textsuperscript{\textregistered}
+Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
+each CPU with 28 physical cores with multithreading enabeled,
 for a total of 112 logical cores per server.
@@
 
+As a baseline, we implemented a similar database system comprised of a
+combination of widely available and out-of-the-box components: MySQL Server 5.7,
+Apache Web Server, and OpenCV.
 Figure \ref{figure:systems} shows a logical view of the difference between the
 interaction of the client application (which wants to retrieve metadata and
-imaged) with VDMS (left) and MySQL + Apache HTTP Server (right).
+imaged) with VDMS (left) and the baseline (right).
 It is worth noting that the images are stored in a shared repository
@@
 
-As a baseline, we implemented a similar database system, depicted in Figure~\ref{fig:systems}, comprised of a combination of widely available and out-of-the-box components: MySQL Server 5.7, Apache Web Server, and OpenCV.
-Each database system used the following configuration: Intel\textsuperscript{\textregistered} Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake) server running Ubuntu 16.04.
-
-We built VDMS and MySQL databases using the YFCC100M dataset with incremental sample sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
-For each dataset size, we created a PMGD graph from the YFCC100m metadata, the YFCC100m autotags associated with each metadata identifier, and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
-With simple queries, we insert 1,570 nodes for the autotags and insert nodes for the YFCC media objects with the appropriate metadata.
-Adding the autotag confidence scores is a recursive process and requires more complex queries than adding the previous nodes.
-For each metadata identifier in autotags, the query must find the associated metadata node and a tag node for an autotag listed for the identifier.
-A connection between the two nodes is created with the confidence score as a property.
-The number of metadata nodes are dependent on the database size and the connections are responsible for 90\% of the elements in each database as shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, autotags, and longitude/latitude coordinates are set as indexes of the database to allow faster retrieval of the metadata.
+We built VDMS and MySQL databases using the YFCC100M dataset with
+incremental sample sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+For each dataset size, we created a PMGD graph from the YFCC100m metadata,
+the YFCC100m autotags associated with each metadata identifier,
+and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
+With simple queries, we insert 1,570 nodes for the autotags and
+insert nodes for the YFCC media objects with the appropriate metadata.
+Adding the autotag confidence scores is a recursive process and
+requires more complex queries than adding the previous nodes.
+For each metadata identifier in autotags, the query must find
+the associated metadata node and a tag node for an autotag listed
+for the identifier.
+A connection between the two nodes is created with
+the confidence score as a property.
+The number of metadata nodes are dependent on the database
+size and the connections are responsible for 90\% of the elements
+in each database as shown in Table~\ref{table:vdmsnodes}.
+It is important to note that the metadata identifier, autotags,
+and longitude/latitude coordinates are set as indexes of the
+database to allow faster retrieval of the metadata.
 
@@
 
-Alternatively, each MySQL database is created from the same data using three tables: taglist, metadata, and autotags.
-By default, MySQL accounts for thread concurrency when processing requests but we increased the buffer pool size to handle the large amounts of data to avoid locking ~\cite{mysql}.
-Using a Python client and simple queries, the taglist table is read from the list of tags with an auto-incremented tagid as an index and the metadata table is read from the YFCC100m metadata using the identifier as an index.
-The autotags table contains the generated autotags and confidence scores for metadata entries of the metadata table.
-To generate the table, we split the autotags data for each database by the metadata identifier and autotag into new files.
-The new files are read into the respective autotags table with the metadata identifier and tagid as indexes.
+Alternatively, each MySQL database is created from the same data using three
+tables: taglist, metadata, and autotags.
+By default, MySQL accounts for thread concurrency when processing requests
+but we increased the buffer pool size to handle the large amounts
+of data to avoid locking ~\cite{mysql}.
+Using a Python client and simple queries, the taglist table is
+read from the list of tags with an auto-incremented tagid as an
+index and the metadata table is read from the YFCC100m metadata
+using the identifier as an index.
+The autotags table contains the generated autotags and confidence
+scores for metadata entries of the metadata table.
+To generate the table, we split the autotags data for each database
+by the metadata identifier and autotag into new files.
+The new files are read into the respective autotags table
+with the metadata identifier and tagid as indexes.
 
@@
 
-An image and metadata search on the YFCC100M databases are very extensive and would benefit from multithreading to process queries simultaneously.
-Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for the \textit{1tag} query using VDMS to investigate the number of concurrent clients to use for the per-query performance analysis.
-This figure shows the metadata performance plateaus at 32 clients and the image performance at 56 clients for a simple query.
-To further investigate, Figure~\ref{fig:q_throughput_32} and Figure~\ref{fig:q_throughput_56} illustrate the throughput performance for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients, respectively.
-On the 56-client experiment, a 10x difference occurs between VDMS and MySQL in metadata throughput while 32 clients is similar but not as high.
-The overall performance for both MySQL and VDMS is better with 56 clients; therefore, we used this value for our per-query analysis.
+An image and metadata search on the YFCC100M databases are very extensive
+and would benefit from multithreading to process queries simultaneously.
+Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for
+the \textit{1tag} query using VDMS to investigate the number of
+concurrent clients to use for the per-query performance analysis.
+This figure shows the metadata performance plateaus at 32 clients and
+the image performance at 56 clients for a simple query.
+To further investigate, Figure~\ref{fig:q_throughput_32} and
+Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
+for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
+respectively.
+On the 56-client experiment, a 10x difference occurs between VDMS and
+MySQL in metadata throughput while 32 clients is similar but not as high.
+The overall performance for both MySQL and VDMS is better with 56 clients;
+therefore, we used this value for our per-query analysis.
 

### 2020-02-01 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Minor changes and reformating

@@
 
-\numberofauthors{8} %  in this sample file, there are a *total*
+\numberofauthors{2} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 
+\section{Acknowledgments}
+
+We have to acknowledge Philip, Vishakha, Nilesh, and any other reviewer.
+
 % \balance
@@
 
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references).
+% \subsection{References}
+% Generated by bibtex from your ~.bib file.  Run latex,
+% then bibtex, then latex twice (to resolve references).
 

### 2020-02-03 yfcc100m/paper/vldb2020/design.tex
More refs and other additions

@@
-\section{Design and Implementation}
+\section{VDMS Desing and Implementation}
 \label{arch}
 
-VDMS implements a client-server design.
-The VDMS Server handles client requests concurrently and coordinates
-request execution across the metadata and data components in order to return a
-unified response. The metadata component is the Persistent Memory Graph
-Database (PMGD). The data component is our custom library, the
-Visual Compute Library (VCL). The VCL enables machine friendly enhancements to
-visual data.
-VDMS and its components are fully available as open source tools
-\footnote{https://github.com/IntelLabs/\{vdms, pmgd, vcl\}}.
+In this section, we briefly describe VDMS design and implementation,
+but most of it is already cover in previous work \cite{vdms-nips}.
+Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
+VDMS implements a client-server architecture that handles client
+requests concurrently and coordinates query execution across
+the metadata and data components in order to return a unified response.
+
+The metadata component is the Persistent Memory Graph
+Database (PMGD). The (visual) data component is our Visual Compute Module.
+The Visual Compute Module enables machine friendly enhancements to
+visual data, exposing high-level abstractions to the Request Server for
+dealing with a variety of images and video formats (through OpenCV),
+and different methods for indexing for Feature Vectors
+(including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
+VDMS and its components are fully available open source
+\footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
+We briefly describe each of the main components as follows:
 
@@
 \textbf{Persistent Memory Graph Database:}
-Recent developments in persistent memory technologies
-like 3D XPoint~\cite{IntelXPoint15}
-promise storage elements  providing  nearly  the  speed  of  DRAM  and  the
-durability of block-oriented storage. To provide an efficient storage
+We use PMGD to provide an efficient storage
 solution addressing the increasing popularity of connected data and
@@
 work that will be publish separately.
+PMGD has been design and optimized for persistent memory technologies
+like 3D XPoint~\cite{IntelXPoint15}, which
+promise storage providing nearly the speed of DRAM and the
+durability of block-oriented storage.
 
-\textbf{Visual Compute Library:} The VCL was designed to provide an interface
-through which users can interact with visual data. For traditional formats,
+\textbf{Visual Compute Module:} This module was designed to provide
+an abstraction layer for interacting with with visual data.
+For traditional formats (jpg, png, tiff, mp4, etc),
 the interface is an abstraction layer over OpenCV. However, it also provides a
@@
 visual analytics.
-The VCL currently provides limited support for videos but we are enhancing
-its capabilities as part of our future work.
+This module also provides support for videos, enabling operations like
+encoding/decoding/transcoding and resize as part of VDMS interface.
 Feature vector support is provided through an implementation based
@@
 
-\begin{comment}
-
-\begin{figure*}[t!]
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.92\linewidth]{figures/query}
-  \caption{Simple Metadata Query.}
-  \label{fig:sfig1}
-\end{subfigure}%
-\begin{subfigure}{.52\textwidth}
-  \centering
-  \includegraphics[width=.94\linewidth]{figures/query2}
-  \caption{Query involving visual transformations.}
-  \label{fig:sfig2}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-  \centering
-  \includegraphics[width=.9\linewidth]{figures/results}
-  \caption{Original image (left) and the result of query in (b)}
-  \label{fig:sfig3}
-\end{subfigure}
-\caption{Snapshot of a Python notebook with sample queries}
-\label{fig:query}
-\end{figure*}
-
-\end{comment}
-
 \textbf{Request Server:}
@@
 applications favor simpler interfaces to access and process data and cannot
-be expected to deal with two different ways of interacting with information (
-metadata and visual data) instead of focusing on the algorithmic parts of their
-pipelines.
+be expected to deal with two different ways of interacting with information
+(metadata and visual data) instead of focusing on the
+algorithmic parts of their pipelines.
 VDMS takes care of coordinating client requests across the metadata and the
@@
 commands and an array or vector of blobs.
-Internally, the library wraps the query string and blob using Google Protobufs
-and sends it to VDMS. It also receives a similarly wrapped response from VDMS
+Internally, the library wraps the query string and blob using
+Google Protobufs \cite{protobufs} and sends it to VDMS.
+It also receives a similarly wrapped response from VDMS
 and returns it to the client. The responses require JSON parsing on the client
@@
 
-\section{VDMS API}
-\label{arch}
-One of our goals with VDMS was to define an API that is easy to use.
-Our API explicitly predefines certain
+\textbf{VDMS API:}
+VDMS API is easy to use and explicitly predefines certain
 visual primitives associated with images, videos, and feature vectors. In
 addition, while we use a graph database to store our metadata, the API is not
-graph specific.
-We have paid particular attention to hide the complexities of our internal
+graph-specific.
+Authors have paid particular attention to hide the complexities of our internal
 implementation and up-level the API to a JSON-based
@@
 which is very popular across various application domains.
-We understand that by defining a new JSON-based API we had to trade-off
-expressiveness (compared to SPARQL or Gremlim) for the ability to design
-native support for visual data, but we believe that, as development of VDMS
-evolves, we will be able to achieve similar levels of expressiveness compared
-to more mature query languages.
-We have developed Python and C++ client
+By defining a new JSON-based API, there is a trade-off between
+expressiveness (compared to SPARQL or Gremlim, or even SQL) and
+the ability to natively support for visual data operations.
+However, it should be possible to achieve similar levels of
+expressiveness compared to more mature query languages over time.
+The current front-end available are a Python and C++ client
 library to provide a simple query function that accepts a JSON string with
@@
 
-\begin{comment}
-% All of the folliwng is commented out.
+While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
+that can be used to store metadata~\cite{memsql, vertica}, and systems that
+manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
+be distinguished from them on the following aspects:
 
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=0.9\columnwidth]{figures/features}
-\caption{Feature Vectors natively supported in VDMS}
-\label{fig:features}
-\end{figure}
-
-We use a medical imaging dataset as a driving example, and use it
-here to showcase our API. More details about this use case are discussed in
-Section~\ref{demo}.
-Figure~\ref{fig:query} shows a snapshot from a Jupyter Notebook with two
-simple queries.
-Figure~\ref{fig:sfig1} shows a simple metadata query
-where information about properties of patients matching some specific
-characteristics (admitted with 85 years of age or more)
-is used to constrain the search.
-The returned JSON structure shows 2 patients found in the database.
-Note that in this examples the queries are expressed as plain strings to
-help the reader understand its structure. In real applications, the queries
-are structured using the preferred method or library
-for handling JSON structures (e.g. Python dictionaries or jsoncpp in C++).
-
-Figure~\ref{fig:sfig2} shows a query involving visual transformations.
-In this case, the query is looking for an image with a certain \textit{id}
-value, and expects  VDMS to return the image twice: once with a thresholding
-transformation (e.g., zero out all pixels less than a threshold), and the
-second time after applying a threshold and then resizing it.
-\ref{fig:sfig3} shows the image originally inserted in VDMS (left),
-along with the returned images corresponding to the query in \ref{fig:sfig2}.
-Adding new operations to VDMS is easy, as our software architecture
-encapsulates all operations in VCL, and any new operation can use
-or wrap around OpenCV, which is used internally.
-
-\end{comment}
-% End of Comment
-
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities or visual objects.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-These vectors can be labeled and classified to build search indexes,
-and there are many in-memory libraries that are designed for
-this task~\cite{flann, faiss}.
-Using the VDMS API, users can manage feature vector indexes,
-query previously inserted elements (images),
-run a k-nearest neighbor search (knn) and express relationships
-between existing images or feature vectors and
-the newly inserted feature vectors.
-By natively supporting feature vectors and knn,
-VDMS allows out-of-the-box classification
-functionalities for many applications.
-Figure~\ref{fig:features} shows an example of how this functionality
-can be integrated in a medical imaging analytics pipeline:
-(1) A feature vector is extracted from an bounding
-box in a brain image and labeled;
-(2) the feature vector is inserted together
-with all the associated metadata (type of cancer, for example)
-and a link to the original image.
-(3) A new feature vector is extracted from a new image,
-(4) a query to VDMS is issued to classify that feature vector based on the
-indexed features in VDMS, and
-(5) VDMS can respond to the query with the label
-associated to that feature vector.
-Note that, even if the example is based on the medical images used for the
-performance evaluation, this methodology is applicable in many other contexts
-and use cases, such as face detection and matching.
-More details about the JSON API for this functionality can be found on
-our Github wiki page.
+\begin{itemize}
+\item {\em Design for analytics and machine learning}: by targeting
+visual data for use cases that require manipulation
+of visual information and associated metadata,
+\item {\em Ease-of-use}: By defining a common API that allows applications to
+combine their complex metadata searches with operations on resulting visual
+data, and together with full support for feature vectors, VDMS goes beyond the
+traditional SQL or OpenCV level interfaces that do one or the other. Given our
+focus on enabling machine learning, we also provide a client API in Python.
+\item {\em Performance}: We show how a unified system such as VDMS can
+outperform an ad-hoc system constructed with well-known discrete components.
+Because of the capabilities we have built into VDMS, it handles complex
+queries significantly better than the ad-hoc system without compromising the
+performance of simple queries.
+\end{itemize}

### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex
More refs and other additions

@@
 has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
-We use MySQL version XXXXX.
 
@@
 combination of widely available and out-of-the-box components: MySQL Server 5.7,
-Apache Web Server, and OpenCV.
+Apache Web Server 2.4.18, and OpenCV 3.3.
 Figure \ref{figure:systems} shows a logical view of the difference between the
@@
 
+Another key differentiating factor of VDMS is that it allows the creation of
+indexes for high-dimensional feature vectors and the insertion of
+these feature vectors associated with entities or visual objects.
+Feature vectors are intermediate results of various machine
+learning or computer vision algorithms when run on visual data.
+These vectors can be labeled and classified to build search indexes,
+and there are many in-memory libraries that are designed for
+this task~\cite{flann, faiss}.
+Using the VDMS API, users can manage feature vector indexes,
+query previously inserted elements (images),
+run a k-nearest neighbor search (knn) and express relationships
+between existing images or feature vectors and
+the newly inserted feature vectors.
+By natively supporting feature vectors and knn,
+VDMS allows out-of-the-box classification
+functionalities for many applications.
+
+% Figure~\ref{fig:features} shows an example of how this functionality
+% can be integrated in a medical imaging analytics pipeline:
+% (1) A feature vector is extracted from an bounding
+% box in a brain image and labeled;
+% (2) the feature vector is inserted together
+% with all the associated metadata (type of cancer, for example)
+% and a link to the original image.
+% (3) A new feature vector is extracted from a new image,
+% (4) a query to VDMS is issued to classify that feature vector based on the
+% indexed features in VDMS, and
+% (5) VDMS can respond to the query with the label
+% associated to that feature vector.
+% Note that, even if the example is based on the medical images used for the
+% performance evaluation, this methodology is applicable in many other contexts
+% and use cases, such as face detection and matching.
+% More details about the JSON API for this functionality can be found on
+% our Github wiki page.
+
 \textbf{\textit{Explain different engines supported, and how the features vectors

### 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
More refs and other additions

@@
 particularly when the individual object is large.
-
-While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
-that can be used to store metadata~\cite{memsql, vertica}, and systems that
-manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
-be distinguished from them on the following aspects:
-
-\begin{itemize}
-\item {\em Design for analytics and machine learning}: by targeting
-visual data for use cases that require manipulation
-of visual information and associated metadata,
-\item {\em Ease-of-use}: By defining a common API that allows applications to
-combine their complex metadata searches with operations on resulting visual
-data, and together with full support for feature vectors, VDMS goes beyond the
-traditional SQL or OpenCV level interfaces that do one or the other. Given our
-focus on enabling machine learning, we also provide a client API in Python.
-\item {\em Performance}: We show how a unified system such as VDMS can
-outperform an ad-hoc system constructed with well-known discrete components.
-Because of the capabilities we have built into VDMS, it handles complex
-queries significantly better than the ad-hoc system without compromising the
-performance of simple queries.
-\end{itemize}

### 2020-02-03 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
More refs and other additions

@@
 \usepackage{amsmath}
+\usepackage{mathtools}
 
@@
 can be used to simplify the data preparation process and we show how we gain
-in efficient just because we are using a tool that was design for the job.
-To demonstrate this, we use the largest available public dataset (YFCC100M),
+in efficiency just because we are using a tool that was design for the job.
+To demonstrate this, we use one of the largest available
+public dataset (YFCC100M),
 with 99+ million images, 800+ thousand videos plus add-ons that include
 machine-generated tags and 4K dimensional feature vectors,
-a total of ± 13TB of data.
+a total of $\sim$13TB of data.
 VDMS differs from existing large scale photo
@@
 We demonstrate how VDMS outperforms well-know and widely used tools for
-data access by a 10x factor in certain queries, while simplifying the process
+data access by a $\sim$10x factor in certain queries,
+while simplifying the process
 of data preparation and data access, and providing functionalities non-existent
@@
 
-We have to acknowledge Philip, Vishakha, Nilesh, and any other reviewer.
-
-% \balance
+We have to acknowledge Philip and Vishakha for help in fixing PMGD bug,
+Nilesh for his help/input, and any other reviewer.
 
@@
 
+\balance
 \begin{appendix}
@@
 
-\section{Final Thoughts on Good Layout}
-Please use readable font sizes in the figures and graphs.
+% \section{Final Thoughts on Good Layout}
+% Please use readable font sizes in the figures and graphs.
 Avoid tempering with the correct border values, and the spacing (and format) of both text and captions of the PVLDB format (e.g.
@@
 
-At the end, please check for an overall pleasant layout, e.g.
+% At the end, please check for an overall pleasant layout, e.g.
 by ensuring a readable and logical positioning of any floating figures and tables.
@@
 
-Use the \texttt{balance} package together with a \texttt{\char'134 balance} command at the end of your document to ensure that the last page has balanced (i.e. same length) columns.
+% Use the \texttt{balance} package together with a \texttt{\char'134 balance} command at the end of your document to ensure that the last page has balanced (i.e. same length) columns.
 

### 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
Update intro and other minor changes

@@
 extract information out of the visual data which can subsequently be used
-for informed decision making.
+for informed decision making \cite{vdms-nips}.
 The insights this information can provide depend on the
@@
 want to see the effect of a specific treatment on the size of a tumor.
+
+There are many applications that can be studied through the use of large
+and public available datasets. But before getting started, there is usually
+a large effort devoted to cleaning up the data. For this work, we use
+the YFCC100m dataset as an example given that we have use the same dataset
+for multiple proof of concepts and applications without our research lab.
+Applications include basic image search functionality (through the use
+of human-generated tags), advanced image search through the use of
+machine generated tags for each image, and video summarization.
+In all cases, it was needed to select a subset of the data before running
+any processing. Selecting subsets of data is by itself a time consuming task,
+as it involves loading all metadata into a solution that enables searched
+based on tags (relational database, graph database, csv files, etc), and
+building the necessary pipelines for querying and retrieving the right data.
+
+Data scientist and machine learning developers usually end up building ad-hoc
+solutions through a combination of databases and file systems to store metadata
+and visual data (images, videos), respectively.
 

### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex
Update intro and other minor changes
Add video explanations and other changes

@@
 has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
+The servers and client are connected through a 1GB wired link.
 
@@
 
-\textbf{\textit{Explanation of the video queries and comparison.
-Explain concurrency.}}
-
-\begin{figure*}[]
+\begin{figure*}[ht!]
 \centering
 \includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Concurrency and Overhead}
+\caption{Concurrency (left) and Overhead (right)}
 \label{fig:video}
 \end{figure*}
+
+VDMS provides full support for video storage and operations.
+This includes support for encoding/decoding and trans-coding of mp4, avi, and
+mov containers, as well as support for xvid, H.263 and H.264 encoders.
+This is supported through the Visual Compute Module that provides an abstraction
+layer on top of OpenCV (and ffmpeg directly).
+All the operations supported for images in VDMS are also supported at the
+video and frame level on the API. On top of that, there are a number of
+video-specific operations that are supported, such as the interval operations,
+that allow users to retrieve clips at different FPS versions of the video.
+
+All this functionality is provided and integrated with the rest of the
+metadata API as part of the comprehensive VDMS interface. This makes it
+possible for users to interact with metadata and video in a transactional
+manner, enabling users to run queries like: "Retrieve all the videos
+where there is a lake with probability higher than X, converting all videos
+to H.264 mp4 of size 224x244".
+In particular, this functionality was used internally to select a subset
+of videos with the right licenses for a video summarization application.
+
+To the best of our knowledge, there is no solution that can provide
+all the functionality mentioned above. Also, implementing a baseline is a very
+complex task as there is a large number of options and parameters that can
+be chosen, which makes it hard to accurately compare against VDMS functionality.
+For this reason, we chose to make a study using VDMS in various scenarios,
+and analysis what is the impact of having the overhead of VDMS' Request
+Server in the overall access time.
+
+Figure \ref{fig:video} shows the analysis of different queries aimed to retrieve
+a video using VDMS interface. We show how VDMS increases throughput of serving
+a video object as the number of simultaneous client increases, as well as the
+overhead VDMS introduces in the overall query execution time.
+The figure on the left compares the number of video transaction per second
+(i.e., number of videos returned per second) when different operations
+are executed as part of the transaction. The upper-bound of this would be
+simply returning the video as-is (without running any encoding/decoding or
+operation), represented by the red line. This query is the upper-limit because
+it essentially translates to reading the video from the file-system and sending
+it over a TCP/IP socket, without any other overhead or operations.
+
+We also run a set of other queries that involve: (a) running a resize operation
+on the video and, consequently, needs a decoding and
+encoding operations as well (blue line),
+(b) transcoding, meaning the use a different container and encoder
+than the one originally used (yellow line), and (c) both resize and transcoding.
+Note that the resize operation performs a downsize, which translates in less
+data being sent over the wire. This is specially noticeable when supporting 32
+simultaneous clients, where the systems provides more videos per second due to
+sending less data to the client, when compares to just transcoding and not resizing (yellow line).
+
+We can see that the system performs best when using all the physical cores.
+COMPLETE THIS BETTER.
+
+Because we see almost 3 orders of magnitude drop in performance when including
+operations as part of the query, we wanted to understand where most of the time
+was spent on the query, and optimize the Request Server and Visual Compute Module
+if necessary. For this, we run the experiment shown at
+Figure \ref{fig:video} (right) which breaks down the different components of the
+queries. This figure shows that more than 97\% of the query execution is spent
+on encoding/decoding operations, which is well-known to be a
+heavy and compute intensive operation.
+On the one hand, this results show that VDMS introduced overhead for
+video operation is minimal. On the other hand, this result means a
+limit on the opportunities for optimization for video queries given
+that biggest time factors are accounted by encoding/decoding, which is
+outside the scope of VDMS.
+This result also was the inspiration point for one optimization we included
+in future versions of VDMS, which involves using ffmpeg C++ API to
+limit the number of frames being encoding/decoding when possible.
 

### 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
Add video explanations and other changes

@@
 
-Data scientist and machine learning developers usually end up building ad-hoc
-solutions through a combination of databases and file systems to store metadata
-and visual data (images, videos), respectively.
-
 Despite this rich and varied usage environment, there has been very little
@@
 
+Data scientist and machine learning developers usually end up building ad-hoc
+solutions through a combination of databases and file systems to store metadata
+and visual data (images, videos), respectively.
+
+In this paper, we also expand on the Video and Feature Vector capabilities of
+VDMS, which are part of the latest additions to the system.
+We analyse different functionalities and trade-offs for this type of data,
+in combination with metadata filtering. To the best of out knowledge, this
+set of functionalities are unique to VDMS and we were not able to find a system
+with similar functionality.
+
+\begin{comment}
+
 We present the Visual Data Management System (VDMS)\cite{darkside},
@@
 particularly when the individual object is large.
+
+\end{comment}

### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex
Add Features figure with similarity search results

@@
 
-The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a
-large collection of 100 million public Flickr media objects.
-It contains approximately 99.2 million images and 0.8 million videos
-with metadata characterized by 25 fields such as the identifier,
-the user, the date the media was taken/uploaded,
-the location in longitude/latitude coordinates,
-and the device the media was captured.
-The metadata also provides information including the URL to
-download the media object, a title, a description, user and
-auto-generated tags, and Creative Commons license information.
-The YFCC100M dataset also used a deep learning approach to generate
-autotags data which is a set of comma-separated concepts such as people,
-scenery, objects, and animals and confidence scores
-generated from 1,570 trained Caffe classifiers \cite{Thomee_2016}.
+The Yahoo!
+Flickr Creative Commons 100m (YFCC100M) dataset is a large collection of 100 million public Flickr media objects created to provide free, sharable multimedia data for research.
+This dataset contains approximately 99.2 million images and 0.8 million videos with metadata characterized by 25 fields such as the identifier, user, date the media was taken/uploaded, location in longitude/latitude coordinates, device the media was captured, URL to download the media object, and Creative Commons license information.
+The YFCC100M dataset also used a deep learning approach to generate autotags which is a set of comma-separated concepts such as people, scenery, objects, and animals and confidence scores generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
 We have also used feature vectors generated for every image and first frame
@@
 Apache Web Server 2.4.18, and OpenCV 3.3.
-Figure \ref{figure:systems} shows a logical view of the difference between the
+Figure~\ref{fig:systems} shows a logical view of the difference between the
 interaction of the client application (which wants to retrieve metadata and
@@
 
-We built VDMS and MySQL databases using the YFCC100M dataset with
-incremental sample sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
-For each dataset size, we created a PMGD graph from the YFCC100m metadata,
-the YFCC100m autotags associated with each metadata identifier,
-and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
-With simple queries, we insert 1,570 nodes for the autotags and
-insert nodes for the YFCC media objects with the appropriate metadata.
-Adding the autotag confidence scores is a recursive process and
-requires more complex queries than adding the previous nodes.
-For each metadata identifier in autotags, the query must find
-the associated metadata node and a tag node for an autotag listed
-for the identifier.
-A connection between the two nodes is created with
-the confidence score as a property.
-The number of metadata nodes are dependent on the database
-size and the connections are responsible for 90\% of the elements
-in each database as shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, autotags,
-and longitude/latitude coordinates are set as indexes of the
-database to allow faster retrieval of the metadata.
+We built VDMS and MySQL databases using the YFCC100M dataset with incremental database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+For each database size, we created a PMGD graph from the YFCC100m metadata, the YFCC100m autotags associated with each metadata identifier, and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
+With simple queries, we insert 1,570 nodes for the autotags and insert nodes for the YFCC media objects with the appropriate metadata.
+Adding the autotag confidence scores is a recursive process and requires more complex queries than adding the previous nodes.
+For each metadata identifier in autotags, the query must find the associated metadata node and tag node.
+A connection between the two nodes is created with the confidence score as a property.
+The number of metadata nodes are dependent on the database size and the connections are responsible for 90\% of the elements in each database as shown in Table~\ref{table:vdmsnodes}.
+It is important to note that the metadata identifier, autotags, and longitude/latitude coordinates are set as indexes of the database to allow faster retrieval of the metadata.
 
@@
 
-Alternatively, each MySQL database is created from the same data using three
-tables: taglist, metadata, and autotags.
-By default, MySQL accounts for thread concurrency when processing requests
-but we increased the buffer pool size to handle the large amounts
-of data to avoid locking ~\cite{mysql}.
-Using a Python client and simple queries, the taglist table is
-read from the list of tags with an auto-incremented tagid as an
-index and the metadata table is read from the YFCC100m metadata
-using the identifier as an index.
-The autotags table contains the generated autotags and confidence
-scores for metadata entries of the metadata table.
-To generate the table, we split the autotags data for each database
-by the metadata identifier and autotag into new files.
-The new files are read into the respective autotags table
-with the metadata identifier and tagid as indexes.
+Each MySQL database is created in a similar manner but the data is represented as three tables: taglist, metadata, and autotags.
+By default, MySQL uses inifinite number of threads inside InnoDB when processing requests using four threads per IO read/write.
+In some cases, when creating large databases data locks may occur to protect the data from concurrent updates.
+To minimize locking in our runs, we increased the InnoDB buffer pool size to increase the amount of memory allocated to in-memory data structures ~\cite{mysql,mysql_blog}.
+Using a Python client and simple queries, the taglist table is read from the list of tags with an auto-incremented tagid as an index and the metadata table is read from the YFCC100m metadata using the identifier as an index.
+The autotags table contains the generated autotags and confidence scores for metadata entries of the metadata table.
+To generate the table, we split the autotags data for each database by the metadata identifier and autotag into new files.
+The new files are read into the respective autotags table with the metadata identifier and tagid as indexes.
 
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown
-in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but
-VDMS outperforms MySQL in build speed.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build
-databases faster than MySQL as the database size grows.
-A key difference in ingestion times is the low-level implementation of
-how MySQL reads data from the files and saves them to the tables
-versus VDMS ability to add components concurrently.
-In comparison, on average, it took MySQL 3.72x hours longer to build
-each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours
-while VDMS only needed 72.5 hours.
-This is a difference of over 7 days of processing data with
-VDMS missing less than 0.001\% images and 0.048\% connections
-from the entire dataset.
-Alternatively, VDMS requires more storage, shown in
-Figure~\ref{fig:db_time_size}, to store relationship information
-about each node/connection.
-This may become a factor if storage is a limitation but for the YFCC databases,
-VDMS required 30-41\% more storage than MySQL.
-
-% \begin{tabular}{c c c c c}
-% \hline\hline
-%  & \multicolumn{2}{c}{Building Metadata Time} & \multicolumn{2}{c}{Metadata Size}\\
-% \cline{2-3} \cline{4-5}%\hline
-% Database Size & MySQL & VDMD & MySQL & VDMS\\
-% \hline
-% 100k & 0.2   & 0.1  & 0.173 & 0.226\\
-% 500k & 1.1   & 0.4  & 0.805 & 1.1\\
-% 1M   & 2.2   & 0.6  & 1.6   & 2.2\\
-% 5M   & 11.7  & 3.0  & 7.8   & 11\\
-% 10M  & 22.1  & 6.0  & 16    & 22\\
-% 50M  & 147.1 & 24.2 & 78    & 109\\
-% 100M & 263.0 & 72.5 & 163   & 228\\
-% \hline
-% \end{tabular}
-% \label{table:metadata}
-% \end{table*}
+The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
+Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
+Key difference in the build times are contributed to the low-level implementation of how MySQL reads and stores data from the files and the optimizations (increased InnoDB pool size, etc.) needed to handle large datasets such as YFCC100m.
+On average, it took MySQL 3.72x hours longer to build each database than VDMS.
+For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
+This is a difference of over 7 days of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
+Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store information about each node/connection.
+This may become a factor if storage is a limitation.
+For our runs, VDMS required 30-41\% more storage than MySQL.
 
 \subsection{Images + Metadata}
-
-\textbf{\textit{Explanation of the evaluation and description of some of the queries.}}
 
@@
 
+WE SHOULD ADD HERE SAMPLES OF THE QUERYS TO SHOW HOW THESE ARE RETRIEVED.
+
 Because we see almost 3 orders of magnitude drop in performance when including
@@
 
+For this work, we wanted to
+
 % Figure~\ref{fig:features} shows an example of how this functionality
@@
 
-\textbf{\textit{Explain different engines supported, and how the features vectors
-were ingested and handled.}}
+\begin{figure*}[]
+\centering
+\includegraphics[width=\textwidth]{figures/feature_img_results}
+\caption{Sample Results of Similarity Search}
+\label{fig:similarity}
+\end{figure*}
 

### 2020-02-03 yfcc100m/paper/vldb2020/eval.tex
Update FeatureVector results

@@
 learning or computer vision algorithms when run on visual data.
-These vectors can be labeled and classified to build search indexes,
+Feature vectors are also known as "descriptors" or "visual descriptors".
+We use these terms interchangeably.
+These descriptors can be labeled and classified to build search indexes,
 and there are many in-memory libraries that are designed for
@@
 run a k-nearest neighbor search (knn) and express relationships
-between existing images or feature vectors and
-the newly inserted feature vectors.
-By natively supporting feature vectors and knn,
+between existing images or descriptors and
+the newly inserted descriptors.
+By natively supporting descriptors and knn,
 VDMS allows out-of-the-box classification
@@
 
-For this work, we wanted to
+For this work, and as part of a comprehensive image search implementation,
+we have used 4096-dimensional descriptors extracted from every image
+(and first frame of every video) and created a collection of these feature
+vectors in VDMS to be able to perform similarity search (i.e., find
+images that are "similar" to an query (input) image).
+"Similarity" in this particular case is defined as closeness
+in a an 4096-dimensional space using euclidean distance as the metric.
 
-% Figure~\ref{fig:features} shows an example of how this functionality
-% can be integrated in a medical imaging analytics pipeline:
-% (1) A feature vector is extracted from an bounding
-% box in a brain image and labeled;
-% (2) the feature vector is inserted together
-% with all the associated metadata (type of cancer, for example)
-% and a link to the original image.
-% (3) A new feature vector is extracted from a new image,
-% (4) a query to VDMS is issued to classify that feature vector based on the
-% indexed features in VDMS, and
-% (5) VDMS can respond to the query with the label
-% associated to that feature vector.
-% Note that, even if the example is based on the medical images used for the
-% performance evaluation, this methodology is applicable in many other contexts
-% and use cases, such as face detection and matching.
-% More details about the JSON API for this functionality can be found on
-% our Github wiki page.
+EXPLAIN HERE DIFFERENT ACCURACY/EXECUTIME TIME TRADEOFF.
+
+The process of loading descriptors in VDMS is simple.
+First, the user has to create a DescriptorsSet, using a single command.
+At creation of the DescriptorSet, the dimensionality of the descriptors
+is specified, together with the desired indexing method and the desired metric
+for computing distances (Euclidean Distance, L2, or Inner Product, IP).
+Once the DescriptorSet is created, descriptors can be inserted to the set.
+After the descriptors are inserted, similarity search can be performed.
+Note that descriptors can be inserted later on, and any search will reflect
+the existence of all descriptors in the set.
 
@@
 
+Figure \ref{fig:similarity} shows 3 examples of a query image (on the left),
+and images returned as "similar" by VDMS.
+The query input is a descriptor generated after a query image. The "query"
+descriptor is sent to VDMS as part of the query, and VDMS use that descriptor
+to find similar one, and retrieve the images associated with those "similar"
+descriptors. We used this as an example and as a visual validation of the
+functionality and applicability in this particular dataset, but we also
+provide an analytical approach to accuracy and trade-offs in our system.
+It is important to note that the accuracy of the results is entirely tied
+to the quality of the descriptors chosen by the applications.
+VDMS is completely agnostic to information contained within the descriptor,
+and simply offers the interface to store and index them, but the quality
+of the similarity result will be tied to the quality of descriptor extraction
+that the application is using.
+
 \begin{figure*}[]
@@
 \includegraphics[width=\textwidth]{figures/features_alternatives}
-\caption{Feature Vector Evaluation}
+\caption{Feature Vector Evaluation: Trade-off between query execution speed
+and accuracy of the results, using ground-truth data for computing accuracy.
+For this evaluation, we query the 10 closest neighbors (k = 10), and compute
+accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
+results that is present within the top 10 computed neighbors).
+We average the query execution time and accuracy for 100 queries (nq = 100).}
 \label{fig:features_eval}
 \end{figure*}
+
+As mentioned before, VDMS provides different levels of customization of the
+indexes created for a descriptor set, that includes the indexing techniques
+and the metric for similarity.
+These different indexing techniques comes with different trade-offs in terms
+of speed of search and accuracy.
+VDMS aims to provide functionality that is agnostics to application-specific
+techniques, enabling functionality that is general to visual data processing
+applications.
+Figure \ref{fig:features_eval} shows an analysis at the different indexing
+techniques provided by VDMS and its trade-off between accuracy and query
+execution speed, for a single threaded client.
+For this evaluation, we query the 10 closest neighbors (k = 10), and compute
+accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
+results that is present within the top 10 computed neighbors).
+We average the query execution time and accuracy for 100 queries (nq = 100).
+The "flat" index (red line) implements exact search and
+represents ground-truth, which explain why the accuracy is always 100\% on the
+right plot. The other indexes implement "approximate search", which trade-offs
+between accuracy and speed of search~\cite{flann, faiss}.
+We have also tried the "ivfflat" index (inverted file index), as well as
+LSH-based indexes using a different number of bits per descriptor.
+\footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
+Results show how "ivfflat" is the fastest option by comes at the trade-off
+of about 30\% lost in accuracy, while simple brute-force search
+is among the slowest options at the expenses of 100\% accuracy,
+meaning exact search.
 
@@
 \end{figure*}
+
+Another important trade-off to be made is w.r.t to space efficiency: Set of
+descriptors can grow very large and hard to manage.
+In this particular case, 4096-dimensional descriptors for 100M elements
+translates into 1TB of data, only in raw floating-point data alone (without
+accounting for any metadata or indexes associated with it).
+This component is very important on the overall analysis because a large set
+of feature vectors may not fit in memory and thus cause a pressure on the IO
+system while retrieving descriptors for computing distance, that severely impact
+the overall query execution time.
+Because of this, when the set of descriptors grows significantly large,
+it may be worth trading off accuracy to speed and space.
+Figure \ref{fig:features_size_does_matter} show the different indexes and
+their size in disk. These indexes already contain all the descriptors (or
+a quantized version of them), and can be loaded in memory directly when it fits.
+Note how, because of quantization of the descriptors,
+LSH provides a significantly lower space foot print, which can be a great option
+for large collections of descriptors when accuracy is not a main factor.

### 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
Update FeatureVector results

@@
 VDMS, which are part of the latest additions to the system.
-We analyse different functionalities and trade-offs for this type of data,
+We analyze different functionalities and trade-offs for this type of data,
 in combination with metadata filtering. To the best of out knowledge, this
@@
 with similar functionality.
+
+We show how VDMS can be used as the single and centralize point for data
+management and data access even when having multiple modalities of data:
+Metadata, Image, Videos, and Feature Vectors, all part of a comprehensive
+dataset that is used for multiple use-cases.
 

### 2020-02-10 yfcc100m/paper/vldb2020/eval.tex
Update final plots
Update fonts on figures

@@
 
-To evaluate the access to metadata and images, we use the following four queries:
+In order to evaluate VDMS and the baseline on certain queries, we
+implement 4 queries that would enable us to retrieve certain specific images.
+We choose these queries because they represent typical use-case where a
+cohort of images is to be retrieve and processed from a large corpus of data.
+We use the metadata associated to the images to filter said images.
+In particular, this dataset contains a set of around 1500 machine generated
+tags to each image.
+Tags includes things like objects (chair, person, people, etc), and
+concepts like places (indoor, outdoor, lake) and others (party, etc).
+Together with each tag, a probability is assign.
+This is, an image can have the tags people, person, party, outdoor, and each
+tag will be accompanied by a probability of that tag in that image.
+We represent that probability
+On average there are 8 tags assigned to each image.
+Also, some images have geo-location information (latitude/longitude).
+
+We use the tags (because they contain information about the content of
+the image), as well as the geo-location (as an example of searchable properties
+of the images).
+
+On top of that, we assume the use case of further processing of images
+through the use of ML, such as Convolutional Neural Networks TODO-CITE HERE.
+For this, we resize the images to 224x224, which is the input layer size for
+a popular neural network for object detection on images: ResNet TODO-CITEHERE.
+
+To evaluate the access to metadata and images,
+we use the following four queries:
 \begin{enumerate}
-\item {\bf {\em 1tag}}: Find metadata/images of alligator.
-\item {\bf {\em 1tag\_resize}}: Find metadata/images of alligator and resize to 224x224.
-\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images of alligator, resize to 224x224, and in a particular geolocation (20 degrees radius).
-\item {\bf {\em 2tag\_resize}}: Find metadata/images of BOTH alligator and lake and resize to 224x224.
+\item {\bf {\em 1tag}}: Find metadata/images with 1 specific tag (i.e. alligator, lake, etc).
+\item {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific tag and resize to 224x224.
+\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific tag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
+\item {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
 \end{enumerate}
+
+It is important to note that then querying for images with certain tags, we
+also apply a filter using the probability. To give an example, we only retrieve
+images with a tag "alligator" and a probability higher than 92\%.
+This probabilities are both present in VDMS (in the form of metadata present
+in the connection between the image and that tag), as well as in MySQL (in
+the form of a column of the table that links images with tags).
 
@@
 \centering
-\includegraphics[width=\columnwidth]{figures/concurrency_vdms}
-\caption{Concurrency Analysis - VDMS}
+\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
+\caption{Concurrency Analysis on Query \"1tag\_resize\"- VDMS vs Baseline}
 \label{fig:concurrency_vdms}

### 2020-02-12 yfcc100m/paper/vldb2020/eval.tex
Minor updates on eval

@@
 
-The Yahoo!
-Flickr Creative Commons 100m (YFCC100M) dataset is a large collection of 100 million public Flickr media objects created to provide free, sharable multimedia data for research.
-This dataset contains approximately 99.2 million images and 0.8 million videos with metadata characterized by 25 fields such as the identifier, user, date the media was taken/uploaded, location in longitude/latitude coordinates, device the media was captured, URL to download the media object, and Creative Commons license information.
-The YFCC100M dataset also used a deep learning approach to generate autotags which is a set of comma-separated concepts such as people, scenery, objects, and animals and confidence scores generated from 1,570 trained Caffe classifiers ~\cite{Thomee_2016}.
+The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
+collection of 100 million public Flickr media objects created to provide free,
+sharable multimedia data for research. This dataset contains approximately
+99.2 million images and 0.8 million videos with metadata characterized by
+25 fields such as the unique identifier, userid,
+date the media was taken/uploaded, location in longitude/latitude coordinates,
+device type the media was captured, URL to download the media object,
+and the Creative Commons license type information.
+The YFCC100M dataset also used a deep learning approach to generate autotags
+which is a set of comma-separated concepts such as people, scenery, objects,
+and animals and confidence scores generated from 1,570 trained
+Caffe classifiers ~\cite{Thomee_2016}. Together with the autotag, there is a
+probability associated to each tag to indicate the certainty level of the tag
+in that image/frame.
 We have also used feature vectors generated for every image and first frame
@@
 Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
-each CPU with 28 physical cores with multithreading enabeled,
+each CPU with 28 physical cores with multithreading enabled,
 for a total of 112 logical cores per server.
@@
 
-We built VDMS and MySQL databases using the YFCC100M dataset with incremental database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
-For each database size, we created a PMGD graph from the YFCC100m metadata, the YFCC100m autotags associated with each metadata identifier, and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
-With simple queries, we insert 1,570 nodes for the autotags and insert nodes for the YFCC media objects with the appropriate metadata.
-Adding the autotag confidence scores is a recursive process and requires more complex queries than adding the previous nodes.
-For each metadata identifier in autotags, the query must find the associated metadata node and tag node.
-A connection between the two nodes is created with the confidence score as a property.
-The number of metadata nodes are dependent on the database size and the connections are responsible for 90\% of the elements in each database as shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, autotags, and longitude/latitude coordinates are set as indexes of the database to allow faster retrieval of the metadata.
+We built VDMS and MySQL databases using the YFCC100M dataset with incremental
+database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+For each database size, we created a PMGD graph from the YFCC100m metadata,
+the YFCC100m autotags associated with each metadata identifier,
+and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
+With simple queries, we insert 1,570 nodes for the autotags and insert nodes
+for the YFCC media objects with the appropriate metadata.
+Adding the autotag confidence scores is a recursive process and requires more
+complex queries than adding the previous nodes.
+For each metadata identifier in autotags, the query must find the associated
+metadata node and tag node.
+A connection between the two nodes is created with the confidence score as a
+property. The number of metadata nodes are dependent on the database size and
+the connections are responsible for 90\% of the elements in each database as
+shown in Table~\ref{table:vdmsnodes}.
+It is important to note that the metadata identifier, autotags, and
+longitude/latitude coordinates are set as indexes of the database to allow
+faster retrieval of the metadata.
 
@@
 
-Each MySQL database is created in a similar manner but the data is represented as three tables: taglist, metadata, and autotags.
-By default, MySQL uses inifinite number of threads inside InnoDB when processing requests using four threads per IO read/write.
-In some cases, when creating large databases data locks may occur to protect the data from concurrent updates.
-To minimize locking in our runs, we increased the InnoDB buffer pool size to increase the amount of memory allocated to in-memory data structures ~\cite{mysql,mysql_blog}.
-Using a Python client and simple queries, the taglist table is read from the list of tags with an auto-incremented tagid as an index and the metadata table is read from the YFCC100m metadata using the identifier as an index.
-The autotags table contains the generated autotags and confidence scores for metadata entries of the metadata table.
-To generate the table, we split the autotags data for each database by the metadata identifier and autotag into new files.
-The new files are read into the respective autotags table with the metadata identifier and tagid as indexes.
+Each MySQL database is created in a similar manner but the data is represented
+as three tables: taglist, metadata, and autotags.
+By default, MySQL uses infinite number of threads inside InnoDB when processing
+requests using four threads per IO read/write.
+In some cases, when creating large databases data locks may occur to protect the
+data from concurrent updates.
+To minimize locking in our runs, we increased the InnoDB buffer pool size to
+increase the amount of memory allocated to in-memory data structures
+~\cite{mysql,mysql_blog}.
+Using a Python client and simple queries, the taglist table is read from the
+list of tags with an auto-incremented tagid as an index and the metadata table
+is read from the YFCC100m metadata using the identifier as an index.
+The autotags table contains the generated autotags and confidence scores for
+metadata entries of the metadata table.
+To generate the table, we split the autotags data for each database by the
+metadata identifier and autotag into new files.
+The new files are read into the respective autotags table with the metadata
+identifier and tagid as indexes.
 
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown in Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables} but VDMS outperforms MySQL in build speed.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases faster than MySQL as the database size grows.
-Key difference in the build times are contributed to the low-level implementation of how MySQL reads and stores data from the files and the optimizations (increased InnoDB pool size, etc.) needed to handle large datasets such as YFCC100m.
+The VDMS and MySQL databases have comparable number of elements as shown in
+Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables},
+but VDMS outperforms MySQL in build speed.
+Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
+faster than MySQL as the database size grows.
+Key difference in the build times are contributed to the low-level
+implementation of how MySQL reads and stores data from the files and the
+optimizations (increased InnoDB pool size, etc.) needed
+to handle large datasets such as YFCC100m.
 On average, it took MySQL 3.72x hours longer to build each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours while VDMS only needed 72.5 hours.
-This is a difference of over 7 days of processing data with VDMS missing less than 0.001\% images and 0.048\% connections from the entire dataset.
-Alternatively, VDMS requires more storage, shown in Figure~\ref{fig:db_time_size}, to store information about each node/connection.
-This may become a factor if storage is a limitation.
-For our runs, VDMS required 30-41\% more storage than MySQL.
+For example, to build the 100M database, MySQL took 263 hours
+while VDMS only needed 72.5 hours.
+This is a difference of over 7 days of processing data with VDMS missing less
+than 0.001\% images and 0.048\% connections from the entire dataset.
+
+Note that VDMS requires more storage for metadata, shown in
+Figure~\ref{fig:db_time_size}.
+This is space used to store information about each node/connection.
+The Graph Database internal to VDMS (called PMGD) was designed for performance,
+specially in environments where persistent memory is present.
+This design decision comes as a trade-off for storage footprint, which is
+noticeable in our results.
+VDMS required 30-41\% more storage than MySQL for storing the same amount
+of metadata.
+This may become a factor if storage is a limitation, but it should also be noted
+that even if we have a 41\% increase in metadata size,
+metadata accounts for less than 2\% of the overall database size.
+Take, for example, the largest database (both metadata and images) we built
+(100M): around 230GB of metadata and 12TB of images.
+On the other hand, in systems were persistent memory is a scarce resource,
+the increase storage foot print of PMGD may represent a challenge.
 

### 2020-02-12 yfcc100m/paper/vldb2020/eval.tex
Update eval

@@
 \label{eval}
+
+We have used the YFCC100M dataset to evaluate different aspects of our system.
+We use the images in the dataset and its associated metadata to implement
+an image-search engine based on properties associated to those images.
+This is a very common use-case we have encountered when attacking
+applications such as smart-retail, sports applications, and video summarization,
+where the starting point is a large set of data that must be curated
+before proceeding with the data processing and neural network training.
+In order to evaluate the different aspects of the performance on the
+image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
+the data search problem, which we describe in Section \ref{images}.
+
+We also cover evaluation we have done using the Video interface VDMS offers,
+in order to asses how the implementation performs and scale when used
+for video handling.
+When it comes to video, ad-hoc solutions have a large number of parameters that
+can be tunned, and complex ways to handle the data. Together with that,
+there is no system that enables transactional operations over videos files
+in the way VDMS does. Because of these two reasons, we focus our efforts in
+understanding variations in the the performance of VDMS and its scalability,
+rather than comparing it to a baseline that would no represent a fair comparison
+for either of the systems. We describe the results in Section \ref{videos}
+
+Similarly, we explore the behaviour of the feature vector functionality in VDMS,
+and an evaluation of the different trade-offs that the systems offers for
+applications developers. For this, we implemented an image-search application
+based on \textit{similarity} search, which we describe in Section \ref{features}
 
@@
 
-For all our experiments, we use 2 servers, one hosting VDMS server and
-another hosting MySQL server. Both servers have a dual-socket
-Intel\textsuperscript{\textregistered}
+As a baseline, and given that there are no other open-source system
+that implements similar functionality and interfaces as VDMS,
+we implemented an equivalent visual data management system comprised of a
+combination of widely available, off-the-shelve components:
+MySQL Server 5.7 (for storing metadata),
+Apache Web Server 2.4.18 (as interface for image access), and
+OpenCV 3.3 (to provide pre-processing operations on images).
+We have implemented a set of client-side applications that takes care
+of retrieving the components from the different systems, and applies
+pre-processing operations when needed.
+For visual data workloads, building an approach like the one
+implemented as a baseline for this work
+is the common practice in the industry \cite{heystack, tao}.
+
+For all our experiments, we use 2 servers, one hosting a VDMS server and
+another hosting a the Baseline implementation.
+Both servers have a dual-socket Intel\textsuperscript{\textregistered}
 Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
@@
 has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
-The servers and client are connected through a 1GB wired link.
+The client application (running the queries and measuring round-trip time)
+is connected to the server through a 1GB wired link through
+a 10GB-backplane switch.
 
-As a baseline, we implemented a similar database system comprised of a
-combination of widely available and out-of-the-box components: MySQL Server 5.7,
-Apache Web Server 2.4.18, and OpenCV 3.3.
 Figure~\ref{fig:systems} shows a logical view of the difference between the
@@
 imaged) with VDMS (left) and the baseline (right).
+The client application was implemented using Python 3 for both VDMS and
+the baseline.
+
 It is worth noting that the images are stored in a shared repository
 (ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
-and VDMS have direct access. In the case of MySQL, metadata is stored in an
-attached SSD disk. Even if VDMS has native support for Optane Persistent Memory,
+and VDMS have direct access. In the case of the baseline, metadata is
+stored in MySQL using an attached SSD disk.
+Even if VDMS has native support for Optane Persistent Memory,
 we do not use it in this experiment because of fairness of comparison w.r.t
-MySQL. The benefits of Persistent Memory on metadata operations is left
+MySQL (which was not design for Persistent Memory type of storage).
+The benefits of Persistent Memory on metadata operations is left
 for another paper, and outside the scope of this evaluation.
-For this experiment then, we simply use a similar attached SSD disk to store
-metadata (even if VDMS Graph store is designed for PM, it can deliver good
-performance when using SSDs directly).
+For this experiment then, in the case of VDMS we simply use a similar
+attached SSD disk to store metadata.
+Even if PMGD, the graph database used by VDMS, is designed for PM,
+it can deliver good performance when using SSDs directly.
 
@@
 
-We built VDMS and MySQL databases using the YFCC100M dataset with incremental
-database sizes 100k, 500k, 1M, 5M, 10M, 50M, and 100M.
+For the metadata, we built VDMS and MySQL databases
+using the YFCC100M dataset with incremental database sizes.
+For simplicity, we named the database based on the number of images
+it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
+The exact number of images/elements on each database is shown in
+Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
 For each database size, we created a PMGD graph from the YFCC100m metadata,
 the YFCC100m autotags associated with each metadata identifier,
-and a list of 1,570 autotags using 100 concurrent VDMS Python clients.
+and the list of 1,570 autotags using 100 concurrent VDMS Python clients.
 With simple queries, we insert 1,570 nodes for the autotags and insert nodes
@@
 \begin{table}[h]
-\caption{Nodes in VDMS database}
+\caption{VDMS Database - Number of Elements}
 \centering
@@
 \hline\hline
-Database Size & Connections & Images & Autotags\\
+Database Name & \# of Images & \# of Connections & \# of Autotags\\
 \hline
-100k & 848,432      & 100,000     & 1,570\\
-500k & 4,249,500    & 500,000     & 1,570\\
-1M   & 8,503,045    & 1,000,000   & 1,570\\
-5M   & 42,505,478   & 5,000,000   & 1,570\\
-10M  & 85,040,404   & 10,000,000  & 1,570\\
-50M  & 425,162,070  & 50,000,000  & 1,570\\
-100M & 895,572,430  & 99,205,984  & 1,570\\
+100k & 100,000     & 848,432      & 1,570\\
+500k & 500,000     & 4,249,500    & 1,570\\
+1M   & 1,000,000   & 8,503,045    & 1,570\\
+5M   & 5,000,000   & 42,505,478   & 1,570\\
+10M  & 10,000,000  & 85,040,404   & 1,570\\
+50M  & 50,000,000  & 425,162,070  & 1,570\\
+100M & 99,205,984  & 895,572,430  & 1,570\\
 \hline
@@
 Each MySQL database is created in a similar manner but the data is represented
-as three tables: taglist, metadata, and autotags.
+as three tables, following the relational model:
+1) \textit{metadata} table, contains one row per image,
+and a column for each property
+associated with the images (some of which are listed in \ref{dataset});
+2) \textit{taglist} table: contains one row per each autotag elements
+(always 1570 rows);
+3) \textit{autotags} table: contains on row per each autotag
+assigned to an image, each row containing a foreign key to the
+image and a foreign key to the tag, and
+the probability assigned to that tag belonging to that image.
+Given that there are 8 autotags, on average, per image, the \textit{autotags}
+table has around 8 times the number of rows present in the
+\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.
+
+THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
 By default, MySQL uses infinite number of threads inside InnoDB when processing
@@
 ~\cite{mysql,mysql_blog}.
+
 Using a Python client and simple queries, the taglist table is read from the
@@
 metadata entries of the metadata table.
+
 To generate the table, we split the autotags data for each database by the
@@
 \begin{table}[h]
-\caption{Rows in MySQL database}
+\caption{MySQL Database - Number of Elements}
 \centering
@@
 \cline{2-4}
-Database Size & Autotags & MetaData & Tag List\\
+Database Name & \# of Images & \# of Image Tags & Tag List\\
 \hline
-100k & 848,912     & 100,000    & 1,570\\
-500k & 4,241,200   & 498,707    & 1,570\\
-1M   & 8,508,380   & 1,000,000  & 1,570\\
-5M   & 42,425,905  & 4,987,379  & 1,570\\
-10M  & 85,095,265  & 10,000,000 & 1,570\\
-50M  & 425,446,208 & 50,000,000 & 1,570\\
-100M & 896,002,496 & 99,206,564 & 1,570\\
+100k & 100,000    & 848,912     & 1,570\\
+500k & 498,707    & 4,241,200   & 1,570\\
+1M   & 1,000,000  & 8,508,380   & 1,570\\
+5M   & 4,987,379  & 42,425,905  & 1,570\\
+10M  & 10,000,000 & 85,095,265  & 1,570\\
+50M  & 50,000,000 & 425,446,208 & 1,570\\
+100M & 99,206,564 & 896,002,496 & 1,570\\
 \hline
@@
 The VDMS and MySQL databases have comparable number of elements as shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables},
-but VDMS outperforms MySQL in build speed.
+Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
+can be attributed to failures in data preparation/loading because of
+incomplete/inconsistent formating, which is common on large datasets.
+The difference is less than 0.1\% in terms of number of elements (images and/or
+metadata information).
+
+\subsubsection{Database Loading Time}
+
+One of the first things we noted is the difference is loading times,
+where VDMS outperforms MySQL by a large margin.
 Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
-faster than MySQL as the database size grows.
-Key difference in the build times are contributed to the low-level
+faster than MySQL, and how the speedup is sustained as the database size grows.
+Key difference in the build times are attributed to the low-level
 implementation of how MySQL reads and stores data from the files and the
@@
 to handle large datasets such as YFCC100m.
-On average, it took MySQL 3.72x hours longer to build each database than VDMS.
+On average, it took MySQL 3.72x longer to build each database than VDMS.
 For example, to build the 100M database, MySQL took 263 hours
 while VDMS only needed 72.5 hours.
-This is a difference of over 7 days of processing data with VDMS missing less
-than 0.001\% images and 0.048\% connections from the entire dataset.
 
-Note that VDMS requires more storage for metadata, shown in
-Figure~\ref{fig:db_time_size}.
+\subsubsection{Database Storage Footprint}
+
+Another aspect important to note is that
+VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
 This is space used to store information about each node/connection.
@@
 
+%=========================================
+
 \subsection{Images + Metadata}
+\label{images}
 
@@
 \includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on Query \"1tag\_resize\"- VDMS vs Baseline}
+\caption{Concurrency Analysis on Query "1tag\_resize\- VDMS vs Baseline}
 \label{fig:concurrency_vdms}
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/queries_throughput_32}
-\caption{Queries - Throughput - 32 Clients}
-\label{fig:q_throughput_32}
-\end{figure}
+% We decided to remove this image because it is redudant with the concurrency
+% analysis.
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
+% \caption{Queries - Throughput - 32 Clients}
+% \label{fig:q_throughput_32}
+% \end{figure}
 
@@
 and would benefit from multithreading to process queries simultaneously.
-Figure~\ref{fig:concurrency_vdms} illustrates a concurrency analysis for
+Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
 the \textit{1tag} query using VDMS to investigate the number of
@@
 the image performance at 56 clients for a simple query.
-To further investigate, Figure~\ref{fig:q_throughput_32} and
-Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
-for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
-respectively.
+% To further investigate, Figure~\ref{fig:q_throughput_32} and
+% Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
+% for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
+% respectively.
 On the 56-client experiment, a 10x difference occurs between VDMS and
@@
 
+%=========================================
+
 \subsection{Videos}
-
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Concurrency (left) and Overhead (right)}
-\label{fig:video}
-\end{figure*}
+\label{videos}
 
@@
 Server in the overall access time.
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=\textwidth]{figures/video_overhead}
+\caption{Concurrency (left) and Overhead (right)}
+\label{fig:video}
+\end{figure*}
 
@@
 
+%=========================================
+
 \subsection{Feature Vectors}
+\label{features}
 
@@
 learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as "descriptors" or "visual descriptors".
-We use these terms interchangeably.
-These descriptors can be labeled and classified to build search indexes,
+Feature vectors are also known as \textit{descriptors}
+or \textit{visual descriptors}. We use these terms interchangeably.
+These descriptors can be classified and labeled and used to build search indexes,
 and there are many in-memory libraries that are designed for
@@
 query previously inserted elements (images),
-run a k-nearest neighbor search (knn) and express relationships
+run a k-nearest neighbor search (knn), and express relationships
 between existing images or descriptors and
@@
 By natively supporting descriptors and knn,
-VDMS allows out-of-the-box classification
-functionalities for many applications.
+VDMS allows out-of-the-box classification functionalities for many applications.
 

### 2020-02-12 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Update intro and other minor changes
Minor updates on eval
Add acks

@@
 
-We have to acknowledge Philip and Vishakha for help in fixing PMGD bug,
-Nilesh for his help/input, and any other reviewer.
+We would like to thank the many people that made this project possible and
+helped us through the process, as this work is the results of many efforts.
+We want to thank our managers Nilesh Jain and Ravi Iyer for their
+full support and input during the duration of the project.
+We want to specially acknowledge Philip Lantz and Vishakha Gupta for their help
+with PMGD, key to load large datasets into VDMS.
+We want to acknowledge the Intel Labs VDMS team for their efforts
+open-sourcing and maintaining the system.
+We want to thank Jason Gardner for his helping in setting up many of the servers
+and infrastructure needed to conduct our experiments.
 

### 2020-02-17 yfcc100m/paper/vldb2020/design.tex
Add video explanations and other changes
Varios modification to paper files

@@
-\section{VDMS Desing and Implementation}
+\section{VDMS Design \& Implementation}
 \label{arch}
 
-In this section, we briefly describe VDMS design and implementation,
-but most of it is already cover in previous work \cite{vdms-nips}.
+In this section, we briefly describe VDMS design principles and implementation,
+(most of it is already cover in previous work \cite{vdms-nips}).
 Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
@@
 
-The metadata component is the Persistent Memory Graph
-Database (PMGD). The (visual) data component is our Visual Compute Module.
+The metadata component is the \textit{Persistent Memory Graph
+Database} (PMGD). The (visual) data component is our Visual Compute Module.
 The Visual Compute Module enables machine friendly enhancements to
-visual data, exposing high-level abstractions to the Request Server for
+visual data, exposing high-level abstractions to the \textit{Request Server} for
 dealing with a variety of images and video formats (through OpenCV),
@@
 A user application can use the VDMS API by defining metadata conforming to the
-query protocol we have defined and passing along blobs. The client side VDMS
-library provides a simple query function that accepts a JSON string with
-commands and an array or vector of blobs.
+query protocol we have defined.
+The client side VDMS library provides a simple query function that
+accepts a JSON string with commands and an array or vector of blobs.
 Internally, the library wraps the query string and blob using
-Google Protobufs \cite{protobufs} and sends it to VDMS.
-It also receives a similarly wrapped response from VDMS
+Google Protobufs \cite{protobufs} and sends it to the VDMS server.
+It also receives a similarly formed response from VDMS
 and returns it to the client. The responses require JSON parsing on the client
-side for the metadata string that indicates how to interpret blobs.
-VDMS provides Python and C++ client libraries with mainly
-one function: a query that passes metadata as a JSON string and,
-optionally, data as a vector of blobs. The VDMS client library manages the
-rest.
+side for the metadata string that indicates how to interpret blobs field.
 
@@
 VDMS API is easy to use and explicitly predefines certain
-visual primitives associated with images, videos, and feature vectors. In
-addition, while we use a graph database to store our metadata, the API is not
-graph-specific.
+visual primitives associated with metadata, images, videos, and feature vectors.
+While we use a graph database to store our metadata,
+the API is not graph-specific.
 Authors have paid particular attention to hide the complexities of our internal
@@
 data, and together with full support for feature vectors, VDMS goes beyond the
-traditional SQL or OpenCV level interfaces that do one or the other. Given our
-focus on enabling machine learning, we also provide a client API in Python.
+traditional SQL or OpenCV level interfaces that do one or the other.
 \item {\em Performance}: We show how a unified system such as VDMS can

### 2020-02-17 yfcc100m/paper/vldb2020/eval.tex
Varios modification to paper files

@@
 there is no system that enables transactional operations over videos files
-in the way VDMS does. Because of these two reasons, we focus our efforts in
-understanding variations in the the performance of VDMS and its scalability,
-rather than comparing it to a baseline that would no represent a fair comparison
-for either of the systems. We describe the results in Section \ref{videos}
+in the way VDMS does. Because of the two reasons above mentioned,
+we focus our efforts in understanding variations in the the performance of
+VDMS and its scalability, rather than comparing it to a baseline that
+would no represent a fair comparison for either of the systems.
+We describe the results in Section \ref{videos}.
 
-Similarly, we explore the behaviour of the feature vector functionality in VDMS,
+Similarly, we explore the behavior of the feature vector functionality in VDMS,
 and an evaluation of the different trade-offs that the systems offers for
@@
 of every video \cite{features} to implement similarity search.
+Given that there is no standard benchmark oriented towards visual data queries,
+we have built a series of queries to filter this dataset that is model after
+our internal use cases for many of the mentioned applications we have worked
+with.
 
@@
 using the YFCC100M dataset with incremental database sizes.
-For simplicity, we named the database based on the number of images
+For simplicity, we named the database based on the approximate number of images
 it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
@@
 Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
-For each database size, we created a PMGD graph from the YFCC100m metadata,
-the YFCC100m autotags associated with each metadata identifier,
-and the list of 1,570 autotags using 100 concurrent VDMS Python clients.
-With simple queries, we insert 1,570 nodes for the autotags and insert nodes
-for the YFCC media objects with the appropriate metadata.
+For each database size, we created an instance of VDMS using YFCC100m metadata,
+the YFCC100m machine-generated \textit{autotags} associated with
+each metadata identifier, and the list of 1,570 \textit{autotags}
+using 100 concurrent VDMS Python clients.
+Internally, that information is represented as a property graph,
+where we have one node for each image, one node for each tag
+(always 1570 tags), and connections between images.
+For instance, if an image has 4 \textit{autotags} assigned,
+there will be 4 connections between that image and
+the different nodes for those \textit{autotags}.
+The information about the probability the \textit{autotag} present in an image
+is expressed as a property in the \textit{connection} between the 2 nodes.
+Figure \ref{fig:graph_representation} shows an example on two images and
+2 \textit{autotags}, and the \textit{connections} between those \textit{autotags}
+and the images.
+Image id: 23143252 has two \textit{autotags} assigned:
+\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
+Image id: 86756231, on the other hand, has a single \textit{autotags} assigned:
+\textit{Alligator} with probability 0.894.
+
+\begin{figure}[]
+\centering
+\includegraphics[width=\columnwidth]{figures/graph_representation}
+\caption{VDMS Data Representation Using a Property Graph:
+Example on two images and 2 \textit{autotags} with
+their respective probabilities expressed in the \textit{connection}.
+Image id: 23143252 has two \textit{autotags} assigned:
+\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
+Similarly, Image id: 86756231 has a single \textit{autotags} assigned:
+\textit{Alligator} with probability 0.894.}
+\label{fig:graph_representation}
+\end{figure}
+
+Because there are on average 8 tags assigned to each image,
+there will be around 8 times more connections than images, as can be seen
+in Table~\ref{table:vdmsnodes}.
+Also, each image node will contain multiple properties associated
+with that object.
+Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
+and then insert nodes for the YFCC media objects (images or videos)
+with the associated metadata.
 Adding the autotag confidence scores is a recursive process and requires more
 complex queries than adding the previous nodes.
-For each metadata identifier in autotags, the query must find the associated
-metadata node and tag node.
+For each metadata identifier in \textit{autotags},
+the query must find the associated metadata node and tag node.
 A connection between the two nodes is created with the confidence score as a
@@
 shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, autotags, and
+It is important to note that the metadata identifier, \textit{autotags}, and
 longitude/latitude coordinates are set as indexes of the database to allow
@@
 \hline\hline
-Database Name & \# of Images & \# of Connections & \# of Autotags\\
+DB Name & \# Images & \# Connections & \# Autotags\\
 \hline
@@
 as three tables, following the relational model:
-1) \textit{metadata} table, contains one row per image,
+1) \textit{images} table, contains one row per image,
 and a column for each property
@@
 
-Using a Python client and simple queries, the taglist table is read from the
-list of tags with an auto-incremented tagid as an index and the metadata table
-is read from the YFCC100m metadata using the identifier as an index.
-The autotags table contains the generated autotags and confidence scores for
-metadata entries of the metadata table.
+Using a Python client and simple queries, the \textit{taglist}
+table is read from the list of tags with an auto-incremented
+\textit{tagid} as a primary key and the metadata table
+is read from the YFCC100m metadata using the identifier as a primary key.
+The \textit{autotags} table contains the generated autotags and
+probabilities for entries of the \textit{images} table.
 
-To generate the table, we split the autotags data for each database by the
-metadata identifier and autotag into new files.
-The new files are read into the respective autotags table with the metadata
-identifier and tagid as indexes.
+To generate the table, we split the \textit{autotags} data for each database
+by the image identifier and autotag into new files.
+The new files are read into the \textit{autotags} table with the image
+identifier and \textit{tagid} as foreign keys.
 
 \begin{table}[h]
-\caption{MySQL Database - Number of Elements}
+\caption{MySQL Database - Number of Rows in each Table}
 \centering
@@
 \cline{2-4}
-Database Name & \# of Images & \# of Image Tags & Tag List\\
+DB Name & images & autotags & taglist\\
 \hline
@@
 
-One of the first things we noted is the difference is loading times,
+One of the first things we noticed is the difference is loading times,
 where VDMS outperforms MySQL by a large margin.
@@
 
-In order to evaluate VDMS and the baseline on certain queries, we
-implement 4 queries that would enable us to retrieve certain specific images.
+In order to evaluate VDMS and the baseline on our use-case queries,
+we implemented 4 queries that that filter and retrieve certain specific images.
 We choose these queries because they represent typical use-case where a
 cohort of images is to be retrieve and processed from a large corpus of data.
+As we mentioned before, we took this approach due to the lack of a standard
+benchmarks that are oriented towards visual data retrieval.
 We use the metadata associated to the images to filter said images.
-In particular, this dataset contains a set of around 1500 machine generated
-tags to each image.
-Tags includes things like objects (chair, person, people, etc), and
+In particular, this dataset contains a set of 1570 machine generated
+tags (\textit{autotags})assigned to the images depending on the content.
+\textit{autotags} includes things like objects (chair, person, people, etc), and
 concepts like places (indoor, outdoor, lake) and others (party, etc).
-Together with each tag, a probability is assign.
-This is, an image can have the tags people, person, party, outdoor, and each
-tag will be accompanied by a probability of that tag in that image.
-We represent that probability
-On average there are 8 tags assigned to each image.
+Together with each \textit{autotag}, a probability is assign.
+This is, an image can have the \textit{autotags} people, person, party,
+outdoor, and each \textit{autotag} assigned will be accompanied by a
+probability of that \textit{autotags} being present in that image.
+As mentioned earlier, there are 8 tags assigned to each image on average.
 Also, some images have geo-location information (latitude/longitude).
 
-We use the tags (because they contain information about the content of
-the image), as well as the geo-location (as an example of searchable properties
-of the images).
+We use the \textit{autotags} (as they contain information about
+the content of the image), as well as the geo-location
+(as an example of properties of the images used used for search and filtering).
 
-On top of that, we assume the use case of further processing of images
-through the use of ML, such as Convolutional Neural Networks TODO-CITE HERE.
+On top of that and for our use case, we would like to extract more information
+about the content of the image through the use of ML,
+such as Convolutional Neural Networks TODO-CITE HERE.
 For this, we resize the images to 224x224, which is the input layer size for
@@
 To evaluate the access to metadata and images,
-we use the following four queries:
+we use the following four queries, modeled after our internal use cases:
 \begin{enumerate}
-\item {\bf {\em 1tag}}: Find metadata/images with 1 specific tag (i.e. alligator, lake, etc).
-\item {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific tag and resize to 224x224.
-\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific tag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
+\item {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
+\item {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
+\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
 \item {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
@@
 
-It is important to note that then querying for images with certain tags, we
-also apply a filter using the probability. To give an example, we only retrieve
-images with a tag "alligator" and a probability higher than 92\%.
-This probabilities are both present in VDMS (in the form of metadata present
-in the connection between the image and that tag), as well as in MySQL (in
-the form of a column of the table that links images with tags).
+It is important to note that when querying for images with certain
+\textit{autotags}, we also apply a filter using the probability.
+For instance, we only retrieve images with a autotag \textit{alligator}
+and a probability higher than 92\%.
+This probabilities are both present in VDMS (in the form of a property
+of the \textit{connection} between the image and that \textit{autotag}),
+as well as in MySQL (in the form of a column in the \textit{autotags} table
+that links images with tags.
 
@@
 \includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on Query "1tag\_resize\- VDMS vs Baseline}
-\label{fig:concurrency_vdms}
+\caption{Concurrency Analysis on Query \textit{1tag\_resize} \- VDMS vs Baseline}
+\label{fig:concurrency_comparison}
 \end{figure}
@@
 
-An image and metadata search on the YFCC100M databases are very extensive
-and would benefit from multithreading to process queries simultaneously.
+Image search based on metadata is very expensive on large databases.
+Because of the large volume of data, the processing of the retrieve images
+is perform in parallel, using multi-core and/or distributed systems.
+For instance, a common implementation of an image processing pipeline
+would involve the use of distributed processing frameworks
+like Hadoop \cite{hadoop} or Spark \cite{spark}.
+Consequently, it is key that the data management system used supports
+concurrency, providing multiple workers with data in parallel.
+The ability to scale with the number of simultaneous clients is key for the
+applicability of visual data management systems like VDMS.
+
 Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
-the \textit{1tag} query using VDMS to investigate the number of
-concurrent clients to use for the per-query performance analysis.
+the \textit{1tag\_resize} query, described above, using both VDMS and
+the baseline to evaluate the scalability of both solutions, as the
+size of the databases grows and as the number of concurrent clients grows.
+
+It is important to note that the size of the result (number of images retrieved)
+is linear with the size of the database. This is, if a query returns 100 images
+for the 1M database, it will return around 1000 images for the 10M database.
+This poses a problem when evaluating performance as database increase,
+and clearly understanding the measurements.
+Because of this reason, we control the number of returned images for all the
+database using the probability of the \textit{autotags}, so that the queries
+in this experiment returns a similar number of images for all database sizes.
+Essentially, as the size of the database increase, we increase the probability
+threshold for the queries. We do this for both VDMS and the baseline, of course.
+
+The first thing to notice is that when it comes to retrieving metadata
+in the case of VDMS, (Figure ~\ref{fig:concurrency_comparison} (top),
+the size of the database (in terms of number of images) does not have
+a strong effect on the performance
+(measured in aggregated transactions per second).
+This means that even as database increase in size, there is no large increase
+in the query time.
+
+ the number of concurrent clients to use for the per-query performance analysis.
 This figure shows the metadata performance plateaus at 32 clients and
@@
 % respectively.
-On the 56-client experiment, a 10x difference occurs between VDMS and
-MySQL in metadata throughput while 32 clients is similar but not as high.
+On the 56-client experiment, a 10x difference can be seen between VDMS and
+baseline in metadata throughput while 32 clients is similar but not as high.
 The overall performance for both MySQL and VDMS is better with 56 clients;
@@
 VDMS provides full support for video storage and operations.
-This includes support for encoding/decoding and trans-coding of mp4, avi, and
-mov containers, as well as support for xvid, H.263 and H.264 encoders.
+This includes support for encoding/decoding and trans-coding of
+\textit{mp4}, \textit{avi}, and \textit{mov} containers,
+as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
 This is supported through the Visual Compute Module that provides an abstraction
-layer on top of OpenCV (and ffmpeg directly).
+layer on top of OpenCV (and \textit{ffmpeg} directly).
 All the operations supported for images in VDMS are also supported at the
@@
 video-specific operations that are supported, such as the interval operations,
-that allow users to retrieve clips at different FPS versions of the video.
+that allow users to retrieve clips at different
+frames-per-second (FPS) versions of the video.
 
@@
 where there is a lake with probability higher than X, converting all videos
-to H.264 mp4 of size 224x244".
+to \textit{H.264} \textit{mp4} of size 224x244".
 In particular, this functionality was used internally to select a subset
@@
 This result also was the inspiration point for one optimization we included
-in future versions of VDMS, which involves using ffmpeg C++ API to
+in future versions of VDMS, which involves using \textit{ffmpeg} C++ API to
 limit the number of frames being encoding/decoding when possible.
@@
 query previously inserted elements (images),
-run a k-nearest neighbor search (knn), and express relationships
+run a k-nearest neighbor search (\textit{knn}), and express relationships
 between existing images or descriptors and
 the newly inserted descriptors.
-By natively supporting descriptors and knn,
+By natively supporting descriptors and \textit{knn},
 VDMS allows out-of-the-box classification functionalities for many applications.
@@
 vectors in VDMS to be able to perform similarity search (i.e., find
-images that are "similar" to an query (input) image).
-"Similarity" in this particular case is defined as closeness
+images that are \textit{similar} to an query (input) image).
+\textit{Similarity} in this particular case is defined as closeness
 in a an 4096-dimensional space using euclidean distance as the metric.
@@
 is specified, together with the desired indexing method and the desired metric
-for computing distances (Euclidean Distance, L2, or Inner Product, IP).
+for computing distances (Euclidean Distance, \textit{L2},
+or Inner Product, \textit{IP}).
 Once the DescriptorSet is created, descriptors can be inserted to the set.
@@
 We average the query execution time and accuracy for 100 queries (nq = 100).
-The "flat" index (red line) implements exact search and
+The \textit{flat} index (red line) implements exact search and
 represents ground-truth, which explain why the accuracy is always 100\% on the
-right plot. The other indexes implement "approximate search", which trade-offs
-between accuracy and speed of search~\cite{flann, faiss}.
-We have also tried the "ivfflat" index (inverted file index), as well as
-LSH-based indexes using a different number of bits per descriptor.
+right plot. The other indexes implement \textit{approximate search},
+which trade-offs between accuracy and speed of search~\cite{flann, faiss}.
+We have also tried the \textit{ivfflat} index (inverted file index), as well as
+\textit{LSH}-based indexes using a different number of bits per descriptor
 \footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
-Results show how "ivfflat" is the fastest option by comes at the trade-off
+Results show how \textit{ivfflat} is the fastest option by comes at the trade-off
 of about 30\% lost in accuracy, while simple brute-force search
@@
 a quantized version of them), and can be loaded in memory directly when it fits.
-Note how, because of quantization of the descriptors,
-LSH provides a significantly lower space foot print, which can be a great option
-for large collections of descriptors when accuracy is not a main factor.
+Note how, because of quantization of the descriptors, \textit{LSH} provides a
+significantly lower space foot print, which can be a great option for
+large collections of descriptors when accuracy is not a main factor.

### 2020-02-17 yfcc100m/paper/vldb2020/intro.tex
Minor updates on eval
Varios modification to paper files

@@
 extract information out of the visual data which can subsequently be used
-for informed decision making \cite{vdms-nips}.
-The insights this information can provide depend on the
-application: a retail vendor might be interested in the amount of time
-shoppers spend in front of a specific product, while a medical expert might
-want to see the effect of a specific treatment on the size of a tumor.
+for informed decision making and analytics\cite{vdms-nips}.
+Visual data can provide important insights for various applications:
+a retail vendor might be interested in the amount of time
+shoppers spend in front of a specific product, a medical expert might
+want to study the effect of a specific treatment on the size of a tumor by
+looking at brain scans, and a media company would like to leverage
+thousands of hours of produced video material for different uses.
 
-There are many applications that can be studied through the use of large
-and public available datasets. But before getting started, there is usually
-a large effort devoted to cleaning up the data. For this work, we use
-the YFCC100m dataset as an example given that we have use the same dataset
-for multiple proof of concepts and applications without our research lab.
+Moreover, there are many applications that can be studied through
+the use of large and public available datasets.
+But before getting started, there is usually
+a large effort devoted to cleaning up the data.
+For this work, we use the YFCC100m dataset because of its comprehensive nature,
+and mainly because we have used it for multiple proof of concepts
+and applications within our research lab at Intel.
 Applications include basic image search functionality (through the use
 of human-generated tags), advanced image search through the use of
-machine generated tags for each image, and video summarization.
+machine generated tags for each image and feature vectors,
+and video summarization.
 In all cases, it was needed to select a subset of the data before running
 any processing. Selecting subsets of data is by itself a time consuming task,
-as it involves loading all metadata into a solution that enables searched
-based on tags (relational database, graph database, csv files, etc), and
-building the necessary pipelines for querying and retrieving the right data.
+as it involves loading all metadata into a solution that enables searches
+based on properties (some of the common tools used for this task are
+relational database, graph database, csv files, etc), and
+building the necessary pipelines and plug-ins
+for querying and retrieving the right data.
 
-Despite this rich and varied usage environment, there has been very little
-research on the management of visual data.
-Most of the current storage solutions are
-an ad-hoc collection of tools combined with custom scripts to tie them
-together, unique not only to a specific discipline but often to individual
-researchers. For example, consider an ML developer constructing a pipeline
+Despite the fact that visual data is becoming a primary component in analytics,
+there has been very little research on the management of visual data, even
+less that resulted in fully functioning and available tools.
+Most of the current storage solutions are an ad-hoc collection of
+tools combined with custom scripts to tie them together, unique not only to
+a specific discipline but often to individual researchers.
+For example, consider an ML developer constructing a pipeline
 for extracting brain tumor information from existing brain images in a
-classic medical imaging use case. This requires assigning consistent
+classic medical imaging use case. This task requires assigning consistent
 identifiers for the scans and adding their metadata in
-some form of relational or key-value database. If the queries require
-search over some patient information, then patients have to be associated
-with their brain scans. Finally, if the ML pipeline needs images that
-are of a size different than the stored ones, there is additional compute
-diverted towards preprocessing after the potentially larger images are
-fetched. All these steps require investigation of different software
-solutions that provide various functionalities that can then be stitched
-together with a script for this specific use case.
-Moreover, if the pipeline identifies
-new metadata to be added for the tumor images, most databases make it
-hard to evolve the schema on the fly.
+some form of relational or key-value database.
+If the queries require search over some patient information,
+then patients have to be associated with their brain scans images.
+Finally, if the ML pipeline needs images that are of a size different
+than the stored ones, there is additional compute diverted
+towards preprocessing after the potentially larger images are fetched.
+All these steps require investigation of different software
+solutions that provide various functionalities that have to be adapted,
+re-purposed, and stitched together with scripts for each specific use case.
+Moreover, most databases make it hard to evolve the schema on the fly and
+require a definition of the representation of the data beforehand, which becomes
+a problem when new forms of data becomes available with the advances on ML based
+data extraction techniques.
 Not only do these ad hoc solutions make replicating experiments
@@
 Addressing the storage and retrieval of visual data necessitates a complete
-overhaul of the storage architecture,
-preferably using emerging breakthroughs in
-heterogeneous memory and storage for efficiency.
+overhaul of the storage architecture, preferably using emerging breakthroughs in
+heterogeneous memory and storage for efficiency, and with a focus to recognizing
+types of data that becoming the center of analytics (i.e., images and videos).
 
-Data scientist and machine learning developers usually end up building ad-hoc
-solutions through a combination of databases and file systems to store metadata
-and visual data (images, videos), respectively.
+In this paper, we show how a new approach for visual data management can
+a very complex task become simple and productive.
+We also show how, buy using a tool that was design for the job,
+inefficiencies that emerge from the re-purpose of multiple system disappear
+and there are important gains in overall data access performance.
+We put a particular emphasis on the scalability and concurrenct aspect of the
+performance evaluation given that large datasets are processed in parallel
+using distributed computing.
 
-In this paper, we also expand on the Video and Feature Vector capabilities of
+We also expand on the Video and Feature Vector capabilities of
 VDMS, which are part of the latest additions to the system.
 We analyze different functionalities and trade-offs for this type of data,
-in combination with metadata filtering. To the best of out knowledge, this
-set of functionalities are unique to VDMS and we were not able to find a system
-with similar functionality.
+in combination with metadata filtering.
+To the best of our knowledge, this set of functionalities are unique to
+VDMS and we were not able to find a system with similar functionality.
 
@@
 
-\begin{comment}
-
-We present the Visual Data Management System (VDMS)\cite{darkside},
-an Open Source project designed to enable efficient access of visual data.
-Since visual data often contains
-rich metadata (such as objects, locations, and time), VDMS stores this
-information in a high performance graph database. Using this metadata, VDMS
-can quickly identify which data is relevant to a given query.
-Additionally, VDMS uses a custom library to store and retrieve visual data,
-which provides an interface for machine friendly formats as well as
-traditional formats. These new formats are designed to support applications
-that are often interested in specific areas of images or videos,
-particularly when the individual object is large.
-
-\end{comment}
+We start by briefly describing VDMS design principles and implementation
+in Section \ref{arch}.
+We then describe the dataset we used for our experiments and explain
+why this is a comprehensive use-case for visual data analytics, as well as
+describe our evaluation methodology in Section \ref{eval}.
+Finally, we show the results of our evaluation in Section \ref{eval}.

### 2020-02-17 yfcc100m/paper/vldb2020/legal.tex
Varios modification to paper files

\scriptsize
\begin{framed}

2020 Intel Corporation.

Intel and Intel Core are trademarks of Intel Corporation in
the U.S. and/or other countries. Other names and brands may be claimed as the property of others.

Performance results are based on testing as of 10/02/2020 and may not reflect
all publicly available security updates.
See configuration disclosure for details.
No product can be absolutely secure.

Software and workloads used in performance tests may have been optimized for
performance only on Intel microprocessors. Performance tests, such as SYSmark
and MobileMark, are measured using specific computer systems, components,
software, operations and functions.
Any change to any of those factors may cause the results to vary.
You should consult other information and performance tests
to assist you in fully evaluating your contemplated purchases, including the
performance of that product when combined with other products.
For more information go to http://www.intel.com/performance.

Intel's compilers may or may not optimize to the same degree for non-Intel
microprocessors for optimizations that are not unique to Intel microprocessors.
These optimizations include SSE2, SSE3, and SSSE3 instruction
sets and other optimizations. Intel does not guarantee the availability,
functionality, or effectiveness of any optimization on microprocessors not
manufactured by Intel. Microprocessor-dependent optimizations in this product
are intended for use with Intel microprocessors. Certain optimizations not
specific to Intel microarchitecture are reserved for Intel microprocessors.
Please refer to the applicable product User and Reference Guides for more
information regarding the specific instruction sets covered by this notice.
\end{framed}

### 2020-02-17 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Varios modification to paper files

@@
 % Include information below and uncomment for camera ready
-\vldbTitle{A Sample Proceedings of the VLDB Endowment Paper in LaTeX Format}
-\vldbAuthors{Ben Trovato, G. K. M. Tobin, Lars Th{\sf{\o}}rv{$\ddot{\mbox{a}}$}ld, Lawrence P. Leipuner, Sean Fogarty, Charles Palmer, John Smith, Julius P.~Kumquat, and Ahmet Sacan}
+\vldbTitle{Using VDMS to Search 100M Images: A Piece of Cake}
+\vldbAuthors{Luis Remis, Chaunt\'e W. Lacewell}
 \vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
@@
 \vldbNumber{xxx}
-\vldbYear{2019}
+\vldbYear{2020}
 
@@
 \alignauthor
-Luis Remis\titlenote{Most of the work was done while at Intel Labs.}\\
+Luis Remis\titlenote{Luis was a Research Scientist at Intel Labs until late 2019. Most of this work was done while at Intel Labs.}\\
 %       \affaddr{Institute for Clarity in Documentation}\\
@@
 The problem grows as the scale of the data that is used for analysis or training
-grows. In this paper, we demonstrate how VDMS (Visual Data Management System)
+grows. In this paper, we demonstrate how VDMS (the Visual Data Management System)
 can be used to simplify the data preparation process and we show how we gain
@@
 public dataset (YFCC100M),
-with 99+ million images, 800+ thousand videos plus add-ons that include
+with 99+ million images, 800+ thousand videos, plus add-ons that include
 machine-generated tags and 4K dimensional feature vectors,
-a total of $\sim$13TB of data.
+for a total of $\sim$13TB of data.
 VDMS differs from existing large scale photo
@@
 We demonstrate how VDMS outperforms well-know and widely used tools for
-data access by a $\sim$10x factor in certain queries,
-while simplifying the process
-of data preparation and data access, and providing functionalities non-existent
-in alternative options.
+data access by a $\sim$10x factor on this dataset most of the queries we
+used for our internal use-cases, and particularly at scale,
+while simplifying the process of data preparation and data access,
+and providing functionalities non-existent in alternative options.
 \end{abstract}
@@
 \input{eval.tex}
-%\input{demo.tex}
 \input{conclusion.tex}
@@
 
+\input{legal.tex}
+
 \clearpage
@@
 
-\input{legal.tex}
-
 % \subsection{References}
@@
 \balance
+
+\begin{comment}
 \begin{appendix}
@@
 \end{appendix}
+\end{comment}
 

### 2020-02-18 yfcc100m/paper/vldb2020/eval.tex
Update query figures

@@
 we use the following four queries, modeled after our internal use cases:
-\begin{enumerate}
-\item {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
-\item {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
-\item {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
-\item {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
-\end{enumerate}
+\begin{itemize}
+\item Query 1 - {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
+\item Query 2 - {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
+\item Query 3 - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
+\item Query 4 - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
+\end{itemize}
 
@@
 
-\begin{figure}[t!]
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=\columnwidth]{figures/queries_throughput_56}
+% \caption{Queries - Throughput - 56 Clients}
+% \label{fig:q_throughput_56}
+% \end{figure}
+
+\begin{figure*}[ht]
+\begin{subfigure}{.5\linewidth}
   \centering
-\includegraphics[width=\columnwidth]{figures/queries_throughput_56}
+  % include first image
+  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
+  \caption{Query 1 (red), and Query 2 (blue).}
+  \label{fig:q1_q2}
+\end{subfigure}
+\begin{subfigure}{.5\linewidth}
+  \centering
+  % include second image
+  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
+  \caption{Query 3 (red), and Query 4 (blue).}
+  \label{fig:q3_q4}
+\end{subfigure}
 \caption{Queries - Throughput - 56 Clients}
 \label{fig:q_throughput_56}
-\end{figure}
+\end{figure*}
 

### 2020-02-25 yfcc100m/paper/vldb2020/acks.tex
Update to overleaf version

--- 2020-02-12 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
@@
-% THIS IS AN EXAMPLE DOCUMENT FOR VLDB 2012
-% based on ACM SIGPROC-SP.TEX VERSION 2.7
-% Modified by  Gerald Weber <gerald@cs.auckland.ac.nz>
-% Removed the requirement to include *bbl file in here. (AhmetSacan, Sep2012)
-% Fixed the equation on page 3 to prevent line overflow. (AhmetSacan, Sep2012)
-
-\documentclass{vldb}
-\usepackage{graphicx}
-\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
-\usepackage{hyperref}       % hyperlinks
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{multicol}
-\usepackage{framed}
-\usepackage{subcaption}
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{amsmath}
-\usepackage{mathtools}
-
-% Include information below and uncomment for camera ready
-\vldbTitle{A Sample Proceedings of the VLDB Endowment Paper in LaTeX Format}
-\vldbAuthors{Ben Trovato, G. K. M. Tobin, Lars Th{\sf{\o}}rv{$\ddot{\mbox{a}}$}ld, Lawrence P. Leipuner, Sean Fogarty, Charles Palmer, John Smith, Julius P.~Kumquat, and Ahmet Sacan}
-\vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
-\vldbVolume{12}
-\vldbNumber{xxx}
-\vldbYear{2019}
-
-\begin{document}
-
-% ****************** TITLE ****************************************
-
-\title{Using VDMS to Search 100M Images: A Piece of Cake}
-
-% possible, but not really needed or used for PVLDB:
-%\subtitle{[Extended Abstract]
-%\titlenote{A full version of this paper is available
-% as\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-% \LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}
-
-% ****************** AUTHORS **************************************
-
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% Authors
-\alignauthor
-Luis Remis\titlenote{Most of the work was done while at Intel Labs.}\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{ApertureData Inc.}\\
-       \email{luis@aperturedata.io}
-\alignauthor
-Chaunt\'e W. Lacewell\\
-%       \affaddr{Institute for Clarity in Documentation}\\
-%       \affaddr{P.O. Box 1212}\\
-       \affaddr{Intel Labs}\\
-       \email{chaunte.w.lacewell@intel.com}
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-%\additionalauthors{Additional authors: John Smith (The Th{\o}rv\"{a}ld Group, {\texttt{jsmith@affiliation.org}}), Julius P.~Kumquat
-%(The \raggedright{Kumquat} Consortium, {\small \texttt{jpkumquat@consortium.net}}), and Ahmet Sacan (Drexel University, {\small \texttt{ahmetdevel@gmail.com}})}
-%\date{30 July 1999}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-
-\begin{abstract}
-Data scientists spend most of their time dealing with data preparation,
-rather than doing what they know best:
-build machine learning models and algorithms to solve problems.
-The problem grows as the scale of the data that is used for analysis or training
-grows. In this paper, we demonstrate how VDMS (Visual Data Management System)
-can be used to simplify the data preparation process and we show how we gain
-in efficiency just because we are using a tool that was design for the job.
-To demonstrate this, we use one of the largest available
-public dataset (YFCC100M),
-with 99+ million images, 800+ thousand videos plus add-ons that include
-machine-generated tags and 4K dimensional feature vectors,
-a total of $\sim$13TB of data.
-VDMS differs from existing large scale photo
-serving, video streaming, and textual big-data management systems due to
-its primary focus on supporting machine learning and
-data analytics pipelines that use visual data
-(images, videos, and feature vectors), treating these as first class citizens.
-We demonstrate how VDMS outperforms well-know and widely used tools for
-data access by a $\sim$10x factor in certain queries,
-while simplifying the process
-of data preparation and data access, and providing functionalities non-existent
-in alternative options.
-\end{abstract}
-
-\input{intro.tex}
-\input{design.tex}
-\input{eval.tex}
-%\input{demo.tex}
-\input{conclusion.tex}
-
 \section{Acknowledgments}
@@
 helped us through the process, as this work is the results of many efforts.
-We want to thank our managers Nilesh Jain and Ravi Iyer for their
+We want to specially thank our senior technologists Nilesh Jain and Ravi Iyer for their
 full support and input during the duration of the project.
@@
 with PMGD, key to load large datasets into VDMS.
+We want to thank Jim Blakley for his input during the various 
+phases of our project, and for advocating and promoting its use.
 We want to acknowledge the Intel Labs VDMS team for their efforts
@@
 and infrastructure needed to conduct our experiments.
-
-\clearpage
-
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{vdms-vldb2020}
-% vldb_sample.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-% and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-
-\input{legal.tex}
-
-% \subsection{References}
-% Generated by bibtex from your ~.bib file.  Run latex,
-% then bibtex, then latex twice (to resolve references).
-
-%APPENDIX is optional.
-% ****************** APPENDIX **************************************
-% Example of an appendix; typically would start on a new page
-%pagebreak
-
-\balance
-\begin{appendix}
-You can use an appendix for optional proofs or details of your evaluation which are not absolutely necessary to the core understanding of your paper.
-
-% \section{Final Thoughts on Good Layout}
-% Please use readable font sizes in the figures and graphs.
-Avoid tempering with the correct border values, and the spacing (and format) of both text and captions of the PVLDB format (e.g.
-captions are bold).
-
-% At the end, please check for an overall pleasant layout, e.g.
-by ensuring a readable and logical positioning of any floating figures and tables.
-Please also check for any line overflows, which are only allowed in extraordinary circumstances (such as wide formulas or URLs where a line wrap would be counterintuitive).
-
-% Use the \texttt{balance} package together with a \texttt{\char'134 balance} command at the end of your document to ensure that the last page has balanced (i.e. same length) columns.
-
-\end{appendix}
-
-\end{document}

### 2020-02-25 yfcc100m/paper/vldb2020/conclusion.tex
Update to overleaf version

@@
 
-We introduced the Visual Data Management System,
-designed to enable efficient access of visual data.
-We presented our rich, JSON-based API, designed to simplify visual data access
-for data scientists and analytics pipelines.
-We compared the performance of our system to a
-baseline of a combination of widely available and used systems.
-Our findings showed that VDMS efficiently deals with complex queries,
-providing a performance improvement
-of up to 2x in the examined medical data search use-case.
-Furthermore, VDMS requires significantly fewer
-lines of code to execute complex queries in complex visual pipelines.
-We intend to continue with the evaluation of
-VDMS performance different use cases, and to identify more
-opportunities to optimize and simplify the access of visual data
-for machine learning and analytics applications.
+In this paper, we show a comprehensive evaluation of VDMS 
+using one of the largest publicly available datasets: 
+The Yahoo Flickr Creative Commons 100M (YFCC100M), 
+together with the expansions packs that include 
+machine-generated labels and feature vectors.
+We compare VDMS against a combination of industry standard systems, 
+all of which are needed to replicate a portion of VDMS' 
+functionality in order to implement an Image Search application. 
+We see improvements up to 35x in certain queries, 
+and an average improvement of about 15x. 
+VDMS also supports video and 
+we provide an analysis of the performance of the video querying
+on a highly concurrent setting. 
+Finally, we analyze the different trade-offs of VDMS' 
+descriptors indexes, a functionality 
+that fully integrates with the rest of VDMS interface, 
+providing a powerful and comprehensive API.
+The design of VDMS, which was conceived as a 
+data management system that treats visual entities 
+as first class citizens, can remove inefficiencies 
+that result from re-purposing and combining solutions
+that were not designed for the job while providing 
+simpler and richer interfaces.

### 2020-02-25 yfcc100m/paper/vldb2020/design.tex
Update to overleaf version

@@
 
-In this section, we briefly describe VDMS design principles and implementation,
-(most of it is already cover in previous work \cite{vdms-nips}).
+In this section, we briefly describe VDMS design principles and implementation, which is already covered in previous work \cite{vdms-nips}.
 Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
 VDMS implements a client-server architecture that handles client
-requests concurrently and coordinates query execution across
+requests concurrently, and coordinates query execution across
 the metadata and data components in order to return a unified response.
@@
 The metadata component is the \textit{Persistent Memory Graph
-Database} (PMGD). The (visual) data component is our Visual Compute Module.
-The Visual Compute Module enables machine friendly enhancements to
-visual data, exposing high-level abstractions to the \textit{Request Server} for
-dealing with a variety of images and video formats (through OpenCV),
-and different methods for indexing for Feature Vectors
+Database} (PMGD) and the (visual) data component is our Visual Compute Module.
+The Visual Compute Module enables machine-friendly enhancements to
+visual data, exposing high-level abstractions to the \textit{Request Server}
+for dealing with a variety of images and video formats (through OpenCV),
+and different methods for indexing for feature vectors
 (including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
@@
 
-\begin{figure}[]
+\begin{figure}
 \centering
@@
 solution addressing the increasing popularity of connected data and
-applications that benefit from graph like processing, we have designed
+applications that benefit from graph-like processing. We have designed
 and implemented an in-persistent-memory graph database, PMGD, optimized
@@
 enhancements to existing metadata over time.
-The specific details on how the data structures are optimized for persistent
-memory, and the performance comparison with other graph databases is on-going
-work that will be publish separately.
-PMGD has been design and optimized for persistent memory technologies
-like 3D XPoint~\cite{IntelXPoint15}, which
+PMGD is designed and optimized for persistent memory technologies
+like Intel Optane~\cite{IntelXPoint15}, which
 promise storage providing nearly the speed of DRAM and the
@@
 \textbf{Visual Compute Module:} This module was designed to provide
-an abstraction layer for interacting with with visual data.
-For traditional formats (jpg, png, tiff, mp4, etc),
+an abstraction layer for interacting with visual data.
+For traditional formats (jpg, png, tiff, mp4, etc.),
 the interface is an abstraction layer over OpenCV. However, it also provides a
@@
 This module also provides support for videos, enabling operations like
-encoding/decoding/transcoding and resize as part of VDMS interface.
+encoding/decoding/transcoding and resize as part of the VDMS interface.
 Feature vector support is provided through an implementation based
-on high-dimensional sparse arrays, also using TileDB, which contains both
-storage and search functionality over feature vectors.
-In addition, the VCL provides a wrapper
+on high-dimensional sparse arrays, also using TileDB.
+In addition, the Visual Compute Library provides a wrapper
 for another high-dimensional index implementation,
@@
 Developers and users of machine learning frameworks and data science
-applications favor simpler interfaces to access and process data and cannot
-be expected to deal with two different ways of interacting with information
+applications favor simpler interfaces to access and process data. They cannot be expected to deal with two different ways of interacting with information
 (metadata and visual data) instead of focusing on the
@@
 VDMS takes care of coordinating client requests across the metadata and the
-data as well as efficiently manages multiple clients through its Request
-Server component, by implementing a JSON-based API.
+data, and managing multiple clients through its Request
+Server component by implementing a JSON-based API.
 It decomposes the command into
@@
 and returns a coherent response to the user after applying any additional
-operations (explained in the next section).
+operations.
 
@@
 query protocol we have defined.
-The client side VDMS library provides a simple query function that
+The client side of the VDMS library provides a simple query function that
 accepts a JSON string with commands and an array or vector of blobs.
@@
 and returns it to the client. The responses require JSON parsing on the client
-side for the metadata string that indicates how to interpret blobs field.
+side for the metadata string that indicates how to interpret the blobs field.
 
 \textbf{VDMS API:}
-VDMS API is easy to use and explicitly predefines certain
+VDMS API is easy to use and explicitly pre-defines certain
 visual primitives associated with metadata, images, videos, and feature vectors.
@@
 expressiveness (compared to SPARQL or Gremlim, or even SQL) and
-the ability to natively support for visual data operations.
+the ability to natively support visual data operations.
 However, it should be possible to achieve similar levels of
 expressiveness compared to more mature query languages over time.
-The current front-end available are a Python and C++ client
+The current front-ends available are a Python and C++ client
 library to provide a simple query function that accepts a JSON string with
-commands and an array or vector of blobs.
+commands and an array (or vector) of blobs.
 
@@
 \begin{itemize}
-\item {\em Design for analytics and machine learning}: by targeting
+\item {\em Design for analytics and machine learning}: By targeting
 visual data for use cases that require manipulation
@@
 combine their complex metadata searches with operations on resulting visual
-data, and together with full support for feature vectors, VDMS goes beyond the
+data, and together with full support for feature vectors. VDMS goes beyond the
 traditional SQL or OpenCV level interfaces that do one or the other.

### 2020-02-25 yfcc100m/paper/vldb2020/eval.tex
Update to overleaf version

@@
 We use the images in the dataset and its associated metadata to implement
-an image-search engine based on properties associated to those images.
-This is a very common use-case we have encountered when attacking
+an image-search engine based on properties associated with those images.
+This is a very common use-case we have encountered when building
 applications such as smart-retail, sports applications, and video summarization,
 where the starting point is a large set of data that must be curated
-before proceeding with the data processing and neural network training.
+before proceeding with the data processing (such as neural network training).
 In order to evaluate the different aspects of the performance on the
-image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
-the data search problem, which we describe in Section \ref{images}.
+image search, we have built a baseline following the methodology
+used in the industry, and following what we have done in the past
+in order to solve the data search problem,
+which we describe in Section \ref{images}.
 
-We also cover evaluation we have done using the Video interface VDMS offers,
-in order to asses how the implementation performs and scale when used
-for video handling.
+We also evaluate the performance and scalability of the video management
+capabilities offered by VDMS.
+Handling video in a general way is a complex task.
+Some of this complexity comes from the existence of a variety of open and
+proprietary implementations, different encoding techniques and container
+formats, and different parameters of the video itself that are
+application-dependent, like frames per second, lossy compression, etc).
 When it comes to video, ad-hoc solutions have a large number of parameters that
-can be tunned, and complex ways to handle the data. Together with that,
-there is no system that enables transactional operations over videos files
-in the way VDMS does. Because of the two reasons above mentioned,
-we focus our efforts in understanding variations in the the performance of
-VDMS and its scalability, rather than comparing it to a baseline that
-would no represent a fair comparison for either of the systems.
+can be tuned.
+Together with that, there is no system that enables transactional
+operations over videos files in the way VDMS does.
+Because of this, we focus our efforts on understanding variations in the
+performance of VDMS and its scalability, rather than comparing
+it to a baseline that would not represent a fair
+comparison for either of the systems.
 We describe the results in Section \ref{videos}.
@@
 and an evaluation of the different trade-offs that the systems offers for
-applications developers. For this, we implemented an image-search application
-based on \textit{similarity} search, which we describe in Section \ref{features}
+application developers. For this, we implemented an image-search application
+based on \textit{similarity} search, which we describe in Section \ref{features}.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/comparison_system}
+\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline system (right).}
+\label{fig:systems}
+\end{figure*}
 
@@
 and the Creative Commons license type information.
-The YFCC100M dataset also used a deep learning approach to generate autotags
-which is a set of comma-separated concepts such as people, scenery, objects,
-and animals and confidence scores generated from 1,570 trained
-Caffe classifiers ~\cite{Thomee_2016}. Together with the autotag, there is a
-probability associated to each tag to indicate the certainty level of the tag
-in that image/frame.
+The YFCC100M dataset also contains \textit{autotags}
+provided as a set of comma-separated concepts such as people, scenery, objects,
+and animals from 1,570 trained machine learning classifiers ~\cite{Thomee_2016}.
+Together with each \textit{autotags}, there is a
+probability associated with each tag to indicate certainty of the classification.
+This is, an image can have the \textit{autotags} "people", "person", "party",
+"outdoor", and each \textit{autotag} assigned will be accompanied by a
+probability of that \textit{autotags} being present in that image/frame.
 We have also used feature vectors generated for every image and first frame
-of every video \cite{features} to implement similarity search.
+of every video \cite{features} to implement a similarity search.
 Given that there is no standard benchmark oriented towards visual data queries,
-we have built a series of queries to filter this dataset that is model after
+we have built a series of queries to filter this dataset that is modeled after
 our internal use cases for many of the mentioned applications we have worked
@@
 \subsection{Experimental Setup}
+\label{setup}
 
-As a baseline, and given that there are no other open-source system
-that implements similar functionality and interfaces as VDMS,
-we implemented an equivalent visual data management system comprised of a
-combination of widely available, off-the-shelve components:
+Given that there are no other open-source systems
+that implement similar functionality and interfaces as VDMS, as a baseline,
+we have implemented an equivalent visual data management system comprised of a
+combination of widely available, off-the-shelf components:
 MySQL Server 5.7 (for storing metadata),
@@
 OpenCV 3.3 (to provide pre-processing operations on images).
-We have implemented a set of client-side applications that takes care
+Note that this implementation only partially replicates the functionalities
+that VDMS offers when it comes to image and metadata handling, built for the
+purpose of an ad-hoc image search implementation.
+This implementation is based upon internal tools used for
+that particular task in the past.
+We have implemented a set of client-side applications that take care
 of retrieving the components from the different systems, and applies
@@
 implemented as a baseline for this work
-is the common practice in the industry \cite{heystack, tao}.
+is the common practice in the industry \cite{haystack, tao}.
 
-For all our experiments, we use 2 servers, one hosting a VDMS server and
-another hosting a the Baseline implementation.
+For all our experiments, we use two servers, one hosting a VDMS server and
+another hosting the baseline implementation.
 Both servers have a dual-socket Intel\textsuperscript{\textregistered}
 Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
-each CPU with 28 physical cores with multithreading enabled,
+each CPU with 28 physical cores with hyper-threading enabled,
 for a total of 112 logical cores per server.
 The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
-has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
-The client application (running the queries and measuring round-trip time)
+has 64GB of DDR4 DRAM.
+We decided to run VDMS server in the machine with less DRAM to make
+sure MySQL had no disadvantage, and because previous evaluation
+indicated smaller footprint in the case of VDMS when
+compared to similar baselines based on MySQL.
+Other than the difference in DRAM space, machines are identical.
+Both servers run Ubuntu 16.04.
+The client application running the queries and measuring round-trip time
 is connected to the server through a 1GB wired link through
-a 10GB-backplane switch.
+a 10GB back-plane switch, same as both servers.
 
 Figure~\ref{fig:systems} shows a logical view of the difference between the
-interaction of the client application (which wants to retrieve metadata and
-imaged) with VDMS (left) and the baseline (right).
+interaction of the client application (retrieves metadata and
+images) with VDMS (left) and the baseline (right).
 The client application was implemented using Python 3 for both VDMS and
@@
 It is worth noting that the images are stored in a shared repository
-(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
-and VDMS have direct access. In the case of the baseline, metadata is
+(ext4 filesystem on a RAID 6 configuration of 16TB) that both 
+Apache WebServer and VDMS have direct access. 
+In the case of the baseline, metadata is
 stored in MySQL using an attached SSD disk.
 Even if VDMS has native support for Optane Persistent Memory,
-we do not use it in this experiment because of fairness of comparison w.r.t
-MySQL (which was not design for Persistent Memory type of storage).
+we do not use it in this experiment because of fairness of
+comparison with respect to MySQL, which was not designed for
+Persistent Memory type of storage.
 The benefits of Persistent Memory on metadata operations is left
 for another paper, and outside the scope of this evaluation.
-For this experiment then, in the case of VDMS we simply use a similar
+For this experiment, in the case of VDMS we simply use a similar
 attached SSD disk to store metadata.
-Even if PMGD, the graph database used by VDMS, is designed for PM,
-it can deliver good performance when using SSDs directly.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/comparison_system}
-\caption{Comparison Systems}
-\label{fig:systems}
-\end{figure*}
+Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
+it can deliver good performance when using SSDs directly, while still
+providing ACID-compliant  transactions.
 
@@
 For simplicity, we named the database based on the approximate number of images
-it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
-The exact number of images/elements on each database is shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
-For each database size, we created an instance of VDMS using YFCC100m metadata,
-the YFCC100m machine-generated \textit{autotags} associated with
-each metadata identifier, and the list of 1,570 \textit{autotags}
-using 100 concurrent VDMS Python clients.
+it contains, as follows: 1M, 5M, 10M, 50M, 100M.
+The VDMS and MySQL databases have comparable number of elements. 
+The exact number of images/elements in each database are shown in
+Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
+The differences can be attributed to failures in data 
+preparation/loading because of incomplete/inconsistent formatting, 
+which is common in large datasets \cite{failures}.
+In our set up, that difference is very small: 
+less than 0.1\% in terms of number of elements (images and/or metadata information).
+
+\subsubsection{Data Representation}
+
+\textbf{VDMS:}
+For each database size, we created an instance of VDMS using the image/video metadata,
+the machine-generated \textit{autotags} associated with
+each image/video identifier, and the list of 1,570 \textit{autotags}.
 Internally, that information is represented as a property graph,
 where we have one node for each image, one node for each tag
-(always 1570 tags), and connections between images.
-For instance, if an image has 4 \textit{autotags} assigned,
-there will be 4 connections between that image and
+(always 1,570 tags), and connections between images.
+For instance, if an image has four \textit{autotags} assigned,
+there will be four connections between that image and
 the different nodes for those \textit{autotags}.
-The information about the probability the \textit{autotag} present in an image
-is expressed as a property in the \textit{connection} between the 2 nodes.
-Figure \ref{fig:graph_representation} shows an example on two images and
-2 \textit{autotags}, and the \textit{connections} between those \textit{autotags}
-and the images.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Image id: 86756231, on the other hand, has a single \textit{autotags} assigned:
+The probability the \textit{autotag} is present in an image
+is expressed as a property in the \textit{connection} between the two nodes.
+Figure~\ref{fig:graph_representation} shows an example on two images,
+two \textit{autotags}, and the \textit{connections} between
+those \textit{autotags} and the images.
+Image id 23143252 has two \textit{autotags} assigned:
+\textit{Alligator} with probability 0.285, and \textit{Lake} with probability 0.872.
+Image id 86756231, on the other hand, has a single \textit{autotags} assigned:
 \textit{Alligator} with probability 0.894.
+On average there are 8 tags assigned to each image so
+there will be around 8 times more connections than images, as shown
+in Table~\ref{table:vdmsnodes}.
+Also, each image node will contain multiple properties associated
+with it (some of which are listed in Section \ref{dataset}). 
+% Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
+% and then insert nodes for the YFCC media objects (images or videos)
+% with the associated metadata (as properties of the image/video).
+% Then, for each image, we add a \textit{connection} between the image node 
+% and the \textit{autotag} node. The probability of that \textit{autotag}
+% being present in that image is stored as a property in that \textit{connection}.
+The number of nodes (representing images and \textit{autotags}) 
+are dependent on the database size and the \textit{connections} are responsible for
+90\% of the elements in each database instances,  
+as shown in Table~\ref{table:vdmsnodes}.
 
-\begin{figure}[]
+It is important to note that we create indexes for the image identifier,
+\textit{autotags} properties, and longitude/latitude coordinates 
+to enable faster retrieval.
+
+\begin{figure}[ht]
 \centering
@@
 their respective probabilities expressed in the \textit{connection}.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Similarly, Image id: 86756231 has a single \textit{autotags} assigned:
+Image id 23143252 has two \textit{autotags} assigned:
+\textit{Alligator} with probability 0.285, and \textit{Lake} with probability 0.872.
+Similarly, Image id 86756231 has a single \textit{autotags} assigned:
 \textit{Alligator} with probability 0.894.}
@@
 
-Because there are on average 8 tags assigned to each image,
-there will be around 8 times more connections than images, as can be seen
-in Table~\ref{table:vdmsnodes}.
-Also, each image node will contain multiple properties associated
-with that object.
-Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
-and then insert nodes for the YFCC media objects (images or videos)
-with the associated metadata.
-Adding the autotag confidence scores is a recursive process and requires more
-complex queries than adding the previous nodes.
-For each metadata identifier in \textit{autotags},
-the query must find the associated metadata node and tag node.
-A connection between the two nodes is created with the confidence score as a
-property. The number of metadata nodes are dependent on the database size and
-the connections are responsible for 90\% of the elements in each database as
-shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, \textit{autotags}, and
-longitude/latitude coordinates are set as indexes of the database to allow
-faster retrieval of the metadata.
-
-\begin{table}[h]
+\begin{table}[ht]
 \caption{VDMS Database - Number of Elements}
@@
 \hline\hline
-DB Name & \# Images & \# Connections & \# Autotags\\
+DB Name & \# Images & \# Connections & \# TagList\\
 \hline
-100k & 100,000     & 848,432      & 1,570\\
-500k & 500,000     & 4,249,500    & 1,570\\
+% 100k & 100,000     & 848,432      & 1,570\\
+% 500k & 500,000     & 4,249,500    & 1,570\\
 1M   & 1,000,000   & 8,503,045    & 1,570\\
@@
 
-Each MySQL database is created in a similar manner but the data is represented
-as three tables, following the relational model:
+\textbf{MySQL/Baseline:}
+Each MySQL database is created in a similar manner as VDMS 
+but the data is represented as three tables, following the relational model:
 1) \textit{images} table, contains one row per image,
 and a column for each property
-associated with the images (some of which are listed in \ref{dataset});
-2) \textit{taglist} table: contains one row per each autotag elements
-(always 1570 rows);
-3) \textit{autotags} table: contains on row per each autotag
-assigned to an image, each row containing a foreign key to the
-image and a foreign key to the tag, and
+associated with the images (some of which are listed in Section \ref{dataset});
+2) \textit{taglist} table: contains one row per autotag element
+(always 1,570 rows);
+3) \textit{autotags} table: contains one row per autotag
+assigned to an image. Each row contains a foreign key to the
+image, a foreign key to the tag, and
 the probability assigned to that tag belonging to that image.
@@
 table has around 8 times the number of rows present in the
-\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.
-
-THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
-By default, MySQL uses infinite number of threads inside InnoDB when processing
-requests using four threads per IO read/write.
-In some cases, when creating large databases data locks may occur to protect the
-data from concurrent updates.
-To minimize locking in our runs, we increased the InnoDB buffer pool size to
-increase the amount of memory allocated to in-memory data structures
-~\cite{mysql,mysql_blog}.
-
+\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
 Using a Python client and simple queries, the \textit{taglist}
 table is read from the list of tags with an auto-incremented
-\textit{tagid} as a primary key and the metadata table
-is read from the YFCC100m metadata using the identifier as a primary key.
+\textit{tagid} as a primary key, and the metadata table
+is read from the YFCC100M metadata using the identifier as a primary key.
 The \textit{autotags} table contains the generated autotags and
 probabilities for entries of the \textit{images} table.
-
 To generate the table, we split the \textit{autotags} data for each database
@@
 
-\begin{table}[h]
+In an attempt to have the best MySQL configuration possible for this use case, 
+we explore several parameters to increase the performance of both loading the data, 
+as well executing the queries.
+In particular, MySQL optimizes threads and transactions out-of-box, 
+but it cannot handle the entire YFCC100M dataset without configuring 
+specific parameters.
+When creating large databases, a data lock may occur to protect the
+data from concurrent updates~\cite{mysql_blog}.
+To avoid this mechanism, we increased the buffer pool size to
+increase the amount of memory allocated to internal data structures.
+It is recommended to set the buffer pool size to 60-80\% of the physical
+memory size ~\cite{mysql,mysql_blog}.
+However, the time to build a database increased. 
+We later changed  the buffer pool size to a multiple of the default value, i.e. 16x,
+which produced the best results for loading time.
+
+By default, MySQL uses the available operating system threads to  
+execute \textit{n} requests in parallel where \textit{n} is 
+the number of background read/write I/O threads.
+Setting the respective parameters in the MySQL configuration file can limit the
+number of concurrent threads and the number of background threads.
+When a limit is set on the number of threads, and no threads are available,
+requests will go into a FIFO queue until threads are available to execute
+the request ~\cite{mysql,mysql_blog}.
+We ran a few experiments investigating the effects of setting a limitation on the
+number of concurrent and background threads.
+We concluded that the default settings perform better for large databases instead of
+setting a limit.
+Therefore, we let MySQL to automatically handle the concurrency.
+
+In the case of VDMS, we did not attempt to tune any
+parameter to avoid unfairness in the comparison against the baseline.
+We use the default parameters provided by the implementation.
+For both VDMS and the baseline, we created indexes over the 
+properties we used for search, such as name of 
+\textit{autotag}, and geo-location values.
+Building indexes for the right properties and objects
+is basic operation that would be present in any real-world deployment,
+and measuring performance without them would lead to useless analysis in our
+real-world applications and use cases.
+
+\begin{table}[ht]
 \caption{MySQL Database - Number of Rows in each Table}
@@
 \hline
-100k & 100,000    & 848,912     & 1,570\\
-500k & 498,707    & 4,241,200   & 1,570\\
+% 100k & 100,000    & 848,912     & 1,570\\
+% 500k & 498,707    & 4,241,200   & 1,570\\
 1M   & 1,000,000  & 8,508,380   & 1,570\\
@@
 
-\begin{figure}[]
+\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/db_time_size}
-\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
+\caption{Time to build and size (in GB) of MySQL and VDMS databases.}
 \label{fig:db_time_size}
@@
 
-The VDMS and MySQL databases have comparable number of elements as shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
-can be attributed to failures in data preparation/loading because of
-incomplete/inconsistent formating, which is common on large datasets.
-The difference is less than 0.1\% in terms of number of elements (images and/or
-metadata information).
-
 \subsubsection{Database Loading Time}
 
-One of the first things we noticed is the difference is loading times,
+One of the first things we noticed is the difference in loading times,
 where VDMS outperforms MySQL by a large margin.
+This analysis includes the metadata, as the images are stored in a shared filesystem. 
 Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
@@
 optimizations (increased InnoDB pool size, etc.) needed
-to handle large datasets such as YFCC100m.
-On average, it took MySQL 3.72x longer to build each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours
-while VDMS only needed 72.5 hours.
+to handle large datasets such as YFCC100M.
+On average, it took MySQL around 3.72x longer to build each database than VDMS.
 
@@
 
-Another aspect important to note is that
+Another important aspect to note is that
 VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
@@
 The Graph Database internal to VDMS (called PMGD) was designed for performance,
-specially in environments where persistent memory is present.
+especially in environments where persistent memory is present.
 This design decision comes as a trade-off for storage footprint, which is
@@
 metadata accounts for less than 2\% of the overall database size.
-Take, for example, the largest database (both metadata and images) we built
-(100M): around 230GB of metadata and 12TB of images.
-On the other hand, in systems were persistent memory is a scarce resource,
-the increase storage foot print of PMGD may represent a challenge.
+For example, the largest database (both metadata and images) we built
+(100M) has around 230GB of metadata and 12TB of images.
+In systems where persistent memory is a scarce resource,
+the increased storage foot print of PMGD may represent a challenge.
+On the other hand, persistent memory is expected to be available 
+in the order of TBs per server, which should fit the 
+metadata of intensive use-cases\cite{IntelXPoint15}.
 
@@
 
-\subsection{Images + Metadata}
+\subsection{Images Search}
 \label{images}
@@
 In order to evaluate VDMS and the baseline on our use-case queries,
-we implemented 4 queries that that filter and retrieve certain specific images.
-We choose these queries because they represent typical use-case where a
-cohort of images is to be retrieve and processed from a large corpus of data.
-As we mentioned before, we took this approach due to the lack of a standard
+we implemented 4 queries that filter and retrieve a specific set of images.
+We chose these queries because they represent typical use-cases where a
+cohort of images is to be retrieved and processed from a large corpus of data.
+As we mentioned before, we took this approach due to the lack of standard
 benchmarks that are oriented towards visual data retrieval.
-We use the metadata associated to the images to filter said images.
-In particular, this dataset contains a set of 1570 machine generated
-tags (\textit{autotags})assigned to the images depending on the content.
-\textit{autotags} includes things like objects (chair, person, people, etc), and
-concepts like places (indoor, outdoor, lake) and others (party, etc).
-Together with each \textit{autotag}, a probability is assign.
-This is, an image can have the \textit{autotags} people, person, party,
-outdoor, and each \textit{autotag} assigned will be accompanied by a
-probability of that \textit{autotags} being present in that image.
-As mentioned earlier, there are 8 tags assigned to each image on average.
-Also, some images have geo-location information (latitude/longitude).
+We use the metadata associated with the images to filter said images.
 
-We use the \textit{autotags} (as they contain information about
-the content of the image), as well as the geo-location
-(as an example of properties of the images used used for search and filtering).
-
-On top of that and for our use case, we would like to extract more information
+We use the \textit{autotags} (as they contain information about the content 
+of the image), and geo-location information (latitude/longitude) 
+of the images for search and filtering.
+Note that, even if we use geo-location for our study, any other property 
+assigned to the images can be used to refine the search 
+in both VDMS and baseline implementations.
+On top of that, and for our use cases, we would like to extract more information
 about the content of the image through the use of ML,
-such as Convolutional Neural Networks TODO-CITE HERE.
+such as Convolutional Neural Networks~\cite{cnn}.
 For this, we resize the images to 224x224, which is the input layer size for
-a popular neural network for object detection on images: ResNet TODO-CITEHERE.
+popular variations of neural networks for object detection on images~\cite{resnet}.
 
@@
 \begin{itemize}
-\item Query 1 - {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
-\item Query 2 - {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
-\item Query 3 - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
-\item Query 4 - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
+\item \textit{q1} - {\bf {\em 1tag}}: Find metadata/images with one specific autotag (i.e. alligator, lake, etc).
+\item \textit{q2} - {\bf {\em 1tag\_resize}}: Find metadata/images with one specific autotag and resize to 224x224.
+\item \textit{q3} - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with one specific autotag, resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
+\item \textit{q4} - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with two specific autotags (i.e.
+alligator AND lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
 \end{itemize}
@@
 \textit{autotags}, we also apply a filter using the probability.
-For instance, we only retrieve images with a autotag \textit{alligator}
+For instance, we only retrieve images with an autotag \textit{alligator}
 and a probability higher than 92\%.
-This probabilities are both present in VDMS (in the form of a property
+These probabilities are both present in VDMS (in the form of a property
 of the \textit{connection} between the image and that \textit{autotag}),
 as well as in MySQL (in the form of a column in the \textit{autotags} table
-that links images with tags.
+that links images with tags).
+In the case of VDMS, the query involves a graph traversal query that starts
+from the \textit{autotag} node and ends in the images node,
+following \textit{connections} between the image and that \textit{autotag}).
+In the case of the baseline implementation,
+the query involves JOIN operations between the 3 tables.
+The implementation of this evaluation, as well as all the queries, are available
+under the benchmarks branch of the VDMS project
+\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
+and some examples of the queries can be found in the appendix of this
+paper.
 
-\begin{figure}[]
-\centering
+\begin{figure}[ht]
 \includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on Query \textit{1tag\_resize} \- VDMS vs Baseline}
+\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_resize}).
+Hardware concurrency (number of physical cores in each system)
+is shown with a blue vertical line (hw-concurrency = 56).
+Top figure shows aggregated throughput (transactions per second)
+when retrieving only metadata associated with the images, as the number of
+concurrent clients increase.
+Bottom figure shows aggregated throughput (images per second) when retrieving
+resized versions of the images, as the number of concurrent clients increases.}
 \label{fig:concurrency_comparison}
@@
 
-% We decided to remove this image because it is redudant with the concurrency
-% analysis.
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
-% \caption{Queries - Throughput - 32 Clients}
-% \label{fig:q_throughput_32}
-% \end{figure}
+Also, note that the size of the result (number of images retrieved)
+is linear with the size of the database. This is, if a query returns 100 images
+for the 1M database, it will return around 1000 images for the 10M database.
+This poses a problem when evaluating performance as the size of the database increase,
+and clearly understanding the measurements.
+Because of this reason, we control the number of returned images for all the
+databases using the probability of the \textit{autotags} 
+(higher probabilities returns less images), so that the queries in 
+this experiment return a similar number of images for all database sizes.
+In other words, as the size of the database increase, we increase the probability
+threshold for the queries. We do this for both VDMS and the baseline, of course.
+This way, we remove bottleneck introduced by network bandwidth that would
+otherwise over-complicate the understanding of the results.
 
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_56}
-% \caption{Queries - Throughput - 56 Clients}
-% \label{fig:q_throughput_56}
-% \end{figure}
-
-\begin{figure*}[ht]
+\begin{figure*}[ht!]
 \begin{subfigure}{.5\linewidth}
@@
   \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
-  \caption{Query 1 (red), and Query 2 (blue).}
+  \caption{\textit{q1}(red), and \textit{q2}(blue).}
   \label{fig:q1_q2}
@@
   \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
-  \caption{Query 3 (red), and Query 4 (blue).}
+  \caption{\textit{q3}(orange), and \textit{q4}(black).}
   \label{fig:q3_q4}
 \end{subfigure}
-\caption{Queries - Throughput - 56 Clients}
+\caption{Performance Analysis using 4 queries from our use-case
+described in the Experimental Setup Section.
+We show 2 queries in each figure for readability reasons.
+Top figures show the throughput of the 4 queries,
+just retrieving metadata associated with the images.
+Bottom figures show the throughput when retrieving both metadata and images,
+plus operations applied to images when applicable (queries 2, 3 and 4).
+The experiments show the performance of both systems (VDMS and baseline) as the
+database size increases.
+These queries were run using 56 simultaneous clients (nthr = 56),
+and averaged out of 10 runs (niter = 10),
+each client running 2 transactions (ntags = 2).}
 \label{fig:q_throughput_56}
@@
 
-Image search based on metadata is very expensive on large databases.
-Because of the large volume of data, the processing of the retrieve images
-is perform in parallel, using multi-core and/or distributed systems.
+Image search based on metadata is very expensive in large databases.
+Because of the large volume of data, the processing of the retrieved images
+is performed in parallel, using multi-core and/or distributed systems.
 For instance, a common implementation of an image processing pipeline
@@
 applicability of visual data management systems like VDMS.
+Because of this, we put emphasis on the analysis of concurrency and throughput,
+rather than latency.
+
+%% ========= Concurrency Comparison Analysis
+
+\subsubsection{Concurrency Analysis}
 
 Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
-the \textit{1tag\_resize} query, described above, using both VDMS and
-the baseline to evaluate the scalability of both solutions, as the
-size of the databases grows and as the number of concurrent clients grows.
+\textit{q2}(\textit{1tag\_resize}), described above,
+using both VDMS and the baseline.
+Here we evaluate the scalability of both systems, as the number of concurrent
+clients grows (x-axis) and as the
+size of the databases grows (each full-line represents a database size
+for VDMS and each dotted-line represents a database size for the baseline).
 
-It is important to note that the size of the result (number of images retrieved)
-is linear with the size of the database. This is, if a query returns 100 images
-for the 1M database, it will return around 1000 images for the 10M database.
-This poses a problem when evaluating performance as database increase,
-and clearly understanding the measurements.
-Because of this reason, we control the number of returned images for all the
-database using the probability of the \textit{autotags}, so that the queries
-in this experiment returns a similar number of images for all database sizes.
-Essentially, as the size of the database increase, we increase the probability
-threshold for the queries. We do this for both VDMS and the baseline, of course.
+We start by analyzing Figure~\ref{fig:concurrency_comparison} (top),
+which shows aggregated throughput (transactions per second)
+when retrieving only metadata associated with the images, as the number of
+concurrent clients increase.
+The first thing to notice is that at low concurrency (2 to 8 concurrent clients),
+both systems show similar performance, except in the 100M case of the baseline.
+Note that, for this particular experiment,
+baseline translates directly to MySQL performance,
+as the metadata-only queries only involve running a query to MySQL.
 
-The first thing to notice is that when it comes to retrieving metadata
-in the case of VDMS, (Figure ~\ref{fig:concurrency_comparison} (top),
-the size of the database (in terms of number of images) does not have
-a strong effect on the performance
-(measured in aggregated transactions per second).
-This means that even as database increase in size, there is no large increase
-in the query time.
+For the baseline system, in the case of 100M, the increase in the
+size of data seems to have a larger impact in performance.
+This result can be attributed to the increase in the complexity of the JOIN
+operation as the number of rows in the tables increases.
 
- the number of concurrent clients to use for the per-query performance analysis.
-This figure shows the metadata performance plateaus at 32 clients and
-the image performance at 56 clients for a simple query.
-% To further investigate, Figure~\ref{fig:q_throughput_32} and
-% Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
-% for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
-% respectively.
-On the 56-client experiment, a 10x difference can be seen between VDMS and
-baseline in metadata throughput while 32 clients is similar but not as high.
-The overall performance for both MySQL and VDMS is better with 56 clients;
-therefore, we used this value for our per-query analysis.
+Another thing to notice is that, as the number of concurrent clients increases,
+VDMS throughput continues to increase up to 56 threads, which is
+the hardware concurrency of the system.
+Also, more parallelism after 56 threads does not increase the delivered throughput,
+and it is actually slightly detrimental (112 concurrent clients case).
+On the other hand, the baseline seems to deliver less aggregated
+throughput after 16 threads, with an increase for 64 threads, but these effects
+are hard to interpret fully as the standard deviation in the measurements is high.
+We noticed throughout many experiments that MySQL results showed higher
+standard deviation, meaning less consistent and noise measurements, when compared
+to VDMS.
+We tried increasing the number of measurements and discarding
+outliers but were not able to get less noisy results.
 
-\begin{figure}[]
+We continue by looking at Figure~\ref{fig:concurrency_comparison} (bottom),
+which shows aggregated images per second delivered by each system.
+Again, we evaluate the scalability of both systems, as the number of
+concurrent clients grows (x-axis) and as the
+size of the databases grows (each full-line represents a database size
+for VDMS and each dotted-line represents a database size for the baseline).
+Note that most of the baseline dotted-lines are very close to each other.
+This is mostly an effect of the log-scale used, which is needed to clearly
+depict the difference between VDMS and the baseline.
+Here, the baseline is the full architecture described in
+Figure~\ref{fig:systems} (right).
+Figure~\ref{fig:concurrency_comparison} (bottom) shows a similar trend as the top
+figure when it comes to low concurrency. The baseline does as good and even better than VDMS with 2 or 4 concurrent clients.
+However, as concurrency increases beyond 4 concurrent clients, the difference
+in throughput becomes clear, with VDMS reaching its peak performance at
+56 concurrent clients.
+This query (as well as \textit{q3} and 4) runs a resize operation on the image,
+an operation that requires decoding, resizing, and encoding the image
+before sending it back to the client.
+These operations are mainly compute bound, and that is the reason for
+the system to stop scaling beyond the number of physical cores.
+In contrast, the baseline does not scale nearly as well as VDMS,
+and we see that even after increasing concurrency, the increase
+in throughput is just about 2x.
+When comparing the case of 56 or 64 concurrent clients,
+VDMS delivers between 8x and 10X the throughput.
+
+There are many reasons why we see this performance improvement, the main being
+that the entire operation (metadata query, image fetching and resizing) happens
+on the server side in the case of VDMS, within a single message
+exchange between the client and the server.
+Many of the inefficiencies that come with combining tools that were designed
+for other use cases simply disappear when building a tool that treats
+visual entities as first class citizens, as it is the case of VDMS.
+Another reason, which is quantifiable in the bottom figures, is that
+VDMS sends resized (smaller) versions to the client instead of the full image
+to be resized on the client side (as is the case in the baseline).
+This is in contrast with the baseline, where 2 rounds of blocking back-and-forth
+communication with the server is needed, as depicted in Figure~\ref{fig:systems}.
+Note that on the point 1), one could argue that the opposite will happen
+when the resize operation retrieves a up-sampled (larger) version of the image
+instead of a down-sampled (smaller) one.
+In practice, retrieving an up-sampled version is not a common use case,
+given that up-sampling the image does not add any extra information that can help,
+for instance, improve the accuracy of a ML model.
+The case of down-sampling the original image is much more common and is the common
+practice when it comes to image processing through CNNs \cite{cnn,resnet}.
+
+%% ========= Concurrency Comparison Analysis - END
+
+%% ========= Queries Analysis
+
+\subsubsection{Query Execution Analysis}
+
+The next step in our analysis involved running a different set of queries
+(described above in this section),
+to better understand the performance of the systems under different query conditions.
+Figure~\ref{fig:q_throughput_56} shows the evaluation of the 4 representative
+queries we analysed for our use case.
+Top figures show the throughput of the 4 queries
+retrieving metadata associated with the images.
+Bottom figures show the throughput when retrieving both metadata and images,
+plus operations applied to images when applicable  (queries 2, 3 and 4).
+The experiments show the performance of both systems (VDMS and baseline) as the
+database size increases in terms of number of images.
+These queries were run using 56 simultaneous clients (nthr = 56),
+and averaged over 10 runs (niter = 10),
+each client running the retrieval of the tag 2 times (ntags = 2).
+The last parameter (ntags) is to avoid having some queries to finish the work
+too fast before other clients can even send the query to the server.
+This ensures that there is enough work to do in a query so that
+all queries execute in parallel on the server side.
+To analyze these plots, one needs to compare the full-line (VDMS) versus the
+dotted-line (baseline), each color representing a different query.
+For example, to compare \textit{q1} performance, one needs to look at
+the full-red line (VDMS) and the dotted-red line (baseline).
+
+\begin{figure*}[ht!]
 \centering
-\includegraphics[width=\columnwidth]{figures/q1_latency}
-\caption{Query 1 - Latency}
-\label{fig:q1_latency}
-\end{figure}
+\includegraphics[width=\textwidth]{figures/summary_images}
+\caption{Summary of performance gains for all queries.
+We see up to 35x speedup (\textit{q2}), and an average of about 15x.
+More importantly, we see that the speedup grow as the database size increases,
+showing that VDMS scales better than the baseline.}
+\label{fig:summary_images}
+\end{figure*}
+
+\subsubsection{Metadata Retrieval}
+
+First, we analyze the top figures, showing metadata transactions per second.
+We evaluate the performance when retrieving only metadata associated with the images,
+and not the images themselves.
+In this particular case, the baseline translates directly into MySQL performance.
+Note that for these figures (top), \textit{q1} and \textit{q2} are
+essentially the same query. This is because the metadata retrieved for both
+queries does not change. The only difference between \textit{q1} and \textit{q2}
+is the presence of the resize operations that does not have any impact on
+analyzing performance of metadata retrieval.
+For \textit{q1} and \textit{q2}, we can appreciate higher performance being delivered by
+VDMS when compared to the baseline, and how this improvement is maintained
+as the size of the database increase.
+For \textit{q3}, we see that MySQL performs best when the database size
+is small (1M images), with VDMS outperforming MySQL as the database increases in size.
+For \textit{q4}, we see MySQL also outperforming VDMS on small size databases, and as
+the database size increase, the gap between the two narrows.
+It is interesting to note that adding filtering by geo-location
+(\textit{q3} and \textit{q4}) slightly increases the performance of MySQL
+for small databases, and decreases it as the database sizes scales.
+In the case of VDMS, we see \textit{q3} performance is comparable to \textit{q1}
+and \textit{q2}, but \textit{q4} suffers significantly when the scale of the database
+increases.
+The reason for that lack of scalability lies on the query implementation: given that
+VDMS does not yet support operators that enable querying images 
+that have both connections to a \textit{tagA} and a \textit{tagB}; 
+we have to implement this transaction by doing 2 retrievals.
+This involves retrieving partial information in the first retrieval, 
+applying an INTERSECTION operation in the client, and doing a second retrieval
+to bring the right metadata and/or images.
+The reason for this is a lack of operations that would enable this query to be
+run entirely on the server is not an inherent limitation to VDMS but rather
+just a missing implementation.
+Future release will add more such operators in order to prevent 
+unnecessary retrievals.
+
+\subsubsection{Image Retrieval}
+
+We continue by analyzing the bottom figures, showing measured throughput,
+as images per second, delivered by each system.
+For the case of VDMS, \textit{q1}(full-red-line) shows less throughput
+than \textit{q2} (full-blue-line) (bottom left figure).
+This is expected as \textit{q1} returns a full size version of the image,
+whereas \textit{q2} returns a resized (smaller) version of the image,
+thus transferring less data over the network.
+In the case of the baseline, both \textit{q1} and \textit{q2} transfer the full size
+version of the image, and as part of \textit{q2}, the resize is performed in the client.
+This is why, contrary to the VDMS case, \textit{q1} performs better (even if slightly)
+when compared with \textit{q2}.
+We can also see that, for VDMS, \textit{q3} perform worst than \textit{q2} because
+of the extra step needed for filtering based on geo-location.
+Moreover, we see a great performance degradation in the case of \textit{q4} as the
+database size increases.
+This is entirely attributed to the 2-round process needed for this query,
+as we explained before.
+From the first 3 queries, we clearly see that VDMS outperforms the baseline 
+when retrieving visual data and applying operations.
+This is one of the most important finding, as it validates the design principles
+of VDMS, which aims to provide scalability and performance acceleration
+at the type of queries that require visual data access and transformations.
+
+%% ========= Queries Analysis - END
+
+%% ========= Summary Analysis
+
+Finally, Figure~\ref{fig:summary_images} summarizes the results.
+We see up to 35x speedup (for the case of \textit{q2}),
+and an average improvement in throughput of about 15x.
+More importantly, we see that the speedup increase as the database size grows,
+showing that VDMS scales better than the baseline.
+We also see how \textit{q4} shows poor performances and scalability 
+when compared to the baseline, and this evaluation served the 
+purpose of understanding the importance of VDMS server side operators 
+that enable more complex queries for our use cases.
+The team will address the missing implementation as part of future work.
+
+%% ========= Summary Analysis - END
 
@@
 
-\subsection{Videos}
+\subsection{Video Search}
 \label{videos}
 
-VDMS provides full support for video storage and operations.
-This includes support for encoding/decoding and trans-coding of
+VDMS provides full support for video storage and operations,
+in a similar way it does for images.
+This includes support for encoding, decoding, and transcoding of
 \textit{mp4}, \textit{avi}, and \textit{mov} containers,
@@
 This is supported through the Visual Compute Module that provides an abstraction
-layer on top of OpenCV (and \textit{ffmpeg} directly).
-All the operations supported for images in VDMS are also supported at the
-video and frame level on the API. On top of that, there are a number of
-video-specific operations that are supported, such as the interval operations,
-that allow users to retrieve clips at different
+layer on top of OpenCV~\cite{opencv} and \textit{libffmpeg}\cite{ffmpeg}.
+All operations supported for images in VDMS are also supported at the
+video and frame level of the API.
+On top of that, there are a number of video-specific operations that
+are supported, such as the interval operations,
+enabling users to retrieve clips at different
 frames-per-second (FPS) versions of the video.
-
-All this functionality is provided and integrated with the rest of the
-metadata API as part of the comprehensive VDMS interface. This makes it
-possible for users to interact with metadata and video in a transactional
-manner, enabling users to run queries like: "Retrieve all the videos
-where there is a lake with probability higher than X, converting all videos
-to \textit{H.264} \textit{mp4} of size 224x244".
-In particular, this functionality was used internally to select a subset
-of videos with the right licenses for a video summarization application.
-
-To the best of our knowledge, there is no solution that can provide
-all the functionality mentioned above. Also, implementing a baseline is a very
-complex task as there is a large number of options and parameters that can
-be chosen, which makes it hard to accurately compare against VDMS functionality.
-For this reason, we chose to make a study using VDMS in various scenarios,
-and analysis what is the impact of having the overhead of VDMS' Request
-Server in the overall access time.
 
@@
 
-Figure \ref{fig:video} shows the analysis of different queries aimed to retrieve
-a video using VDMS interface. We show how VDMS increases throughput of serving
-a video object as the number of simultaneous client increases, as well as the
-overhead VDMS introduces in the overall query execution time.
+All this functionality is provided and integrated with the rest of the
+metadata API as part of the comprehensive VDMS interface.
+This makes it possible for users to interact with metadata and video in 
+a transactional manner, enabling users to run queries like: 
+"Retrieve all the videos where there is a \textit{lake} with 
+probability higher than 0.86, converting all videos to \textit{H.264} 
+\textit{mp4} of size 224x224".
+Appendix shows a sample of how this query would be implemented using the VDMS API
+\footnote{https://github.com/IntelLabs/vdms/wiki/FindVideo}.
+In particular, this functionality was used internally to select a subset
+of videos with the right licenses for a video summarization application.
+
+To the best of our knowledge, there is no solution that can provide
+all the functionality mentioned above, behind a single interface
+that also allows users to interact with images and metadata.
+Implementing a baseline, like we did for images, is significantly more complex
+due to the parametrization of video encondings and containers, 
+as explained at the beginning of this section.
+For this reason, we chose to make a study using VDMS in various scenarios,
+and analysis of scalability and the impact of having the overhead of VDMS' Request
+Server in the overall access time and throughput.
+
+Figure~\ref{fig:video} shows the analysis of different queries aimed
+at retrieving a video using the VDMS interface.
+We show how VDMS throughput increases when serving
+a video object as the number of simultaneous clients increases, as well as the
+overhead operations introduced in the overall query execution time.
 The figure on the left compares the number of video transaction per second
@@
 
-We also run a set of other queries that involve: (a) running a resize operation
-on the video and, consequently, needs a decoding and
-encoding operations as well (blue line),
-(b) transcoding, meaning the use a different container and encoder
-than the one originally used (yellow line), and (c) both resize and transcoding.
-Note that the resize operation performs a downsize, which translates in less
-data being sent over the wire. This is specially noticeable when supporting 32
-simultaneous clients, where the systems provides more videos per second due to
-sending less data to the client, when compares to just transcoding and not resizing (yellow line).
+We also run a set of other queries that involve, showed in Figure~\ref{fig:video}:
+(a) running a resize operation on the video and, consequently,
+decoding and encoding operations as well (blue line),
+(b) transcoding, meaning the use of a different container and encoder
+than the one originally used (yellow line), and
+(c) both resize and transcoding.
+Note that the resize operation (blue and green lines) performs a downsize,
+which translates in less data being sent over the wire.
+This is specially noticeable when supporting 32 simultaneous clients, 
+where the system provides more videos per second due to sending less data to 
+the client, when compared to just transcoding and not resizing (yellow line).
+We can see that the system performs best when using all the physical cores,
+and this can be attributed to the compute-bound nature of video
+encoding, decoding, and processing.
 
-We can see that the system performs best when using all the physical cores.
-COMPLETE THIS BETTER.
-
-WE SHOULD ADD HERE SAMPLES OF THE QUERYS TO SHOW HOW THESE ARE RETRIEVED.
-
-Because we see almost 3 orders of magnitude drop in performance when including
-operations as part of the query, we wanted to understand where most of the time
-was spent on the query, and optimize the Request Server and Visual Compute Module
-if necessary. For this, we run the experiment shown at
-Figure \ref{fig:video} (right) which breaks down the different components of the
-queries. This figure shows that more than 97\% of the query execution is spent
-on encoding/decoding operations, which is well-known to be a
-heavy and compute intensive operation.
-On the one hand, this results show that VDMS introduced overhead for
-video operation is minimal. On the other hand, this result means a
-limit on the opportunities for optimization for video queries given
-that biggest time factors are accounted by encoding/decoding, which is
-outside the scope of VDMS.
-This result also was the inspiration point for one optimization we included
-in future versions of VDMS, which involves using \textit{ffmpeg} C++ API to
-limit the number of frames being encoding/decoding when possible.
-
-%=========================================
-
-\subsection{Feature Vectors}
-\label{features}
-
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities or visual objects.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as \textit{descriptors}
-or \textit{visual descriptors}. We use these terms interchangeably.
-These descriptors can be classified and labeled and used to build search indexes,
-and there are many in-memory libraries that are designed for
-this task~\cite{flann, faiss}.
-Using the VDMS API, users can manage feature vector indexes,
-query previously inserted elements (images),
-run a k-nearest neighbor search (\textit{knn}), and express relationships
-between existing images or descriptors and
-the newly inserted descriptors.
-By natively supporting descriptors and \textit{knn},
-VDMS allows out-of-the-box classification functionalities for many applications.
-
-For this work, and as part of a comprehensive image search implementation,
-we have used 4096-dimensional descriptors extracted from every image
-(and first frame of every video) and created a collection of these feature
-vectors in VDMS to be able to perform similarity search (i.e., find
-images that are \textit{similar} to an query (input) image).
-\textit{Similarity} in this particular case is defined as closeness
-in a an 4096-dimensional space using euclidean distance as the metric.
-
-EXPLAIN HERE DIFFERENT ACCURACY/EXECUTIME TIME TRADEOFF.
-
-The process of loading descriptors in VDMS is simple.
-First, the user has to create a DescriptorsSet, using a single command.
-At creation of the DescriptorSet, the dimensionality of the descriptors
-is specified, together with the desired indexing method and the desired metric
-for computing distances (Euclidean Distance, \textit{L2},
-or Inner Product, \textit{IP}).
-Once the DescriptorSet is created, descriptors can be inserted to the set.
-After the descriptors are inserted, similarity search can be performed.
-Note that descriptors can be inserted later on, and any search will reflect
-the existence of all descriptors in the set.
-
-\begin{figure*}[]
+\begin{figure*}[ht!]
 \centering
@@
 
-Figure \ref{fig:similarity} shows 3 examples of a query image (on the left),
-and images returned as "similar" by VDMS.
-The query input is a descriptor generated after a query image. The "query"
-descriptor is sent to VDMS as part of the query, and VDMS use that descriptor
-to find similar one, and retrieve the images associated with those "similar"
-descriptors. We used this as an example and as a visual validation of the
-functionality and applicability in this particular dataset, but we also
-provide an analytical approach to accuracy and trade-offs in our system.
+It is important to note an almost 3 orders of magnitude drop in performance
+when including operations as part of the query.
+We wanted to understand where most of the time was spent on the queries,
+and optimize the Request Server and Visual Compute Module
+if necessary. For this, we run the experiment shown at
+Figure~\ref{fig:video} (right) which breaks down the different components of the
+queries. This figure shows that more than 97\% of the query execution is spent
+on encoding/decoding operations, which is well-known to be a
+compute intensive operation\cite{videosgoogle}.
+On the one hand, this result shows that VDMS barely introduces any overhead. 
+On the other hand, this result means a limit on the opportunities 
+for optimization for video queries given that biggest time factors 
+are accounted by encoding/decoding, which is outside the scope of VDMS.
+This result was the call to action for one optimization we will include
+in future releases of VDMS, which involves using \textit{ffmpeg} C++ API to
+limit the number of frames being encoded/decoded when possible.
+This functionality will prevent encoding/decoding to happen on all frames
+when users only need to retrieve a subset of the frames in the video.
+
+%=========================================
+
+\subsection{Similarity Search}
+\label{features}
+
+Another key differentiating factor of VDMS is that it allows the creation of
+indexes for high-dimensional feature vectors and the insertion of
+these feature vectors associated with entities, images, and/or videos.
+Feature vectors are intermediate results of various machine
+learning or computer vision algorithms when run on visual data.
+Feature vectors are also known as \textit{descriptors}
+or \textit{visual descriptors}. We use these terms interchangeably.
+These descriptors can be classified, labeled, and used to build search
+indexes. There are many in-memory libraries that are designed for
+this task~\cite{flann, faiss}.
+Using the VDMS API, users can manage feature vector indexes,
+query previously inserted elements,
+run a k-nearest neighbor search (\textit{knn}), and express relationships
+between existing images or descriptors and
+the newly inserted descriptors.
+By natively supporting descriptors and \textit{knn},
+VDMS allows out-of-the-box classification functionalities for many applications
+\footnote{https://github.com/IntelLabs/vdms/wiki/ClassifyDescriptor}.
+
+For this work, and as part of a comprehensive image search implementation,
+we have used 4096-dimensional descriptors extracted from every image
+(and first frame of every video) from the YFCC100M dataset
+and created a collection of these feature vectors in VDMS to
+perform similarity search (i.e., find images that are
+\textit{similar} to an query (input) image).
+\textit{Similarity} in this particular case is defined as closeness
+in a 4096-dimensional space using euclidean distance as the metric.
+
+The process of loading descriptors in VDMS is simple.
+First, the user has to create a DescriptorSet, using a single command.
+At creation of the DescriptorSet, the dimensionality of the descriptors
+is specified, together with the desired indexing method and the desired metric
+for computing distances (Euclidean Distance, \textit{L2},
+or Inner Product, \textit{IP}).
+Once the DescriptorSet is created, descriptors can be inserted to the set.
+After the descriptors are inserted, a similarity search can be performed.
+
+Figure~\ref{fig:similarity} shows 3 examples of a query image (on the left),
+and images returned as \textit{similar} by VDMS.
+The input is a descriptor generated after a query image.
+The \textit{query input} descriptor is sent to VDMS as part of the query,
+VDMS uses that descriptor to find similar ones,
+and retrieves the images associated with those \textit{similar} descriptors.
+We show this as an example of the functionality and to depict
+how the feature vectors provided by the dataset can be used,
+but we also provide an analytical approach to
+the trade-off between accuracy and execution time in our system.
 It is important to note that the accuracy of the results is entirely tied
 to the quality of the descriptors chosen by the applications.
-VDMS is completely agnostic to information contained within the descriptor,
-and simply offers the interface to store and index them, but the quality
-of the similarity result will be tied to the quality of descriptor extraction
-that the application is using.
+The quality of the similarity result will be tied to the quality
+of the descriptor extraction that the application is using.
 
-\begin{figure*}[]
+\begin{figure*}
 \centering
@@
 
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=\textwidth]{figures/features_disksize}
+\caption{Feature Collection Size in Disk}
+\label{fig:features_size_does_matter}
+\end{figure*}
+
 As mentioned before, VDMS provides different levels of customization of the
@@
 and the metric for similarity.
-These different indexing techniques comes with different trade-offs in terms
-of speed of search and accuracy.
+These different indexing techniques come with different trade-offs in terms
+of speed of search and accuracy of the computation.
 VDMS aims to provide functionality that is agnostics to application-specific
-techniques, enabling functionality that is general to visual data processing
+techniques, enabling features that are generic to visual data processing
 applications.
-Figure \ref{fig:features_eval} shows an analysis at the different indexing
-techniques provided by VDMS and its trade-off between accuracy and query
+Figure~\ref{fig:features_eval} shows an analysis at the different indexing
+techniques provided by VDMS and its trades-off between accuracy and query
 execution speed, for a single threaded client.
@@
 The \textit{flat} index (red line) implements exact search and
-represents ground-truth, which explain why the accuracy is always 100\% on the
-right plot. The other indexes implement \textit{approximate search},
-which trade-offs between accuracy and speed of search~\cite{flann, faiss}.
+represents ground-truth, which explain why the
+accuracy is always 100\% in the plot on the right.
+The other indexes implement \textit{approximate search},
+which trades-off between accuracy and speed of search~\cite{flann, faiss}.
 We have also tried the \textit{ivfflat} index (inverted file index), as well as
@@
 \footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
-Results show how \textit{ivfflat} is the fastest option by comes at the trade-off
-of about 30\% lost in accuracy, while simple brute-force search
+Results show how \textit{ivfflat} is the fastest option but it comes with a trade-off
+of about 30\% loss in accuracy, while simple brute-force search
 is among the slowest options at the expenses of 100\% accuracy,
@@
 
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/features_disksize}
-\caption{Feature Collection Size in Disk}
-\label{fig:features_size_does_matter}
-\end{figure*}
-
-Another important trade-off to be made is w.r.t to space efficiency: Set of
-descriptors can grow very large and hard to manage.
+Another important trade-off to be made is with respect to space efficiency:
+The DescriptorSet can grow very large and expensive to load and manage.
 In this particular case, 4096-dimensional descriptors for 100M elements
-translates into 1TB of data, only in raw floating-point data alone (without
-accounting for any metadata or indexes associated with it).
-This component is very important on the overall analysis because a large set
-of feature vectors may not fit in memory and thus cause a pressure on the IO
-system while retrieving descriptors for computing distance, that severely impact
-the overall query execution time.
-Because of this, when the set of descriptors grows significantly large,
-it may be worth trading off accuracy to speed and space.
-Figure \ref{fig:features_size_does_matter} show the different indexes and
+translates into 1TB of data, only in raw floating-point data alone
+(without accounting for any metadata or indexes associated with it).
+This component is very important for the overall analysis on which
+index structure to use because a large set of descriptors may not fit in memory
+and thus cause a pressure on the IO system while retrieving descriptors
+for computing distance.
+This can severely impact the overall query execution time.
+When the DescriptorSet grows significantly large,
+it may be worth trading off accuracy for speed and space.
+Figure~\ref{fig:features_size_does_matter} shows the different indexes and
 their size in disk. These indexes already contain all the descriptors (or
-a quantized version of them), and can be loaded in memory directly when it fits.
+a quantized version of them in the case of LSH~\cite{lsh}),
+and can be loaded in memory directly when it fits.
 Note how, because of quantization of the descriptors, \textit{LSH} provides a
@@
 large collections of descriptors when accuracy is not a main factor.
+It is not uncommon to sacrifice accuracy as images and videos are captured
+using a noise sensor (i.e., the camera), and an approximate search
+in many cases can provide the necessary accuracy for applications
+to achieve their goals.

### 2020-02-25 yfcc100m/paper/vldb2020/intro.tex
Update to overleaf version

--- 2020-02-03 yfcc100m/paper/vldb2020/intro.tex
@@
 have become prolific across a wide range of application domains.
-This is in part due to the growing ability of machine learning techniques to
-extract information out of the visual data which can subsequently be used
+This is in part due to the growing ability of machine learning (ML) techniques to
+extract information from the visual data which can subsequently be used
 for informed decision making \cite{vdms-nips}.
@@
 application: a retail vendor might be interested in the amount of time
-shoppers spend in front of a specific product, while a medical expert might
 want to see the effect of a specific treatment on the size of a tumor.
-
-There are many applications that can be studied through the use of large
-and public available datasets. But before getting started, there is usually
-a large effort devoted to cleaning up the data. For this work, we use
-the YFCC100m dataset as an example given that we have use the same dataset
-for multiple proof of concepts and applications without our research lab.
-Applications include basic image search functionality (through the use
-of human-generated tags), advanced image search through the use of
-machine generated tags for each image, and video summarization.
-In all cases, it was needed to select a subset of the data before running
-any processing. Selecting subsets of data is by itself a time consuming task,
-as it involves loading all metadata into a solution that enables searched
-based on tags (relational database, graph database, csv files, etc), and
-building the necessary pipelines for querying and retrieving the right data.
 
@@
 research on the management of visual data.
-Most of the current storage solutions are
-an ad-hoc collection of tools combined with custom scripts to tie them
-together, unique not only to a specific discipline but often to individual
-researchers. For example, consider an ML developer constructing a pipeline
+Most of the current storage solutions for visual data are
+an ad-hoc collection of tools and systems. 
+For example, consider a ML developer constructing a pipeline
 for extracting brain tumor information from existing brain images in a
-classic medical imaging use case. This requires assigning consistent
+classic medical imaging use case. 
+This requires assigning consistent
 identifiers for the scans and adding their metadata in
-some form of relational or key-value database. If the queries require
-search over some patient information, then patients have to be associated
-with their brain scans. Finally, if the ML pipeline needs images that
+some form of relational or key-value database. 
+If the queries require a search over some patient information, 
+then patients have to be associated with their brain scans. 
+Finally, if the ML pipeline needs images that
 are of a size different than the stored ones, there is additional compute
-diverted towards preprocessing after the potentially larger images are
-fetched. All these steps require investigation of different software
+diverted towards pre-processing after the potentially larger images are fetched. 
+All these steps require investigation of different software
 solutions that provide various functionalities that can then be stitched
@@
 hard to evolve the schema on the fly.
-Not only do these ad hoc solutions make replicating experiments
-difficult, they do not scale well to real-world applications.
-Addressing the storage and retrieval of visual data necessitates a complete
-overhaul of the storage architecture,
-preferably using emerging breakthroughs in
-heterogeneous memory and storage for efficiency.
+As another example, many applications that can be studied through the use of large
+and publicly available datasets. 
+Applications include basic image search functionality (through the use
+of human-generated tags), advanced image search through the use of
+machine-generated tags and feature vectors\cite{imagesearch} 
+for each image, and video summarization.
+For these use-cases, the usual first step consists on selecting a 
+subset of the data before running any processing, and a large effort 
+is devoted to cleaning up and pre-processing the data.
+Selecting subsets of data is by itself a time consuming task,
+as it involves loading all metadata into a solution that enables searching
+based on tags (relational database, graph database, csv files, etc), and
+building the necessary pipelines for querying and retrieving the right data.
 
-Data scientist and machine learning developers usually end up building ad-hoc
-solutions through a combination of databases and file systems to store metadata
-and visual data (images, videos), respectively.
+More generally, data scientists and machine learning developers 
+usually end up building an ad-hoc solution that results in a 
+combination of databases and file systems to store 
+metadata and visual data (images, videos), respectively. 
+This is integrated with a set of custom scripts that tie multiple systems together, 
+unique not only to a specific application/discipline but often to individual researchers.
+These ad-hoc solutions make replicating experiments difficult, 
+and more importantly, they do not scale well when deployed in the real-world.
+The reason behind such complexity is the lack of a one-system 
+that can be used to store and access all the data the application needs.
 
-In this paper, we also expand on the Video and Feature Vector capabilities of
+In this paper, we show how VDMS~\cite{vdms-nips} provides a comprehensive solutions 
+to the data management for applications that heavily rely on visual data. 
+VDMS is an Open Source project designed to enable efficient access of visual data.
+We also expand on the video and feature vector capabilities of
 VDMS, which are part of the latest additions to the system.
-We analyse different functionalities and trade-offs for this type of data,
-in combination with metadata filtering. To the best of out knowledge, this
-set of functionalities are unique to VDMS and we were not able to find a system
-with similar functionality.
+We analyze different functionalities and trade-offs for this type of data,
+in combination with metadata filtering. 
+To the best of our knowledge, this set of functionalities, 
+provided behind an integrated API, are unique to VDMS and 
+we were unable to find a system with similar functionality.
+We show how VDMS can be used as the single and centralize point for data
+management and data access even when having multiple modalities of data:
+Metadata, Image, Videos, and Feature Vectors.
 
-\begin{comment}
-
-We present the Visual Data Management System (VDMS)\cite{darkside},
-an Open Source project designed to enable efficient access of visual data.
-Since visual data often contains
-rich metadata (such as objects, locations, and time), VDMS stores this
-information in a high performance graph database. Using this metadata, VDMS
-can quickly identify which data is relevant to a given query.
-Additionally, VDMS uses a custom library to store and retrieve visual data,
-which provides an interface for machine friendly formats as well as
-traditional formats. These new formats are designed to support applications
-that are often interested in specific areas of images or videos,
-particularly when the individual object is large.
-
-\end{comment}
+For this work, we use the YFCC100M dataset\cite{Thomee_2016}. 
+The YFCC100M is the largest publicly multimedia collection. 
+It contains the metadata of around 99.2 million photos 
+and 0.8 million videos from Flickr,
+plus expansion packs that include a variety of multidimensional data,
+all of which were shared under one of the various Creative Commons licenses.
+We have used this dataset
+for multiple proof of concepts and applications within our research lab.

### 2020-02-25 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Update to overleaf version

--- 2020-02-03 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
@@
 \usepackage{mathtools}
+\usepackage{minted}
 
 % Include information below and uncomment for camera ready
-\vldbTitle{A Sample Proceedings of the VLDB Endowment Paper in LaTeX Format}
-\vldbAuthors{Ben Trovato, G. K. M. Tobin, Lars Th{\sf{\o}}rv{$\ddot{\mbox{a}}$}ld, Lawrence P. Leipuner, Sean Fogarty, Charles Palmer, John Smith, Julius P.~Kumquat, and Ahmet Sacan}
+\vldbTitle{Using VDMS to Index and Search 100M Images}
+\vldbAuthors{Luis Remis, Chaunt\'e W. Lacewell}
 \vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
@@
 
-\title{Using VDMS to Search 100M Images: A Piece of Cake}
+\title{Using Visual Data Management System (VDMS) to\\ Index and Search 100M Images}
 
@@
 build machine learning models and algorithms to solve problems.
-The problem grows as the scale of the data that is used for analysis or training
-grows. In this paper, we demonstrate how VDMS (Visual Data Management System)
-can be used to simplify the data preparation process and we show how we gain
-in efficiency just because we are using a tool that was design for the job.
+The problem grows as the scale of the data used for analysis or training
+grows. In this paper, we demonstrate how VDMS (the Visual Data Management System)
+can be used to simplify the data preparation process and how we gain
+in efficiency simply because we are using a system designed for the job.
 To demonstrate this, we use one of the largest available
-public dataset (YFCC100M),
-with 99+ million images, 800+ thousand videos plus add-ons that include
+public datasets (YFCC100M),
+with 99+ million images, 800+ thousand videos, plus add-ons that include
 machine-generated tags and 4K dimensional feature vectors,
-a total of $\sim$13TB of data.
+for a total of $\sim$13TB of data.
 VDMS differs from existing large scale photo
@@
 (images, videos, and feature vectors), treating these as first class citizens.
-We demonstrate how VDMS outperforms well-know and widely used tools for
-data access by a $\sim$10x factor in certain queries,
-while simplifying the process
-of data preparation and data access, and providing functionalities non-existent
-in alternative options.
+We demonstrate how VDMS outperforms well-known and widely used 
+systems for data management by up to $\sim$35x, with 
+an average improvement of about 15x for the queries we
+used in our use-cases, and particularly at scale.
+At the same time, VDMS simplifies the process of data preparation and data access,
+and provides functionalities non-existent in alternative options.
 \end{abstract}
@@
 \input{eval.tex}
-%\input{demo.tex}
 \input{conclusion.tex}
-
-\section{Acknowledgments}
-
-We have to acknowledge Philip and Vishakha for help in fixing PMGD bug,
-Nilesh for his help/input, and any other reviewer.
+\input{acks.tex}
 
@@
 
-\balance
-\begin{appendix}
-You can use an appendix for optional proofs or details of your evaluation which are not absolutely necessary to the core understanding of your paper.
+\input{legal.tex}
+
+\clearpage
+\input{appendix}
 
@@
 
-\end{appendix}
-
 \end{document}

### 2020-02-27 yfcc100m/paper/vldb2020/acks.tex
Paper submitted Feb 26

@@
 helped us through the process, as this work is the results of many efforts.
-We want to specially thank our senior technologists Nilesh Jain and Ravi Iyer for their
-full support and input during the duration of the project.
+We want to specially thank our senior technologists Nilesh Jain and 
+Ravi Iyer for their full support and input during the duration of the project.
 We want to specially acknowledge Philip Lantz and Vishakha Gupta for their help
-with PMGD, key to load large datasets into VDMS.
+with PMGD, key to loading large datasets into VDMS.
 We want to thank Jim Blakley for his input during the various 
-phases of our project, and for advocating and promoting its use.
+phases of our project, and for advocating and promoting VDMS.
 We want to acknowledge the Intel Labs VDMS team for their efforts
 open-sourcing and maintaining the system.
-We want to thank Jason Gardner for his helping in setting up many of the servers
-and infrastructure needed to conduct our experiments.
+We want to thank Jason Gardner for his helping in setting up many of 
+the servers and infrastructure needed to conduct our experiments.

### 2020-02-27 yfcc100m/paper/vldb2020/eval.tex
Paper submitted Feb 26

@@
 formats, and different parameters of the video itself that are
-application-dependent, like frames per second, lossy compression, etc).
+application-dependent, like frames per second, lossy compression, etc.
 When it comes to video, ad-hoc solutions have a large number of parameters that
@@
 are dependent on the database size and the \textit{connections} are responsible for
-90\% of the elements in each database instances,  
+90\% of the elements in each database instance,  
 as shown in Table~\ref{table:vdmsnodes}.
@@
 but the data is represented as three tables, following the relational model:
-1) \textit{images} table, contains one row per image,
+1) \textit{images} table: contains one row per image,
 and a column for each property
@@
 we explore several parameters to increase the performance of both loading the data, 
-as well executing the queries.
+as well as executing the queries.
 In particular, MySQL optimizes threads and transactions out-of-box, 
@@
 \includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Concurrency (left) and Overhead (right)}
+\caption{Analysis of video operations.
+The left figure shows the video throughput (videos per sec) as the number of concurrent clients increase and the right figure breaks down the different components of the
+queries using 32 clients.}
 \label{fig:video}

### 2020-02-27 yfcc100m/paper/vldb2020/intro.tex
Paper submitted Feb 26

@@
 hard to evolve the schema on the fly.
-As another example, many applications that can be studied through the use of large
+As another example, many applications can be studied through the use of large
 and publicly available datasets. 
@@
 For this work, we use the YFCC100M dataset\cite{Thomee_2016}. 
-The YFCC100M is the largest publicly multimedia collection. 
+The YFCC100M is the largest public multimedia collection. 
 It contains the metadata of around 99.2 million photos 

### 2020-02-27 yfcc100m/paper/vldb2020/legal.tex
Update to overleaf version
Paper submitted Feb 26

@@
 \begin{framed}
+\noindent \textbf{Notices \& Disclaimers}
 
-2020 Intel Corporation.
+\copyright  Intel Corporation.  Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.  Other names and brands may be claimed as the property of others.
 
-Intel and Intel Core are trademarks of Intel Corporation in
-the U.S. and/or other countries. Other names and brands may be claimed as the property of others.
+Performance results are based on testing as of dates shown in configurations and may not reflect all publicly available security updates.
+See backup for configuration details.
+No product or component can be absolutely secure.
 
-Performance results are based on testing as of 10/02/2020 and may not reflect
-all publicly available security updates.
-See configuration disclosure for details.
-No product can be absolutely secure.
+Your costs and results may vary.
 
-Software and workloads used in performance tests may have been optimized for
-performance only on Intel microprocessors. Performance tests, such as SYSmark
-and MobileMark, are measured using specific computer systems, components,
-software, operations and functions.
+Intel technologies may require enabled hardware, software or service activation.
+
+Software and workloads used in performance tests may have been optimized for performance only on Intel microprocessors.
+Performance tests, such as SYSmark and MobileMark, are measured using specific computer systems, components, software, operations and functions.
 Any change to any of those factors may cause the results to vary.
-You should consult other information and performance tests
-to assist you in fully evaluating your contemplated purchases, including the
-performance of that product when combined with other products.
-For more information go to http://www.intel.com/performance.
+You should consult other information and performance tests to assist you in fully evaluating your contemplated purchases, including the performance of that product when combined with other products.
+For more complete information visit www.intel.com/benchmarks.
 
-Intel's compilers may or may not optimize to the same degree for non-Intel
-microprocessors for optimizations that are not unique to Intel microprocessors.
-These optimizations include SSE2, SSE3, and SSSE3 instruction
-sets and other optimizations. Intel does not guarantee the availability,
-functionality, or effectiveness of any optimization on microprocessors not
-manufactured by Intel. Microprocessor-dependent optimizations in this product
-are intended for use with Intel microprocessors. Certain optimizations not
-specific to Intel microarchitecture are reserved for Intel microprocessors.
-Please refer to the applicable product User and Reference Guides for more
-information regarding the specific instruction sets covered by this notice.
+Intel's compilers may or may not optimize to the same degree for non-Intel microprocessors for optimizations that are not unique to Intel microprocessors.
+These optimizations include SSE2, SSE3, and SSSE3 instruction sets and other optimizations.
+Intel does not guarantee the availability, functionality, or effectiveness of any optimization on microprocessors not manufactured by Intel.
+Microprocessor-dependent optimizations in this product are intended for use with Intel microprocessors.
+Certain optimizations not specific to Intel microarchitecture are reserved for Intel microprocessors.
+Please refer to the applicable product User and Reference Guides for more information regarding the specific instruction sets covered by this notice.
+Notice Revision #20110804
 \end{framed}

### 2020-03-30 yfcc100m/python/eval/mysql_eval/MySQLQuery.py
Updates on Queries, Add OR to VDMSQuery

@@ -29,22 +29,22 @@ class MySQL(object):
         self.db = self.get_connection(params)
         # creating database_cursor to perform SQL operation
         self.db_cursor = self.db.cursor()
-        
+
     def close_connection(self):
         self.db.close()
-        self.db_cursor.close()         
+        self.db_cursor.close()
 
     def get_connection(self, params):
         return mysql.connector.connect(host=params.db_host, user=params.db_user, passwd=params.db_pswd, port=params.db_port, database="yfcc_" + params.db_name)
 
     def get_metadata_by_tags(self, tags, probs, lat=-1, long=-1, range_dist=0, return_response=True):
-        
+
         if lat != -1:
             location_qstr = '''(latitude >= {} AND latitude <= {} ) AND (longitude >= {} AND longitude <= {})'''.format(lat-range_dist*1.0,lat+range_dist*1.0,long-range_dist*1.0,long+range_dist*1.0)
             qstr = ['''id IN (select id from (test_metadata INNER JOIN test_autotags a on test_metadata.id=a.metadataid and a.probability >={} AND {} and a.tagid=(select test_taglist.tagid from test_taglist where test_taglist.tag='{}' limit 1)))'''.format(prob, location_qstr, tag) for prob,tag in zip(probs, tags)]
         else:
             qstr = ['''id IN (select id from (test_metadata INNER JOIN test_autotags a on test_metadata.id=a.metadataid and a.probability >={} and a.tagid=(select test_taglist.tagid from test_taglist where test_taglist.tag='{}' limit 1)))'''.format(prob, tag) for prob,tag in zip(probs, tags)]
-        
+
         query = '''SELECT line_number, download_url, id, latitude, longitude, license_name FROM test_metadata WHERE {}'''.format(''' AND '''.join(qstr))
 
         start_t = time.time()
@@ -76,9 +76,6 @@ class MySQL(object):
         start = time.time()
         metadata = self.get_metadata_by_tags(tags, probs, lat, long, range_dist)
 
-        height = operations[0]["height"]
-        width = operations[0]["width"]
-
         img_array = []
         cols = ['line_number', 'download_url', 'id', 'latitude', 'longitude', 'license_name']
 
@@ -94,30 +91,37 @@ class MySQL(object):
             imgPath = "http://"+IMG_HOST+"/images/" + urlparse(res[cols.index('download_url')]).path
 
             try:
-                # img = np.frombuffer(open(imgPath, 'rb').read(), dtype='uint8')
                 imgdata = requests.get(imgPath)
[…]

### 2020-03-30 yfcc100m/python/eval/vdms_eval/VDMSQuery.py

@@ -7,6 +7,8 @@ import random
 
 import vdms
 
+OR_OPERATION = True
+
 def create_dir(folder):
     if not os.path.isdir(folder):
         os.mkdir(folder)
@@ -246,38 +248,83 @@ class VDMSQuery(object):
                            lat=-1, long=-1, range_dist=0, return_images=True):
 
         if len(tags) > 1:
-            start = time.time()
-            results = self.get_metadata_by_tags(tags, probs,
-                                                lat, long, range_dist)
 
-            all_cmds = []
+            if OR_OPERATION:
+                all_cmds = []
 
-            for ele in results:
-                fI = {
-                    "FindImage": {
-                        "constraints": {
-                            "ID":  ["==", ele['ID']]
-                        },
-                        # "results": {
-                        #     "list": ["ID"]
-                        # }
+                ref = 1
+                for (tag, prob) in zip(tags,probs):
+                    fE = {
+                        "FindEntity": {
+                            "_ref": ref,
+                            "class": "autotags",
+                            "constraints": {
+                                "name": ["==", tag]
+                            }
+                        }
                     }
-                }
 
-                if len(operations) >= 0:
-                    fI["FindImage"]["operations"] = operations
+                    fI = {
+                        "FindImage": {
+                            "link": {
+                                "ref": ref,
+                                "constraints": {
+                                    "tag_prob": [">=", prob]
+                                }
+                            }
+                        }
+                    }
 
-                all_cmds.append(fI)
+                    if (lat != -1):
+                        fI["FindImage"]["constraints"] = {
+                            "Latitude":  [">=", lat  - range_dist*1.0,
+                                          "<=", lat  + range_dist*1.0  ],
+                            "Longitude": [">=", long - range_dist*1.0,
+                                          "<=", long + range_dist*1.0  ]
+                        }
 
-            # start = time.time()
-            responses, blobs = self.db.query(all_cmds)
-            end_time = time.time() - start
+                    all_cmds.append(fE)
+                    all_cmds.append(fI)
 
-            # if lat == -1:
-            #     out_file = open("perf_results/vdms_img_list.txt", 'w')
-            #     for ent in responses:
-            #         out_file.write(str(ent["FindImage"]["entities"][0]["ID"]))
-            #         out_file.write("\n")
+                    ref += 1
+
+                start = time.time()
+                responses, blobs = self.db.query(all_cmds)
[…]

### 2020-03-30 yfcc100m/paper/vldb2020/design.tex

@@
 By defining a new JSON-based API, there is a trade-off between
-expressiveness (compared to SPARQL or Gremlim, or even SQL) and
+expressiveness (compared to SPARQL, Gremlim, or even SQL) and
 the ability to natively support visual data operations.

### 2020-05-22 yfcc100m/python/eval/plot_performance.py
Update plotting script

@@ -60,9 +60,9 @@ newpath = str(params.outfile.parents[0])
 if not os.path.exists(newpath):
     os.makedirs(newpath)
 
-color = ['red', 'blue', 'orange', 'red', 'blue', 'orange']
-linestyles = ['-', '-', '-', '--', '--', '--', ]
-markers = ['o', 'o', 'o', '*', '*', '*']
+# color = ['red', 'blue', 'red', 'blue', 'orange', 'orange']
+# linestyles = ['-', '-', '--', '--', '-', '--', ]
+# markers = ['o', 'o', '*', '*', 'o', '*']
 plot = "all"
 line_counter = 0
 
@@ -139,6 +139,28 @@ imgs_std = imgs_std.transpose()
 
 x_pos = [value_to_float(i) for i in db_sizes]
 
+if len(imgs_[0, :]) == 2:
+    color = ['orange', 'black']
+    linestyles = ['-', '--']
+    markers = ['o', '*']
+elif len(imgs_[0, :]) == 4:
+    color = ['red', 'blue', 'red', 'blue']
+    # color = ['orange', 'black', 'orange', 'black']
+    linestyles = ['-', '-', '--', '--']
+    markers = ['o', 'o', '*', '*']
+elif len(imgs_[0, :]) == 6:
+    color = ['red', 'blue', 'orange', 'red', 'blue', 'orange']
+    linestyles = ['-', '-', '-', '--', '--', '--', ]
+    markers = ['o', 'o', 'o', '*', '*', '*']
+elif len(imgs_[0, :]) == 8:
+    color = ['red', 'blue', 'orange', 'black', 'red', 'blue', 'orange', 'black']
+    linestyles = ['-', '-', '-', '-', '--', '--', '--', '--',]
+    markers = ['o', 'o', 'o', 'o', '*', '*', '*', '*']
+
+color      = ['red', 'blue', 'orange', 'black', 'pink', 'brown', 'violet', 'green']
+linestyles = ['-', '--']
+markers    = ['o', '*']
+
 """
 Plot Tx/sec
 """
@@ -152,23 +174,27 @@ else:
 
 plt.rc('lines', linewidth=1)
 
+total_queries = int(len(imgs_sec[0, :]) / 2)
 for i in range(0, len(tx_sec[0, :])):
     ax0.errorbar(x_pos, tx_sec[:, i],
                  yerr=tx_sec_std[:, i],
                  label=query_name[i],
[…]

### 2020-05-22 yfcc100m/paper/osdi2020/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for USENIX papers.
%
% History:
%
% - TEMPLATE for Usenix papers, specifically to meet requirements of
%   USENIX '05. originally a template for producing IEEE-format
%   articles using LaTeX. written by Matthew Ward, CS Department,
%   Worcester Polytechnic Institute. adapted by David Beazley for his
%   excellent SWIG paper in Proceedings, Tcl 96. turned into a
%   smartass generic template by De Clarke, with thanks to both the
%   above pioneers. Use at your own risk. Complaints to /dev/null.
%   Make it two column with no page numbering, default is 10 point.
%
% - Munged by Fred Douglis <douglis@research.att.com> 10/97 to
%   separate the .sty file from the LaTeX source template, so that
%   people can more easily include the .sty file into an existing
%   document. Also changed to more closely follow the style guidelines
%   as represented by the Word sample file.
%
% - Note that since 2010, USENIX does not require endnotes. If you
%   want foot of page notes, don't include the endnotes package in the
%   usepackage command, below.
% - This version uses the latex2e styles, not the very ancient 2.09
%   stuff.
%
% - Updated July 2018: Text block size changed from 6.5" to 7"
%
% - Updated Dec 2018 for ATC'19:
%
%   * Revised text to pass HotCRP's auto-formatting check, with
%     hotcrp.settings.submission_form.body_font_size=10pt, and
%     hotcrp.settings.submission_form.line_height=12pt
%
%   * Switched from \endnote-s to \footnote-s to match Usenix's policy.
%
%   * \section* => \begin{abstract} ... \end{abstract}
%
%   * Make template self-contained in terms of bibtex entires, to allow
%     this file to be compiled. (And changing refs style to 'plain'.)
%
%   * Make template self-contained in terms of figures, to
%     allow this file to be compiled.
%
%   * Added packages for hyperref, embedding fonts, and improving
%     appearance.
%
%   * Removed outdated text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2019_v3}

% to be able to draw some self-contained figs
\usepackage{graphicx}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{hyperref}       % hyperlinks
\usepackage{graphicx}
\usepackage{comment}
\usepackage{multicol}
\usepackage{framed}
\usepackage{subcaption}
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{minted}

% inlined bib file
\usepackage{filecontents}

%-------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------

%don't want date printed
\date{}

% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Using VDMS to Index and Search 100M Images}

%for single author (just remove % characters)
\author{
{\rm Luis Remis}\\
ApertureData Inc
\and
{\rm Chaunt\'e W. Lacewel}\\
Intel Labs
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle

%-------------------------------------------------------------------------------
\begin{abstract}
%-------------------------------------------------------------------------------
Data scientists spend most of their time dealing with data preparation,
rather than doing what they know best:
build machine learning models and algorithms to solve problems.
The problem grows as the scale of the data used for analysis or training
grows. In this paper, we demonstrate how VDMS (the Visual Data Management System)
can be used to simplify the data preparation process and how we gain
in efficiency simply because we are using a system designed for the job.
To demonstrate this, we use one of the largest available
public datasets (YFCC100M),
with 99+ million images, 800+ thousand videos, plus add-ons that include
machine-generated tags and 4K dimensional feature vectors,
for a total of $\sim$13TB of data.
VDMS differs from existing large scale photo
serving, video streaming, and textual big-data management systems due to
its primary focus on supporting machine learning and
data analytics pipelines that use visual data
(images, videos, and feature vectors), treating these as first class citizens.
We demonstrate how VDMS outperforms well-known and widely used
systems for data management by up to $\sim$35x, with
an average improvement of about 15x for the queries we
used in our use-cases, and particularly at scale.
At the same time, VDMS simplifies the process of data preparation and data access,
and provides functionalities non-existent in alternative options.
\end{abstract}

\input{intro.tex}
\input{design.tex}
\input{eval.tex}
\input{conclusion.tex}
\input{acks.tex}

\clearpage

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{plain}
\bibliography{main}

\input{legal.tex}

\clearpage
\input{appendix}

%-------------------------------------------------------------------------------
\section*{Availability}
%-------------------------------------------------------------------------------

USENIX program committees give extra points to submissions that are
backed by artifacts that are publicly available. If you made your code
or data available, it's worth mentioning this fact in a dedicated
section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
%%  LocalWords:  pdflatex acks

### 2021-01-20 yfcc100m/paper/vldb2021/eval-video.tex
Add vldb2021 paper template

--- 2020-02-18 yfcc100m/paper/vldb2020/eval.tex
@@
-\section{Evaluation}
-\label{eval}
-
-We have used the YFCC100M dataset to evaluate different aspects of our system.
-We use the images in the dataset and its associated metadata to implement
-an image-search engine based on properties associated to those images.
-This is a very common use-case we have encountered when attacking
-applications such as smart-retail, sports applications, and video summarization,
-where the starting point is a large set of data that must be curated
-before proceeding with the data processing and neural network training.
-In order to evaluate the different aspects of the performance on the
-image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
-the data search problem, which we describe in Section \ref{images}.
-
-We also cover evaluation we have done using the Video interface VDMS offers,
-in order to asses how the implementation performs and scale when used
-for video handling.
-When it comes to video, ad-hoc solutions have a large number of parameters that
-can be tunned, and complex ways to handle the data. Together with that,
-there is no system that enables transactional operations over videos files
-in the way VDMS does. Because of the two reasons above mentioned,
-we focus our efforts in understanding variations in the the performance of
-VDMS and its scalability, rather than comparing it to a baseline that
-would no represent a fair comparison for either of the systems.
-We describe the results in Section \ref{videos}.
-
-Similarly, we explore the behavior of the feature vector functionality in VDMS,
-and an evaluation of the different trade-offs that the systems offers for
-applications developers. For this, we implemented an image-search application
-based on \textit{similarity} search, which we describe in Section \ref{features}
-
-\subsection{YFCC100M Dataset}
-\label{dataset}
-
-The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
-collection of 100 million public Flickr media objects created to provide free,
-sharable multimedia data for research. This dataset contains approximately
-99.2 million images and 0.8 million videos with metadata characterized by
-25 fields such as the unique identifier, userid,
-date the media was taken/uploaded, location in longitude/latitude coordinates,
-device type the media was captured, URL to download the media object,
-and the Creative Commons license type information.
-The YFCC100M dataset also used a deep learning approach to generate autotags
-which is a set of comma-separated concepts such as people, scenery, objects,
-and animals and confidence scores generated from 1,570 trained
-Caffe classifiers ~\cite{Thomee_2016}. Together with the autotag, there is a
-probability associated to each tag to indicate the certainty level of the tag
-in that image/frame.
-We have also used feature vectors generated for every image and first frame
-of every video \cite{features} to implement similarity search.
-Given that there is no standard benchmark oriented towards visual data queries,
-we have built a series of queries to filter this dataset that is model after
-our internal use cases for many of the mentioned applications we have worked
-with.
-
-\subsection{Experimental Setup}
-
-As a baseline, and given that there are no other open-source system
-that implements similar functionality and interfaces as VDMS,
-we implemented an equivalent visual data management system comprised of a
-combination of widely available, off-the-shelve components:
-MySQL Server 5.7 (for storing metadata),
-Apache Web Server 2.4.18 (as interface for image access), and
-OpenCV 3.3 (to provide pre-processing operations on images).
-We have implemented a set of client-side applications that takes care
-of retrieving the components from the different systems, and applies
-pre-processing operations when needed.
-For visual data workloads, building an approach like the one
-implemented as a baseline for this work
-is the common practice in the industry \cite{heystack, tao}.
-
-For all our experiments, we use 2 servers, one hosting a VDMS server and
-another hosting a the Baseline implementation.
-Both servers have a dual-socket Intel\textsuperscript{\textregistered}
-Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
-each CPU with 28 physical cores with multithreading enabled,
-for a total of 112 logical cores per server.
-The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
-has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
-The client application (running the queries and measuring round-trip time)
-is connected to the server through a 1GB wired link through
-a 10GB-backplane switch.
-
-Figure~\ref{fig:systems} shows a logical view of the difference between the
-interaction of the client application (which wants to retrieve metadata and
-imaged) with VDMS (left) and the baseline (right).
-The client application was implemented using Python 3 for both VDMS and
-the baseline.
-
-It is worth noting that the images are stored in a shared repository
-(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
-and VDMS have direct access. In the case of the baseline, metadata is
-stored in MySQL using an attached SSD disk.
-Even if VDMS has native support for Optane Persistent Memory,
-we do not use it in this experiment because of fairness of comparison w.r.t
-MySQL (which was not design for Persistent Memory type of storage).
-The benefits of Persistent Memory on metadata operations is left
-for another paper, and outside the scope of this evaluation.
-For this experiment then, in the case of VDMS we simply use a similar
-attached SSD disk to store metadata.
-Even if PMGD, the graph database used by VDMS, is designed for PM,
-it can deliver good performance when using SSDs directly.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/comparison_system}
-\caption{Comparison Systems}
-\label{fig:systems}
-\end{figure*}
-
-For the metadata, we built VDMS and MySQL databases
-using the YFCC100M dataset with incremental database sizes.
-For simplicity, we named the database based on the approximate number of images
-it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
-The exact number of images/elements on each database is shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
-For each database size, we created an instance of VDMS using YFCC100m metadata,
-the YFCC100m machine-generated \textit{autotags} associated with
-each metadata identifier, and the list of 1,570 \textit{autotags}
-using 100 concurrent VDMS Python clients.
-Internally, that information is represented as a property graph,
-where we have one node for each image, one node for each tag
-(always 1570 tags), and connections between images.
-For instance, if an image has 4 \textit{autotags} assigned,
-there will be 4 connections between that image and
-the different nodes for those \textit{autotags}.
-The information about the probability the \textit{autotag} present in an image
-is expressed as a property in the \textit{connection} between the 2 nodes.
-Figure \ref{fig:graph_representation} shows an example on two images and
-2 \textit{autotags}, and the \textit{connections} between those \textit{autotags}
-and the images.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Image id: 86756231, on the other hand, has a single \textit{autotags} assigned:
-\textit{Alligator} with probability 0.894.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/graph_representation}
-\caption{VDMS Data Representation Using a Property Graph:
-Example on two images and 2 \textit{autotags} with
-their respective probabilities expressed in the \textit{connection}.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Similarly, Image id: 86756231 has a single \textit{autotags} assigned:
-\textit{Alligator} with probability 0.894.}
-\label{fig:graph_representation}
-\end{figure}
-
-Because there are on average 8 tags assigned to each image,
-there will be around 8 times more connections than images, as can be seen
-in Table~\ref{table:vdmsnodes}.
-Also, each image node will contain multiple properties associated
-with that object.
-Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
-and then insert nodes for the YFCC media objects (images or videos)
-with the associated metadata.
-Adding the autotag confidence scores is a recursive process and requires more
-complex queries than adding the previous nodes.
-For each metadata identifier in \textit{autotags},
-the query must find the associated metadata node and tag node.
-A connection between the two nodes is created with the confidence score as a
-property. The number of metadata nodes are dependent on the database size and
-the connections are responsible for 90\% of the elements in each database as
-shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, \textit{autotags}, and
-longitude/latitude coordinates are set as indexes of the database to allow
-faster retrieval of the metadata.
-
-\begin{table}[h]
-\caption{VDMS Database - Number of Elements}
-\centering
-\begin{tabular}{c c c c}
-\hline\hline
-DB Name & \# Images & \# Connections & \# Autotags\\
-\hline
-100k & 100,000     & 848,432      & 1,570\\
-500k & 500,000     & 4,249,500    & 1,570\\
-1M   & 1,000,000   & 8,503,045    & 1,570\\
-5M   & 5,000,000   & 42,505,478   & 1,570\\
-10M  & 10,000,000  & 85,040,404   & 1,570\\
-50M  & 50,000,000  & 425,162,070  & 1,570\\
-100M & 99,205,984  & 895,572,430  & 1,570\\
-\hline
-\end{tabular}
-\label{table:vdmsnodes}
-\end{table}
-
-Each MySQL database is created in a similar manner but the data is represented
-as three tables, following the relational model:
-1) \textit{images} table, contains one row per image,
-and a column for each property
-associated with the images (some of which are listed in \ref{dataset});
-2) \textit{taglist} table: contains one row per each autotag elements
-(always 1570 rows);
-3) \textit{autotags} table: contains on row per each autotag
-assigned to an image, each row containing a foreign key to the
-image and a foreign key to the tag, and
-the probability assigned to that tag belonging to that image.
-Given that there are 8 autotags, on average, per image, the \textit{autotags}
-table has around 8 times the number of rows present in the
-\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.
-
-THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
-By default, MySQL uses infinite number of threads inside InnoDB when processing
-requests using four threads per IO read/write.
-In some cases, when creating large databases data locks may occur to protect the
-data from concurrent updates.
-To minimize locking in our runs, we increased the InnoDB buffer pool size to
-increase the amount of memory allocated to in-memory data structures
-~\cite{mysql,mysql_blog}.
-
-Using a Python client and simple queries, the \textit{taglist}
-table is read from the list of tags with an auto-incremented
-\textit{tagid} as a primary key and the metadata table
-is read from the YFCC100m metadata using the identifier as a primary key.
-The \textit{autotags} table contains the generated autotags and
-probabilities for entries of the \textit{images} table.
-
-To generate the table, we split the \textit{autotags} data for each database
-by the image identifier and autotag into new files.
-The new files are read into the \textit{autotags} table with the image
-identifier and \textit{tagid} as foreign keys.
-
-\begin{table}[h]
-\caption{MySQL Database - Number of Rows in each Table}
-\centering
-\begin{tabular}{c c c c}
-\hline\hline
- & \multicolumn{3}{c}{Table}\\
-\cline{2-4}
-DB Name & images & autotags & taglist\\
-\hline
-100k & 100,000    & 848,912     & 1,570\\
-500k & 498,707    & 4,241,200   & 1,570\\
-1M   & 1,000,000  & 8,508,380   & 1,570\\
-5M   & 4,987,379  & 42,425,905  & 1,570\\
-10M  & 10,000,000 & 85,095,265  & 1,570\\
-50M  & 50,000,000 & 425,446,208 & 1,570\\
-100M & 99,206,564 & 896,002,496 & 1,570\\
-\hline
-\end{tabular}
-\label{table:mysqltables}
-\end{table}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/db_time_size}
-\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
-\label{fig:db_time_size}
-\end{figure}
-
-The VDMS and MySQL databases have comparable number of elements as shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
-can be attributed to failures in data preparation/loading because of
-incomplete/inconsistent formating, which is common on large datasets.
-The difference is less than 0.1\% in terms of number of elements (images and/or
-metadata information).
-
-\subsubsection{Database Loading Time}
-
-One of the first things we noticed is the difference is loading times,
-where VDMS outperforms MySQL by a large margin.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
-faster than MySQL, and how the speedup is sustained as the database size grows.
-Key difference in the build times are attributed to the low-level
-implementation of how MySQL reads and stores data from the files and the
-optimizations (increased InnoDB pool size, etc.) needed
-to handle large datasets such as YFCC100m.
-On average, it took MySQL 3.72x longer to build each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours
-while VDMS only needed 72.5 hours.
-
-\subsubsection{Database Storage Footprint}
-
-Another aspect important to note is that
-VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
-This is space used to store information about each node/connection.
-The Graph Database internal to VDMS (called PMGD) was designed for performance,
-specially in environments where persistent memory is present.
-This design decision comes as a trade-off for storage footprint, which is
-noticeable in our results.
-VDMS required 30-41\% more storage than MySQL for storing the same amount
-of metadata.
-This may become a factor if storage is a limitation, but it should also be noted
-that even if we have a 41\% increase in metadata size,
-metadata accounts for less than 2\% of the overall database size.
-Take, for example, the largest database (both metadata and images) we built
-(100M): around 230GB of metadata and 12TB of images.
-On the other hand, in systems were persistent memory is a scarce resource,
-the increase storage foot print of PMGD may represent a challenge.
-
 %=========================================
-
-\subsection{Images + Metadata}
-\label{images}
-
-In order to evaluate VDMS and the baseline on our use-case queries,
-we implemented 4 queries that that filter and retrieve certain specific images.
-We choose these queries because they represent typical use-case where a
-cohort of images is to be retrieve and processed from a large corpus of data.
-As we mentioned before, we took this approach due to the lack of a standard
-benchmarks that are oriented towards visual data retrieval.
-We use the metadata associated to the images to filter said images.
-In particular, this dataset contains a set of 1570 machine generated
-tags (\textit{autotags})assigned to the images depending on the content.
-\textit{autotags} includes things like objects (chair, person, people, etc), and
-concepts like places (indoor, outdoor, lake) and others (party, etc).
-Together with each \textit{autotag}, a probability is assign.
-This is, an image can have the \textit{autotags} people, person, party,
-outdoor, and each \textit{autotag} assigned will be accompanied by a
-probability of that \textit{autotags} being present in that image.
-As mentioned earlier, there are 8 tags assigned to each image on average.
-Also, some images have geo-location information (latitude/longitude).
-
-We use the \textit{autotags} (as they contain information about
-the content of the image), as well as the geo-location
-(as an example of properties of the images used used for search and filtering).
-
-On top of that and for our use case, we would like to extract more information
-about the content of the image through the use of ML,
-such as Convolutional Neural Networks TODO-CITE HERE.
-For this, we resize the images to 224x224, which is the input layer size for
-a popular neural network for object detection on images: ResNet TODO-CITEHERE.
-
-To evaluate the access to metadata and images,
-we use the following four queries, modeled after our internal use cases:
-\begin{itemize}
-\item Query 1 - {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
-\item Query 2 - {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
-\item Query 3 - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
-\item Query 4 - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
-\end{itemize}
-
-It is important to note that when querying for images with certain
-\textit{autotags}, we also apply a filter using the probability.
-For instance, we only retrieve images with a autotag \textit{alligator}
-and a probability higher than 92\%.
-This probabilities are both present in VDMS (in the form of a property
-of the \textit{connection} between the image and that \textit{autotag}),
-as well as in MySQL (in the form of a column in the \textit{autotags} table
-that links images with tags.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on Query \textit{1tag\_resize} \- VDMS vs Baseline}
-\label{fig:concurrency_comparison}
-\end{figure}
-
-% We decided to remove this image because it is redudant with the concurrency
-% analysis.
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
-% \caption{Queries - Throughput - 32 Clients}
-% \label{fig:q_throughput_32}
-% \end{figure}
-
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_56}
-% \caption{Queries - Throughput - 56 Clients}
-% \label{fig:q_throughput_56}
-% \end{figure}
-
-\begin{figure*}[ht]
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include first image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
-  \caption{Query 1 (red), and Query 2 (blue).}
-  \label{fig:q1_q2}
-\end{subfigure}
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include second image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
-  \caption{Query 3 (red), and Query 4 (blue).}
-  \label{fig:q3_q4}
-\end{subfigure}
-\caption{Queries - Throughput - 56 Clients}
-\label{fig:q_throughput_56}
-\end{figure*}
-
-Image search based on metadata is very expensive on large databases.
-Because of the large volume of data, the processing of the retrieve images
-is perform in parallel, using multi-core and/or distributed systems.
-For instance, a common implementation of an image processing pipeline
-would involve the use of distributed processing frameworks
-like Hadoop \cite{hadoop} or Spark \cite{spark}.
-Consequently, it is key that the data management system used supports
-concurrency, providing multiple workers with data in parallel.
-The ability to scale with the number of simultaneous clients is key for the
-applicability of visual data management systems like VDMS.
-
-Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
-the \textit{1tag\_resize} query, described above, using both VDMS and
-the baseline to evaluate the scalability of both solutions, as the
-size of the databases grows and as the number of concurrent clients grows.
-
-It is important to note that the size of the result (number of images retrieved)
-is linear with the size of the database. This is, if a query returns 100 images
-for the 1M database, it will return around 1000 images for the 10M database.
-This poses a problem when evaluating performance as database increase,
-and clearly understanding the measurements.
-Because of this reason, we control the number of returned images for all the
-database using the probability of the \textit{autotags}, so that the queries
-in this experiment returns a similar number of images for all database sizes.
-Essentially, as the size of the database increase, we increase the probability
-threshold for the queries. We do this for both VDMS and the baseline, of course.
-
-The first thing to notice is that when it comes to retrieving metadata
-in the case of VDMS, (Figure ~\ref{fig:concurrency_comparison} (top),
-the size of the database (in terms of number of images) does not have
-a strong effect on the performance
-(measured in aggregated transactions per second).
-This means that even as database increase in size, there is no large increase
-in the query time.
-
- the number of concurrent clients to use for the per-query performance analysis.
-This figure shows the metadata performance plateaus at 32 clients and
-the image performance at 56 clients for a simple query.
-% To further investigate, Figure~\ref{fig:q_throughput_32} and
-% Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
-% for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
-% respectively.
-On the 56-client experiment, a 10x difference can be seen between VDMS and
-baseline in metadata throughput while 32 clients is similar but not as high.
-The overall performance for both MySQL and VDMS is better with 56 clients;
-therefore, we used this value for our per-query analysis.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/q1_latency}
-\caption{Query 1 - Latency}
-\label{fig:q1_latency}
-\end{figure}
-
-%=========================================
-
-\subsection{Videos}
+\subsection{Video Search}
 \label{videos}
 
-VDMS provides full support for video storage and operations.
-This includes support for encoding/decoding and trans-coding of
+VDMS provides full support for video storage and operations,
+in a similar way it does for images.
+This includes support for encoding, decoding, and transcoding of
 \textit{mp4}, \textit{avi}, and \textit{mov} containers,
@@
 This is supported through the Visual Compute Module that provides an abstraction
-layer on top of OpenCV (and \textit{ffmpeg} directly).
-All the operations supported for images in VDMS are also supported at the
-video and frame level on the API. On top of that, there are a number of
-video-specific operations that are supported, such as the interval operations,
-that allow users to retrieve clips at different
+layer on top of OpenCV~\cite{opencv} and \textit{libffmpeg}\cite{ffmpeg}.
+All operations supported for images in VDMS are also supported at the
+video and frame level of the API.
+On top of that, there are a number of video-specific operations that
+are supported, such as the interval operations,
+enabling users to retrieve clips at different
 frames-per-second (FPS) versions of the video.
 
+We evaluate the performance and scalability of the video management
+capabilities offered by VDMS.
+Handling video in a general way is a complex task.
+Some of this complexity comes from the existence of a variety of open and
+proprietary implementations, different encoding techniques and container
+formats, and different parameters of the video itself that are
+application-dependent, like frames per second, lossy compression, etc.
+When it comes to video, ad-hoc solutions have a large number of parameters that
+can be tuned.
+Together with that, there is no system that enables transactional
+operations over videos files in the way VDMS does.
+Because of this, we focus our efforts on understanding variations in the
+performance of VDMS and its scalability, rather than comparing
+it to a baseline that would not represent a fair
+comparison for either of the systems.
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=\textwidth]{figures/video_overhead}
+\caption{Analysis of video operations.
+The left figure shows the video throughput (videos per sec) as the number of concurrent clients increase and the right figure breaks down the different components of the
+queries using 32 clients.}
+\label{fig:video}
+\end{figure*}
+
 All this functionality is provided and integrated with the rest of the
-metadata API as part of the comprehensive VDMS interface. This makes it
-possible for users to interact with metadata and video in a transactional
-manner, enabling users to run queries like: "Retrieve all the videos
-where there is a lake with probability higher than X, converting all videos
-to \textit{H.264} \textit{mp4} of size 224x244".
+metadata API as part of the comprehensive VDMS interface.
+This makes it possible for users to interact with metadata and video in 
+a transactional manner, enabling users to run queries like: 
+"Retrieve all the videos where there is a \textit{lake} with 
+probability higher than 0.86, converting all videos to \textit{H.264} 
+\textit{mp4} of size 224x224".
+Appendix shows a sample of how this query would be implemented using the VDMS API
+\footnote{https://github.com/\{Not shown during submission\}}.
+% \footnote{https://github.com/IntelLabs/vdms/wiki/FindVideo}.
 In particular, this functionality was used internally to select a subset
@@
 To the best of our knowledge, there is no solution that can provide
-all the functionality mentioned above. Also, implementing a baseline is a very
-complex task as there is a large number of options and parameters that can
-be chosen, which makes it hard to accurately compare against VDMS functionality.
+all the functionality mentioned above, behind a single interface
+that also allows users to interact with images and metadata.
+Implementing a baseline, like we did for images, is significantly more complex
+due to the parametrization of video encodings and containers, 
+as explained at the beginning of this section.
 For this reason, we chose to make a study using VDMS in various scenarios,
-and analysis what is the impact of having the overhead of VDMS' Request
-Server in the overall access time.
+and analysis of scalability and the impact of having the overhead of VDMS' Request
+Server in the overall access time and throughput.
 
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Concurrency (left) and Overhead (right)}
-\label{fig:video}
-\end{figure*}
-
-Figure \ref{fig:video} shows the analysis of different queries aimed to retrieve
-a video using VDMS interface. We show how VDMS increases throughput of serving
-a video object as the number of simultaneous client increases, as well as the
-overhead VDMS introduces in the overall query execution time.
+Figure~\ref{fig:video} shows the analysis of different queries aimed
+at retrieving a video using the VDMS interface.
+We show how VDMS throughput increases when serving
+a video object as the number of simultaneous clients increases, as well as the
+overhead operations introduced in the overall query execution time.
 The figure on the left compares the number of video transaction per second
@@
 
-We also run a set of other queries that involve: (a) running a resize operation
-on the video and, consequently, needs a decoding and
-encoding operations as well (blue line),
-(b) transcoding, meaning the use a different container and encoder
-than the one originally used (yellow line), and (c) both resize and transcoding.
-Note that the resize operation performs a downsize, which translates in less
-data being sent over the wire. This is specially noticeable when supporting 32
-simultaneous clients, where the systems provides more videos per second due to
-sending less data to the client, when compares to just transcoding and not resizing (yellow line).
+\begin{listing}[ht!]
+\begin{minted}[frame=single,
+               framesep=3mm,
+               linenos=true,
+               xleftmargin=21pt,
+               tabsize=4]{js}
+"FindEntity"{     
+    "class": "autotag",
+    "constraints": { 
+        "name": ["==", "lake"]
+    }
+    "_ref" : 1
+},
+"FindVideo":{
+    "container": "mp4",
+    "codec": "h.264",
+    "link": {
+        "ref":1,
+        "constraints": {
+            "prob": [">=", 0.86]
+        }
+    }
+    "operations": [{
+        "type": "resize",
+        "height": 1080,
+        "width":  1920,
+    }]
+}
 
-We can see that the system performs best when using all the physical cores.
-COMPLETE THIS BETTER.
+\end{minted}
+\caption{Sample Query for Video - 
+The query expresses the following: 
+Find all videos connected to the autotag \textit{lake} 
+with probability higher than 0.86, apply a resize operation
+to make the video 1920×1080, and convert to "mp4" file, 
+using H.264 encoding.} 
+\label{findvideo}
+\end{listing}
 
-WE SHOULD ADD HERE SAMPLES OF THE QUERYS TO SHOW HOW THESE ARE RETRIEVED.
+We also run a set of other queries that involve, showed in Figure~\ref{fig:video}:
+(a) running a resize operation on the video and, consequently,
+decoding and encoding operations as well (blue line),
+(b) transcoding, meaning the use of a different container and encoder
+than the one originally used (yellow line), and
+(c) both resize and transcoding.
+Note that the resize operation (blue and green lines) performs a downsize,
+which translates in less data being sent over the wire.
+This is specially noticeable when supporting 32 simultaneous clients, 
+where the system provides more videos per second due to sending less data to 
+the client, when compared to just transcoding and not resizing (yellow line).
+We can see that the system performs best when using all the physical cores,
+and this can be attributed to the compute-bound nature of video
+encoding, decoding, and processing.
 
-Because we see almost 3 orders of magnitude drop in performance when including
-operations as part of the query, we wanted to understand where most of the time
-was spent on the query, and optimize the Request Server and Visual Compute Module
+It is important to note an almost 3 orders of magnitude drop in performance
+when including operations as part of the query.
+We wanted to understand where most of the time was spent on the queries,
+and optimize the Request Server and Visual Compute Module
 if necessary. For this, we run the experiment shown at
-Figure \ref{fig:video} (right) which breaks down the different components of the
+Figure~\ref{fig:video} (right) which breaks down the different components of the
 queries. This figure shows that more than 97\% of the query execution is spent
 on encoding/decoding operations, which is well-known to be a
-heavy and compute intensive operation.
-On the one hand, this results show that VDMS introduced overhead for
-video operation is minimal. On the other hand, this result means a
-limit on the opportunities for optimization for video queries given
-that biggest time factors are accounted by encoding/decoding, which is
-outside the scope of VDMS.
-This result also was the inspiration point for one optimization we included
-in future versions of VDMS, which involves using \textit{ffmpeg} C++ API to
-limit the number of frames being encoding/decoding when possible.
+compute intensive operation\cite{videosgoogle}.
+On the one hand, this result shows that VDMS barely introduces any overhead. 
+On the other hand, this result means a limit on the opportunities 
+for optimization for video queries given that biggest time factors 
+are accounted by encoding/decoding, which is outside the scope of VDMS.
+This result was the call to action for one optimization we will include
+in future releases of VDMS, which involves using \textit{ffmpeg} C++ API to
+limit the number of frames being encoded/decoded when possible.
+This functionality will prevent encoding/decoding to happen on all frames
+when users only need to retrieve a subset of the frames in the video.
 
 %=========================================
-
-\subsection{Feature Vectors}
-\label{features}
-
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities or visual objects.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as \textit{descriptors}
-or \textit{visual descriptors}. We use these terms interchangeably.
-These descriptors can be classified and labeled and used to build search indexes,
-and there are many in-memory libraries that are designed for
-this task~\cite{flann, faiss}.
-Using the VDMS API, users can manage feature vector indexes,
-query previously inserted elements (images),
-run a k-nearest neighbor search (\textit{knn}), and express relationships
-between existing images or descriptors and
-the newly inserted descriptors.
-By natively supporting descriptors and \textit{knn},
-VDMS allows out-of-the-box classification functionalities for many applications.
-
-For this work, and as part of a comprehensive image search implementation,
-we have used 4096-dimensional descriptors extracted from every image
-(and first frame of every video) and created a collection of these feature
-vectors in VDMS to be able to perform similarity search (i.e., find
-images that are \textit{similar} to an query (input) image).
-\textit{Similarity} in this particular case is defined as closeness
-in a an 4096-dimensional space using euclidean distance as the metric.
-
-EXPLAIN HERE DIFFERENT ACCURACY/EXECUTIME TIME TRADEOFF.
-
-The process of loading descriptors in VDMS is simple.
-First, the user has to create a DescriptorsSet, using a single command.
-At creation of the DescriptorSet, the dimensionality of the descriptors
-is specified, together with the desired indexing method and the desired metric
-for computing distances (Euclidean Distance, \textit{L2},
-or Inner Product, \textit{IP}).
-Once the DescriptorSet is created, descriptors can be inserted to the set.
-After the descriptors are inserted, similarity search can be performed.
-Note that descriptors can be inserted later on, and any search will reflect
-the existence of all descriptors in the set.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/feature_img_results}
-\caption{Sample Results of Similarity Search}
-\label{fig:similarity}
-\end{figure*}
-
-Figure \ref{fig:similarity} shows 3 examples of a query image (on the left),
-and images returned as "similar" by VDMS.
-The query input is a descriptor generated after a query image. The "query"
-descriptor is sent to VDMS as part of the query, and VDMS use that descriptor
-to find similar one, and retrieve the images associated with those "similar"
-descriptors. We used this as an example and as a visual validation of the
-functionality and applicability in this particular dataset, but we also
-provide an analytical approach to accuracy and trade-offs in our system.
-It is important to note that the accuracy of the results is entirely tied
-to the quality of the descriptors chosen by the applications.
-VDMS is completely agnostic to information contained within the descriptor,
-and simply offers the interface to store and index them, but the quality
-of the similarity result will be tied to the quality of descriptor extraction
-that the application is using.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/features_alternatives}
-\caption{Feature Vector Evaluation: Trade-off between query execution speed
-and accuracy of the results, using ground-truth data for computing accuracy.
-For this evaluation, we query the 10 closest neighbors (k = 10), and compute
-accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
-results that is present within the top 10 computed neighbors).
-We average the query execution time and accuracy for 100 queries (nq = 100).}
-\label{fig:features_eval}
-\end{figure*}
-
-As mentioned before, VDMS provides different levels of customization of the
-indexes created for a descriptor set, that includes the indexing techniques
-and the metric for similarity.
-These different indexing techniques comes with different trade-offs in terms
-of speed of search and accuracy.
-VDMS aims to provide functionality that is agnostics to application-specific
-techniques, enabling functionality that is general to visual data processing
-applications.
-Figure \ref{fig:features_eval} shows an analysis at the different indexing
-techniques provided by VDMS and its trade-off between accuracy and query
-execution speed, for a single threaded client.
-For this evaluation, we query the 10 closest neighbors (k = 10), and compute
-accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
-results that is present within the top 10 computed neighbors).
-We average the query execution time and accuracy for 100 queries (nq = 100).
-The \textit{flat} index (red line) implements exact search and
-represents ground-truth, which explain why the accuracy is always 100\% on the
-right plot. The other indexes implement \textit{approximate search},
-which trade-offs between accuracy and speed of search~\cite{flann, faiss}.
-We have also tried the \textit{ivfflat} index (inverted file index), as well as
-\textit{LSH}-based indexes using a different number of bits per descriptor
-\footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
-Results show how \textit{ivfflat} is the fastest option by comes at the trade-off
-of about 30\% lost in accuracy, while simple brute-force search
-is among the slowest options at the expenses of 100\% accuracy,
-meaning exact search.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/features_disksize}
-\caption{Feature Collection Size in Disk}
-\label{fig:features_size_does_matter}
-\end{figure*}
-
-Another important trade-off to be made is w.r.t to space efficiency: Set of
-descriptors can grow very large and hard to manage.
-In this particular case, 4096-dimensional descriptors for 100M elements
-translates into 1TB of data, only in raw floating-point data alone (without
-accounting for any metadata or indexes associated with it).
-This component is very important on the overall analysis because a large set
-of feature vectors may not fit in memory and thus cause a pressure on the IO
-system while retrieving descriptors for computing distance, that severely impact
-the overall query execution time.
-Because of this, when the set of descriptors grows significantly large,
-it may be worth trading off accuracy to speed and space.
-Figure \ref{fig:features_size_does_matter} show the different indexes and
-their size in disk. These indexes already contain all the descriptors (or
-a quantized version of them), and can be loaded in memory directly when it fits.
-Note how, because of quantization of the descriptors, \textit{LSH} provides a
-significantly lower space foot print, which can be a great option for
-large collections of descriptors when accuracy is not a main factor.

### 2021-01-20 yfcc100m/paper/vldb2021/eval.tex
Add vldb2021 paper template

--- 2020-02-27 yfcc100m/paper/vldb2020/eval.tex
@@
 This is a very common use-case we have encountered when building
-applications such as smart-retail, sports applications, and video summarization,
-where the starting point is a large set of data that must be curated
+applications such as smart-retail, sports applications, and video summarization.
+For these type of applications, the starting point is usually 
+a large set of data that must be curated
 before proceeding with the data processing (such as neural network training).
@@
 which we describe in Section \ref{images}.
-
-We also evaluate the performance and scalability of the video management
-capabilities offered by VDMS.
-Handling video in a general way is a complex task.
-Some of this complexity comes from the existence of a variety of open and
-proprietary implementations, different encoding techniques and container
-formats, and different parameters of the video itself that are
-application-dependent, like frames per second, lossy compression, etc.
-When it comes to video, ad-hoc solutions have a large number of parameters that
-can be tuned.
-Together with that, there is no system that enables transactional
-operations over videos files in the way VDMS does.
-Because of this, we focus our efforts on understanding variations in the
-performance of VDMS and its scalability, rather than comparing
-it to a baseline that would not represent a fair
-comparison for either of the systems.
-We describe the results in Section \ref{videos}.
-
-Similarly, we explore the behavior of the feature vector functionality in VDMS,
-and an evaluation of the different trade-offs that the systems offers for
-application developers. For this, we implemented an image-search application
-based on \textit{similarity} search, which we describe in Section \ref{features}.
+We have also included performance evaluation on the Video and Feature Vectors 
+functionality in the appendix for further reference, together with a 
+description of the key aspects of those functionalities.
 
@@
 collection of 100 million public Flickr media objects created to provide free,
-sharable multimedia data for research. This dataset contains approximately
+shareable multimedia data for research. This dataset contains approximately
 99.2 million images and 0.8 million videos with metadata characterized by
@@
 Given that there are no other open-source systems
-that implement similar functionality and interfaces as VDMS, as a baseline,
-we have implemented an equivalent visual data management system comprised of a
-combination of widely available, off-the-shelf components:
+that provides similar functionality and interfaces as VDMS,
+we have implemented an equivalent visual data management system as a baseline, 
+comprised of a combination of widely available, off-the-shelf components:
 MySQL Server 5.7 (for storing metadata),
@@
 OpenCV 3.3 (to provide pre-processing operations on images).
-Note that this implementation only partially replicates the functionalities
+We decided to use a relational database instead of a non-relational 
+database because~\cite{Jatana2012, li_2019}: 
+\begin{itemize}
+    \item Relational databases support atomicity, consistency, isolation, 
+    and durability (ACID) while non-relational may compromise some ACID properties. 
+    \item The YFCC100m data is clearly structured.
+    \item We need to efficiently collate and return metadata records.
+\end{itemize}
+The baseline implementation only partially replicates the functionalities
 that VDMS offers when it comes to image and metadata handling, built for the
@@
 This implementation is based upon internal tools used for
-that particular task in the past.
+ML-based pipelines for media.
 We have implemented a set of client-side applications that take care
@@
 a 10GB back-plane switch, same as both servers. 
-
 Figure~\ref{fig:systems} shows a logical view of the difference between the
@@
 Apache WebServer and VDMS have direct access. 
+In the case of Videos, only the first frame is used for the image search.
+More information and evaluation on video-specific functionalities can be found
+in the appendix of this work.
 In the case of the baseline, metadata is
@@
 Persistent Memory type of storage.
-The benefits of Persistent Memory on metadata operations is left
-for another paper, and outside the scope of this evaluation.
+The benefits of Persistent Memory for metadata and a full evaluation of 
+the PMGD subsystem is material for another paper, 
+and outside the scope of this evaluation.
 For this experiment, in the case of VDMS we simply use a similar
@@
 where we have one node for each image, one node for each tag
-(always 1,570 tags), and connections between images.
+(always 1,570 tags), and a connection between each image and its respective tag(s).
 For instance, if an image has four \textit{autotags} assigned,
@@
 as shown in Table~\ref{table:vdmsnodes}. 
-
-It is important to note that we create indexes for the image identifier,
+It is also important to note that we create indexes for the image identifier,
 \textit{autotags} properties, and longitude/latitude coordinates 
@@
 setting a limit.
-Therefore, we let MySQL to automatically handle the concurrency.
+Therefore, we let MySQL automatically handle the concurrency.
 
@@
 parameter to avoid unfairness in the comparison against the baseline.
-We use the default parameters provided by the implementation.
+Unlike the baseline, VDMS can handle the entire YFCC100m dataset using the default parameters provided by the implementation.
 For both VDMS and the baseline, we created indexes over the 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 
-\subsubsection{Database Loading Time}
+\subsubsection{Database Building Time}
 
-One of the first things we noticed is the difference in loading times,
-where VDMS outperforms MySQL by a large margin.
-This analysis includes the metadata, as the images are stored in a shared filesystem. 
+One of the first things we noticed is the difference in the time needed to 
+build each database, where VDMS outperforms MySQL by a large margin.
+This analysis includes only the metadata, as the images are stored in a shared filesystem. 
 Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
@@
 Key difference in the build times are attributed to the low-level
-implementation of how MySQL reads and stores data from the files and the
+implementation of how MySQL reads and stores data from the files, the
 optimizations (increased InnoDB pool size, etc.) needed
-to handle large datasets such as YFCC100M.
+to handle large datasets such as YFCC100M, and the efficiency of PMGD.  
 On average, it took MySQL around 3.72x longer to build each database than VDMS.
@@
 In order to evaluate VDMS and the baseline on our use-case queries,
-we implemented 4 queries that filter and retrieve a specific set of images.
+we implemented 5 queries that filter and retrieve a specific set of images.
 We chose these queries because they represent typical use-cases where a
@@
 To evaluate the access to metadata and images,
-we use the following four queries, modeled after our internal use cases:
+we use the following queries, modeled after our internal use cases:
 \begin{itemize}
@@
 \item \textit{q3} - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with one specific autotag, resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
-\item \textit{q4} - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with two specific autotags (i.e.
+\item \textit{q4} - {\bf {\em 2tag\_resize\_geo\_and}}: Find metadata/images with two specific autotags (i.e.
 alligator AND lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
+\item \textit{q5} - {\bf {\em 2tag\_resize\_geo\_or}}: Find metadata/images with either of two specific autotags (i.e.
+alligator OR lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
 \end{itemize}
@@
 under the benchmarks branch of the VDMS project
-\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
+\footnote{https://github.com/\{Not shown during submission\}}
+% \footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
 and some examples of the queries can be found in the appendix of this
@@
   % include first image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
-  \caption{\textit{q1}(red), and \textit{q2}(blue).}
+  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2_q3}
+  \caption{\textit{q1}(red), \textit{q2}(blue),and \textit{q3}(orange).}
   \label{fig:q1_q2}
@@
   % include second image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
-  \caption{\textit{q3}(orange), and \textit{q4}(black).}
+  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q4_q5}
+  \caption{\textit{q4}(black), and \textit{q5}(pink).}
   \label{fig:q3_q4}
 \end{subfigure}
-\caption{Performance Analysis using 4 queries from our use-case
+\caption{Performance Analysis using 5 queries from our use-case
 described in the Experimental Setup Section.
-We show 2 queries in each figure for readability reasons.
-Top figures show the throughput of the 4 queries,
+We show queries in different figures for readability reasons.
+Top figures show the throughput of the queries for
 just retrieving metadata associated with the images.
 Bottom figures show the throughput when retrieving both metadata and images,
-plus operations applied to images when applicable (queries 2, 3 and 4).
+plus operations applied to images when applicable (queries 2, 3, 4 and 5).
 The experiments show the performance of both systems (VDMS and baseline) as the
@@
 as the metadata-only queries only involve running a query to MySQL.
-
 For the baseline system, in the case of 100M, the increase in the
@@
 56 concurrent clients.
-This query (as well as \textit{q3} and 4) runs a resize operation on the image,
+This query (as well as \textit{q3}, \textit{q4} and \textit{q5}) runs a resize operation on the image,
 an operation that requires decoding, resizing, and encoding the image
@@
 to better understand the performance of the systems under different query conditions.
-Figure~\ref{fig:q_throughput_56} shows the evaluation of the 4 representative
+Figure~\ref{fig:q_throughput_56} shows the evaluation of the 
 queries we analysed for our use case.
-Top figures show the throughput of the 4 queries
+Top figures show the throughput of these queries
 retrieving metadata associated with the images.
 Bottom figures show the throughput when retrieving both metadata and images,
-plus operations applied to images when applicable  (queries 2, 3 and 4).
+plus operations applied to images when applicable (queries 2, 3, 4,and 5).
 The experiments show the performance of both systems (VDMS and baseline) as the
@@
 For \textit{q4}, we see MySQL also outperforming VDMS on small size databases, and as
-the database size increase, the gap between the two narrows.
+the database size increase, the gap between the two narrows, 
+similar to what we see for \textit{q5}.
 It is interesting to note that adding filtering by geo-location
-(\textit{q3} and \textit{q4}) slightly increases the performance of MySQL
+(\textit{q3}, and \textit{q4}) slightly increases the performance of MySQL
 for small databases, and decreases it as the database sizes scales. 
+The behaviour is different in the case of \textit{q5}, 
+which also filters by geo-location. 
+This is attributed to the fact that OR queries involves processing larger results, 
+and thus does not benefit from the filtering of the AND operation as \textit{q4}.
 In the case of VDMS, we see \textit{q3} performance is comparable to \textit{q1}
-and \textit{q2}, but \textit{q4} suffers significantly when the scale of the database
-increases.
+and \textit{q2}, but \textit{q4} suffers significantly 
+when the scale of the database increases.
 The reason for that lack of scalability lies on the query implementation: given that
@@
 just a missing implementation.
-Future release will add more such operators in order to prevent 
-unnecessary retrievals.
+In the case of \textit{q5}, we see a similar effect, although it is worth noting 
+that for the OR operation there is no need for 2 retrievals. 
+Rather, a single retrieval is performed and the result filtered on the client.
+Future release will add more of such operators (AND, OR, etc) in order to prevent 
+unnecessary retrievals and extra filtering on the client side.
 
@@
 thus transferring less data over the network.
-In the case of the baseline, both \textit{q1} and \textit{q2} transfer the full size
-version of the image, and as part of \textit{q2}, the resize is performed in the client.
+In the case of the baseline, both \textit{q1} and \textit{q2} transfer 
+the full size version of the image, and as part of \textit{q2}, 
+the resize is performed in the client.
 This is why, contrary to the VDMS case, \textit{q1} performs better (even if slightly)
@@
 as we explained before.
-From the first 3 queries, we clearly see that VDMS outperforms the baseline 
+From the first 3 queries, as well as \textit{q5}, 
+we clearly see that VDMS outperforms the baseline 
 when retrieving visual data and applying operations.
@@
 and an average improvement in throughput of about 15x.
-More importantly, we see that the speedup increase as the database size grows,
+More importantly, we see that the speedup increases as the database size grows,
 showing that VDMS scales better than the baseline.
@@
 The team will address the missing implementation as part of future work.
+Most of the performance improvements can be attributed to the design 
+principles of VDMS, which aims to eliminate the need of combining and 
+re-purposing systems that were designed to handle types of date other than visual. 
+VDMS, by design, eliminates all the inefficiencies that result 
+from a forced integration of components designed for a 
+different range of applications.
 
@@
 
-%=========================================
+%% ========= Video Eval
 
-\subsection{Video Search}
-\label{videos}
+% \include{eval-video}
 
-VDMS provides full support for video storage and operations,
-in a similar way it does for images.
-This includes support for encoding, decoding, and transcoding of
-\textit{mp4}, \textit{avi}, and \textit{mov} containers,
-as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
-This is supported through the Visual Compute Module that provides an abstraction
-layer on top of OpenCV~\cite{opencv} and \textit{libffmpeg}\cite{ffmpeg}.
-All operations supported for images in VDMS are also supported at the
-video and frame level of the API.
-On top of that, there are a number of video-specific operations that
-are supported, such as the interval operations,
-enabling users to retrieve clips at different
-frames-per-second (FPS) versions of the video.
+%% ========= Video Eval - END
 
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Analysis of video operations.
-The left figure shows the video throughput (videos per sec) as the number of concurrent clients increase and the right figure breaks down the different components of the
-queries using 32 clients.}
-\label{fig:video}
-\end{figure*}
+%% ========= Descriptors Eval
 
-All this functionality is provided and integrated with the rest of the
-metadata API as part of the comprehensive VDMS interface.
-This makes it possible for users to interact with metadata and video in 
-a transactional manner, enabling users to run queries like: 
-"Retrieve all the videos where there is a \textit{lake} with 
-probability higher than 0.86, converting all videos to \textit{H.264} 
-\textit{mp4} of size 224x224".
-Appendix shows a sample of how this query would be implemented using the VDMS API
-\footnote{https://github.com/IntelLabs/vdms/wiki/FindVideo}.
-In particular, this functionality was used internally to select a subset
-of videos with the right licenses for a video summarization application.
+% \include{eval-descriptors}
 
-To the best of our knowledge, there is no solution that can provide
-all the functionality mentioned above, behind a single interface
-that also allows users to interact with images and metadata.
-Implementing a baseline, like we did for images, is significantly more complex
-due to the parametrization of video encondings and containers, 
-as explained at the beginning of this section.
-For this reason, we chose to make a study using VDMS in various scenarios,
-and analysis of scalability and the impact of having the overhead of VDMS' Request
-Server in the overall access time and throughput.
-
-Figure~\ref{fig:video} shows the analysis of different queries aimed
-at retrieving a video using the VDMS interface.
-We show how VDMS throughput increases when serving
-a video object as the number of simultaneous clients increases, as well as the
-overhead operations introduced in the overall query execution time.
-The figure on the left compares the number of video transaction per second
-(i.e., number of videos returned per second) when different operations
-are executed as part of the transaction. The upper-bound of this would be
-simply returning the video as-is (without running any encoding/decoding or
-operation), represented by the red line. This query is the upper-limit because
-it essentially translates to reading the video from the file-system and sending
-it over a TCP/IP socket, without any other overhead or operations.
-
-We also run a set of other queries that involve, showed in Figure~\ref{fig:video}:
-(a) running a resize operation on the video and, consequently,
-decoding and encoding operations as well (blue line),
-(b) transcoding, meaning the use of a different container and encoder
-than the one originally used (yellow line), and
-(c) both resize and transcoding.
-Note that the resize operation (blue and green lines) performs a downsize,
-which translates in less data being sent over the wire.
-This is specially noticeable when supporting 32 simultaneous clients, 
-where the system provides more videos per second due to sending less data to 
-the client, when compared to just transcoding and not resizing (yellow line).
-We can see that the system performs best when using all the physical cores,
-and this can be attributed to the compute-bound nature of video
-encoding, decoding, and processing.
-
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/feature_img_results}
-\caption{Sample Results of Similarity Search}
-\label{fig:similarity}
-\end{figure*}
-
-It is important to note an almost 3 orders of magnitude drop in performance
-when including operations as part of the query.
-We wanted to understand where most of the time was spent on the queries,
-and optimize the Request Server and Visual Compute Module
-if necessary. For this, we run the experiment shown at
-Figure~\ref{fig:video} (right) which breaks down the different components of the
-queries. This figure shows that more than 97\% of the query execution is spent
-on encoding/decoding operations, which is well-known to be a
-compute intensive operation\cite{videosgoogle}.
-On the one hand, this result shows that VDMS barely introduces any overhead. 
-On the other hand, this result means a limit on the opportunities 
-for optimization for video queries given that biggest time factors 
-are accounted by encoding/decoding, which is outside the scope of VDMS.
-This result was the call to action for one optimization we will include
-in future releases of VDMS, which involves using \textit{ffmpeg} C++ API to
-limit the number of frames being encoded/decoded when possible.
-This functionality will prevent encoding/decoding to happen on all frames
-when users only need to retrieve a subset of the frames in the video.
-
-%=========================================
-
-\subsection{Similarity Search}
-\label{features}
-
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities, images, and/or videos.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as \textit{descriptors}
-or \textit{visual descriptors}. We use these terms interchangeably.
-These descriptors can be classified, labeled, and used to build search
-indexes. There are many in-memory libraries that are designed for
-this task~\cite{flann, faiss}.
-Using the VDMS API, users can manage feature vector indexes,
-query previously inserted elements,
-run a k-nearest neighbor search (\textit{knn}), and express relationships
-between existing images or descriptors and
-the newly inserted descriptors.
-By natively supporting descriptors and \textit{knn},
-VDMS allows out-of-the-box classification functionalities for many applications
-\footnote{https://github.com/IntelLabs/vdms/wiki/ClassifyDescriptor}.
-
-For this work, and as part of a comprehensive image search implementation,
-we have used 4096-dimensional descriptors extracted from every image
-(and first frame of every video) from the YFCC100M dataset
-and created a collection of these feature vectors in VDMS to
-perform similarity search (i.e., find images that are
-\textit{similar} to an query (input) image).
-\textit{Similarity} in this particular case is defined as closeness
-in a 4096-dimensional space using euclidean distance as the metric.
-
-The process of loading descriptors in VDMS is simple.
-First, the user has to create a DescriptorSet, using a single command.
-At creation of the DescriptorSet, the dimensionality of the descriptors
-is specified, together with the desired indexing method and the desired metric
-for computing distances (Euclidean Distance, \textit{L2},
-or Inner Product, \textit{IP}).
-Once the DescriptorSet is created, descriptors can be inserted to the set.
-After the descriptors are inserted, a similarity search can be performed.
-
-Figure~\ref{fig:similarity} shows 3 examples of a query image (on the left),
-and images returned as \textit{similar} by VDMS.
-The input is a descriptor generated after a query image.
-The \textit{query input} descriptor is sent to VDMS as part of the query,
-VDMS uses that descriptor to find similar ones,
-and retrieves the images associated with those \textit{similar} descriptors.
-We show this as an example of the functionality and to depict
-how the feature vectors provided by the dataset can be used,
-but we also provide an analytical approach to
-the trade-off between accuracy and execution time in our system.
-It is important to note that the accuracy of the results is entirely tied
-to the quality of the descriptors chosen by the applications.
-The quality of the similarity result will be tied to the quality
-of the descriptor extraction that the application is using.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/features_alternatives}
-\caption{Feature Vector Evaluation: Trade-off between query execution speed
-and accuracy of the results, using ground-truth data for computing accuracy.
-For this evaluation, we query the 10 closest neighbors (k = 10), and compute
-accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
-results that is present within the top 10 computed neighbors).
-We average the query execution time and accuracy for 100 queries (nq = 100).}
-\label{fig:features_eval}
-\end{figure*}
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=\textwidth]{figures/features_disksize}
-\caption{Feature Collection Size in Disk}
-\label{fig:features_size_does_matter}
-\end{figure*}
-
-As mentioned before, VDMS provides different levels of customization of the
-indexes created for a descriptor set, that includes the indexing techniques
-and the metric for similarity.
-These different indexing techniques come with different trade-offs in terms
-of speed of search and accuracy of the computation.
-VDMS aims to provide functionality that is agnostics to application-specific
-techniques, enabling features that are generic to visual data processing
-applications.
-Figure~\ref{fig:features_eval} shows an analysis at the different indexing
-techniques provided by VDMS and its trades-off between accuracy and query
-execution speed, for a single threaded client.
-For this evaluation, we query the 10 closest neighbors (k = 10), and compute
-accuracy using recall at 4 (r\_k = 4) (i.e. percentage of the top 4 ground-truth
-results that is present within the top 10 computed neighbors).
-We average the query execution time and accuracy for 100 queries (nq = 100).
-The \textit{flat} index (red line) implements exact search and
-represents ground-truth, which explain why the
-accuracy is always 100\% in the plot on the right.
-The other indexes implement \textit{approximate search},
-which trades-off between accuracy and speed of search~\cite{flann, faiss}.
-We have also tried the \textit{ivfflat} index (inverted file index), as well as
-\textit{LSH}-based indexes using a different number of bits per descriptor
-\footnote{https://github.com/facebookresearch/faiss/wiki/Faiss-indexes}.
-Results show how \textit{ivfflat} is the fastest option but it comes with a trade-off
-of about 30\% loss in accuracy, while simple brute-force search
-is among the slowest options at the expenses of 100\% accuracy,
-meaning exact search.
-
-Another important trade-off to be made is with respect to space efficiency:
-The DescriptorSet can grow very large and expensive to load and manage.
-In this particular case, 4096-dimensional descriptors for 100M elements
-translates into 1TB of data, only in raw floating-point data alone
-(without accounting for any metadata or indexes associated with it).
-This component is very important for the overall analysis on which
-index structure to use because a large set of descriptors may not fit in memory
-and thus cause a pressure on the IO system while retrieving descriptors
-for computing distance.
-This can severely impact the overall query execution time.
-When the DescriptorSet grows significantly large,
-it may be worth trading off accuracy for speed and space.
-Figure~\ref{fig:features_size_does_matter} shows the different indexes and
-their size in disk. These indexes already contain all the descriptors (or
-a quantized version of them in the case of LSH~\cite{lsh}),
-and can be loaded in memory directly when it fits.
-Note how, because of quantization of the descriptors, \textit{LSH} provides a
-significantly lower space foot print, which can be a great option for
-large collections of descriptors when accuracy is not a main factor.
-It is not uncommon to sacrifice accuracy as images and videos are captured
-using a noise sensor (i.e., the camera), and an approximate search
-in many cases can provide the necessary accuracy for applications
-to achieve their goals.
+%% ========= Descriptors Eval - END

### 2021-01-20 yfcc100m/paper/vldb2021/intro.tex
Add vldb2021 paper template

\section{Introduction}
\label{intro}

Visual computing workloads performing analytics on videos and/or images
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning (ML) techniques to
extract information from the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: retail vendors might be interested in knowing 
which are the most visited areas of their stores using security 
video feeds as input, or a doctor might want know the effect of a
specific treatment by looking at the changes in size of a tumor from a brain scan.

Despite the increasing use of visual data processing, 
there has been very little research on the \textit{management} of visual data.
Most of the current storage solutions for visual data are
an ad-hoc collection of tools and systems, that are re-purposed and 
adapted to work with visual data.
The approach of re-purposing and integrating solutions not designed for a task
results in resource utilization inefficiencies\cite{rasdaman}. 
For example, consider a ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. 
This requires assigning consistent
identifiers for the scans and adding their metadata in
a relational or key-value database. 
If the queries require a search over patient information, 
then patients are associated with their brain scans. 
Finally, if the ML pipeline needs images with a different resolution than the original, there is additional compute
diverted towards pre-processing the original images which are typically larger. 
All these steps require understanding different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to change the schema on the fly.
As another example, many applications can be studied through the use of large
and publicly available datasets. 
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine-generated tags and feature vectors\cite{imagesearch} 
for each image, and video summarization.
For these use-cases, the usual first step consists on selecting a 
subset of the data before running any processing, and a large effort 
is devoted to cleaning and pre-processing the data.
Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searching
based on tags (relational database, graph database, csv files, etc), and
building the necessary pipelines for querying and retrieving the right data.

More generally, data scientists and machine learning developers 
usually end up building an ad-hoc solution that results in a 
combination of databases and file systems to store 
metadata and visual data (images, videos), respectively. 
This is integrated with a set of custom scripts that tie multiple systems together, 
unique not only to a specific application/discipline but often 
to individual researchers or development teams.
These ad-hoc solutions make replicating experiments difficult, 
and more importantly, they do not scale well when deployed in the real-world.
The reason behind such complexity is \textit{the lack of a system 
that can be used to store and access all the data the application needs, 
including metadata, images, videos, and feature vectors}.

In this paper, we describe VDMS and show how it provides a 
comprehensive solution to the data management for applications 
that heavily rely on visual data. 
VDMS is a completely Open Source project designed to enable 
efficient access of visual data.
To the best of our knowledge, a rich set of functionalities 
designed for visual data management, provided behind an integrated API, 
is unique to VDMS and we were unable to find a system with similar functionality.
While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
that can be used to store metadata~\cite{memsql, vertica}, and systems that
manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
be distinguished from them on the following aspects:

\begin{itemize}
\item {\em Design for analytics and machine learning}: By targeting
visual data for use cases that require manipulation
of visual information and associated metadata,
\item {\em Ease-of-use}: By defining a common API that allows applications to
combine their complex metadata searches with operations on resulting visual
data, and together with full support for feature vectors. VDMS goes beyond the
traditional SQL or OpenCV level interfaces that do one or the other.
\item {\em Performance}: We show how a unified system such as VDMS can
outperform an ad-hoc system constructed with well-known discrete components.
Because of the capabilities we have built into VDMS, it handles complex
queries significantly better than the ad-hoc system without compromising the
performance of simple queries.
\end{itemize}

In order to evaluate VDMS in a realistic use case, 
we use the YFCC100M dataset\cite{Thomee_2016}. 
The YFCC100M is the largest public multimedia collection. 
It contains the metadata of around 99.2 million photos 
and 0.8 million videos from Flickr,
plus expansion packs that include a variety of multidimensional data,
all of which were shared under one of the various Creative Commons licenses.
We have used this dataset
for multiple proof of concepts and applications within our research lab.

We show how VDMS can be used as a centralize point for data
management and data access even when having multiple modalities of data:
Metadata, Image, Videos, and Feature Vectors.

### 2021-02-03 yfcc100m/python/eval/find_probs.py
Update probabilities generator scripts

@@ -1,6 +1,7 @@
 import argparse
 import time
 from threading import Thread
+import json
 
 import pandas as pd
 import numpy as np
@@ -9,37 +10,42 @@ from mysql_eval  import MySQLQuery
 from memsql_eval import MemSQLQuery
 from vdms_eval   import VDMSQuery
 
+from DBEvalFramework import EvalFramework
+
 VDMS_PORT_MAPPING = {'100k': 55500, '500k': 55405,
                      '1M':   55501, '5M':   55450,
                      '10M':  55510, '50M':  55000,
                      '100M': 50000}
 
-# SIZE_PROB_MAPPING = {'100k': 0.21, '500k': 0.41,
-#                      '1M':   0.55, '5M':   0.65,
-#                      '10M':  0.70, '50M':  0.75,
-#                      '100M': 0.82}
+QUERY_PARAMS = [
+            {'key': '1tag',
+             'tags': ["alligator"], 'probs': [0.98],
+             'operations': [],
+             'comptype': "and" },
+            {'key': '1tag_loc20',
+             'tags': ["alligator"], 'probs': [0.5],
+             'lat': -14.354356, 'long': -39.002567,
+             'range_dist': 20,
+             'operations': [],
+             'comptype': "and"  },
+            {'key': '2tag_and',
+             'tags': ["alligator", "lake"], 'probs': [0.95, 0.95],
+             'comptype': "and" },
+            {'key': '2tag_or',
+             'tags': ["alligator", "lake"], 'probs': [0.9, 0.9],
+             'comptype': "or" },
+            {'key': '2tag_loc20_and',
+             'tags': ["alligator", "lake"], 'probs': [0.55, 0.55],
+             'lat': -14.354356, 'long': -39.002567,
+             'range_dist': 20,
+             'comptype': "and" },
+            {'key': '2tag_loc20_or',
+             'tags': ["alligator", "lake"], 'probs': [0.95, 0.95],
+             'lat': -14.354356, 'long': -39.002567,
+             'range_dist': 20,
+             'comptype': "or" },
 
-SIZE_PROB_MAPPING = {'100k': 0.81, '500k': 0.81,
-                     '1M':   0.81, '5M':   0.81,
-                     '10M':  0.81, '50M':  0.81,
-                     '100M': 0.81}
-
-RESIZE = {"type": "resize", "width": 224, "height": 224}
-QUERY_PARAMS = [{'key': '_1tag_resize',
-                 'tags': ["alligator"],
[…]

### 2021-02-03 yfcc100m/python/eval/probs_run.sh

@@ -1,14 +1,10 @@
 #!/bin/bash
-db_sizes="100k,500k,1M,5M,10M,50M,100M"
 
-for db in vdms
+for size in 1M 5M 10M 50M 100M
 do
-    for size in 1M 5M 10M 50M 100M
-    do
-        # Run VDMS Queries
-        echo "Running $db ${size}..."
-        python3 find_probs.py \
-                -db_type=$db \
-                -db_name=$size
-    done
+    # Run VDMS Queries
+    echo "Running $db ${size}..."
+    python3 find_probs.py \
+            -db_type=vdms \
+            -db_name=$size
 done

### 2021-02-03 yfcc100m/python/eval/mysql_eval/MySQLQuery.py
Update query scripts for mysql and vdms

@@ -140,7 +140,8 @@ class MySQL(object):
     def get_metadata(self, tags, probs,
                              lat=-1, long=-1,
                              range_dist=0,
-                             comptype='and'):
+                             comptype='and',
+                             return_responses=False):
 
         if lat != -1:
             location_qstr = '''(latitude >= {} AND latitude <= {} ) AND (longitude >= {} AND longitude <= {})'''.format(lat-range_dist*1.0,lat+range_dist*1.0,long-range_dist*1.0,long+range_dist*1.0)
@@ -168,6 +169,9 @@ class MySQL(object):
         #         out_file.write(str(res[cols.index('id')]))
         #         out_file.write("\n")
 
+        if return_responses:
+            return response
+
         out_dict = {'response_len':len(response),'response_time':endtime}
 
         return out_dict
@@ -181,11 +185,16 @@ class MySQL(object):
     def get_images(self, tags, probs, operations = [],
                            lat=-1, long=-1, range_dist=0, return_images=False, comptype='and'):
         start = time.time()
-        metadata = self.get_metadata(tags, probs, lat, long, range_dist, comptype=comptype)
+        metadata = self.get_metadata(tags, probs,
+                                     lat, long, range_dist,
+                                     comptype=comptype,
+                                     return_responses=True)
 
         img_array = []
         cols = ['line_number', 'download_url', 'id',
                 'latitude', 'longitude', 'license_name']
+
+        # print(metadata)
 
         for res in metadata:
 
@@ -193,10 +202,10 @@ class MySQL(object):
             # # imgPath = FOLDER_CHOICES[quotient] + urlparse(res[cols.index('download_url')]).path
             # imgPath = "http://"+IMG_HOST+FOLDER_CHOICES[quotient] + urlparse(res[cols.index('download_url')]).path
 
-            # print(imgPath)
             # print(res[cols.index('id')])
 
             imgPath = "http://"+IMG_HOST+"/images/" + urlparse(res[cols.index('download_url')]).path
+            # print(imgPath)
 
             try:
                 imgdata = requests.get(imgPath)
@@ -239,11 +248,9 @@ class MySQL(object):
 
         total_time = time.time() - start
 
-        decoded_images = [img for img in img_array if img]
-
         out_dict = {
-            'images_len':len(decoded_images),
-            'images_time':total_time
+            'response_len':  len(img_array),
+            'response_time': total_time
         }
 
         if return_images:

### 2021-02-03 yfcc100m/python/eval/vdms_eval/VDMSQuery.py

@@ -338,6 +338,8 @@ class VDMSQuery(object):
 
         endtime = time.time() - start
 
+        # print("GetMetadataVDMS Query:", endtime)
+
         # This is for the case of internal call from within this calss
         if return_response:
             return results
@@ -481,11 +483,13 @@ class VDMSQuery(object):
             total_time = time.time() - start
 
         decoded_images = []
+        counter_bad_img = 0
         try:
             for im in blobs:
                 if len(im) == 0:
-                    print("WARNING - Returned blob of size 0")
+                    # print("WARNING - Returned blob of size 0")
                     dec_img = None
+                    counter_bad_img += 1
                 else:
                     img = np.frombuffer(im, dtype='uint8')
                     dec_img = cv2.imdecode(img, cv2.IMREAD_COLOR)
@@ -495,6 +499,10 @@ class VDMSQuery(object):
             print("ERROR: Error decoding image result from VDMS")
             raise
             decoded_images = []
+
+        if counter_bad_img > 1:
+            print("WARNING:", counter_bad_img * 100 / len(decoded_images),
+                  "% of the images are empty")
 
         out_dict = {
             'response_len':  len(decoded_images),

### 2021-02-03 yfcc100m/python/eval/performance.py
Update performance experiment scripts

@@ -16,10 +16,56 @@ VDMS_PORT_MAPPING = {'100k': 55500, '500k': 55405,
                      '10M':  55510, '50M':  55000,
                      '100M': 50000}
 
-SIZE_PROB_MAPPING = {'100k': 0.60,       '500k': 0.06,
-                     '1M':   0.92175,    '5M':   0.955,
-                     '10M':  0.972499,   '50M':  0.99375,
-                     '100M': 0.997656249}
+# SIZE_PROB_MAPPING = {'100k': 0.60,       '500k': 0.06,
+#                      '1M':   0.92175,    '5M':   0.955,
+#                      '10M':  0.952499,   '50M':  0.95375,
+#                      '100M': 0.967656249}
+
+# These are pre-computed probabilities so that queries returned
+# a controlled number of results, to prevent the throughput to be
+# affected by higher number of results
+PROB_MAPPING = {
+            '1M': {
+              "1tag":           0.735,
+              "1tag_loc20":     0.0625,
+              "2tag_loc20_or":  0.059375,
+              "2tag_or":        0.975,
+              "2tag_loc20_and": 0.06875,
+              "2tag_and":       0.059375,
+            },
+            '5M': {
+              "1tag":           0.949375,
+              "1tag_loc20":     0.0625,
+              "2tag_or":        0.984375,
+              "2tag_and":       0.59375,
+              "2tag_loc20_or":  0.71249,
+              "2tag_loc20_and": 0.06875,
+            },
+            '10M': {
+              "2tag_loc20_or":  0.8312499,
+              "2tag_and":       0.653125,
+              "1tag_loc20":     0.0625,
+              "2tag_or":        0.9875,
+              "2tag_loc20_and": 0.06875,
+              "1tag":           0.96468749
+            },
+            '50M':  {
+              "1tag":           0.995,
+              "2tag_and":       0.74218749,
+              "1tag_loc20":     0.0625,
+              "2tag_loc20_or":  0.95,
+              "2tag_loc20_and": 0.06875,
+              "2tag_or":        0.9953125
+            },
+            '100M': {
+              "1tag":           0.9975,
+              "1tag_loc20":     0.75,
+              "2tag_and":       0.8015625,
+              "2tag_loc20_or":  0.9781249,
+              "2tag_loc20_and": 0.06875,
+              "2tag_or":        0.997656249
+            }
+        }
 
 # These are operations using VDMS API format / params.
[…]

### 2021-02-03 yfcc100m/python/eval/run_performance.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
-
-experiment_name=perf_metadata_8_50M
+NITER=5
+experiment_name=perf_all_threads_images
 
 running_folder=running_experiment
 info=${running_folder}/info.txt
@@ -14,11 +14,11 @@ echo $experiment_name "started..." >> $info
 
 date >> $info
 
-# for th in 2 4 8 16 32 56 64 112
-for th in 8
+for th in 2 4 8 16 32 56 64 112
+# for th in 4 8 16
 do
-    # for size in 5M 10M
-    for size in 1M 5M 10M 50M
+    # for size in 1M 5M 10M 50M
+    for size in 1M 5M 10M 50M 100M
     do
         for db in vdms mysql
         do
@@ -28,6 +28,7 @@ do
                     -db_type=$db \
                     -db_name=$size \
                     -numthreads=$th \
+                    -numiters=$NITER \
                     -out_folder=$running_folder \
                     >> $running_folder/log.log
                     2>> $running_folder/log_error.log

### 2021-02-03 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Update plotting scripts

@@ -137,6 +137,7 @@ class EvalFramework(object):
             self.plot_query_time_speedup(i)
             self.plot_n_results(i)
 
+
     def plot_results_throughput_parallelism_dbsizes(self, q,
                                             result_type="results"):
 
@@ -324,6 +325,7 @@ class EvalFramework(object):
                           filename=filename,
                           ylabel=result_type + "/s")
 
+
     def plot_query_throughput(self, n_threads):
 
         # Plot query times:
@@ -372,6 +374,7 @@ class EvalFramework(object):
                           filename=filename,
                           ylabel="Queries per second")
 
+
     def plot_query_time(self, n_threads):
 
         # Plot query times:
@@ -415,6 +418,7 @@ class EvalFramework(object):
 
         return
 
+
     def plot_n_results(self, n_threads):
 
         # Plot query times:
@@ -451,12 +455,13 @@ class EvalFramework(object):
         filename += "plot_" + str(n_threads) + "_n_results_mosaic.pdf"
 
         title = "Returned Results for different queries"
-        p.plot_lines_all_mosaic(queries, db_sizes, engines, values, log="both",
+        p.plot_lines_all_mosaic(queries, db_sizes, engines, values, log="x",
                           title=title,
                           filename=filename,
                           ylabel="Number of Results")
 
         return
+
 
     # Bar plot
     def plot_query_time_speedup(self, n_threads):

### 2021-02-03 yfcc100m/python/eval/DBEvalFramework/Plotting.py

@@ -243,11 +243,13 @@ class Plotting(object):
 
         ax0.set_title(title)
 
-        ax0.set_xticks(index)
+        ax0.set_xticks(index + (len(db_sizes) / 2 + 1) * bar_width)
         ax0.set_xticklabels(db_sizes, fontsize=10)
         ax0.tick_params(axis='y', labelsize=10)
 
-        ax0.set_ylim(0,30)
+        limit = values[:,0:len(db_sizes)].max() * 1.5
+
+        ax0.set_ylim(0, limit)
 
         plt.ylabel('Speedup', fontsize=12)
 

### 2021-02-03 yfcc100m/paper/osdi2020/acks.tex
Update OSDI submission files

--- 2020-02-27 yfcc100m/paper/vldb2020/acks.tex
@@
 
+Not shown during submission.
+
+\begin{comment}
 We would like to thank the many people that made this project possible and
@@
 the servers and infrastructure needed to conduct our experiments.
+\end{comment}

### 2021-02-03 yfcc100m/paper/osdi2020/conclusion.tex
Update OSDI submission files

\section{Conclusion}
In this paper, we described VDMS design and implementation and 
show a comprehensive evaluation on our Image Search Application. 
We use one of the largest publicly available datasets: 
The Yahoo Flickr Creative Commons 100M (YFCC100M), 
together with the expansions packs that include 
machine-generated labels and feature vectors.
We show how VDMS compares against a combination of 
industry standard systems, all of which are needed to 
replicate a only portion of VDMS' functionality. 
We see improvements up to 35x in certain queries, 
and an average improvement of about 15x. 
% Finally, we analyze the different trade-offs of VDMS' 
% descriptors indexes, a functionality 
% that fully integrates with the rest of VDMS interface, 
% providing a powerful and comprehensive API.
The design of VDMS, which was conceived as a 
data management system that treats visual entities 
as first class citizens, can remove inefficiencies 
that result from re-purposing and combining solutions
that were not designed for the job while providing 
simpler and richer interfaces. 
VDMS' easy-to-use interfaces outperform industry standard systems
with a set of functionalities which, to the best of our knowledge,
are not available in any other single data management solution for visual data.
VDMS was designed for analytics and it can efficiently handle complex queries 
which can simplify the design of future applications that rely on visual data.

### 2021-02-03 yfcc100m/paper/osdi2020/design.tex
Update OSDI submission files

\section{VDMS Design \& Implementation}
\label{arch}

In this section, we describe VDMS design principles and implementation, 
which was briefly introduced in previous work \cite{vdms-nips}.
Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
VDMS implements a client-server architecture that handles client
requests concurrently, and coordinates query execution across
the metadata and visual data components in order to return a unified response.
Users interact with both metadata and visual data 
(i.e., images, videos, feature vectors) using a unified API, 
in a transactional manner.

The metadata component is the \textit{Persistent Memory Graph
Database} (PMGD) and the (visual) data component is our Visual Compute Module.
The Visual Compute Module enables machine-friendly enhancements to
visual data, exposing high-level abstractions to the \textit{Request Server}
for dealing with a variety of images and video formats (through OpenCV and ffmpeg),
and different methods for indexing for feature vectors
(including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
Finally, a main component of VDMS is in charge of implementing the API and orchestrating between the PMGD and the Visual Compute Module
to serve client's requests. This component is the \textit{Request Server}.

VDMS and its components are fully available as open source projects
% \footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
\footnote{https://github.com/\{Not shown during submission\}}.
We briefly describe each of the main components as follows:

\begin{figure}
\centering
\includegraphics[width=1\columnwidth]{figures/vdms_arch.pdf}
\caption{VDMS Architecture}
\label{fig:arch}
\end{figure}

\subsection{Persistent Memory Graph Database}

We use the Persistent Memory Graph Database (PMGD) to provide an efficient storage
solution addressing the increasing popularity of connected data and
applications that benefit from graph-like processing. 
PMGD implements an in-persistent-memory graph database, PMGD, optimized
to run on a platform equipped with persistent memory.
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability 
(ACID) properties expected from databases. 
Graph represents an easier abstraction to model complex problems \cite{tao}.
Moreover, the graph model makes it very suitable for the data and
access patterns shown by visual metadata, which can be easily mapped 
into application-level abstractions by developers. 
For instance, abstractions like \textit{BoundingBoxes} associated to 
images or videos can be easily represented using nodes and edges.
With its natural ability to extend the schema very
easily (due to the use of a property graph model),
we can support new developments in machine learning that can lead to
enhancements to existing metadata over time.
These are the reasons the team chose a graph database over a
relational database as the metadata management for the implementation of VDMS.
PMGD is designed and optimized for persistent memory technologies
like Intel Optane~\cite{IntelXPoint15}, which
promise storage providing nearly the speed of DRAM and the
durability of block-oriented storage.

\subsection{Visual Compute Module}

The Visual Compute Module was designed and implemented to provide
an internal abstraction layer for interacting with visual data.
It enables simple visual data handling and processing (i.e., basic building block 
operations like crop, resize, etc).
For traditional formats (jpg, png, tiff, mp4, etc.),
the interface is an abstraction layer over OpenCV.
However, it also provides a way to use novel formats that are 
better suited for visual analytics: a novel, array-based lossless image format. 
This format is built on the array data manager TileDB~\cite{TileDB} and 
is well suited for images that are used in visual analytics.

VDMS provides full support for video storage and operations,
in a similar way it does for images.
This includes support for encoding, decoding, and transcoding of
\textit{mp4}, \textit{avi}, and \textit{mov} containers,
as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
This is supported through the use of either OpenCV~\cite{opencv} 
or \textit{libffmpeg}\cite{ffmpeg}, or both.
All operations supported for images in VDMS are also supported at the
video and frame level of the API.
On top of that, there are a number of video-specific operations that
are supported, such as the interval operations,
enabling users to retrieve clips at different
frames-per-second (FPS) versions of the video.

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities, images, and/or videos.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}. 
We use these terms interchangeably.
These descriptors can be classified, labeled, and used to build search indexes.
Feature Vectors support is provided through our implementation based
on high-dimensional sparse arrays, also using TileDB.
In addition, the Visual Compute Library provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.

\subsection{Request Server}

Developers and users of machine learning frameworks and data science
applications favor simpler interfaces to access and process data. 
They cannot be expected to deal with two different ways of interacting 
with information (metadata and visual data) instead of focusing on the
algorithmic parts of their pipelines.
VDMS takes care of coordinating client requests across the metadata and the
visual data, and managing multiple clients through the \textit{Request Server}.
The \textit{Request Server} is a key component of the system, 
as it implements VDMS' API and coordinates request and responses from 
the PMGD and Visual Compute Module subsystems.
It decomposes the user queries into
metadata and visual data requests, invokes the relevant calls behind the scene,
and returns a coherent and unified response that is easy for the user to parse 
and interpret.

\subsection{VDMS API}

One of the most important differentiating factor of VDMS is its interface.
VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
as first class citizens. Thus, VDMS provides an API that revolves 
around visual data operations and retrieval. 
VDMS API is easy to use and explicitly pre-defines certain
primitives associated with metadata, images, videos, and feature vectors.
Authors have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based API,
which is very popular across various application domains.
By defining a new JSON-based API, there is a trade-off between expressiveness 
(compared to well-established query languages like SPARQL, Gremlim, or even SQL) 
and the ability to natively support visual data operations.
However, we believe it is possible for our API to achieve similar levels of
expressiveness compared to more mature query languages over time.

Listing~\ref{findimagegeo} shows a sample query. 
In this particular example, the transaction retrieves 
all the images of \textit{alligators} 
with probability higher than 0.66,
filter by latitude and longitude within 1 degree, 
apply a resize operation to make the images 224x224, 
rotate the images 45.34 degrees, 
and return the images as "png" files. 
It is important to note how the API natively supports basic building 
blocks of visual data processing, like resize, rotation, or transcoding
(changing output formats and encodings).
The API allows interaction with metadata, images, videos, 
and feature vectors in a similar fashion, and it is fully 
documented on the project's Github 
wiki~\footnote{https://github.com/\{Not shown during submission\}}.
% wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.

\begin{listing}[ht!]
\begin{minted}[frame=single,
              framesep=3mm,
              linenos=true,
              xleftmargin=21pt,
              tabsize=4]{js}
"FindEntity"{     
    "class": "autotag",
    "constraints": { 
        "name": ["==", "alligator"]
    }
    "_ref" : 1
},
"FindImage":{
    "format": "png",
    "link": {
        "ref":1,
        "constraints": {
            "prob": [">=", 0.66]
        }
    }, 
    "constraints": {
        "latitude": [">=", 36.23433, 
                     "<=", 38.23433]
        "longitude":[">=", -114.80666, 
                     "<=", -116.80666]
    },
    "operations": [{
        "type": "resize",
        "height": 224,
        "width":  224,
    }, {
        "type": "rotate",
        "angle": 45.34
    }]
}

\end{minted}
\caption{Sample Query for Image Retrieval - 
The query expresses the following: 
Find all the images connected to the autotag \textit{alligator} 
with probability higher than 0.66, 
filter the images by latitude and longitude within 1 degree, 
apply a resize operation to make the images 224x224, 
rotate the image 45.34 degrees, 
and return the images as "png" files.} 
\label{findimagegeo}
\end{listing}

\subsection{Client Library}

The client library implements TCP/IP based connectors to the VDMS Server, 
similar to most database\cite{memsql, mysql}.
Users can connect to VDMS and implement queries using VDMS' API 
by defining JSON commands conforming to the query protocol we have defined.
The client library provides a simple method that
accepts a JSON string and an array or vector of blobs.
Internally, the library wraps the query string and blob using
Google Protobufs \cite{protobufs} and sends it to the VDMS server.
It also receives a similarly formed response from VDMS
and returns it to the client. 
The responses require JSON parsing on the client
side for the metadata string that indicates how to interpret the blobs field.
Currently, client libraries are implemented for Python and C++ client. 
The client libraries are lightweight, as they simply implement the communication
protocol between the client and the server. 
This makes it easier for developers to implement similar client libraries using 
any other programming language of their choice.

### 2021-02-03 yfcc100m/paper/osdi2020/eval-descriptors.tex
Update OSDI submission files

--- 2020-02-18 yfcc100m/paper/vldb2020/eval.tex
@@
-\section{Evaluation}
-\label{eval}
-
-We have used the YFCC100M dataset to evaluate different aspects of our system.
-We use the images in the dataset and its associated metadata to implement
-an image-search engine based on properties associated to those images.
-This is a very common use-case we have encountered when attacking
-applications such as smart-retail, sports applications, and video summarization,
-where the starting point is a large set of data that must be curated
-before proceeding with the data processing and neural network training.
-In order to evaluate the different aspects of the performance on the
-image search, we built a baseline following the methodology used in the industry, and following what we have done in the past in order to solve
-the data search problem, which we describe in Section \ref{images}.
-
-We also cover evaluation we have done using the Video interface VDMS offers,
-in order to asses how the implementation performs and scale when used
-for video handling.
-When it comes to video, ad-hoc solutions have a large number of parameters that
-can be tunned, and complex ways to handle the data. Together with that,
-there is no system that enables transactional operations over videos files
-in the way VDMS does. Because of the two reasons above mentioned,
-we focus our efforts in understanding variations in the the performance of
-VDMS and its scalability, rather than comparing it to a baseline that
-would no represent a fair comparison for either of the systems.
-We describe the results in Section \ref{videos}.
-
-Similarly, we explore the behavior of the feature vector functionality in VDMS,
-and an evaluation of the different trade-offs that the systems offers for
-applications developers. For this, we implemented an image-search application
-based on \textit{similarity} search, which we describe in Section \ref{features}
-
-\subsection{YFCC100M Dataset}
-\label{dataset}
-
-The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
-collection of 100 million public Flickr media objects created to provide free,
-sharable multimedia data for research. This dataset contains approximately
-99.2 million images and 0.8 million videos with metadata characterized by
-25 fields such as the unique identifier, userid,
-date the media was taken/uploaded, location in longitude/latitude coordinates,
-device type the media was captured, URL to download the media object,
-and the Creative Commons license type information.
-The YFCC100M dataset also used a deep learning approach to generate autotags
-which is a set of comma-separated concepts such as people, scenery, objects,
-and animals and confidence scores generated from 1,570 trained
-Caffe classifiers ~\cite{Thomee_2016}. Together with the autotag, there is a
-probability associated to each tag to indicate the certainty level of the tag
-in that image/frame.
-We have also used feature vectors generated for every image and first frame
-of every video \cite{features} to implement similarity search.
-Given that there is no standard benchmark oriented towards visual data queries,
-we have built a series of queries to filter this dataset that is model after
-our internal use cases for many of the mentioned applications we have worked
-with.
-
-\subsection{Experimental Setup}
-
-As a baseline, and given that there are no other open-source system
-that implements similar functionality and interfaces as VDMS,
-we implemented an equivalent visual data management system comprised of a
-combination of widely available, off-the-shelve components:
-MySQL Server 5.7 (for storing metadata),
-Apache Web Server 2.4.18 (as interface for image access), and
-OpenCV 3.3 (to provide pre-processing operations on images).
-We have implemented a set of client-side applications that takes care
-of retrieving the components from the different systems, and applies
-pre-processing operations when needed.
-For visual data workloads, building an approach like the one
-implemented as a baseline for this work
-is the common practice in the industry \cite{heystack, tao}.
-
-For all our experiments, we use 2 servers, one hosting a VDMS server and
-another hosting a the Baseline implementation.
-Both servers have a dual-socket Intel\textsuperscript{\textregistered}
-Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
-each CPU with 28 physical cores with multithreading enabled,
-for a total of 112 logical cores per server.
-The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
-has 64GB of DDR4 DRAM. Both servers run Ubuntu 16.04.
-The client application (running the queries and measuring round-trip time)
-is connected to the server through a 1GB wired link through
-a 10GB-backplane switch.
-
-Figure~\ref{fig:systems} shows a logical view of the difference between the
-interaction of the client application (which wants to retrieve metadata and
-imaged) with VDMS (left) and the baseline (right).
-The client application was implemented using Python 3 for both VDMS and
-the baseline.
-
-It is worth noting that the images are stored in a shared repository
-(ext4 filesystem on a RAID 6 configuration of 16TB) that both Apache WebServer
-and VDMS have direct access. In the case of the baseline, metadata is
-stored in MySQL using an attached SSD disk.
-Even if VDMS has native support for Optane Persistent Memory,
-we do not use it in this experiment because of fairness of comparison w.r.t
-MySQL (which was not design for Persistent Memory type of storage).
-The benefits of Persistent Memory on metadata operations is left
-for another paper, and outside the scope of this evaluation.
-For this experiment then, in the case of VDMS we simply use a similar
-attached SSD disk to store metadata.
-Even if PMGD, the graph database used by VDMS, is designed for PM,
-it can deliver good performance when using SSDs directly.
-
-\begin{figure*}[]
-\centering
-\includegraphics[width=\textwidth]{figures/comparison_system}
-\caption{Comparison Systems}
-\label{fig:systems}
-\end{figure*}
-
-For the metadata, we built VDMS and MySQL databases
-using the YFCC100M dataset with incremental database sizes.
-For simplicity, we named the database based on the approximate number of images
-it contains, as follows: 100k, 500k, 1M, 5M, 10M, 50M, 100M.
-The exact number of images/elements on each database is shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}
-For each database size, we created an instance of VDMS using YFCC100m metadata,
-the YFCC100m machine-generated \textit{autotags} associated with
-each metadata identifier, and the list of 1,570 \textit{autotags}
-using 100 concurrent VDMS Python clients.
-Internally, that information is represented as a property graph,
-where we have one node for each image, one node for each tag
-(always 1570 tags), and connections between images.
-For instance, if an image has 4 \textit{autotags} assigned,
-there will be 4 connections between that image and
-the different nodes for those \textit{autotags}.
-The information about the probability the \textit{autotag} present in an image
-is expressed as a property in the \textit{connection} between the 2 nodes.
-Figure \ref{fig:graph_representation} shows an example on two images and
-2 \textit{autotags}, and the \textit{connections} between those \textit{autotags}
-and the images.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Image id: 86756231, on the other hand, has a single \textit{autotags} assigned:
-\textit{Alligator} with probability 0.894.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/graph_representation}
-\caption{VDMS Data Representation Using a Property Graph:
-Example on two images and 2 \textit{autotags} with
-their respective probabilities expressed in the \textit{connection}.
-Image id: 23143252 has two \textit{autotags} assigned:
-\textit{Alligator} with probability 0.285 and \textit{Lake} with probability 0.872.
-Similarly, Image id: 86756231 has a single \textit{autotags} assigned:
-\textit{Alligator} with probability 0.894.}
-\label{fig:graph_representation}
-\end{figure}
-
-Because there are on average 8 tags assigned to each image,
-there will be around 8 times more connections than images, as can be seen
-in Table~\ref{table:vdmsnodes}.
-Also, each image node will contain multiple properties associated
-with that object.
-Using the Python Client Module, we insert 1,570 nodes for the \textit{autotags},
-and then insert nodes for the YFCC media objects (images or videos)
-with the associated metadata.
-Adding the autotag confidence scores is a recursive process and requires more
-complex queries than adding the previous nodes.
-For each metadata identifier in \textit{autotags},
-the query must find the associated metadata node and tag node.
-A connection between the two nodes is created with the confidence score as a
-property. The number of metadata nodes are dependent on the database size and
-the connections are responsible for 90\% of the elements in each database as
-shown in Table~\ref{table:vdmsnodes}.
-It is important to note that the metadata identifier, \textit{autotags}, and
-longitude/latitude coordinates are set as indexes of the database to allow
-faster retrieval of the metadata.
-
-\begin{table}[h]
-\caption{VDMS Database - Number of Elements}
-\centering
-\begin{tabular}{c c c c}
-\hline\hline
-DB Name & \# Images & \# Connections & \# Autotags\\
-\hline
-100k & 100,000     & 848,432      & 1,570\\
-500k & 500,000     & 4,249,500    & 1,570\\
-1M   & 1,000,000   & 8,503,045    & 1,570\\
-5M   & 5,000,000   & 42,505,478   & 1,570\\
-10M  & 10,000,000  & 85,040,404   & 1,570\\
-50M  & 50,000,000  & 425,162,070  & 1,570\\
-100M & 99,205,984  & 895,572,430  & 1,570\\
-\hline
-\end{tabular}
-\label{table:vdmsnodes}
-\end{table}
-
-Each MySQL database is created in a similar manner but the data is represented
-as three tables, following the relational model:
-1) \textit{images} table, contains one row per image,
-and a column for each property
-associated with the images (some of which are listed in \ref{dataset});
-2) \textit{taglist} table: contains one row per each autotag elements
-(always 1570 rows);
-3) \textit{autotags} table: contains on row per each autotag
-assigned to an image, each row containing a foreign key to the
-image and a foreign key to the tag, and
-the probability assigned to that tag belonging to that image.
-Given that there are 8 autotags, on average, per image, the \textit{autotags}
-table has around 8 times the number of rows present in the
-\textit{metadata} table, as can be seen in~\ref{table:mysqltables}.
-
-THIS NEEDS FIX, NOT SURE I UNDERSTAND WHAT YOU MEAN:
-By default, MySQL uses infinite number of threads inside InnoDB when processing
-requests using four threads per IO read/write.
-In some cases, when creating large databases data locks may occur to protect the
-data from concurrent updates.
-To minimize locking in our runs, we increased the InnoDB buffer pool size to
-increase the amount of memory allocated to in-memory data structures
-~\cite{mysql,mysql_blog}.
-
-Using a Python client and simple queries, the \textit{taglist}
-table is read from the list of tags with an auto-incremented
-\textit{tagid} as a primary key and the metadata table
-is read from the YFCC100m metadata using the identifier as a primary key.
-The \textit{autotags} table contains the generated autotags and
-probabilities for entries of the \textit{images} table.
-
-To generate the table, we split the \textit{autotags} data for each database
-by the image identifier and autotag into new files.
-The new files are read into the \textit{autotags} table with the image
-identifier and \textit{tagid} as foreign keys.
-
-\begin{table}[h]
-\caption{MySQL Database - Number of Rows in each Table}
-\centering
-\begin{tabular}{c c c c}
-\hline\hline
- & \multicolumn{3}{c}{Table}\\
-\cline{2-4}
-DB Name & images & autotags & taglist\\
-\hline
-100k & 100,000    & 848,912     & 1,570\\
-500k & 498,707    & 4,241,200   & 1,570\\
-1M   & 1,000,000  & 8,508,380   & 1,570\\
-5M   & 4,987,379  & 42,425,905  & 1,570\\
-10M  & 10,000,000 & 85,095,265  & 1,570\\
-50M  & 50,000,000 & 425,446,208 & 1,570\\
-100M & 99,206,564 & 896,002,496 & 1,570\\
-\hline
-\end{tabular}
-\label{table:mysqltables}
-\end{table}
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/db_time_size}
-\caption{Time (in hours) to build and size (in GB) of MySQL and VDMS databases}
-\label{fig:db_time_size}
-\end{figure}
-
-The VDMS and MySQL databases have comparable number of elements as shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}, and the differences
-can be attributed to failures in data preparation/loading because of
-incomplete/inconsistent formating, which is common on large datasets.
-The difference is less than 0.1\% in terms of number of elements (images and/or
-metadata information).
-
-\subsubsection{Database Loading Time}
-
-One of the first things we noticed is the difference is loading times,
-where VDMS outperforms MySQL by a large margin.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
-faster than MySQL, and how the speedup is sustained as the database size grows.
-Key difference in the build times are attributed to the low-level
-implementation of how MySQL reads and stores data from the files and the
-optimizations (increased InnoDB pool size, etc.) needed
-to handle large datasets such as YFCC100m.
-On average, it took MySQL 3.72x longer to build each database than VDMS.
-For example, to build the 100M database, MySQL took 263 hours
-while VDMS only needed 72.5 hours.
-
-\subsubsection{Database Storage Footprint}
-
-Another aspect important to note is that
-VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
-This is space used to store information about each node/connection.
-The Graph Database internal to VDMS (called PMGD) was designed for performance,
-specially in environments where persistent memory is present.
-This design decision comes as a trade-off for storage footprint, which is
-noticeable in our results.
-VDMS required 30-41\% more storage than MySQL for storing the same amount
-of metadata.
-This may become a factor if storage is a limitation, but it should also be noted
-that even if we have a 41\% increase in metadata size,
-metadata accounts for less than 2\% of the overall database size.
-Take, for example, the largest database (both metadata and images) we built
-(100M): around 230GB of metadata and 12TB of images.
-On the other hand, in systems were persistent memory is a scarce resource,
-the increase storage foot print of PMGD may represent a challenge.
-
-%=========================================
-
-\subsection{Images + Metadata}
-\label{images}
-
-In order to evaluate VDMS and the baseline on our use-case queries,
-we implemented 4 queries that that filter and retrieve certain specific images.
-We choose these queries because they represent typical use-case where a
-cohort of images is to be retrieve and processed from a large corpus of data.
-As we mentioned before, we took this approach due to the lack of a standard
-benchmarks that are oriented towards visual data retrieval.
-We use the metadata associated to the images to filter said images.
-In particular, this dataset contains a set of 1570 machine generated
-tags (\textit{autotags})assigned to the images depending on the content.
-\textit{autotags} includes things like objects (chair, person, people, etc), and
-concepts like places (indoor, outdoor, lake) and others (party, etc).
-Together with each \textit{autotag}, a probability is assign.
-This is, an image can have the \textit{autotags} people, person, party,
-outdoor, and each \textit{autotag} assigned will be accompanied by a
-probability of that \textit{autotags} being present in that image.
-As mentioned earlier, there are 8 tags assigned to each image on average.
-Also, some images have geo-location information (latitude/longitude).
-
-We use the \textit{autotags} (as they contain information about
-the content of the image), as well as the geo-location
-(as an example of properties of the images used used for search and filtering).
-
-On top of that and for our use case, we would like to extract more information
-about the content of the image through the use of ML,
-such as Convolutional Neural Networks TODO-CITE HERE.
-For this, we resize the images to 224x224, which is the input layer size for
-a popular neural network for object detection on images: ResNet TODO-CITEHERE.
-
-To evaluate the access to metadata and images,
-we use the following four queries, modeled after our internal use cases:
-\begin{itemize}
-\item Query 1 - {\bf {\em 1tag}}: Find metadata/images with 1 specific autotag (i.e. alligator, lake, etc).
-\item Query 2 - {\bf {\em 1tag\_resize}}: Find metadata/images with 1 specific autotag and resize to 224x224.
-\item Query 3 - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with 1 specific autotag, resize to 224x224, and in a particular geolocation (with a 20 degrees radius in latitude and longitude).
-\item Query 4 - {\bf {\em 2tag\_resize\_geo}}: Find metadata/images with 2 specific tags (i.e. alligator AND lake) and resize to 224x224.
-\end{itemize}
-
-It is important to note that when querying for images with certain
-\textit{autotags}, we also apply a filter using the probability.
-For instance, we only retrieve images with a autotag \textit{alligator}
-and a probability higher than 92\%.
-This probabilities are both present in VDMS (in the form of a property
-of the \textit{connection} between the image and that \textit{autotag}),
-as well as in MySQL (in the form of a column in the \textit{autotags} table
-that links images with tags.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on Query \textit{1tag\_resize} \- VDMS vs Baseline}
-\label{fig:concurrency_comparison}
-\end{figure}
-
-% We decided to remove this image because it is redudant with the concurrency
-% analysis.
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_32}
-% \caption{Queries - Throughput - 32 Clients}
-% \label{fig:q_throughput_32}
-% \end{figure}
-
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/queries_throughput_56}
-% \caption{Queries - Throughput - 56 Clients}
-% \label{fig:q_throughput_56}
-% \end{figure}
-
-\begin{figure*}[ht]
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include first image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2}
-  \caption{Query 1 (red), and Query 2 (blue).}
-  \label{fig:q1_q2}
-\end{subfigure}
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include second image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q3_q4}
-  \caption{Query 3 (red), and Query 4 (blue).}
-  \label{fig:q3_q4}
-\end{subfigure}
-\caption{Queries - Throughput - 56 Clients}
-\label{fig:q_throughput_56}
-\end{figure*}
-
-Image search based on metadata is very expensive on large databases.
-Because of the large volume of data, the processing of the retrieve images
-is perform in parallel, using multi-core and/or distributed systems.
-For instance, a common implementation of an image processing pipeline
-would involve the use of distributed processing frameworks
-like Hadoop \cite{hadoop} or Spark \cite{spark}.
-Consequently, it is key that the data management system used supports
-concurrency, providing multiple workers with data in parallel.
-The ability to scale with the number of simultaneous clients is key for the
-applicability of visual data management systems like VDMS.
-
-Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
-the \textit{1tag\_resize} query, described above, using both VDMS and
-the baseline to evaluate the scalability of both solutions, as the
-size of the databases grows and as the number of concurrent clients grows.
-
-It is important to note that the size of the result (number of images retrieved)
-is linear with the size of the database. This is, if a query returns 100 images
-for the 1M database, it will return around 1000 images for the 10M database.
-This poses a problem when evaluating performance as database increase,
-and clearly understanding the measurements.
-Because of this reason, we control the number of returned images for all the
-database using the probability of the \textit{autotags}, so that the queries
-in this experiment returns a similar number of images for all database sizes.
-Essentially, as the size of the database increase, we increase the probability
-threshold for the queries. We do this for both VDMS and the baseline, of course.
-
-The first thing to notice is that when it comes to retrieving metadata
-in the case of VDMS, (Figure ~\ref{fig:concurrency_comparison} (top),
-the size of the database (in terms of number of images) does not have
-a strong effect on the performance
-(measured in aggregated transactions per second).
-This means that even as database increase in size, there is no large increase
-in the query time.
-
- the number of concurrent clients to use for the per-query performance analysis.
-This figure shows the metadata performance plateaus at 32 clients and
-the image performance at 56 clients for a simple query.
-% To further investigate, Figure~\ref{fig:q_throughput_32} and
-% Figure~\ref{fig:q_throughput_56} illustrate the throughput performance
-% for the \textit{1tag} and \textit{1tag\_resize} queries for 32 and 56 clients,
-% respectively.
-On the 56-client experiment, a 10x difference can be seen between VDMS and
-baseline in metadata throughput while 32 clients is similar but not as high.
-The overall performance for both MySQL and VDMS is better with 56 clients;
-therefore, we used this value for our per-query analysis.
-
-\begin{figure}[]
-\centering
-\includegraphics[width=\columnwidth]{figures/q1_latency}
-\caption{Query 1 - Latency}
-\label{fig:q1_latency}
-\end{figure}
-
-%=========================================
-
-\subsection{Videos}
-\label{videos}
-
-VDMS provides full support for video storage and operations.
-This includes support for encoding/decoding and trans-coding of
-\textit{mp4}, \textit{avi}, and \textit{mov} containers,
-as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
-This is supported through the Visual Compute Module that provides an abstraction
-layer on top of OpenCV (and \textit{ffmpeg} directly).
-All the operations supported for images in VDMS are also supported at the
-video and frame level on the API. On top of that, there are a number of
-video-specific operations that are supported, such as the interval operations,
-that allow users to retrieve clips at different
-frames-per-second (FPS) versions of the video.
-
-All this functionality is provided and integrated with the rest of the
-metadata API as part of the comprehensive VDMS interface. This makes it
-possible for users to interact with metadata and video in a transactional
-manner, enabling users to run queries like: "Retrieve all the videos
-where there is a lake with probability higher than X, converting all videos
-to \textit{H.264} \textit{mp4} of size 224x244".
-In particular, this functionality was used internally to select a subset
-of videos with the right licenses for a video summarization application.
-
-To the best of our knowledge, there is no solution that can provide
-all the functionality mentioned above. Also, implementing a baseline is a very
-complex task as there is a large number of options and parameters that can
-be chosen, which makes it hard to accurately compare against VDMS functionality.
-For this reason, we chose to make a study using VDMS in various scenarios,
-and analysis what is the impact of having the overhead of VDMS' Request
-Server in the overall access time.
-
 \begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/video_overhead}
-\caption{Concurrency (left) and Overhead (right)}
-\label{fig:video}
-\end{figure*}
-
-Figure \ref{fig:video} shows the analysis of different queries aimed to retrieve
-a video using VDMS interface. We show how VDMS increases throughput of serving
-a video object as the number of simultaneous client increases, as well as the
-overhead VDMS introduces in the overall query execution time.
-The figure on the left compares the number of video transaction per second
-(i.e., number of videos returned per second) when different operations
-are executed as part of the transaction. The upper-bound of this would be
-simply returning the video as-is (without running any encoding/decoding or
-operation), represented by the red line. This query is the upper-limit because
-it essentially translates to reading the video from the file-system and sending
-it over a TCP/IP socket, without any other overhead or operations.
-
-We also run a set of other queries that involve: (a) running a resize operation
-on the video and, consequently, needs a decoding and
-encoding operations as well (blue line),
-(b) transcoding, meaning the use a different container and encoder
-than the one originally used (yellow line), and (c) both resize and transcoding.
-Note that the resize operation performs a downsize, which translates in less
-data being sent over the wire. This is specially noticeable when supporting 32
-simultaneous clients, where the systems provides more videos per second due to
-sending less data to the client, when compares to just transcoding and not resizing (yellow line).
-
-We can see that the system performs best when using all the physical cores.
-COMPLETE THIS BETTER.
-
-WE SHOULD ADD HERE SAMPLES OF THE QUERYS TO SHOW HOW THESE ARE RETRIEVED.
-
-Because we see almost 3 orders of magnitude drop in performance when including
-operations as part of the query, we wanted to understand where most of the time
-was spent on the query, and optimize the Request Server and Visual Compute Module
-if necessary. For this, we run the experiment shown at
-Figure \ref{fig:video} (right) which breaks down the different components of the
-queries. This figure shows that more than 97\% of the query execution is spent
-on encoding/decoding operations, which is well-known to be a
-heavy and compute intensive operation.
-On the one hand, this results show that VDMS introduced overhead for
-video operation is minimal. On the other hand, this result means a
-limit on the opportunities for optimization for video queries given
-that biggest time factors are accounted by encoding/decoding, which is
-outside the scope of VDMS.
-This result also was the inspiration point for one optimization we included
-in future versions of VDMS, which involves using \textit{ffmpeg} C++ API to
-limit the number of frames being encoding/decoding when possible.
-
-%=========================================
-
-\subsection{Feature Vectors}
-\label{features}
-
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities or visual objects.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as \textit{descriptors}
-or \textit{visual descriptors}. We use these terms interchangeably.
-These descriptors can be classified and labeled and used to build search indexes,
-and there are many in-memory libraries that are designed for
-this task~\cite{flann, faiss}.
-Using the VDMS API, users can manage feature vector indexes,
-query previously inserted elements (images),
-run a k-nearest neighbor search (\textit{knn}), and express relationships
-between existing images or descriptors and
-the newly inserted descriptors.
-By natively supporting descriptors and \textit{knn},
-VDMS allows out-of-the-box classification functionalities for many applications.
-
-For this work, and as part of a comprehensive image search implementation,
-we have used 4096-dimensional descriptors extracted from every image
-(and first frame of every video) and created a collection of these feature
-vectors in VDMS to be able to perform similarity search (i.e., find
-images that are \textit{similar} to an query (input) image).
-\textit{Similarity} in this particular case is defined as closeness
-in a an 4096-dimensional space using euclidean distance as the metric.
-
-EXPLAIN HERE DIFFERENT ACCURACY/EXECUTIME TIME TRADEOFF.
-
-The process of loading descriptors in VDMS is simple.
-First, the user has to create a DescriptorsSet, using a single command.
-At creation of the DescriptorSet, the dimensionality of the descriptors
-is specified, together with the desired indexing method and the desired metric
-for computing distances (Euclidean Distance, \textit{L2},
-or Inner Product, \textit{IP}).
-Once the DescriptorSet is created, descriptors can be inserted to the set.
-After the descriptors are inserted, similarity search can be performed.
-Note that descriptors can be inserted later on, and any search will reflect
-the existence of all descriptors in the set.
-
-\begin{figure*}[]
 \centering
@@
 
-Figure \ref{fig:similarity} shows 3 examples of a query image (on the left),
-and images returned as "similar" by VDMS.
-The query input is a descriptor generated after a query image. The "query"
-descriptor is sent to VDMS as part of the query, and VDMS use that descriptor
-to find similar one, and retrieve the images associated with those "similar"
-descriptors. We used this as an example and as a visual validation of the
-functionality and applicability in this particular dataset, but we also
-provide an analytical approach to accuracy and trade-offs in our system.
+\subsection{Similarity Search}
+\label{features}
+
+Another key differentiating factor of VDMS is that it allows the creation of
+indexes for high-dimensional feature vectors and the insertion of
+these feature vectors associated with entities, images, and/or videos.
+Feature vectors are intermediate results of various machine
+learning or computer vision algorithms when run on visual data.
+Feature vectors are also known as \textit{descriptors}
+or \textit{visual descriptors}. We use these terms interchangeably.
+These descriptors can be classified, labeled, and used to build search
+indexes. There are many in-memory libraries that are designed for
+this task~\cite{flann, faiss}.
+
+% We analyze the behavior of the feature vector functionality in VDMS,
+% and an evaluation of the different trade-offs that the systems offers for
+% application developers. For this, we implemented an image-search application
+% based on \textit{similarity} search.
+
+Using the VDMS API, users can manage feature vector indexes,
+query previously inserted elements,
+run a k-nearest neighbor search (\textit{knn}), and express relationships
+between existing images or descriptors and
+the newly inserted descriptors.
+By natively supporting descriptors and \textit{knn},
+VDMS allows out-of-the-box classification functionalities for many applications
+\footnote{https://github.com/\{Not shown during submission\}}.
+% \footnote{https://github.com/IntelLabs/vdms/wiki/ClassifyDescriptor}.
+
+For this work, and as part of a comprehensive image search implementation,
+we have used 4096-dimensional descriptors extracted from every image
+(and first frame of every video) from the YFCC100M dataset
+and created a collection of these feature vectors in VDMS to
+perform similarity search (i.e., find images that are
+\textit{similar} to an query (input) image).
+\textit{Similarity} in this particular case is defined as closeness
+in a 4096-dimensional space using euclidean distance as the metric.
+
+The process of loading descriptors in VDMS is simple.
+First, the user has to create a DescriptorSet, using a single command.
+At creation of the DescriptorSet, the dimensionality of the descriptors
+is specified, together with the desired indexing method and the desired metric
+for computing distances (Euclidean Distance, \textit{L2},
+or Inner Product, \textit{IP}).
+Once the DescriptorSet is created, descriptors can be inserted to the set.
+After the descriptors are inserted, a similarity search can be performed.
+
+Figure~\ref{fig:similarity} shows 3 examples of a query image (on the left),
+and images returned as \textit{similar} by VDMS.
+The input is a descriptor generated after a query image.
+The \textit{query input} descriptor is sent to VDMS as part of the query,
+VDMS uses that descriptor to find similar ones,
+and retrieves the images associated with those \textit{similar} descriptors.
+We show this as an example of the functionality and to depict
+how the feature vectors provided by the dataset can be used,
+but we also provide an analytical approach to
+the trade-off between accuracy and execution time in our system.
 It is important to note that the accuracy of the results is entirely tied
 to the quality of the descriptors chosen by the applications.
-VDMS is completely agnostic to information contained within the descriptor,
-and simply offers the interface to store and index them, but the quality
-of the similarity result will be tied to the quality of descriptor extraction
-that the application is using.
+The quality of the similarity result will be tied to the quality
+of the descriptor extraction that the application is using.
 
-\begin{figure*}[]
+\begin{figure*}[ht]
 \centering

### 2021-02-03 yfcc100m/paper/osdi2020/main.tex
Update OSDI submission files

@@
 \usepackage{minted}
+\usepackage{todonotes}
 
@@
 %for single author (just remove % characters)
-\author{
-{\rm Luis Remis}\\
-ApertureData Inc
-\and
-{\rm Chaunt\'e W. Lacewel}\\
-Intel Labs
-% copy the following lines to add more authors
+% \author{
+% {\rm Luis Remis}\\
+% ApertureData Inc
 % \and
-% {\rm Name}\\
-%Name Institution
-} % end author
+% {\rm Chaunt\'e W. Lacewel}\\
+% Intel Labs
+% % copy the following lines to add more authors
+% % \and
+% % {\rm Name}\\
+% %Name Institution
+% } % end author
 
@@
 rather than doing what they know best:
-build machine learning models and algorithms to solve problems.
-The problem grows as the scale of the data used for analysis or training
-grows. In this paper, we demonstrate how VDMS (the Visual Data Management System)
-can be used to simplify the data preparation process and how we gain
-in efficiency simply because we are using a system designed for the job.
+build machine learning models and algorithms to solve previously unsolvable problems.
+In this paper, we describe the Visual Data Management System (VDMS),
+and demonstrate how it can be used to simplify the data preparation process 
+and consequently gain in efficiency simply because 
+we are using a system designed for the job.
 To demonstrate this, we use one of the largest available
 public datasets (YFCC100M),
-with 99+ million images, 800+ thousand videos, plus add-ons that include
+with 99+ million images and videos, plus add-ons that include
 machine-generated tags and 4K dimensional feature vectors,
 for a total of $\sim$13TB of data.
-VDMS differs from existing large scale photo
-serving, video streaming, and textual big-data management systems due to
-its primary focus on supporting machine learning and
-data analytics pipelines that use visual data
-(images, videos, and feature vectors), treating these as first class citizens.
+VDMS differs from existing data management systems
+due to its focus on supporting machine learning and
+data analytics pipelines that rely on images, videos, and feature vectors, 
+treating these as first class citizens.
 We demonstrate how VDMS outperforms well-known and widely used
 systems for data management by up to $\sim$35x, with
-an average improvement of about 15x for the queries we
-used in our use-cases, and particularly at scale.
+an average improvement of about 15x for our use-cases, and particularly at scale.
 At the same time, VDMS simplifies the process of data preparation and data access,
@@
 \bibliography{main}
-
 \input{legal.tex}
@@
 %-------------------------------------------------------------------------------
-\section*{Availability}
-%-------------------------------------------------------------------------------
+% \section*{Availability}
+% %-------------------------------------------------------------------------------
 
-USENIX program committees give extra points to submissions that are
-backed by artifacts that are publicly available. If you made your code
-or data available, it's worth mentioning this fact in a dedicated
-section.
+% USENIX program committees give extra points to submissions that are
+% backed by artifacts that are publicly available. If you made your code
+% or data available, it's worth mentioning this fact in a dedicated
+% section.
 

### 2021-02-04 yfcc100m/python/eval/performance.py
Update with PostgreSQL

@@ -5,6 +5,7 @@ from threading import Thread
 import pandas as pd
 import numpy as np
 
+from postgresql_eval  import PostgreSQLQuery
 from mysql_eval  import MySQLQuery
 from memsql_eval import MemSQLQuery
 from vdms_eval   import VDMSQuery
@@ -15,11 +16,6 @@ VDMS_PORT_MAPPING = {'100k': 55500, '500k': 55405,
                      '1M':   55501, '5M':   55450,
                      '10M':  55510, '50M':  55000,
                      '100M': 50000}
-
-# SIZE_PROB_MAPPING = {'100k': 0.60,       '500k': 0.06,
-#                      '1M':   0.92175,    '5M':   0.955,
-#                      '10M':  0.952499,   '50M':  0.95375,
-#                      '100M': 0.967656249}
 
 # These are pre-computed probabilities so that queries returned
 # a controlled number of results, to prevent the throughput to be
@@ -145,8 +141,8 @@ def get_args():
 
     # Database Info
     obj.add_argument('-db_type', type=str, default='vdms',
-                     choices=["vdms", "memsql", "mysql"],
-                     help='Database names: vdms, memsql, mysql')
+                     choices=["vdms", "memsql", "mysql", "postgresql"],
+                     help='Database names: vdms, memsql, mysql, postgresql')
     obj.add_argument('-db_name', type=str, default='100k',
                      choices=VDMS_PORT_MAPPING.keys(),
                      help='Database names: 100k, 1M, 10M')
@@ -177,13 +173,20 @@ def get_args():
     elif params.db_type == "mysql":
         params.db_host = "127.0.0.1" #TODO: Get to work with hostname
         params.db_port = 3360
+    elif params.db_type == "postgresql":
+        if params.db_name == '50M':
+            params.db_host = "sky4.local"
+        else:
+            params.db_host = "127.0.0.1"
+        params.db_port = 5432
+        params.db_user = "root"
+        params.db_pswd = "password"
     else:
         # VDMS needs port mapping
         params.db_host = "sky4.local"
         params.db_port = VDMS_PORT_MAPPING[params.db_name]
 
     return params
-
 
 def run_query_thread(obj, params, index, results, query_arguments):
 
@@ -205,6 +208,9 @@ def run_query(params, query_arguments):
         for i in range(params.numthreads):
             list_of_objs.append(
                         VDMSQuery.VDMSQuery(params.db_host, params.db_port))
[…]

### 2021-02-04 yfcc100m/python/eval/postgresql_eval/PostgreSQLQuery.py

@@ -1,0 +1,230 @@
+import psycopg2
+from urllib.parse import urlparse
+import time
+import os
+import cv2
+import random
+import numpy as np
+import requests
+
+IMG_HOST = 'sky4.local'
+
+def create_dir(folder):
+    if not os.path.isdir(folder):
+        os.mkdir(folder)
+
+class PostgreSQL(object):
+    def __init__(self, params):
+        self.db = self.get_connection(params)
+        self.db_cursor = self.db.cursor()
+
+    def close_connection(self):
+        self.db.close()
+        self.db_cursor.close()
+
+    def get_connection(self, params):
+        return psycopg2.connect(host=params.db_host, user=params.db_user,
+                            password=params.db_pswd, port=params.db_port,
+                            dbname="yfcc_" + params.db_name)
+
+
+    # Runs a query, described in the "query" object param
+    def run_query(self, query):
+
+        key   = query['key']
+        tags  = query['tags']
+        probs = query['probs']
+        lat   = query['lat']  if 'lat'  in query else -1
+        long  = query['long'] if 'long' in query else -1
+        range_dist = query['range_dist'] if 'range_dist' in query else 0
+        operations = query['operations'] if 'operations' in query else []
+        comptype   = query["comptype"]   if 'comptype'   in query else "or"
+
+        # Metadata only queries
+
+        if key == "1tag":
+            return self.get_metadata(tags, probs,
+                                     comptype=comptype)
+
+        if key == "1tag_loc20":
+            return self.get_metadata(tags, probs,
+                                     lat=lat,
+                                     long=long,
+                                     range_dist=range_dist,
+                                     comptype=comptype,
+                                     )
+
+        if key == "2tag_and":
+            return self.get_metadata(tags, probs, comptype=comptype, )
+
+        if key == "2tag_or":
+            return self.get_metadata(tags, probs, comptype=comptype, )
+
+        if key == "2tag_loc20_and":
+            return self.get_metadata(tags, probs,
+                                     lat=lat,
+                                     long=long,
+                                     range_dist=range_dist,
+                                     comptype=comptype,
+                                     )
+
+        if key == "2tag_loc20_or":
+            return self.get_metadata(tags, probs,
+                                     lat=lat,
+                                     long=long,
+                                     range_dist=range_dist,
+                                     comptype=comptype,
+                                     )
+
+        # Image queries
+
+        if key == "1tag_resize":
+            return self.get_images(tags, probs,
[…]

### 2021-02-04 yfcc100m/python/eval/postgresql_eval/build_dbs.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+set -e
+
+# Build dbs on 127.0.0.1 or sky4.local
+
+# for size in 100k
+#for size in 1M 5M 10M 50M 100M
+for size in 100M
+#for size in 1M 5M 10M
+do
+case "$size" in
+  50M)
+    host=sky4.local
+    ;;
+  *)
+    host=127.0.0.1
+    ;;
+esac
+echo "HOST: ${host}"
+echo "Processing yfcc_${size} Database"
+sudo -u root psql -h $host -c "drop database if exists \"yfcc_${size}\";"
+sudo -u root psql -h $host -c "create database \"yfcc_${size}\" OWNER root;"
+python3 build_yfcc_db_postgres.py \
+    -data_file \
+    /mnt/yfcc100m/metadata/processed/yfcc100m_photo_dataset_${size} \
+    -tag_file \
+    /mnt/yfcc100m/metadata/processed/yfcc100m_photo_autotags_${size}_extended \
+    -tag_list /mnt/yfcc100m/metadata/processed/autotag_list.txt \
+    -db_name "${size}" \
+    -db_host "${host}" \
+    -db_port 5432 \
+    -db_user "root" \
+    -db_pswd "password"
+done

### 2021-02-04 yfcc100m/python/eval/postgresql_eval/build_yfcc_db_postgres.py

@@ -1,0 +1,229 @@
+"""
+Create Databases in PostgreSQL
+"""
+
+from pathlib import Path
+import pandas as pd
+import numpy as np
+import time
+import math
+import logging
+import argparse
+import threading
+import psycopg2
+import sys, os
+import csv
+import PostgreSQLQuery
+
+connection_batch_limit = 100
+property_names = ['Line number', 'ID', 'Hash', 'User NSID',
+                  'User nickname', 'Date taken', 'Date uploaded',
+                  'Capture device',
+                  'Title', 'Description', 'User tags', 'Machine tags',
+                  'Longitude', 'Latitude', 'Coord. Accuracy', 'Page URL',
+                  'Download URL', 'License name', 'License URL',
+                  'Server ID', 'Farm ID', 'Secret', 'Secret original',
+                  'Extension', 'Marker']
+property_names_sql = ['line_number','id', 'hash', 'user_nsid', 'user_nickname',
+            'date_taken', 'date_uploaded', 'capture_device', 'title', 'description',
+            'user_tags', 'machine_tags', 'longitude', 'latitude', 'coord_accuracy',
+            'page_url', 'download_url', 'license_name', 'license_url', 'server_id',
+            'farm_id', 'secret', 'secret_original', 'extension', 'marker',
+            'imgPath', 'imgBlob', 'format']
+
+def get_args():
+    parserobj = argparse.ArgumentParser()
+    parserobj.add_argument('-num_threads', type=int, default=100,
+                           help='Number of threads to use [default: 100]')
+    parserobj.add_argument('-batch_size', type=int, default=100,
+                           help='Number of entries per thread for autotags and images [default: 100; connections: max {}]'.format(connection_batch_limit))
+    parserobj.add_argument('-data_file', type=lambda s: Path(s),
+                           default='/mnt/yfcc100m/metadata/processed/yfcc100m_photo_dataset',
+                           help='YFCC metadata [default: ' +
+                           '/mnt/yfcc100m/metadata/processed/yfcc100m_photo_dataset')
+    parserobj.add_argument('-tag_list', type=lambda s: Path(s),
+                            default='../yfcc_parse_labels/autotag_list.txt',
[…]

### 2021-02-04 yfcc100m/python/eval/postgresql_eval/postgres_queries.py

@@ -1,0 +1,79 @@
+import PostgreSQLQuery
+from collections import namedtuple
+import os
+import time
+
+import psycopg2
+
+def display_images(imgs):
+    from IPython.display import Image, display
+    img_dir = 'images/'
+    if not os.path.exists(img_dir):
+        os.makedirs(img_dir)
+    counter = 0
+    for im in imgs:
+        img_file = img_dir + 'results_' + str(counter) + '.jpg'
+        counter = counter + 1
+        fd = open(img_file, 'wb+')
+        fd.write(im)
+        fd.close()
+        display(Image(img_file))
+
+
+args = {'db_name':'1M',
+        'db_host':"127.0.0.1",
+        'db_port': 5432,
+        'db_user': 'root',
+        'db_pswd': 'password',
+        'get_imgs': True}
+
+resize = {
+    "type": "resize",
+    "width": 224,
+    "height": 224
+}
+PORT_MAPPING = {'yfcc_100k': '100k', 'yfcc_1M': '1M', 'yfcc_10M': '10M'}
+params = namedtuple("Arguments", args.keys())(*args.values())
+qh = PostgreSQLQuery.PostgreSQL(params)
+
+probability = 0.9
+latitude  = -14.354356
+longitude = -39.002567
+
+print('\nQuery metadata with autotags: alligator>=', probability, 'OR lake>=', probability)
+print('Num IDs: ',len(qh.get_metadata_by_tags(["alligator", "lake"], [probability, probability], comptype='or')))
+if args['get_imgs']:
+    print('Query images with autotags: alligator>=', probability, 'OR lake>=', probability)
+    res = qh.get_images_by_tags(["alligator", "lake"], [probability, probability], [resize], comptype='or')
+    print(len(res))
+
+print('\nQuery metadata with autotags: alligator>=', probability, 'AND lake>=', probability)
+print('Num IDs: ',len(qh.get_metadata_by_tags(["alligator", "lake"], [probability, probability])))
+if args['get_imgs']:
+    print('Query images with autotags: alligator>=', probability, 'AND lake>=', probability)
+    qh.get_images_by_tags(["alligator", "lake"], [probability, probability], [resize])
+
+print('\nQuery metadata with autotags: alligator>=', probability, 'AND lake>=', probability, 'within 20 of lat',  latitude, ', long',  longitude)
+print('Num IDs: ',len(qh.get_metadata_by_tags(["alligator", "lake"], [probability, probability], latitude, longitude, 20)))
[…]

### 2021-02-04 yfcc100m/python/eval/run_performance.sh

@@ -20,7 +20,7 @@ do
     # for size in 1M 5M 10M 50M
     for size in 1M 5M 10M 50M 100M
     do
-        for db in vdms mysql
+        for db in vdms postgresql
         do
             # Run VDMS Queries
             echo "Running $th $db ${size}..." >> $info

### 2021-02-09 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Fix stds plotting, rename mosaics

@@ -155,12 +155,12 @@ class EvalFramework(object):
 
                 times     = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
                                                             "query_time_avg")
+                times_std = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
+                                                            "query_time_std")
                 threads   = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
                                                             "n_threads")
                 n_results = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
                                                             "n_results")
-                stds      = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
-                                                            "n_results_std")
 
                 from numpy import inf
                 from numpy import nan
@@ -168,13 +168,8 @@ class EvalFramework(object):
                 # Compute Results per second
                 rps   = 1/times * threads * n_results
                 rps[rps == inf] = 0
-                # where_are_NaNs = np.isnan(rps)
-                # rps[where_are_NaNs] = 0
 
-                stds  = 1/stds  * threads * n_results
-                stds[stds == inf] = 0
-                # where_are_NaNs = np.isnan(stds)
-                # stds[where_are_NaNs] = 0
+                stds  = rps * (times_std / times)
 
                 values = np.vstack((values, np.append(rps, stds)))
 
@@ -193,7 +188,7 @@ class EvalFramework(object):
                           ylabel=result_type + "/s")
 
         filename  = self.plot_folder + "plot_conc_"
-        filename += str(q) + "_results_throughput_db_size_mosaic.pdf"
+        filename += str(q) + "_mosaic_results_throughput_db_size.pdf"
 
         title = "Throughput as " + result_type + " per second for different queries"
         p.plot_lines_all_mosaic(db_sizes, threads, engines, values,
@@ -219,6 +214,8 @@ class EvalFramework(object):
 
                 times     = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
                                                             "query_time_avg")
+                times_std = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
+                                                            "query_time_std")
                 threads   = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
                                                             "n_threads")
                 n_results = self.get_arr_for_eng_q_dbsize(eng, q, db_size,
[…]

### 2021-02-09 yfcc100m/python/eval/performance.py
Add warning message when rejecting too many outliers

@@ -277,9 +277,14 @@ def main(params):
 
         print("\n")
 
+        len_all_times = len(all_times)
+        len_all_len   = len(all_len)
         # Reject outliers outside 2 std on measurements.
         all_times = reject_outliers(np.array(all_times))
         all_len   = reject_outliers(np.array(all_len))
+
+        if (len(all_times) / len_all_times) < 0.4:
+            print("WARNING: Rejecting too many samples!")
 
         evfw.add_row(query_args["key"],
                      params.db_type,

### 2021-02-11 yfcc100m/paper/vldb2021/design.tex
Update figures, other minor fixes

--- 2021-02-03 yfcc100m/paper/osdi2020/design.tex
@@
 VDMS and its components are fully available as open source projects
-% \footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
-\footnote{https://github.com/\{Not shown during submission\}}.
+\footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
 We briefly describe each of the main components as follows:
@@
 documented on the project's Github
-wiki~\footnote{https://github.com/\{Not shown during submission\}}.
-% wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.
+wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.
 

### 2021-02-11 yfcc100m/paper/vldb2021/eval.tex
Update figures, other minor fixes

@@
 In an attempt to have the best MySQL configuration possible for this use case,
-we explore several parameters to increase the performance of both loading the data, 
-as well as executing the queries.
+we explore several parameters to increase the performance
+of both loading the data, as well as executing the queries.
 In particular, MySQL optimizes threads and transactions out-of-box,
@@
 under the benchmarks branch of the VDMS project
-\footnote{https://github.com/\{Not shown during submission\}}
-% \footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
+\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
 and some examples of the queries can be found in the appendix of this
@@
 
-\begin{figure}[ht]
-\includegraphics[width=\columnwidth]{figures/concurrency_comparison}
-\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_resize}).
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{figures/plot_conc_1tag_loc20_resize_mosaic_results_throughput_db_size}
+\caption{Concurrency Analysis on \textit{q3} (\textit{1tag\_resize\_geo}).
 Hardware concurrency (number of physical cores in each system)
@@
 \label{fig:concurrency_comparison}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{figures/plot_conc_10M_mosaic_results_throughput}
+\caption{Concurrency Analysis for all the queries on the 10M database).
+Hardware concurrency (number of physical cores in each system)
+is shown with a blue vertical line (hw-concurrency = 56).
+Top figure shows aggregated throughput (transactions per second)
+when retrieving only metadata associated with the images, as the number of
+concurrent clients increase.
+Bottom figure shows aggregated throughput (images per second) when retrieving
+resized versions of the images, as the number of concurrent clients increases.}
+\label{fig:concurrency_comparison_10M}
+\end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include first image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q1_q2_q3}
-  \caption{\textit{q1}(red), \textit{q2}(blue),and \textit{q3}(orange).}
-  \label{fig:q1_q2}
-\end{subfigure}
-\begin{subfigure}{.5\linewidth}
-  \centering
-  % include second image
-  \includegraphics[width=\columnwidth]{figures/queries_throughput_56_q4_q5}
-  \caption{\textit{q4}(black), and \textit{q5}(pink).}
-  \label{fig:q3_q4}
-\end{subfigure}
-\caption{Performance Analysis using 5 queries from our use-case
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{figures/plot_56_mosaic_query_times}
+\caption{Performance Analysis using all queries from our use-case
 described in the Experimental Setup Section.
@@
 These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged out of 10 runs (niter = 10),
-each client running 2 transactions (ntags = 2).}
+and averaged out of 10 runs.}
+\label{fig:q_performance_56}
+\end{figure*}
+
+\begin{figure*}[ht]
+\includegraphics[width=\textwidth]{figures/plot_56_mosaic_results_throughput}
+\caption{Throughput Analysis using all queries from our use-case
+described in the Experimental Setup Section.
+We show queries in different figures for readability reasons.
+Top figures show the throughput of the queries for
+just retrieving metadata associated with the images.
+Bottom figures show the throughput when retrieving both metadata and images,
+plus operations applied to images when applicable (queries 2, 3, 4 and 5).
+The experiments show the performance of both systems (VDMS and baseline) as the
+database size increases.
+These queries were run using 56 simultaneous clients (nthr = 56),
+and averaged out of 10 runs.}
 \label{fig:q_throughput_56}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/summary_images}
+\includegraphics[width=\textwidth]{figures/plot_56_query_times_speedup}
 \caption{Summary of performance gains for all queries.

### 2021-02-13 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Add plot per queries for each n threads

@@ -1,5 +1,9 @@
 import os
+
 import numpy as np
+from numpy import inf
+from numpy import nan
+
 import pandas as pd
 
 from DBEvalFramework import Plotting
@@ -88,6 +92,15 @@ class EvalFramework(object):
 
         return arr
 
+    def get_arr_for_eng_q_threads(self, engine, query, threads, key):
+
+        arr = self.data[self.data["query"] == query]
+        arr = arr[self.data["engine"]      == engine]
+        arr = arr[self.data["n_threads"]   == threads]
+        arr = arr[key].to_numpy()
+
+        return arr
+
     def plot_all(self, folder="plots"):
         self.plot_folder=folder
         self.plot_folder += "/"
@@ -96,12 +109,27 @@ class EvalFramework(object):
             os.makedirs(self.plot_folder)
 
         # print(self.data)
+        self.plot_all_for_all_queries()
         self.plot_all_for_db_size()
         self.plot_all_for_n_clients()
 
+    def plot_all_for_all_queries(self):
+
+        threads  = self.get_unique("n_threads")
+        queries  = self.get_unique("query")
+
+        if len(queries) == 1:
+            return
+
+        print("Plotting plot_all_for_all_queries...")
+
+        for q in queries:
+            self.plot_results_throughput_parallelism_threads(q, result_type="Images")
+
+
     def plot_all_for_n_clients(self):
 
-        threads = self.get_unique("n_threads")
+        threads  = self.get_unique("n_threads")
         queries  = self.get_unique("query")
 
         if len(threads) == 1:
@@ -137,6 +165,67 @@ class EvalFramework(object):
             self.plot_query_time_speedup(i)
             self.plot_n_results(i)
 
+    def plot_results_throughput_parallelism_threads(self, q,
+                                            result_type="results"):
+
+        # Plot query times:
+        threads  = self.get_unique("n_threads")
+        queries  = self.get_unique("query")
+        engines  = self.get_unique("engine")
+        db_sizes = self.get_unique("db_size")
+
+        # Todo make general for more db_sizes
+        values = np.zeros(len(db_sizes) * 2)
+
+        for eng in engines:
+            for th in threads:
+
+                times     = self.get_arr_for_eng_q_threads(eng, q, th,
+                                                            "query_time_avg")
+                times_std = self.get_arr_for_eng_q_threads(eng, q, th,
+                                                            "query_time_std")
+                n_results = self.get_arr_for_eng_q_threads(eng, q, th,
[…]

### 2021-02-13 yfcc100m/python/eval/DBEvalFramework/Plotting.py

@@ -121,7 +121,18 @@ class Plotting(object):
         plt.xlabel(xlabel, fontsize=12)
         plt.ylabel(ylabel, fontsize=12)
 
-        plt.legend(loc="best", ncol=len(engines), shadow=True, fancybox=True)
+        if len(engines) > 2 or len(n_queries > 2):
+            plt.legend(bbox_to_anchor=(0, -0.15),
+                       loc="upper left",
+                       # mode="expand",
+                       borderaxespad=0,
+                       ncol=len(engines),
+                       fancybox=True,
+                       )
+        else:
+            plt.legend(loc="best",
+                       ncol=len(engines),
+                       shadow=True, fancybox=True)
 
         plt.savefig(filename, format="pdf", bbox_inches='tight')
         plt.close()
@@ -196,7 +207,8 @@ class Plotting(object):
 
             plt.xticks(x_pos, db_sizes, fontsize=10)
 
-            plt.legend(loc="best", ncol=1, shadow=True, fancybox=True)
+            if i == 0:
+                plt.legend(loc="best", ncol=1, shadow=True, fancybox=True)
 
             if i % side == 0:
                 plt.ylabel(ylabel, fontsize=12)
@@ -255,8 +267,18 @@ class Plotting(object):
 
         plt.axhline(y=1)
 
-        plt.legend(ncol=7, shadow=True, fancybox=True,
-                   fontsize=9.5, loc="upper left")
+        if len(queries) > 4:
+            plt.legend(bbox_to_anchor=(0, -0.15),
+                       loc="upper left",
+                       # mode="expand",
+                       borderaxespad=0,
+                       ncol=4,
+                       fancybox=True,
+                       fontsize=9.5,
+                       )
+        else:
+            plt.legend(ncol=7, shadow=True, fancybox=True,
+                       fontsize=9.5, loc="upper left")
 
         plt.savefig(filename, format="pdf", bbox_inches='tight')
         plt.close()

### 2021-02-13 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Fix Speedup plot for all engines

@@ -109,9 +109,9 @@ class EvalFramework(object):
             os.makedirs(self.plot_folder)
 
         # print(self.data)
-        self.plot_all_for_all_queries()
         self.plot_all_for_db_size()
         self.plot_all_for_n_clients()
+        self.plot_all_for_all_queries()
 
     def plot_all_for_all_queries(self):
 
@@ -532,50 +532,56 @@ class EvalFramework(object):
 
         # Plot query times:
         db_sizes = self.get_unique("db_size")
-        queries  = self.get_unique("query")
         engines  = self.get_unique("engine")
 
         if len(engines) < 2:
             print("Single engine, no speedup plot generated")
             return
 
-        values = np.zeros(len(db_sizes) * 2)
+        # This will only compute speedup of eng[0] vs eng[i > 0].
 
-        # This will only compute speedup of eng[0] vs eng[1].
-        # TODO: Iterate through engines here to generate
-        # eng[0] vs all others. Should be easy.
+        for sub_eng in range(1,len(engines)):
 
-        for eng in engines:
-            for q in queries:
+            # We need to get queries every times because we append "avg"
+            queries  = self.get_unique("query")
 
-                times = self.get_arr_for_eng_q_threads(eng,q, n_threads,
+            s_engines = [engines[0], engines[sub_eng]]
+            values = np.zeros(len(db_sizes) * 2)
+
+            for eng in s_engines:
+                for q in queries:
+
+                    times = self.get_arr_for_eng_q_threads(eng, q, n_threads,
                                                             "query_time_avg")
-                stds  = self.get_arr_for_eng_q_threads(eng, q, n_threads,
+                    stds  = self.get_arr_for_eng_q_threads(eng, q, n_threads,
                                                             "query_time_std")
 
-                values = np.vstack((values, np.append(times, stds)))
+                    values = np.vstack((values, np.append(times, stds)))
 
-        values = values[1:,:] # remove initial row of zeros
+            values = values[1:,:] # remove initial row of zeros
 
-        # Compute speedup
-        for i in range(len(queries)):
-            values[i,:] = values[len(queries)+i, :] / values[i, :]
+            # Compute speedup
+            for i in range(len(queries)):
+                values[i,:] = values[len(queries)+i, :] / values[i, :]
 
-        values = values[0:len(queries),:]
+            values = values[0:len(queries),:]
 
-        # computer average and add "avg" row to queries
-        avgs = np.mean(values, axis=0)
[…]

### 2021-02-13 yfcc100m/python/eval/DBEvalFramework/Plotting.py

@@ -222,7 +222,8 @@ class Plotting(object):
                   title="Generic (but Awesome) Bar Plot",
                   filename="plot_summary.pdf",
                   xlabel="Database Size",
-                  ylabel="Speedup ?"):
+                  ylabel="Speedup ?",
+                  log="none"):
 
         fig, ax0 = plt.subplots(nrows=1)
         fig.set_size_inches(12, 3)
@@ -261,9 +262,14 @@ class Plotting(object):
 
         limit = values[:,0:len(db_sizes)].max() * 1.5
 
-        ax0.set_ylim(0, limit)
+        # ax0.set_ylim(0.1, limit)
 
         plt.ylabel('Speedup', fontsize=12)
+
+        if log == "x" or log == "both":
+            ax0.set_xscale('log')
+        if log == "y" or log == "both":
+            ax0.set_yscale('log')
 
         plt.axhline(y=1)
 

### 2021-02-13 yfcc100m/paper/vldb2021/eval.tex
Update queries on eval.tex

@@
 we use the following queries, modeled after our internal use cases:
+
 \begin{itemize}
-\item \textit{q1} - {\bf {\em 1tag}}: Find metadata/images with one specific autotag (i.e. alligator, lake, etc).
-\item \textit{q2} - {\bf {\em 1tag\_resize}}: Find metadata/images with one specific autotag and resize to 224x224.
-\item \textit{q3} - {\bf {\em 1tag\_resize\_geo}}: Find metadata/images with one specific autotag, resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
-\item \textit{q4} - {\bf {\em 2tag\_resize\_geo\_and}}: Find metadata/images with two specific autotags (i.e.
-alligator AND lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
-\item \textit{q5} - {\bf {\em 2tag\_resize\_geo\_or}}: Find metadata/images with either of two specific autotags (i.e.
-alligator OR lake), resize to 224x224, and in a particular geo-location (with a 20 degrees radius in latitude and longitude).
+\item \textit{q1} - {\bf {\em 1tag\_resize}}: Retrieve images with one specific autotag and resize to 224x224.
+
+\item \textit{q2} - {\bf {\em 1tag\_geo\_resize}}: Retrieve images with one specific autotag, in a particular geo-location, and resize to 224x224,.
+
+\item \textit{q3} - {\bf {\em 2tag\_and\_resize}}: Retrieve images with two specific autotags (i.e. alligator AND lake), and resize to 224x224.
+
+\item \textit{q4} - {\bf {\em 2tag\_or\_resize}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), and resize to 224x224.
+
+\item \textit{q5} - {\bf {\em 2tag\_resize\_geo\_and}}: Retrieve images with two specific autotags (i.e. alligator AND lake), in a particular geo-location, and resize to 224x224.
+
+\item \textit{q6} - {\bf {\em 2tag\_resize\_geo\_or}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), in a particular geo-location, and resize to 224x224.
+
 \end{itemize}

### 2021-02-13 yfcc100m/paper/vldb2021/eval.tex
Add build plot and update exp setup

@@
 that provides similar functionality and interfaces as VDMS,
-we have implemented an equivalent visual data management system as a baseline,
-comprised of a combination of widely available, off-the-shelf components:
+we have implemented two equivalent visual data management system as baselines,
+comprised of a combination of widely available, off-the-shelf components.
+The first baseline uses
 MySQL Server 5.7 (for storing metadata),
@@
 OpenCV 3.3 (to provide pre-processing operations on images).
-We decided to use a relational database instead of a non-relational
-database because~\cite{Jatana2012, li_2019}:
+The other baseline system replaces the MySQL Server 5.7 with the
+most advanced open source relational database, PostgreSQL 9.5 \cite{postgresql}.
+We decided to use relational databases in the baseline systems instead of non-relational
+databases because~\cite{Jatana2012, li_2019}:
 \begin{itemize}
@@
 \end{itemize}
-The baseline implementation only partially replicates the functionalities
+The baseline implementations only partially replicate the functionalities
 that VDMS offers when it comes to image and metadata handling, built for the
@@
 pre-processing operations when needed.
-For visual data workloads, building an approach like the one
-implemented as a baseline for this work
+For visual data workloads, building a system like the baselines for this work
 is the common practice in the industry \cite{haystack, tao}.
 
-For all our experiments, we use two servers, one hosting a VDMS server and
-another hosting the baseline implementation.
+For all our experiments, we use two servers with Ubuntu 16.04, one hosting a VDMS server and
+another hosting the baseline implementations.
 Both servers have a dual-socket Intel\textsuperscript{\textregistered}
@@
 for a total of 112 logical cores per server.
-The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
+The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM, while the server hosting VDMS
 has 64GB of DDR4 DRAM.
 We decided to run VDMS server in the machine with less DRAM to make
-sure MySQL had no disadvantage, and because previous evaluation
+sure MySQL and PostgreSQL had no disadvantage, and because previous evaluation
 indicated smaller footprint in the case of VDMS when
 compared to similar baselines based on MySQL.
-Other than the difference in DRAM space, machines are identical.
-Both servers run Ubuntu 16.04.
+To build both MySQL and PostgreSQL on the same machine,
+we stored each relational database on separate SSD drives.
+Other than the difference in DRAM space and storage needed for two baselines, the machines are identical.
 The client application running the queries and measuring round-trip time
@@
 a 10GB back-plane switch, same as both servers.
+The client application was implemented using Python 3 for both VDMS and the baselines.
 Figure~\ref{fig:systems} shows a logical view of the difference between the
 interaction of the client application (retrieves metadata and
-images) with VDMS (left) and the baseline (right).
-The client application was implemented using Python 3 for both VDMS and
-the baseline.
+images) with VDMS (left) and the MySQL baseline (right).
 
@@
 in the appendix of this work.
-In the case of the baseline, metadata is
-stored in MySQL using an attached SSD disk.
+In the case of the baselines, metadata is
+stored in MySQL and PostgreSQL using an attached SSD disk.
 Even if VDMS has native support for Optane Persistent Memory,
 we do not use it in this experiment because of fairness of
-comparison with respect to MySQL, which was not designed for
+comparison with respect to MySQL and PostgreSQL, which were not designed for
 Persistent Memory type of storage.
@@
 
-For the metadata, we built VDMS and MySQL databases
+For the metadata, we built VDMS, MySQL, and PostgreSQL databases
 using the YFCC100M dataset with incremental database sizes.
@@
 it contains, as follows: 1M, 5M, 10M, 50M, 100M.
-The VDMS and MySQL databases have comparable number of elements.
+The databases have comparable number of elements.
 The exact number of images/elements in each database are shown in
@@
 In our set up, that difference is very small:
-less than 0.1\% in terms of number of elements (images and/or metadata information).
+less than 0.253\% in terms of number of elements (images and/or metadata information).
 
@@
 
-\textbf{MySQL/Baseline:}
+\textbf{MySQL/Baseline 1:}
 Each MySQL database is created in a similar manner as VDMS
@@
 
+\textbf{PostgreSQL/Baseline 2:}
+Each PostgreSQL database is created in the exact same manner as MySQL
+where data is represented as three tables:
+\textit{images}, \textit{taglist}, and \textit{taglist}.
+PostgreSQL works well out-of-box and optimizations were not necessary,
+as in mySQL, to load or execute queries.
+When executing queries for larger databases,
+it is pertinent to have enough disk space for both
+the database and tablespace.
+The tablespace is associated with a database and stores the temporary files
+created within the database object ~\cite{postgresql}.
+When processing the larger databases, additional
+storage may be needed for the tablespace or a diskfull error may occur. To avoid this,
+we allocated a separate SSD disk specifically for the tablespace of all databases.
+
+By default, PostgreSQL can manage concurrent access to data well using Multiversion
+Concurrency Control (MVCC) and a Serializable Snapshot Isolation (SSI) level of transcation isolation.
+MVCC provides better performance by using SSI to guarantee querying data never
+blocks writing data, and vice versa ~\cite{postgresql}.
+The PostgreSQL configuration file initially limits the maximum
+number of concurrent connections to the database server to 100.
+We increased this limit to 200 to complete the concurency analysis in \label{concurrency analysis}.
+
 \begin{table}[ht]
-\caption{MySQL Database - Number of Rows in each Table}
+\caption{MySQL and PostgreSQL Databases - Number of Rows in each Table}
 \centering
@@
 \centering
-\includegraphics[width=\columnwidth]{figures/db_time_size}
-\caption{Time to build and size (in GB) of MySQL and VDMS databases.}
+\includegraphics[width=\columnwidth]{figures/all_build_times_plot}
+\caption{Time to build and size (in GB) of MySQL, PostgreSQL, and VDMS databases.}
 \label{fig:db_time_size}
@@
 
-Another important aspect to note is that
-VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
-This is space used to store information about each node/connection.
+Another important aspect to note is the storage footprint of the databases.
+PostgreSQL and VDMS require the same amount of storage for metadata which
+is more than required by mySQL, shown in Figure~\ref{fig:db_time_size}. In the case of VDMS,
+this is space used to store information about each node/connection.
 The Graph Database internal to VDMS (called PMGD) was designed for performance,
@@
 This may become a factor if storage is a limitation, but it should also be noted
-that even if we have a 41\% increase in metadata size,
-metadata accounts for less than 2\% of the overall database size.
+that metadata accounts for less than 2\% of the overall database size
+even if we have a 41\% increase in metadata size.
 For example, the largest database (both metadata and images) we built
@@
 
-In order to evaluate VDMS and the baseline on our use-case queries,
+In order to evaluate VDMS and the baselines on our use-case queries,
 we implemented 5 queries that filter and retrieve a specific set of images.
@@
 of the \textit{connection} between the image and that \textit{autotag}),
-as well as in MySQL (in the form of a column in the \textit{autotags} table
+as well as in MySQL and PostgreSQL (in the form of a column in the \textit{autotags} table
 that links images with tags).
@@
 following \textit{connections} between the image and that \textit{autotag}).
-In the case of the baseline implementation,
+In the case of the baseline implementations,
 the query involves JOIN operations between the 3 tables.
@@
 In other words, as the size of the database increase, we increase the probability
-threshold for the queries. We do this for both VDMS and the baseline, of course.
+threshold for the queries. We do this for both VDMS and the baselines, of course.
 This way, we remove bottleneck introduced by network bandwidth that would
@@
 plus operations applied to images when applicable (queries 2, 3, 4 and 5).
-The experiments show the performance of both systems (VDMS and baseline) as the
+The experiments show the performance of all systems (VDMS and two baselines) as the
 database size increases.
@@
 plus operations applied to images when applicable (queries 2, 3, 4 and 5).
-The experiments show the performance of both systems (VDMS and baseline) as the
+The experiments show the performance of all systems (VDMS and the two baselines) as the
 database size increases.
@@
 
-\subsubsection{Concurrency Analysis}
+\subsubsection{Concurrency Analysis} \label{concurrency analysis}
 

### 2021-02-16 yfcc100m/paper/vldb2021/eval.tex
Update intro of eval section, and comparison systems figure

@@
 In order to evaluate the different aspects of the performance on the
-image search, we have built a baseline following the methodology
+image search, we have built a different baselines following the methodology
 used in the industry, and following what we have done in the past
@@
 which we describe in Section \ref{images}.
-We have also included performance evaluation on the Video and Feature Vectors
-functionality in the appendix for further reference, together with a
-description of the key aspects of those functionalities.
+% We have also included performance evaluation on the Video and Feature Vectors
+% functionality in the appendix for further reference, together with a
+% description of the key aspects of those functionalities.
 
@@
 \includegraphics[width=\textwidth]{figures/comparison_system}
-\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline system (right).}
+\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline systems (right).
+The baseline systems are either based on MySQL or PostgreSQL for operating and querying metadata.}
 \label{fig:systems}
@@
 collection of 100 million public Flickr media objects created to provide free,
-shareable multimedia data for research. This dataset contains approximately
-99.2 million images and 0.8 million videos with metadata characterized by
-25 fields such as the unique identifier, userid,
+shareable multimedia data for research.
+This dataset contains approximately 99.2 million images and 0.8 million videos
+with metadata characterized by 25 fields such as the unique identifier, userid,
 date the media was taken/uploaded, location in longitude/latitude coordinates,
@@
 and animals from 1,570 trained machine learning classifiers ~\cite{Thomee_2016}.
-Together with each \textit{autotags}, there is a
-probability associated with each tag to indicate certainty of the classification.
+Together with each \textit{autotags}, there is a probability associated with
+each tag to indicate certainty of the classification.
 This is, an image can have the \textit{autotags} "people", "person", "party",
@@
 
-Given that there are no other open-source systems
-that provides similar functionality and interfaces as VDMS,
-we have implemented two equivalent visual data management system as baselines,
-comprised of a combination of widely available, off-the-shelf components.
-The first baseline uses
-MySQL Server 5.7 (for storing metadata),
+Given that there are no other open-source systems that provides similar
+functionality and interfaces as VDMS, we have implemented two equivalent
+visual data management system as baselines, comprised of a combination of
+widely available, off-the-shelf components.
+The first baseline uses MySQL Server 5.7 (for storing metadata),
 Apache Web Server 2.4.18 (as interface for image access), and
@@
 most advanced open source relational database, PostgreSQL 9.5 \cite{postgresql}.
-We decided to use relational databases in the baseline systems instead of non-relational
-databases because~\cite{Jatana2012, li_2019}:
+We decided to use relational databases in the baseline systems instead of
+non-relational databases because~\cite{Jatana2012, li_2019}:
+
 \begin{itemize}
@@
 \end{itemize}
+
 The baseline implementations only partially replicate the functionalities
-that VDMS offers when it comes to image and metadata handling, built for the
-purpose of an ad-hoc image search implementation.
-This implementation is based upon internal tools used for
-ML-based pipelines for media.
+that VDMS offers when it comes to image and metadata handling, and it was
+built for the purpose of an ad-hoc image search implementation.
+The baseline implementations are based upon internal tools used for ML-based
+pipelines for media, is the common practice in the industry \cite{haystack, tao}.
 We have implemented a set of client-side applications that take care
@@
 pre-processing operations when needed.
-For visual data workloads, building a system like the baselines for this work
-is the common practice in the industry \cite{haystack, tao}.
 
-For all our experiments, we use two servers with Ubuntu 16.04, one hosting a VDMS server and
-another hosting the baseline implementations.
+For all our experiments, we use two servers with Ubuntu 20.04,
+one hosting a VDMS server and another hosting the baseline implementations.
 Both servers have a dual-socket Intel\textsuperscript{\textregistered}
@@
 for a total of 112 logical cores per server.
-The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM, while the server hosting VDMS
-has 64GB of DDR4 DRAM.
+The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM,
+while the server hosting VDMS has 194GB of DDR4 DRAM.
 We decided to run VDMS server in the machine with less DRAM to make
@@
 we stored each relational database on separate SSD drives.
-Other than the difference in DRAM space and storage needed for two baselines, the machines are identical.
+Other than the difference in DRAM space and storage needed for two baselines,
+the machines are identical.
 The client application running the queries and measuring round-trip time
@@
 interaction of the client application (retrieves metadata and
-images) with VDMS (left) and the MySQL baseline (right).
+images) with VDMS (left) and the MySQL/PostgreSQL baselines (right).
 
@@
 In the case of Videos, only the first frame is used for the image search.
-More information and evaluation on video-specific functionalities can be found
-in the appendix of this work.
+% More information and evaluation on video-specific functionalities can be found
+% in the appendix of this work.
 In the case of the baselines, metadata is
@@
 Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
-The differences can be attributed to failures in data
+The differences can be attributed to minor failures in data
 preparation/loading because of incomplete/inconsistent formatting,
 which is common in large datasets \cite{failures}.
-In our set up, that difference is very small:
-less than 0.253\% in terms of number of elements (images and/or metadata information).
+In our set up, that difference is very small: less than 0.253\% in terms of
+number of elements (images and/or metadata information).
 
@@
 When processing the larger databases, additional
-storage may be needed for the tablespace or a diskfull error may occur. To avoid this,
-we allocated a separate SSD disk specifically for the tablespace of all databases.
+storage may be needed for the tablespace or a diskfull error may occur.
+To avoid this, we allocated a separate SSD disk specifically
+for the tablespace of all databases.
 
-By default, PostgreSQL can manage concurrent access to data well using Multiversion
-Concurrency Control (MVCC) and a Serializable Snapshot Isolation (SSI) level of transcation isolation.
+By default, PostgreSQL can manage concurrent access to data well using
+Multiversion Concurrency Control (MVCC) and a Serializable Snapshot Isolation
+(SSI) level of transaction isolation.
 MVCC provides better performance by using SSI to guarantee querying data never
@@
 number of concurrent connections to the database server to 100.
-We increased this limit to 200 to complete the concurency analysis in \label{concurrency analysis}.
+We increased this limit to 200 to complete the concurrency analysis
+in \label{concurrency analysis}.
 
@@
 build each database, where VDMS outperforms MySQL by a large margin.
-This analysis includes only the metadata, as the images are stored in a shared filesystem.
-Figure~\ref{fig:db_time_size} illustrates how VDMS can build databases
-faster than MySQL, and how the speedup is sustained as the database size grows.
+This analysis includes only the metadata, as the images are stored
+in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates how
+VDMS can build databases faster than MySQL, and how the speedup
+is sustained as the database size grows.
 Key difference in the build times are attributed to the low-level
@@
 PostgreSQL and VDMS require the same amount of storage for metadata which
-is more than required by mySQL, shown in Figure~\ref{fig:db_time_size}. In the case of VDMS,
-this is space used to store information about each node/connection.
+is more than required by mySQL, shown in Figure~\ref{fig:db_time_size}.
+In the case of VDMS, this is space used to store information about
+each node/connection.
 The Graph Database internal to VDMS (called PMGD) was designed for performance,

### 2021-02-16 yfcc100m/paper/vldb2021/eval.tex
Update eval figures

@@
 In order to evaluate VDMS and the baselines on our use-case queries,
-we implemented 5 queries that filter and retrieve a specific set of images.
+we implemented 6 queries that filter and retrieve a specific set of images.
 We chose these queries because they represent typical use-cases where a
 cohort of images is to be retrieved and processed from a large corpus of data.
-As we mentioned before, we took this approach due to the lack of standard
+As we mentioned before, we took this approach mainly because we wanted to
+replicate systems we have built internally for our use cases,
+and also because of the lack of standard
 benchmarks that are oriented towards visual data retrieval.
+
 We use the metadata associated with the images to filter said images.
-
 We use the \textit{autotags} (as they contain information about the content
@@
 popular variations of neural networks for object detection on images~\cite{resnet}.
+We used both ResNet and Yolov3 for object detection, both of which
+have the requirement of a downsized, lower resolution image as input.
 
@@
 of the \textit{connection} between the image and that \textit{autotag}),
-as well as in MySQL and PostgreSQL (in the form of a column in the \textit{autotags} table
-that links images with tags).
+as well as in MySQL and PostgreSQL (in the form of a column in the
+\textit{autotags} table that links images with tags).
 In the case of VDMS, the query involves a graph traversal query that starts
@@
 under the benchmarks branch of the VDMS project
-\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
-and some examples of the queries can be found in the appendix of this
-paper.
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m},
+and some examples of the queries can be found in the appendix of this paper.
 
 \begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_conc_1tag_loc20_resize_mosaic_results_throughput_db_size}
+\includegraphics[width=\textwidth]{figures/plot_conc_q_1tag_loc20_resize_mosaic_results_throughput_db_size}
 \caption{Concurrency Analysis on \textit{q3} (\textit{1tag\_resize\_geo}).
@@
 \begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_conc_10M_mosaic_results_throughput}
+\includegraphics[width=\textwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
 \caption{Concurrency Analysis for all the queries on the 10M database).
@@
 \begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_56_mosaic_query_times}
+\includegraphics[width=\textwidth]{figures/plot_th_56_mosaic_query_times}
 \caption{Performance Analysis using all queries from our use-case
@@
 \begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_56_mosaic_results_throughput}
+\includegraphics[width=\textwidth]{figures/plot_th_56_mosaic_results_throughput}
 \caption{Throughput Analysis using all queries from our use-case
@@
 \centering
-\includegraphics[width=\textwidth]{figures/plot_56_query_times_speedup}
+\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_mysql}
 \caption{Summary of performance gains for all queries.
@@
 More importantly, we see that the speedup grow as the database size increases,
-showing that VDMS scales better than the baseline.}
-\label{fig:summary_images}
+showing that VDMS scales better than the MySQL baseline.}
+\label{fig:summary_mysql}
+\end{figure*}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_postgresql}
+\caption{Summary of performance gains for all queries.
+We see up to 35x speedup (\textit{q2}), and an average of about 15x.
+More importantly, we see that the speedup grow as the database size increases,
+showing that VDMS scales better than the PostGreSQL baseline.}
+\label{fig:summary_postgresql}
 \end{figure*}
@@
 
-Finally, Figure~\ref{fig:summary_images} summarizes the results.
+Finally, Figure~\ref{fig:summary_mysql} and
+Figure~\ref{fig:summary_postgresql}summarizes the results.
 We see up to 35x speedup (for the case of \textit{q2}),

### 2021-02-16 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Minor updates on plots

@@ -217,6 +217,7 @@ class EvalFramework(object):
         filename += str(q) + "_mosaic_results_throughput_threads.pdf"
 
         title = "Throughput as " + result_type + " per second for " + q
+        title += " as Database Size increases"
         threads = ["clients: " + str(a) for a in threads]
         p.plot_lines_all_mosaic(threads, db_sizes, engines, values,
                           filename=filename,
@@ -277,6 +278,7 @@ class EvalFramework(object):
         filename += str(q) + "_mosaic_results_throughput_db_size.pdf"
 
         title = "Throughput as " + result_type + " per second for " + q
+        title += " as number of concurrent clients increases"
         p.plot_lines_all_mosaic(db_sizes, threads, engines, values,
                           filename=filename,
                           title=title,
@@ -335,6 +337,7 @@ class EvalFramework(object):
         filename += str(db_size) + "_mosaic_results_throughput.pdf"
 
         title = "Throughput as " + result_type + " per second for " + db_size
+        title += " as number of concurrent clients increases"
         p.plot_lines_all_mosaic(queries, threads, engines, values,
                           filename=filename,
                           title=title,

### 2021-02-21 yfcc100m/paper/vldb2021/appendix-eval-plots.tex
Update to eval and add appendix

%=========================================
\subsection{Additional Plots}
This section provides additional plots illustrating the performance of the experiments.

Figure~\ref{fig:q_throughput_16} shows the measured throughput,
as images per second, delivered by each system for all queries and databases.
In comparison to figure~\ref{fig:q_throughput_56}, when using fewer concurrent 
clients (nthr = 16), the throughput for VDMS for queries \textit{q4} (\textit{2tag\_or\_resize}), 
\textit{q5} (\textit{2tag\_resize\_geo\_and}), and
\textit{q6} (\textit{2tag\_resize\_geo\_or}) improved by a large margin 
for the 50M and 100M databases. 
In the case of \textit{q4} and \textit{q6}, the throughput for 50M (and 100M for \textit{q4})
increase by a large margin which outperforms both baselines while in figure~\ref{fig:q_throughput_56}
PostgreSQL slightly performed better than VDMS. 
In the case of \textit{q5}, all systems perform a little differently for databases
larger than 5M.  The gaps between the throughput of each system narrows at 10M.  
At a database size of 50M, the performance of the baselines drastically drop
while the throughput of VDMS has a large improvement in comparison.

\begin{figure}[ht]
\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
\caption{Throughput Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
The experiments show the performance of all systems (VDMS and the two baselines) as the
database size increases.
These queries were run using 16 simultaneous clients (nthr = 16),
and averaged out of 5 runs.}
\label{fig:q_throughput_16}
\end{figure}

Figure~\ref{fig:summary_mysql} summarizes the results
comparing VDMS and MySQL. We see up to 96x speedup (for the case of \textit{q2}),
and an average improvement in throughput of about 31x.
We also see how \textit{q3} and \textit{q5} have poor performances and scalability
as the database size grows same as shown in Figure~\ref{fig:summary_postgresql}.
Considering all other queries (\textit{q1}, \textit{q2}, \textit{q4}, and \textit{q6}), 
VDMS provides at least 3x speedup in throughput.

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_mysql}
\caption{Summary of performance gains for all queries.
We see up to 96x speedup (\textit{q2}), and an average of about 31x.
More importantly, we see that the speedup grow as the database size increases,
showing that VDMS scales better than the MySQL baseline.}
\label{fig:summary_mysql}
\end{figure*}
%=========================================

### 2021-02-21 yfcc100m/paper/vldb2021/eval.tex
Update to eval and add appendix

@@
 In order to evaluate the different aspects of the performance on the
-image search, we have built a different baselines following the methodology
+image search, we have built different baselines following the methodology
 used in the industry, and following what we have done in the past
@@
 
-Given that there are no other open-source systems that provides similar
+Given that there are no other open-source systems that provide similar
 functionality and interfaces as VDMS, we have implemented two equivalent
@@
 The baseline implementations are based upon internal tools used for ML-based
-pipelines for media, is the common practice in the industry \cite{haystack, tao}.
+pipelines for media, the common practice in the industry \cite{haystack, tao}.
 We have implemented a set of client-side applications that take care
@@
 while the server hosting VDMS has 194GB of DDR4 DRAM.
-We decided to run VDMS server in the machine with less DRAM to make
+We decided to run the VDMS server in the machine with less DRAM to make
 sure MySQL and PostgreSQL had no disadvantage, and because previous evaluation
@@
 Apache WebServer and VDMS have direct access.
-In the case of Videos, only the first frame is used for the image search.
+In the case of videos, only the first frame is used for the image search.
 % More information and evaluation on video-specific functionalities can be found
@@
 The exact number of images/elements in each database are shown in
-Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
+Table \ref{table:vdmsnodes} and \ref{table:mysqltables}.
 The differences can be attributed to minor failures in data
@@
 
-In the case of VDMS, we did not attempt to tune any
-parameter to avoid unfairness in the comparison against the baseline.
-Unlike the baseline, VDMS can handle the entire YFCC100m dataset using the default parameters provided by the implementation.
-For both VDMS and the baseline, we created indexes over the
-properties we used for search, such as name of
-\textit{autotag}, and geo-location values.
-Building indexes for the right properties and objects
-is basic operation that would be present in any real-world deployment,
-and measuring performance without them would lead to useless analysis in our
-real-world applications and use cases.
-
-\textbf{PostgreSQL/Baseline 2:}
+\textbf{PostgreSQL (Baseline 2):}
 Each PostgreSQL database is created in the exact same manner as MySQL
@@
 PostgreSQL works well out-of-box and optimizations were not necessary,
-as in mySQL, to load or execute queries.
+as in MySQL, to load or execute queries.
 When executing queries for larger databases,
@@
 We increased this limit to 200 to complete the concurrency analysis
-in \label{concurrency analysis}.
+in Section \ref{concurrency_analysis}.
+
+In the case of VDMS, we did not attempt to tune any
+parameter to avoid unfairness in the comparison against the baselines.
+Unlike the baselines, VDMS can handle the entire YFCC100m dataset using the default parameters provided by the implementation.
+For all systems, we created indexes over the
+properties we used for search, such as name of
+\textit{autotag}, and geo-location values.
+Building indexes for the right properties and objects
+is basic operation that would be present in any real-world deployment,
+and measuring performance without them would lead to useless analysis in our
+real-world applications and use cases.
 
@@
 One of the first things we noticed is the difference in the time needed to
-build each database, where VDMS outperforms MySQL by a large margin.
+build each database, where VDMS outperforms MySQL by a large margin
+but PostgreSQL outperforms VDMS by a large margin as well.
 This analysis includes only the metadata, as the images are stored
-in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates how
-VDMS can build databases faster than MySQL, and how the speedup
+in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates the
+comparison in time to build the databases, and how the speedup
 is sustained as the database size grows.
@@
 to handle large datasets such as YFCC100M, and the efficiency of PMGD.
-On average, it took MySQL around 3.72x longer to build each database than VDMS.
+PostgreSQL Python API can efficiently copy data from files to the database
+which provides an advantage over VDMS which must insert multiple nodes/connections per image.
+On average, it took MySQL around $\sim$4x longer to build each database than VDMS,
+and PostgreSQL was around $\sim$21x faster than VDMS.
 
@@
 Another important aspect to note is the storage footprint of the databases.
-PostgreSQL and VDMS require the same amount of storage for metadata which
-is more than required by mySQL, shown in Figure~\ref{fig:db_time_size}.
+PostgreSQL and VDMS require comparable amount of storage for metadata which
+is more than required by MySQL, shown in Figure~\ref{fig:db_time_size}.
 In the case of VDMS, this is space used to store information about
@@
 noticeable in our results.
-VDMS required 30-41\% more storage than MySQL for storing the same amount
-of metadata.
-This may become a factor if storage is a limitation, but it should also be noted
-that metadata accounts for less than 2\% of the overall database size
-even if we have a 41\% increase in metadata size.
+VDMS required 37-41\% more storage than MySQL for storing the same amount
+of metadata. On the other hand, PostgreSQL required 1-4\% more storage than VDMS.
+The storage footprint may become a factor if storage is a limitation, but
+it should also be noted that metadata accounts for less than 2\% of the
+overall database size even if we have a 41\% increase in metadata size.
 For example, the largest database (both metadata and images) we built
@@
 
-In order to evaluate VDMS and the baselines on our use-case queries,
+In order to evaluate VDMS and the baseline systems on our use-case queries,
 we implemented 6 queries that filter and retrieve a specific set of images.
@@
 
-We use the metadata associated with the images to filter said images.
+We use the metadata associated with the images to filter the images.
 We use the \textit{autotags} (as they contain information about the content
@@
 
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_conc_q_1tag_loc20_resize_mosaic_results_throughput_db_size}
-\caption{Concurrency Analysis on \textit{q3} (\textit{1tag\_resize\_geo}).
-Hardware concurrency (number of physical cores in each system)
-is shown with a blue vertical line (hw-concurrency = 56).
-Top figure shows aggregated throughput (transactions per second)
-when retrieving only metadata associated with the images, as the number of
-concurrent clients increase.
-Bottom figure shows aggregated throughput (images per second) when retrieving
-resized versions of the images, as the number of concurrent clients increases.}
-\label{fig:concurrency_comparison}
-\end{figure*}
-
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
-\caption{Concurrency Analysis for all the queries on the 10M database).
-Hardware concurrency (number of physical cores in each system)
-is shown with a blue vertical line (hw-concurrency = 56).
-Top figure shows aggregated throughput (transactions per second)
-when retrieving only metadata associated with the images, as the number of
-concurrent clients increase.
-Bottom figure shows aggregated throughput (images per second) when retrieving
-resized versions of the images, as the number of concurrent clients increases.}
-\label{fig:concurrency_comparison_10M}
-\end{figure*}
-
 Also, note that the size of the result (number of images retrieved)
@@
 otherwise over-complicate the understanding of the results.
-
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_th_56_mosaic_query_times}
-\caption{Performance Analysis using all queries from our use-case
-described in the Experimental Setup Section.
-We show queries in different figures for readability reasons.
-Top figures show the throughput of the queries for
-just retrieving metadata associated with the images.
-Bottom figures show the throughput when retrieving both metadata and images,
-plus operations applied to images when applicable (queries 2, 3, 4 and 5).
-The experiments show the performance of all systems (VDMS and two baselines) as the
-database size increases.
-These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged out of 10 runs.}
-\label{fig:q_performance_56}
-\end{figure*}
-
-\begin{figure*}[ht]
-\includegraphics[width=\textwidth]{figures/plot_th_56_mosaic_results_throughput}
-\caption{Throughput Analysis using all queries from our use-case
-described in the Experimental Setup Section.
-We show queries in different figures for readability reasons.
-Top figures show the throughput of the queries for
-just retrieving metadata associated with the images.
-Bottom figures show the throughput when retrieving both metadata and images,
-plus operations applied to images when applicable (queries 2, 3, 4 and 5).
-The experiments show the performance of all systems (VDMS and the two baselines) as the
-database size increases.
-These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged out of 10 runs.}
-\label{fig:q_throughput_56}
-\end{figure*}
 
@@
 
-\subsubsection{Concurrency Analysis} \label{concurrency analysis}
+\subsubsection{Concurrency Analysis} \label{concurrency_analysis}
 
-Figure~\ref{fig:concurrency_comparison} illustrates a concurrency analysis for
-\textit{q2}(\textit{1tag\_resize}), described above,
-using both VDMS and the baseline.
-Here we evaluate the scalability of both systems, as the number of concurrent
-clients grows (x-axis) and as the
-size of the databases grows (each full-line represents a database size
-for VDMS and each dotted-line represents a database size for the baseline).
+\begin{figure}[ht]
+% WAS: begin{figure*}[ht]; textwidth
+\includegraphics[width=\columnwidth]{figures/plot_conc_q_1tag_loc20_resize_mosaic_results_throughput_db_size}
+\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_geo\_resize}) from our use-case
+described in the Experimental Setup Section.
+We show each database in different figures for readability reasons.
+The figures show the throughput (images per second) when retrieving
+resized versions of the images, as the number of concurrent clients increases
+Hardware concurrency (number of physical cores in each system)
+is shown with a blue vertical line (hw-concurrency = 56).}
+\label{fig:concurrency_comparison_q2}
+\end{figure}
 
-We start by analyzing Figure~\ref{fig:concurrency_comparison} (top),
-which shows aggregated throughput (transactions per second)
-when retrieving only metadata associated with the images, as the number of
+Figure~\ref{fig:concurrency_comparison_q2} illustrates a concurrency analysis for
+\textit{q2} (\textit{1tag\_geo\_resize}), described above,
+using VDMS and the baselines.
+Here we evaluate the scalability of all systems, as the number of concurrent
+clients grows (x-axis) and as the size of the databases grow.
+Figure~\ref{fig:concurrency_comparison_q2} shows throughput (images per second)
+when retrieving resized versions of the images, as the number of
 concurrent clients increase.
-The first thing to notice is that at low concurrency (2 to 8 concurrent clients),
-both systems show similar performance, except in the 100M case of the baseline.
-Note that, for this particular experiment,
-baseline translates directly to MySQL performance,
-as the metadata-only queries only involve running a query to MySQL.
-For the baseline system, in the case of 100M, the increase in the
+The first thing to notice is that VDMS outperforms both baselines for all
+databases for this query by a large margin.
+For the baseline systems, in the case of 100M, the increase in the
 size of data seems to have a larger impact in performance.
@@
 the hardware concurrency of the system.
-Also, more parallelism after 56 threads does not increase the delivered throughput,
-and it is actually slightly detrimental (112 concurrent clients case).
-On the other hand, the baseline seems to deliver less aggregated
-throughput after 16 threads, with an increase for 64 threads, but these effects
-are hard to interpret fully as the standard deviation in the measurements is high.
-We noticed throughout many experiments that MySQL results showed higher
-standard deviation, meaning less consistent and noise measurements, when compared
+Also, more parallelism after 56 threads does not increase the delivered throughput
+for the larger databases (50M and 100M) but there is a slight improvement in the
+other databases (1M, 5M, and 10M).
+On the other hand, both baselines seem to deliver less aggregated
+throughput after 16 threads except for the 100M case.
+In this case, PostgreSQL has a performance spike at 32 clients
+and the performance for both baselines is less stable when compared
 to VDMS.
-We tried increasing the number of measurements and discarding
-outliers but were not able to get less noisy results.
+Note that most of the throughput for the baseline systems are very close to each other.
+This is mostly an effect of the log-scale used, which is needed to clearly
+depict the difference between VDMS and the baselines.
+Here, the baselines are the full architecture described in
+Figure~\ref{fig:systems} (right).
 
-We continue by looking at Figure~\ref{fig:concurrency_comparison} (bottom),
-which shows aggregated images per second delivered by each system.
-Again, we evaluate the scalability of both systems, as the number of
-concurrent clients grows (x-axis) and as the
-size of the databases grows (each full-line represents a database size
-for VDMS and each dotted-line represents a database size for the baseline).
-Note that most of the baseline dotted-lines are very close to each other.
-This is mostly an effect of the log-scale used, which is needed to clearly
-depict the difference between VDMS and the baseline.
-Here, the baseline is the full architecture described in
-Figure~\ref{fig:systems} (right).
-Figure~\ref{fig:concurrency_comparison} (bottom) shows a similar trend as the top
-figure when it comes to low concurrency. The baseline does as good and even better than VDMS with 2 or 4 concurrent clients.
-However, as concurrency increases beyond 4 concurrent clients, the difference
-in throughput becomes clear, with VDMS reaching its peak performance at
-56 concurrent clients.
-This query (as well as \textit{q3}, \textit{q4} and \textit{q5}) runs a resize operation on the image,
+All queries run a resize operation on the image,
 an operation that requires decoding, resizing, and encoding the image
@@
 
-There are many reasons why we see this performance improvement, the main being
+\begin{figure}[ht]
+% WAS: begin{figure*}[ht]; textwidth
+\includegraphics[width=\columnwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
+\caption{Concurrency Analysis for all the queries using the 10M database).
+We show queries in different figures for readability reasons.
+Each figure shows the throughput (images per second) when retrieving
+resized versions of the images as the number of concurrent clients increases.
+Hardware concurrency (number of physical cores in each system)
+is shown with a blue vertical line (hw-concurrency = 56).}
+\label{fig:concurrency_comparison_10M}
+\end{figure}
+
+We continue by looking at Figure~\ref{fig:concurrency_comparison_10M} 
+which shows the images per second delivered by each system 
+for all queries using the 10M database, as the number of concurrent 
+clients grows (x-axis). This figure takes a closer look at the 
+concurrency for each of the six queries. 
+It is obvious that VDMS consistently outperforms each of the baseline
+systems for queries \textit{q1}, \textit{q2}, \textit{q4}, 
+and \textit{q6} as concurrency grows. This figure 
+shows a similar trend as Figure~\ref{fig:concurrency_comparison_q2} 
+when it comes to concurrency.  The
+VDMS throughput continues to increase up to 56 concurrent clients, which is
+the hardware concurrency of the system, for all queries except \textit{q3}.
+In this query, the throughput seems to decrease after 32 clients 
+and both baselines outperform VDMS after 8 concurrent clients.
+Prior to 8 clients, VDMS has a slight advantage over PostgreSQL which has
+a major performance improvement at 16 concurrent clients. On the other hand,
+MySQL maintains its improvement in throughput over PostgreSQL and VDMS over all
+concurrent clients. In the case of \textit{q5}, the throughput of the baselines
+are less aggregated than those of VDMS up to 16 concurrent clients. When it comes 
+to low concurrency, the baselines do as good and even better than VDMS with 2 or 
+4 concurrent clients.  At 16 concurrent client, the performance of the MySQL baseline
+begin to stabilizes while the throughput of PostgreSQL begins to degrade. However, as 
+concurrency increases beyond 16 concurrent clients, the difference
+in throughput becomes clear, with VDMS reaching its peak performance at
+112 concurrent clients.
+
+In the case of VDMS, we see the performance in the case of \textit{q3}
+and \textit{q5} suffers significantly in comparison to the other queries.
+The reason for that lack of scalability lies on the query implementation: given that
+VDMS does not yet support operators that enable querying images
+that have both connections to a \textit{tagA} and a \textit{tagB};
+we have to implement this transaction by doing 2 retrievals.
+This involves retrieving partial information in the first retrieval,
+applying an INTERSECTION operation in the client, and doing a second retrieval
+to bring the right metadata and/or images.
+The reason for this is a lack of operations that would enable this query to be
+run entirely on the server is not an inherent limitation to VDMS but rather
+just a missing implementation.
+In the case of \textit{q4} and \textit{q6}, it is worth noting
+that for the OR operation, there is no need for 2 retrievals.
+Rather, a single retrieval is performed and the result filtered on the client.
+Future release will add more of such operators (AND, OR, etc) in order to prevent
+unnecessary retrievals and extra filtering on the client side.
+
+% WAS: begin{figure*}[ht]; textwidth
+\begin{figure}[ht]
+\includegraphics[width=\columnwidth]{figures/plot_q_2tag_resize_or_mosaic_results_throughput_threads}
+\caption{Concurrency Analysis on \textit{q4} (\textit{2tag\_or\_resize}) from our use-case
+described in the Experimental Setup Section.
+We show different number of concurrent clients in different figures for readability reasons.
+The experiments show the throughput as images per second for
+all systems (VDMS and two baselines) as the database size increases.}
+\label{fig:concurrency_comparison_q4}
+\end{figure}
+
+In the case of \textit{q4} and \textit{q6}, we see the performance for MySQL 
+degrades up to 64 concurrent clients, then there's an improvement with 112 concurrent clients.
+To further evaluate the concurrency of these queries with an OR operation, we continue by
+looking at Figure~\ref{fig:concurrency_comparison_q4}. This figure illustrates a concurrency analysis for
+\textit{q4} (\textit{2tag\_resize\_or}) which shows the images per second 
+delivered by each system as the database size grows (x-axis). We show different number 
+of concurrent clients in different figures for readability reasons.  For this query,
+VDMS delivers higher throughput for databases 1M through 10M for all concurrent clients.
+As the number of clients and the size of database grows, the performance for this query
+drops and the performance of PostgreSQL becomes more comparable to VDMS. Lets look at the
+case with 32 concurrent clients. In this case, the throughput of VDMS drops drastically with
+the 100M database and PostgreSQL outperforms by a small margin. As the clients increase, we 
+see this trade-off occurs with the 50M database instead of 100M while in the 100M database, 
+VDMS outperforms both baselines. In this case, the throughput of PostgreSQL drastically degrades
+for the 100M database and the performance of the baseline systems are more comparable.
+
+When considering Figure~\ref{fig:concurrency_comparison_q2} through Figure~\ref{fig:concurrency_comparison_q4},
+there are many reasons why we generally see performance improvement, the main being
 that the entire operation (metadata query, image fetching and resizing) happens
@@
 visual entities as first class citizens, as it is the case of VDMS.
-Another reason, which is quantifiable in the bottom figures, is that
+Another reason, which is quantifiable in the figures, is that
 VDMS sends resized (smaller) versions to the client instead of the full image
-to be resized on the client side (as is the case in the baseline).
-This is in contrast with the baseline, where 2 rounds of blocking back-and-forth
+to be resized on the client side (as is the case in the baselines).
+This is in contrast with the baseline implementations, where 2 rounds of blocking back-and-forth
 communication with the server is needed, as depicted in Figure~\ref{fig:systems}.
-Note that on the point 1), one could argue that the opposite will happen
-when the resize operation retrieves a up-sampled (larger) version of the image
+Note that one could argue that the opposite will happen
+when the resize operation retrieves a an up-sampled (larger) version of the image
 instead of a down-sampled (smaller) one.
@@
 
-The next step in our analysis involved running a different set of queries
+\begin{figure}[ht]
+% WAS: begin{figure*}[ht]; textwidth
+\includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput}
+\caption{Throughput Analysis using all queries from our use-case
+described in the Experimental Setup Section.
+We show queries in different figures for readability reasons.
+The experiments show the performance of all systems (VDMS and the two baselines) as the
+database size increases.
+These queries were run using 56 simultaneous clients (nthr = 56),
+and averaged out of 5 runs.}
+\label{fig:q_throughput_56}
+\end{figure}
+
+The next step in our analysis involved running a set of queries
 (described above in this section),
@@
 queries we analysed for our use case.
-Top figures show the throughput of these queries
-retrieving metadata associated with the images.
-Bottom figures show the throughput when retrieving both metadata and images,
-plus operations applied to images when applicable (queries 2, 3, 4,and 5).
-The experiments show the performance of both systems (VDMS and baseline) as the
+Each figure shows the throughput when retrieving images,
+plus operations applied to images when applicable.
+The experiments show the performance of all systems (VDMS and two baselines) as the
 database size increases in terms of number of images.
 These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged over 10 runs (niter = 10),
-each client running the retrieval of the tag 2 times (ntags = 2).
-The last parameter (ntags) is to avoid having some queries to finish the work
-too fast before other clients can even send the query to the server.
-This ensures that there is enough work to do in a query so that
-all queries execute in parallel on the server side.
+and averaged over 5 runs (niter = 5).
 To analyze these plots, one needs to compare the full-line (VDMS) versus the
-dotted-line (baseline), each color representing a different query.
-For example, to compare \textit{q1} performance, one needs to look at
-the full-red line (VDMS) and the dotted-red line (baseline).
+dotted-lines (two baselines), each plot representing a different query.
 
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_mysql}
-\caption{Summary of performance gains for all queries.
-We see up to 35x speedup (\textit{q2}), and an average of about 15x.
-More importantly, we see that the speedup grow as the database size increases,
-showing that VDMS scales better than the MySQL baseline.}
-\label{fig:summary_mysql}
-\end{figure*}
+For \textit{q1} and \textit{q2}, we can appreciate higher performance being delivered by
+VDMS when compared to the baselines, and how this improvement is maintained
+as the size of the database increase while
+on the other hand, the baseline implementations have a drop in performance
+for the 100M database. For \textit{q3}, we see that VDMS performs best when the database size
+is small (1M images), but as the database size increases, the performance degrades as well.  
+For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS for databases larger than 1M.
+This also occurs for \textit{q5} but only for databases larger than 10M.  
+This is entirely attributed to the 2-round process needed for this query,
+as we explained before which is more visible in the larger databases.
+It is interesting to note that adding filtering by geo-location
+(\textit{q5}) decreases the performance as the database sizes scales.
+The behaviour is different in the case of \textit{q6}, which also filters by geo-location.
+This is attributed to the fact that OR queries involves processing larger results,
+and thus does not benefit from the filtering which is evident when compared to \textit{q4}.
 
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_postgresql}
-\caption{Summary of performance gains for all queries.
-We see up to 35x speedup (\textit{q2}), and an average of about 15x.
-More importantly, we see that the speedup grow as the database size increases,
-showing that VDMS scales better than the PostGreSQL baseline.}
-\label{fig:summary_postgresql}
-\end{figure*}
-
-\subsubsection{Metadata Retrieval}
-
-First, we analyze the top figures, showing metadata transactions per second.
-We evaluate the performance when retrieving only metadata associated with the images,
-and not the images themselves.
-In this particular case, the baseline translates directly into MySQL performance.
-Note that for these figures (top), \textit{q1} and \textit{q2} are
-essentially the same query. This is because the metadata retrieved for both
-queries does not change. The only difference between \textit{q1} and \textit{q2}
-is the presence of the resize operations that does not have any impact on
-analyzing performance of metadata retrieval.
-For \textit{q1} and \textit{q2}, we can appreciate higher performance being delivered by
-VDMS when compared to the baseline, and how this improvement is maintained
-as the size of the database increase.
-For \textit{q3}, we see that MySQL performs best when the database size
-is small (1M images), with VDMS outperforming MySQL as the database increases in size.
-For \textit{q4}, we see MySQL also outperforming VDMS on small size databases, and as
-the database size increase, the gap between the two narrows,
-similar to what we see for \textit{q5}.
-It is interesting to note that adding filtering by geo-location
-(\textit{q3}, and \textit{q4}) slightly increases the performance of MySQL
-for small databases, and decreases it as the database sizes scales.
-The behaviour is different in the case of \textit{q5},
-which also filters by geo-location.
-This is attributed to the fact that OR queries involves processing larger results,
-and thus does not benefit from the filtering of the AND operation as \textit{q4}.
-In the case of VDMS, we see \textit{q3} performance is comparable to \textit{q1}
-and \textit{q2}, but \textit{q4} suffers significantly
-when the scale of the database increases.
-The reason for that lack of scalability lies on the query implementation: given that
-VDMS does not yet support operators that enable querying images
-that have both connections to a \textit{tagA} and a \textit{tagB};
-we have to implement this transaction by doing 2 retrievals.
-This involves retrieving partial information in the first retrieval,
-applying an INTERSECTION operation in the client, and doing a second retrieval
-to bring the right metadata and/or images.
-The reason for this is a lack of operations that would enable this query to be
-run entirely on the server is not an inherent limitation to VDMS but rather
-just a missing implementation.
-In the case of \textit{q5}, we see a similar effect, although it is worth noting
-that for the OR operation there is no need for 2 retrievals.
-Rather, a single retrieval is performed and the result filtered on the client.
-Future release will add more of such operators (AND, OR, etc) in order to prevent
-unnecessary retrievals and extra filtering on the client side.
-
-\subsubsection{Image Retrieval}
-
-We continue by analyzing the bottom figures, showing measured throughput,
-as images per second, delivered by each system.
-For the case of VDMS, \textit{q1}(full-red-line) shows less throughput
-than \textit{q2} (full-blue-line) (bottom left figure).
-This is expected as \textit{q1} returns a full size version of the image,
-whereas \textit{q2} returns a resized (smaller) version of the image,
-thus transferring less data over the network.
-In the case of the baseline, both \textit{q1} and \textit{q2} transfer
-the full size version of the image, and as part of \textit{q2},
-the resize is performed in the client.
-This is why, contrary to the VDMS case, \textit{q1} performs better (even if slightly)
-when compared with \textit{q2}.
-We can also see that, for VDMS, \textit{q3} perform worst than \textit{q2} because
-of the extra step needed for filtering based on geo-location.
-Moreover, we see a great performance degradation in the case of \textit{q4} as the
-database size increases.
-This is entirely attributed to the 2-round process needed for this query,
-as we explained before.
-From the first 3 queries, as well as \textit{q5},
-we clearly see that VDMS outperforms the baseline
+From the first 2 queries, as well as \textit{q4} and
+\textit{q6} (with exception to the 50M case),
+we clearly see that VDMS outperforms the baseline systems
 when retrieving visual data and applying operations.
@@
 
-Finally, Figure~\ref{fig:summary_mysql} and
-Figure~\ref{fig:summary_postgresql}summarizes the results.
-We see up to 35x speedup (for the case of \textit{q2}),
-and an average improvement in throughput of about 15x.
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_postgresql}
+\caption{Summary of performance gains for all queries.
+We see up to 364x speedup (\textit{q1}), and an average of about 85x.
+More importantly, we see that the speedup grow as the database size increases,
+showing that VDMS scales better than the PostGreSQL baseline.}
+\label{fig:summary_postgresql}
+\end{figure*}
+
+Finally, Figure~\ref{fig:summary_postgresql} summarizes the results
+comparing VDMS and PostgreSQL. A comparison with MySQL is
+provided in Figure~\ref{fig:summary_mysql} of the appendix.
+In comparison to PostgreSQL, VDMS provides up to 364x speedup 
+(for the case of \textit{q1}),
+and an average improvement in throughput of about 85x. This is
+an impressive improvement over the two baseline systems.
 More importantly, we see that the speedup increases as the database size grows,
-showing that VDMS scales better than the baseline.
-We also see how \textit{q4} shows poor performances and scalability
-when compared to the baseline, and this evaluation served the
+showing that VDMS scales better than the baseline systems.
+We also see how \textit{q3} and \textit{q5} have poor performances and scalability
+when compared to the baselines as the database size grows.
+This evaluation served the
 purpose of understanding the importance of VDMS server side operators
@@
 %% ========= Summary Analysis - END
-
-%% ========= Video Eval
-
-% \include{eval-video}
-
-%% ========= Video Eval - END
-
-%% ========= Descriptors Eval
-
-% \include{eval-descriptors}
-
-%% ========= Descriptors Eval - END

### 2021-02-22 yfcc100m/paper/vldb2021/design.tex
Update intro and design, minor changes in eval

@@
 
-In this section, we describe VDMS design principles and implementation,
+In this section, we further describe VDMS design principles and implementation,
 which was briefly introduced in previous work \cite{vdms-nips}.
-Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
+
 VDMS implements a client-server architecture that handles client
@@
 in a transactional manner.
+Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
 
@@
 The client library implements TCP/IP based connectors to the VDMS Server,
-similar to most database\cite{memsql, mysql}.
+similar to most database\cite{memsql, mysql, postgresql}.
 Users can connect to VDMS and implement queries using VDMS' API
@@
 accepts a JSON string and an array or vector of blobs.
-Internally, the library wraps the query string and blob using
+Internally, the library wraps the query string and blobs using
 Google Protobufs \cite{protobufs} and sends it to the VDMS server.

### 2021-02-23 yfcc100m/paper/vldb2021/intro.tex
Update intro and design, minor changes in eval
Update intro/eval design, add more references

@@
 identifiers for the scans and adding their metadata in
-a relational or key-value database. 
+a relational or key-value database\cite{kumar2017data}.
 If the queries require a search over patient information,
 then patients are associated with their brain scans.
-Finally, if the ML pipeline needs images with a different resolution than the original, there is additional compute
+Finally, if the ML pipeline needs images with a different resolution
+than the original, there is additional compute
 diverted towards pre-processing the original images which are typically larger.
@@
 together with a script for this specific use case.
-Moreover, if the pipeline identifies
-new metadata to be added for the tumor images, most databases make it
-hard to change the schema on the fly.
+Moreover, if the pipeline identifies new metadata to be added for
+the tumor images, most databases make it hard to change the schema on the fly.
 As another example, many applications can be studied through the use of large
@@
 of human-generated tags), advanced image search through the use of
-machine-generated tags and feature vectors\cite{imagesearch} 
+machine-generated tags and feature vectors\cite{imagesearch, qin2020similarity}
 for each image, and video summarization.
@@
 combination of databases and file systems to store
-metadata and visual data (images, videos), respectively. 
+metadata and visual data (images, videos), respectively \cite{sculley2015hidden}.
 This is integrated with a set of custom scripts that tie multiple systems together,
 unique not only to a specific application/discipline but often
-to individual researchers or development teams.
+to individual researchers or development teams
+\cite{mayer2020scalable, sculley2015hidden}.
 These ad-hoc solutions make replicating experiments difficult,
@@
 that heavily rely on visual data.
-VDMS is a completely Open Source project designed to enable 
+VDMS is an Open Source project designed to enable
 efficient access of visual data.
@@
 While there are a number of big-data frameworks~\cite{spark, hadoop}, systems
-that can be used to store metadata~\cite{memsql, vertica}, and systems that
-manipulate a specific category of visual data~\cite{scidb, rasdaman}, VDMS can
-be distinguished from them on the following aspects:
+that can be used to store metadata~\cite{memsql, vertica, mysql, postgresql},
+and systems that manipulate a specific category of
+visual data~\cite{scidb, rasdaman},
+VDMS can be distinguished from them on the following aspects:
 
@@
 all of which were shared under one of the various Creative Commons licenses.
-We have used this dataset
-for multiple proof of concepts and applications within our research lab.
+We have used this dataset for multiple proof of concepts and
+applications within our research lab.
 

### 2021-02-24 yfcc100m/paper/vldb2021/eval.tex
Update intro and design, minor changes in eval
Update intro/eval design, add more references
changed plot to only build sizes

@@
 parameter to avoid unfairness in the comparison against the baselines.
-Unlike the baselines, VDMS can handle the entire YFCC100m dataset using the default parameters provided by the implementation.
+Unlike the baselines, VDMS can handle the entire YFCC100m dataset
+using the default parameters provided by the implementation.
 For all systems, we created indexes over the
@@
 
+% \begin{figure}[ht]
+% \centering
+% \includegraphics[width=\columnwidth]{figures/all_build_times_plot}
+% \caption{Time to build and size (in GB) of MySQL, PostgreSQL, and VDMS databases.}
+% \label{fig:db_time_size}
+% \end{figure}
+
 \begin{figure}[ht]
 \centering
-\includegraphics[width=\columnwidth]{figures/all_build_times_plot}
-\caption{Time to build and size (in GB) of MySQL, PostgreSQL, and VDMS databases.}
-\label{fig:db_time_size}
+\includegraphics[width=\columnwidth]{figures/all_build_sizes_plot}
+\caption{Data footprint (in GB) of MySQL, PostgreSQL, and VDMS databases.}
+\label{fig:db_sizes}
 \end{figure}
 
-\subsubsection{Database Building Time}
+% \subsubsection{Database Building Time}
 
-One of the first things we noticed is the difference in the time needed to
-build each database, where VDMS outperforms MySQL by a large margin
-but PostgreSQL outperforms VDMS by a large margin as well.
-This analysis includes only the metadata, as the images are stored
-in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates the
-comparison in time to build the databases, and how the speedup
-is sustained as the database size grows.
-Key difference in the build times are attributed to the low-level
-implementation of how MySQL reads and stores data from the files, the
-optimizations (increased InnoDB pool size, etc.) needed
-to handle large datasets such as YFCC100M, and the efficiency of PMGD.
-PostgreSQL Python API can efficiently copy data from files to the database
-which provides an advantage over VDMS which must insert multiple nodes/connections per image.
-On average, it took MySQL around $\sim$4x longer to build each database than VDMS,
-and PostgreSQL was around $\sim$21x faster than VDMS.
+% One of the first things we noticed is the difference in the time needed to
+% build each database, where VDMS outperforms MySQL by a large margin
+% but PostgreSQL outperforms VDMS by a large margin as well.
+% This analysis includes only the metadata, as the images are stored
+% in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates the
+% comparison in time to build the databases, and how the speedup
+% is sustained as the database size grows.
+% Key difference in the build times are attributed to the low-level
+% implementation of how MySQL reads and stores data from the files, the
+% optimizations (increased InnoDB pool size, etc.) needed
+% to handle large datasets such as YFCC100M, and the efficiency of PMGD.
+% PostgreSQL Python API can efficiently copy data from files to the database
+% which provides an advantage over VDMS which must insert
+% multiple nodes/connections per image.
+% On average, it took MySQL around $\sim$4x longer to build each database than VDMS,
+% and PostgreSQL was around $\sim$21x faster than VDMS.
 
@@
 
-Another important aspect to note is the storage footprint of the databases.
+% Another important aspect to note is the storage footprint of the databases.
 PostgreSQL and VDMS require comparable amount of storage for metadata which
-is more than required by MySQL, shown in Figure~\ref{fig:db_time_size}.
+is more than required by MySQL, shown in Figure~\ref{fig:db_sizes}.
 In the case of VDMS, this is space used to store information about
@@
 % WAS: begin{figure*}[ht]; textwidth
-\includegraphics[width=\columnwidth]{figures/plot_conc_q_1tag_loc20_resize_mosaic_results_throughput_db_size}
+\includegraphics[width=\columnwidth]{figures/plot_conc_q_1tag_geo_resize_mosaic_results_throughput_db_size}
 \caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_geo\_resize}) from our use-case
@@
 drops and the performance of PostgreSQL becomes more comparable to VDMS. Lets look at the
-case with 32 concurrent clients. In this case, the throughput of VDMS drops drastically with
-the 100M database and PostgreSQL outperforms by a small margin. As the clients increase, we 
-see this trade-off occurs with the 50M database instead of 100M while in the 100M database, 
-VDMS outperforms both baselines. In this case, the throughput of PostgreSQL drastically degrades
+case with 32 concurrent clients.
+In this case, the throughput of VDMS drops drastically with
+the 100M database and PostgreSQL outperforms by a small margin.
+As the clients increase, we see this trade-off occurs with the 50M database
+instead of 100M while in the 100M database, VDMS outperforms both baselines.
+In this case, the throughput of PostgreSQL drastically degrades
 for the 100M database and the performance of the baseline systems are more comparable.
 
-When considering Figure~\ref{fig:concurrency_comparison_q2} through Figure~\ref{fig:concurrency_comparison_q4},
+When considering Figure~\ref{fig:concurrency_comparison_q2}
+through Figure~\ref{fig:concurrency_comparison_q4},
 there are many reasons why we generally see performance improvement, the main being
@@
 to be resized on the client side (as is the case in the baselines).
-This is in contrast with the baseline implementations, where 2 rounds of blocking back-and-forth
-communication with the server is needed, as depicted in Figure~\ref{fig:systems}.
+This is in contrast with the baseline implementations,
+where 2 rounds of blocking back-and-forth communication with the server is needed,
+as depicted in Figure~\ref{fig:systems}.
 Note that one could argue that the opposite will happen
@@
 on the other hand, the baseline implementations have a drop in performance
-for the 100M database. For \textit{q3}, we see that VDMS performs best when the database size
-is small (1M images), but as the database size increases, the performance degrades as well.  
-For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS for databases larger than 1M.
+for the 100M database.
+For \textit{q3}, we see that VDMS performs best when the database size
+is small (1M images), but as the database size increases,
+the performance degrades as well.
+For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS
+for databases larger than 1M.
 This also occurs for \textit{q5} but only for databases larger than 10M.
@@
 (\textit{q5}) decreases the performance as the database sizes scales.
-The behaviour is different in the case of \textit{q6}, which also filters by geo-location.
-This is attributed to the fact that OR queries involves processing larger results,
-and thus does not benefit from the filtering which is evident when compared to \textit{q4}.
+The behaviour is different in the case of \textit{q6},
+which also filters by geo-location.
+This is attributed to the fact that OR queries involves processing
+larger results, and thus does not benefit from the filtering
+which is evident when compared to \textit{q4}.
 

### 2021-02-25 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Update plots to remove super-titles

@@ -216,12 +216,12 @@ class EvalFramework(object):
         filename  = self.plot_folder + "plot_q_"
         filename += str(q) + "_mosaic_results_throughput_threads.pdf"
 
-        title = "Throughput as " + result_type + " per second for " + q
-        title += " as Database Size increases"
         threads = ["clients: " + str(a) for a in threads]
+        # title = "Throughput for " + q
+        # title += " as Database Size increases"
         p.plot_lines_all_mosaic(threads, db_sizes, engines, values,
                           filename=filename,
-                          title=title,
+                          # title=title,
                           log="both",
                           xlabel="Database Size",
                           ylabel=result_type + "/s")
@@ -267,7 +267,7 @@ class EvalFramework(object):
         filename  = self.plot_folder + "plot_conc_q_"
         filename += str(q) + "_results_throughput_db_size.pdf"
 
-        title = "Throughput as " + result_type + " per second Summary"
+        title = "Throughput as " + result_type + " per second - Summary"
         p.plot_lines_all(db_sizes, threads, engines, values,
                           title=title,
                           filename=filename,
@@ -277,11 +277,11 @@ class EvalFramework(object):
         filename  = self.plot_folder + "plot_conc_q_"
         filename += str(q) + "_mosaic_results_throughput_db_size.pdf"
 
-        title = "Throughput as " + result_type + " per second for " + q
-        title += " as number of concurrent clients increases"
+        # title = "Throughput for " + q
+        # title += " as number of concurrent clients increases"
         p.plot_lines_all_mosaic(db_sizes, threads, engines, values,
                           filename=filename,
-                          title=title,
+                          # title=title,
                           xlabel="# of concurrent clients",
                           ylabel=result_type + "/s")
 
@@ -326,7 +326,7 @@ class EvalFramework(object):
         filename  = self.plot_folder + "plot_conc_dbsize_"
         filename += str(db_size) + "_results_throughput.pdf"
 
-        title = "Throughput as " + result_type + " per second Summary"
+        title = "Throughput as " + result_type + " per second - Summary"
         p.plot_lines_all(queries, threads, engines, values,
                           title=title,
                           filename=filename,
@@ -336,11 +336,11 @@ class EvalFramework(object):
         filename  = self.plot_folder + "plot_conc_dbsize_"
         filename += str(db_size) + "_mosaic_results_throughput.pdf"
 
-        title = "Throughput as " + result_type + " per second for " + db_size
-        title += " as number of concurrent clients increases"
[…]

### 2021-02-25 yfcc100m/python/eval/DBEvalFramework/Plotting.py

@@ -141,8 +141,8 @@ class Plotting(object):
                         log="y",
                         title="",
                         filename="plot_unnamed.pdf",
-                        xlabel="Database Size",
-                        ylabel="None"):
+                        xlabel="",
+                        ylabel=""):
 
         # print(len(values))
         # print(len(values[1:]))
@@ -194,7 +194,7 @@ class Plotting(object):
                              values[n_queries*j + i,0:len(db_sizes)],
                              yerr=values[n_queries*j + i,len(db_sizes):],
                              label=engines[j],
-                             color="black",
+                             color=color[j],
                              linestyle=linestyles[j % len(linestyles)],
                              marker=markers[j % len(markers)],
                              )
@@ -213,7 +213,8 @@ class Plotting(object):
             if i % side == 0:
                 plt.ylabel(ylabel, fontsize=12)
 
-            # plt.xlabel(xlabel, fontsize=12)
+            if i >= (n_queries - side):
+                plt.xlabel(xlabel, fontsize=12)
 
         plt.savefig(filename, format="pdf", bbox_inches='tight')
         plt.close()

### 2021-02-25 yfcc100m/paper/vldb2021/legal.tex

--- 2020-02-27 yfcc100m/paper/vldb2020/legal.tex
@@
+\begin{framed}
 \scriptsize
-\begin{framed}
+
 \noindent \textbf{Notices \& Disclaimers}
 
-\copyright  Intel Corporation.  Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.  Other names and brands may be claimed as the property of others.
+\noindent Intel technologies may require enabled hardware, software or service activation.
 
-Performance results are based on testing as of dates shown in configurations and may not reflect all publicly available security updates.
+\noindent No product or component can be absolutely secure.
+
+\noindent Your costs and results may vary.
+
+\noindent \copyright Intel Corporation.
+Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.
+Other names and brands may be claimed as the property of others.
+
+\noindent Intel does not control or audit third-party data.  You should consult other sources to evaluate accuracy.
+
+\noindent Intel disclaims all express and implied warranties, including without limitation, the implied warranties of merchantability, fitness for a particular purpose, and non-infringement, as well as any warranty arising from course of performance, course of dealing, or usage in trade.
+
+% \copyright  Intel Corporation.
+Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.
+Other names and brands may be claimed as the property of others.
+
+% Performance results are based on testing as of dates shown in configurations and may not reflect all publicly available security updates.
 See backup for configuration details.
@@
 
-Your costs and results may vary.
+% Your costs and results may vary.
 
-Intel technologies may require enabled hardware, software or service activation.
+% Intel technologies may require enabled hardware, software or service activation.
 
-Software and workloads used in performance tests may have been optimized for performance only on Intel microprocessors.
+% Software and workloads used in performance tests may have been optimized for performance only on Intel microprocessors.
 Performance tests, such as SYSmark and MobileMark, are measured using specific computer systems, components, software, operations and functions.
@@
 
-Intel's compilers may or may not optimize to the same degree for non-Intel microprocessors for optimizations that are not unique to Intel microprocessors.
+% Intel's compilers may or may not optimize to the same degree for non-Intel microprocessors for optimizations that are not unique to Intel microprocessors.
 These optimizations include SSE2, SSE3, and SSSE3 instruction sets and other optimizations.
@@
 Please refer to the applicable product User and Reference Guides for more information regarding the specific instruction sets covered by this notice.
-Notice Revision #20110804
+Notice Revision \#20110804
 \end{framed}

### 2021-02-25 yfcc100m/paper/vldb2021/intro.tex
Add related work and references

@@
 Metadata, Image, Videos, and Feature Vectors.
+
+\subsection{Related Work}
+
+There are several databases and data management systems that focus one enabling
+analytics or combining transactional and analytic workloads over large
+scale data, such as SciDB~\cite{brown2010overview},
+BigTable~\cite{chang2008bigtable}, Shark~\cite{xin2013shark},
+and Vertica~\cite{lamb2012vertica}.
+While these systems do not focus on visual processing as a primary entity,
+they are valuable resources in distributing and analyzing very large scale data.
+A more concrete application and case of study include
+the Facebook architecture for photos that combines Tao~\cite{tao},
+Haystack~\cite{beaver2010finding}, and f4~\cite{muralidhar2014f4} for metadata,
+hot/recent data, and warm data respectively.
+While the social aspect of Facebook is logically suited to a graph,
+the bulk of their data is already stored in MySQL~\cite{mysql},
+so they chose to develop a graph-aware cache.
+A similar approach was taken by Grail~\cite{fan2015case},
+which argued that a syntactic layer written on top of a relational database
+could answer graph queries.
+Diamond~\cite{huston2004diamond,satyanarayanan2010unique} exploited active
+disks to perform discard-based searching order to shrink the amount of data
+returned to a reasonable size.
+We accomplish the same goal by exploiting persistent memory to store a
+graph database for metadata.

### 2021-02-25 yfcc100m/paper/vldb2021/design.tex
Expand on API on design section

@@
 (including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
-Finally, a main component of VDMS is in charge of implementing the API and orchestrating between the PMGD and the Visual Compute Module
+Finally, a main component of VDMS is in charge of implementing the
+API and orchestrating between the PMGD and the Visual Compute Module
 to serve client's requests. This component is the \textit{Request Server}.
@@
 
-Listing~\ref{findimagegeo} shows a sample query.
-In this particular example, the transaction retrieves
-all the images of \textit{alligators}
-with probability higher than 0.66,
+Listing~\ref{addimageandautotag} shows a sample query for inserting an image
+and its associated metadata to VDMS.
+In this example, the transaction inserts an application-defined "Entity" of
+the class "autotag", with the "name" property being \textit{alligator},
+inserts an Image with its "latitude" and "longitude",
+stores the image  as a JPG (VDMS will transcode if needed),
+and creates a connection between the Image and the Entity, with a "prob"
+property (which indicates the probability of that image
+containing an object of type  \textit{alligator}) with a specific value.
+Note that no schema needs to be defined in advance.
+The Entity of class "autotag", with its properties, are declared and added
+at insertion time, without any need to define a schema of objects
+(and its properties) before hand.
+
+\begin{listing}[ht!]
+\begin{minted}[frame=single,
+              framesep=3mm,
+              linenos=true,
+              xleftmargin=21pt,
+              tabsize=4]{js}
+"AddEntity"{
+    "_ref" : 1,
+    "class": "autotag",
+    "properties": {
+        "name": "alligator"
+    }
+},
+"AddImage":{
+    "_ref": 2,
+    "properties": {
+        "latitude":  36.23433,
+        "longitude": -116.80666
+    },
+    "format": "jpg"
+}
+"AddConnection": {
+    "ref1": 1,
+    "ref2": 2,
+    "properties": {
+        "prob": 0.7653
+    }
+}
+
+\end{minted}
+\caption{Sample Query for Image Insertion -
+The query expresses the following:
+Insert an Entity of the class "autotag", with the "name" property being
+\textit{alligator}, insert an Image with its "latitude" and "longitude",
+store the image  as a JPG, and create a connection between
+the image and the "autotag",  with a property "prob"
+(which indicates the probability of that image
+containing an object of type  \textit{alligator}).}
+\label{addimageandautotag}
+\end{listing}
+
+Listing~\ref{findimagegeo} shows another sample query, in this case, for retrieval.
+In this particular example, the transaction retrieves all the images
+of \textit{alligators} with probability higher than 0.66,
 filter by latitude and longitude within 1 degree,
 apply a resize operation to make the images 224x224,
-rotate the images 45.34 degrees,
-and return the images as "png" files.
+rotate the images 45.34 degrees, and return the images as "png" files.
 It is important to note how the API natively supports basic building

### 2021-02-25 yfcc100m/paper/vldb2021/eval.tex
Add reference to other results
Update Comparison systems figures and eval description

@@
 \includegraphics[width=\textwidth]{figures/comparison_system}
-\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline systems (right).
-The baseline systems are either based on MySQL or PostgreSQL for operating and querying metadata.}
+\caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline systems based on Apache Web Server and PostgreSQL (middle), or MySQL (right).
+The image repository is shared.}
 \label{fig:systems}
@@
 Given that there are no other open-source systems that provide similar
-functionality and interfaces as VDMS, we have implemented two equivalent
+functionality and interfaces as VDMS (i.e., transactionally dealing with images
+and metadata behind a single interface), we have implemented two equivalent
 visual data management system as baselines, comprised of a combination of
@@
 OpenCV 3.3 (to provide pre-processing operations on images).
-The other baseline system replaces the MySQL Server 5.7 with the
-most advanced open source relational database, PostgreSQL 9.5 \cite{postgresql}.
+The other baseline system replaces the MySQL Server 5.7 with a
+most advanced open-source relational database, PostgreSQL 9.5 \cite{postgresql}.
 We decided to use relational databases in the baseline systems instead of
-non-relational databases because~\cite{Jatana2012, li_2019}:
+non-relational databases because of their madurity, and also because it is
+the most common tools used for storing and querying metadata
+~\cite{Jatana2012, li_2019}:
 
@@
 
+Because of space constraints, we are only showing a subset of our evaluation.
+
+We are making available the all available  of For more details about our comprehensive evaluation.
+available
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}
+% TODO. Upload the final version of the plots once we are sure we are not making
+% more modifications.
+
 %% ========= Summary Analysis - END

### 2021-02-25 yfcc100m/paper/vldb2021/design.tex
Update intro and design sections. Move appendix to eval.

@@
 which was briefly introduced in previous work \cite{vdms-nips}.
-
-VDMS implements a client-server architecture that handles client
-requests concurrently, and coordinates query execution across
-the metadata and visual data components in order to return a unified response.
-Users interact with both metadata and visual data
-(i.e., images, videos, feature vectors) using a unified API,
-in a transactional manner.
-Figure \ref{fig:arch} depicts the high-level architecture of VDMS.
-
-The metadata component is the \textit{Persistent Memory Graph
-Database} (PMGD) and the (visual) data component is our Visual Compute Module.
-The Visual Compute Module enables machine-friendly enhancements to
-visual data, exposing high-level abstractions to the \textit{Request Server}
-for dealing with a variety of images and video formats (through OpenCV and ffmpeg),
-and different methods for indexing for feature vectors
-(including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
-Finally, a main component of VDMS is in charge of implementing the
-API and orchestrating between the PMGD and the Visual Compute Module
-to serve client's requests. This component is the \textit{Request Server}.
-
-VDMS and its components are fully available as open source projects
-\footnote{https://github.com/IntelLabs/\{vdms, pmgd\}}.
-We briefly describe each of the main components as follows:
+VDMS implementation is fully open-sourced
+\footnote{https://github.com/IntelLabs/vdms}.
 
@@
 
-\subsection{Persistent Memory Graph Database}
+VDMS implements a the typical client-server architecture that handles
+client queries transactionally and concurrently, as most of the most common
+relational and non-relational database systems
+\cite{mysql, postgresql, chang2008bigtable}.
+The main difference between other data management systems and VDMS is that
+it goes beyond the typical supported data types (string, integers, floats,
+blobs, JSON-documents, etc), recognizing visual entities (image, videos,
+feature vectors, etc) as first class citizens.
+VDMS API enable users to insert, index, process, and query visual data,
+as well as provides full support for inserting, indexing and querying
+user-defined metadata.
+Users interact with both metadata and visual data using a unified API,
+in a transactional manner.
 
-We use the Persistent Memory Graph Database (PMGD) to provide an efficient storage
-solution addressing the increasing popularity of connected data and
-applications that benefit from graph-like processing.
-PMGD implements an in-persistent-memory graph database, PMGD, optimized
-to run on a platform equipped with persistent memory.
-PMGD provides a property graph model of data storage with the traditional
+VDMS provides a property \textit{graph model} abstraction with the traditional
 atomicity, consistency, isolation, and durability
 (ACID) properties expected from databases.
-Graph represents an easier abstraction to model complex problems \cite{tao}.
-Moreover, the graph model makes it very suitable for the data and
-access patterns shown by visual metadata, which can be easily mapped
-into application-level abstractions by developers.
+This is, users interact with their objects (metadata, images, videos, etc),
+as if these objects were in a connected graph.
+Graph represents an easier abstraction to model complex problems,
+making it very suitable for the data and access patterns shown by visual metadata,
+which can be easily mapped into application-level abstractions by
+developers~\cite{tao}.
 For instance, abstractions like \textit{BoundingBoxes} associated to
-images or videos can be easily represented using nodes and edges.
-With its natural ability to extend the schema very
-easily (due to the use of a property graph model),
-we can support new developments in machine learning that can lead to
-enhancements to existing metadata over time.
+images or videos can be easily represented using nodes and edges in a graph.
 These are the reasons the team chose a graph database over a
 relational database as the metadata management for the implementation of VDMS.
-PMGD is designed and optimized for persistent memory technologies
-like Intel Optane~\cite{IntelXPoint15}, which
-promise storage providing nearly the speed of DRAM and the
-durability of block-oriented storage.
 
-\subsection{Visual Compute Module}
+VDMS API provides mechanism to insert and connect Images, Videos, BoundingBoxes,
+Frames, and Descriptors (feature vectors), together with any metadata
+associated with the objects.
+Each object is a node in the graph.
+The metadata associated to each object is is modeled as
+"properties" of the node in the graph.
+Users can query, filter, and retrieve this objects based
+on its metadata (the "properties" associated with them).
+VDMS does not simply treat these objects as binary blobs of data, but rather
+understands them and the type of processing that is common for them,
+providing the ability to run processing on-the-fly, both at insertion and
+at retrieval type.
+Users can, for instance, retrieve a resized version of an image,
+or retrieve only a clip from a large video.
+This is one of the main differentiation aspect when compared to other
+database systems, including relational databases.
+In a relational database, for instance, one can query \textit{and compute}
+over values in a column only for basic data types (strings, integers, floats),
+and over the abstractions the relational model support (tables, columns).
+A SQL query can then retrieve the \textit{computed average} "salary" of all
+the employees in a company (stored using the table/columns abstraction),
+but cannot perform any computation over data stored as a blob or binary object.
+VDMS, because it recognized the nature of visual objects by design,
+provides the ability to \textit{compute} on these visual objects
+(image, videos, etc).
 
-The Visual Compute Module was designed and implemented to provide
-an internal abstraction layer for interacting with visual data.
-It enables simple visual data handling and processing (i.e., basic building block
-operations like crop, resize, etc).
-For traditional formats (jpg, png, tiff, mp4, etc.),
-the interface is an abstraction layer over OpenCV.
-However, it also provides a way to use novel formats that are
-better suited for visual analytics: a novel, array-based lossless image format.
-This format is built on the array data manager TileDB~\cite{TileDB} and
-is well suited for images that are used in visual analytics.
+VDMS API also allows users to insert application-defined "Entities",
+that enable applications to model any use-case specific metadata.
+An "Entity" object (and its properties) is a node in the graph.
+For example, a user can define an Entity object of a class "Person",
+and \textit{connect} this person to one or multiple image objects.
+Later, the user can retrieve the Entity object corresponding to a person,
+together with all the images \textit{connected} to it.
 
-VDMS provides full support for video storage and operations,
-in a similar way it does for images.
-This includes support for encoding, decoding, and transcoding of
-\textit{mp4}, \textit{avi}, and \textit{mov} containers,
-as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
-This is supported through the use of either OpenCV~\cite{opencv}
-or \textit{libffmpeg}\cite{ffmpeg}, or both.
-All operations supported for images in VDMS are also supported at the
-video and frame level of the API.
-On top of that, there are a number of video-specific operations that
-are supported, such as the interval operations,
-enabling users to retrieve clips at different
-frames-per-second (FPS) versions of the video.
+By providing the ability to store visual data objects together with
+application-defined entities and its properties, VDMS can manage
+all the data the application need behind a single, unified API.
+This is in contrast with current applications that rely on a combination
+of multiple systems and APIs to access different portions of the data they need
+\cite{tao, sculley2015hidden, mayer2020scalable, sculley2015hidden}.
 
-Another key differentiating factor of VDMS is that it allows the creation of
-indexes for high-dimensional feature vectors and the insertion of
-these feature vectors associated with entities, images, and/or videos.
-Feature vectors are intermediate results of various machine
-learning or computer vision algorithms when run on visual data.
-Feature vectors are also known as \textit{descriptors}
-or \textit{visual descriptors}.
-We use these terms interchangeably.
-These descriptors can be classified, labeled, and used to build search indexes.
-Feature Vectors support is provided through our implementation based
-on high-dimensional sparse arrays, also using TileDB.
-In addition, the Visual Compute Library provides a wrapper
-for another high-dimensional index implementation,
-Facebook's Faiss~\cite{faiss}.
-
-\subsection{Request Server}
-
-Developers and users of machine learning frameworks and data science
-applications favor simpler interfaces to access and process data.
-They cannot be expected to deal with two different ways of interacting
-with information (metadata and visual data) instead of focusing on the
-algorithmic parts of their pipelines.
-VDMS takes care of coordinating client requests across the metadata and the
-visual data, and managing multiple clients through the \textit{Request Server}.
-The \textit{Request Server} is a key component of the system,
-as it implements VDMS' API and coordinates request and responses from
-the PMGD and Visual Compute Module subsystems.
-It decomposes the user queries into
-metadata and visual data requests, invokes the relevant calls behind the scene,
-and returns a coherent and unified response that is easy for the user to parse
-and interpret.
+Figure \ref{fig:arch} depicts the high-level architecture of VDMS, which
+is composed by several subcomponents, and a Query Engine that implements
+the unified API and hides all the complexity underneath.
+We first describe the VDMS API, and then describe its subcomponents
+and design decisions.
 
@@
 
-One of the most important differentiating factor of VDMS is its interface.
+One of the most important differentiating aspects of VDMS is its API.
 VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
-as first class citizens. Thus, VDMS provides an API that revolves
-around visual data operations and retrieval.
+as first class citizens.
+Thus, VDMS' API revolves around visual data operations and retrieval.
 VDMS API is easy to use and explicitly pre-defines certain
@@
 
+\subsection{Graph Engine}
+
+The Graph Engine used for in VDMS is the
+\textit{Persistent Memory Graph Database} (PMGD).
+PMGD was developed at Intel Labs, and was designed and optimized for
+persistent memory technologies like Intel Optane~\cite{IntelXPoint15}, which
+promise storage providing nearly the speed of DRAM and the
+durability of block-oriented storage.
+PMGD is also fully open-sourced \footnote{https://github.com/IntelLabs/pmgd}.
+PMGD implements an in-persistent-memory graph database, PMGD, optimized
+to run on a platform equipped with persistent memory, but
+also provides the option to run on SSD and ext4 filesystem
+using OS support to provide transactional guarantees (through \textit{msync}).
+PMGD provides a property graph model of data storage with the traditional
+atomicity, consistency, isolation, and durability
+(ACID) properties expected from databases.
+PMGD comes as a C++ library that VDMS uses for its internal query engine.
+Because PMGD already supports most of the graph abstractions we wanted
+to see on VDMS API, it was our preferred option.
+Also, our internal evaluation shows that PMGD is faster
+than other graph databases, including Neo4j\cite{miller2013graph}.
+PMGD performance evaluation will be published separately.
+
+\subsection{Visual Compute Module}
+
+% The Visual Compute Module enables machine-friendly enhancements to
+% visual data, exposing high-level abstractions to the \textit{Request Server}
+% for dealing with a variety of images and video formats (through OpenCV and ffmpeg),
+% and different methods for indexing for feature vectors
+% (including Facebook's Faiss \cite{faiss}, TileDB \cite{TileDB}).
+% Finally, a main component of VDMS is in charge of implementing the
+% API and orchestrating between the PMGD and the Visual Compute Module
+% to serve client's requests. This component is the \textit{Request Server}.
+
+The Visual Compute Module was designed and implemented to provide
+an internal abstraction layer for interacting with visual data.
+It enables simple visual data handling and processing
+(i.e., basic building block operations like crop, resize, etc).
+For traditional formats (jpg, png, tiff, mp4, etc.),
+the interface is an abstraction layer over OpenCV.
+However, it also provides a way to use novel formats that are
+better suited for visual analytics: a novel, array-based lossless image format.
+This format is built on the array data manager TileDB~\cite{TileDB} and
+is well suited for images that are used in visual analytics.
+Note that, even if VDMS currently provides support for array-based lossless
+image format, we do not use it as part of this evaluation.
+This work focuses on a more direct comparison with a combination of alternatives,
+and thus we use the traditional format (jpg and png).
+The performance comparison between the array-based lossless image format
+we develop and other similar formats (like PNG or TIFF) are outside of the
+scope of this evaluation.
+
+VDMS also provides full support for video storage and operations,
+in a similar way it does for images.
+This includes support for encoding, decoding, and transcoding of
+\textit{mp4}, \textit{avi}, and \textit{mov} containers,
+as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
+This is supported through the use of either OpenCV~\cite{opencv}
+or \textit{libffmpeg}\cite{ffmpeg}, or both.
+All operations supported for images in VDMS are also supported at the
+video and frame level of the API.
+On top of that, there are a number of video-specific operations that
+are supported, such as the interval operations,
+enabling users to retrieve clips at different
+frames-per-second (FPS) versions of the video.
+
+Another key differentiating factor of VDMS is that it allows the creation of
+indexes for high-dimensional feature vectors and the insertion of
+these feature vectors associated with entities, images, and/or videos.
+Feature vectors are intermediate results of various machine
+learning or computer vision algorithms when run on visual data.
+Feature vectors are also known as \textit{descriptors}
+or \textit{visual descriptors}.
+We use these terms interchangeably.
+These descriptors can be classified, labeled, and used to build search indexes.
+Feature Vectors support is provided through our implementation based
+on high-dimensional sparse arrays, also using array-based approaches.
+In addition, the Visual Compute Library provides a wrapper
+for another high-dimensional index implementation,
+Facebook's Faiss~\cite{faiss}.
+
 \subsection{Client Library}

### 2021-02-25 yfcc100m/paper/vldb2021/eval.tex
Update intro and design sections. Move appendix to eval.

@@
 \centering
-\includegraphics[width=\columnwidth]{figures/all_build_sizes_plot}
+\includegraphics[width=0.8\columnwidth]{figures/all_build_sizes_plot}
 \caption{Data footprint (in GB) of MySQL, PostgreSQL, and VDMS databases.}
@@
 
+Figure~\ref{fig:q_throughput_16} shows the measured throughput,
+as images per second, delivered by each system for all queries and databases.
+In comparison to figure~\ref{fig:q_throughput_56},
+when using fewer concurrent clients (nthr = 16),
+the throughput for VDMS for queries \textit{q4} (\textit{2tag\_or\_resize}),
+\textit{q5} (\textit{2tag\_resize\_geo\_and}), and
+\textit{q6} (\textit{2tag\_resize\_geo\_or}) improved by a large margin
+for the 50M and 100M databases.
+In the case of \textit{q4} and \textit{q6},
+the throughput for 50M (and 100M for \textit{q4})
+increase by a large margin which outperforms both baselines while in
+Figure~\ref{fig:q_throughput_56} PostgreSQL slightly performed better than VDMS.
+In the case of \textit{q5}, all systems perform a little differently for databases
+larger than 5M.
+The gaps between the throughput of each system narrows at 10M.
+At a database size of 50M, the performance of the baselines drastically drop
+while the throughput of VDMS has a large improvement in comparison.
+
+\begin{figure}[ht!]
+\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
+\caption{Throughput Analysis using all queries from our use-case
+described in the Experimental Setup Section.
+We show queries in different figures for readability reasons.
+The experiments show the performance of all systems (VDMS and the two baselines) as the
+database size increases.
+These queries were run using 16 simultaneous clients (nthr = 16),
+and averaged out of 5 runs.}
+\label{fig:q_throughput_16}
+\end{figure}
+
 %% ========= Queries Analysis - END

### 2021-02-25 yfcc100m/paper/vldb2021/intro.tex
Expand on API on design section
Update intro and design sections. Move appendix to eval.

@@
 
-Visual computing workloads performing analytics on videos and/or images
+Visual computing workloads performing analytics on images and/or videos
 have become prolific across a wide range of application domains.
-This is in part due to the growing ability of machine learning (ML) techniques to
-extract information from the visual data which can subsequently be used
+This is in part due to the growing ability of machine learning (ML) techniques
+to extract information from the visual data which can subsequently be used
 for informed decision making \cite{vdms-nips}.
 The insights this information can provide depend on the
-application: retail vendors might be interested in knowing
-which are the most visited areas of their stores using security
-video feeds as input, or a doctor might want know the effect of a
-specific treatment by looking at the changes in size of a tumor from a brain scan.
+application: retail vendors might be interested in knowing which are the most
+visited areas of their stores using security video feeds as input,
+or a doctor might want know the effect of a specific treatment by
+looking at the changes in size of a tumor from a brain scan.
 
@@
 there has been very little research on the \textit{management} of visual data.
-Most of the current storage solutions for visual data are
-an ad-hoc collection of tools and systems, that are re-purposed and
-adapted to work with visual data.
+Most of the current storage solutions for visual data are an ad-hoc collection
+of tools and systems, that are re-purposed and adapted to work with visual data.
 The approach of re-purposing and integrating solutions not designed for a task
 results in resource utilization inefficiencies\cite{rasdaman}.
-For example, consider a ML developer constructing a pipeline
-for extracting brain tumor information from existing brain images in a
-classic medical imaging use case.
-This requires assigning consistent
-identifiers for the scans and adding their metadata in
-a relational or key-value database\cite{kumar2017data}.
+For example, consider a ML developer constructing a pipeline for extracting
+brain tumor information from existing brain images in a classic medical
+imaging use case.
+This requires assigning consistent identifiers for the scans and adding their
+metadata in a relational or key-value database\cite{kumar2017data}.
 If the queries require a search over patient information,
@@
 
-We show how VDMS can be used as a centralize point for data
-management and data access even when having multiple modalities of data:
-Metadata, Image, Videos, and Feature Vectors.
+Overall, this paper makes the following contributions:
 
-\subsection{Related Work}
+\begin{itemize}
+\item We present the design and implementation of the Visual Data Management System.
+To the best of our knowledge, there are no other system capable of
+managing visual data and its metadata behind a unified API
+and that, at the same time, enable applications to operate over visual data
+(images, videos, feature vectors).
+
+\item We introduce the VDMS API in detail, designed for Machine Learning and visual
+data analytics at scale.
+
+\item We use our internal use-cases and implementations of visual data processing
+pipeline to make a comprehensive evaluation between common ad-hoc solutions
+that rely on multiple systems and VDMS.
+
+\item We use the largest public available image dataset, YFCC100m, together
+with machine generated tags for our real-world use case, providing a performance
+evaluation both from the point of view of high concurrency as well as
+scale (the database size, including images and metadata is 12TB in size).
+\end{itemize}
+
+\section{Related Work}
 
@@
 they are valuable resources in distributing and analyzing very large scale data.
-A more concrete application and case of study include
+A more concrete application involving visual data include
 the Facebook architecture for photos that combines Tao~\cite{tao},

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Update eval

@@
 We decided to use relational databases in the baseline systems instead of
-non-relational databases because of their madurity, and also because it is
+non-relational databases because of their maturity, and also because it is
 the most common tools used for storing and querying metadata
@@
 We decided to run the VDMS server in the machine with less DRAM to make
-sure MySQL and PostgreSQL had no disadvantage, and because previous evaluation
+sure MySQL and PostgreSQL had no disadvantage. Previous evaluation
 indicated smaller footprint in the case of VDMS when
@@
 interaction of the client application (retrieves metadata and
-images) with VDMS (left) and the MySQL/PostgreSQL baselines (right).
+images) with VDMS (left), the PostgreSQL baseline (middle), and the MySQL baseline (right).
 
@@
 In systems where persistent memory is a scarce resource,
-the increased storage foot print of PMGD may represent a challenge.
+the increased storage footprint of PMGD may represent a challenge.
 On the other hand, persistent memory is expected to be available
@@
 The figures show the throughput (images per second) when retrieving
-resized versions of the images, as the number of concurrent clients increases
+resized versions of the images, as the number of concurrent clients increases.
 Hardware concurrency (number of physical cores in each system)
-is shown with a blue vertical line (hw-concurrency = 56).}
+is 56.}
 \label{fig:concurrency_comparison_q2}
@@
 depict the difference between VDMS and the baselines.
-Here, the baselines are the full architecture described in
-Figure~\ref{fig:systems} (right).
+Here, the baselines are the full architectures described in
+Figure~\ref{fig:systems} (middle and right).
 
@@
 Hardware concurrency (number of physical cores in each system)
-is shown with a blue vertical line (hw-concurrency = 56).}
+is 56.}
 \label{fig:concurrency_comparison_10M}
@@
 4 concurrent clients.  At 16 concurrent client, the performance of the MySQL baseline
-begin to stabilizes while the throughput of PostgreSQL begins to degrade. However, as
+begin to stabilize while the throughput of PostgreSQL begins to degrade. However, as
 concurrency increases beyond 16 concurrent clients, the difference
@@
 VDMS when compared to the baselines, and how this improvement is maintained
-as the size of the database increase while
+as the size of the database increase while,
 on the other hand, the baseline implementations have a drop in performance
@@
 Because of space constraints, we are only showing a subset of our evaluation.
-
-We are making available the all available  of For more details about our comprehensive evaluation.
-available
+More details about our comprehensive evaluation are available.
 \footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}

### 2021-02-26 yfcc100m/paper/vldb2021/intro.tex
minor update to intro

@@
 
-There are several databases and data management systems that focus one enabling
+There are several databases and data management systems that focus on enabling
 analytics or combining transactional and analytic workloads over large
@@
 A more concrete application involving visual data include
-the Facebook architecture for photos that combines Tao~\cite{tao},
+the Facebook architecture for photos that combine Tao~\cite{tao},
 Haystack~\cite{beaver2010finding}, and f4~\cite{muralidhar2014f4} for metadata,

### 2021-02-26 yfcc100m/paper/vldb2021/design.tex
Minor updates to design

@@
 
-VDMS API provides mechanism to insert and connect Images, Videos, BoundingBoxes,
+VDMS API provides a mechanism to insert and connect Images, Videos, BoundingBoxes,
 Frames, and Descriptors (feature vectors), together with any metadata
@@
 Each object is a node in the graph.
-The metadata associated to each object is is modeled as
+The metadata associated to each object is modeled as
 "properties" of the node in the graph.
@@
 providing the ability to run processing on-the-fly, both at insertion and
-at retrieval type.
+retrieval.
 Users can, for instance, retrieve a resized version of an image,
@@
 over values in a column only for basic data types (strings, integers, floats),
-and over the abstractions the relational model support (tables, columns).
+and over the abstractions the relational model supports (tables, columns).
 A SQL query can then retrieve the \textit{computed average} "salary" of all
@@
 
-The Graph Engine used for in VDMS is the
+The Graph Engine used in VDMS is the
 \textit{Persistent Memory Graph Database} (PMGD).
@@
 The client library implements TCP/IP based connectors to the VDMS Server,
-similar to most database\cite{memsql, mysql, postgresql}.
+similar to most databases\cite{memsql, mysql, postgresql}.
 Users can connect to VDMS and implement queries using VDMS' API

### 2021-02-26 yfcc100m/paper/vldb2021/conclusion.tex
Update results in conclusion

--- 2021-02-03 yfcc100m/paper/osdi2020/conclusion.tex
@@
 industry standard systems, all of which are needed to 
-replicate a only portion of VDMS' functionality. 
-We see improvements up to 35x in certain queries, 
-and an average improvement of about 15x. 
+replicate only a portion of VDMS' functionality. 
+We see improvements up to 364x in certain queries, 
+and an average improvement of about 85x when compared to PostgreSQL.
+When compared to MySQL, we see up to 96x speedup
+and and average improvement of 31x. 
 % Finally, we analyze the different trade-offs of VDMS' 

### 2021-02-26 yfcc100m/paper/vldb2021/design.tex
Minor fixes

@@
 
-VDMS implements a the typical client-server architecture that handles
-client queries transactionally and concurrently, as most of the most common
+VDMS implements the typical client-server architecture that handles
+client queries transactionally and concurrently, similar to what most common
 relational and non-relational database systems
-\cite{mysql, postgresql, chang2008bigtable}.
+\cite{mysql, postgresql, chang2008bigtable} use.
 The main difference between other data management systems and VDMS is that
@@
 
-VDMS provides a property \textit{graph model} abstraction with the traditional
+VDMS provides a \textit{graph model} abstraction with the traditional
 atomicity, consistency, isolation, and durability
@@
 images or videos can be easily represented using nodes and edges in a graph.
-These are the reasons the team chose a graph database over a
-relational database as the metadata management for the implementation of VDMS.
+This is the main reason why the team chose a graph model over a
+relational one for the implementation of VDMS API.
 
@@
 Each object is a node in the graph.
-The metadata associated to each object is modeled as
+The information associated to each visual object (image, video, etc) is modeled as
 "properties" of the node in the graph.
-Users can query, filter, and retrieve this objects based
-on its metadata (the "properties" associated with them).
+Users can query, filter, and retrieve this objects based on its properties.
 VDMS does not simply treat these objects as binary blobs of data, but rather
@@
 providing the ability to run processing on-the-fly, both at insertion and
-retrieval.
+retrieval time.
 Users can, for instance, retrieve a resized version of an image,
@@
 and over the abstractions the relational model supports (tables, columns).
-A SQL query can then retrieve the \textit{computed average} "salary" of all
+This is, a SQL query can retrieve the \textit{computed average} "salary" of all
 the employees in a company (stored using the table/columns abstraction),
@@
 This is in contrast with current applications that rely on a combination
-of multiple systems and APIs to access different portions of the data they need
+of multiple data management systems and APIs to access different
+portions of the data they need
 \cite{tao, sculley2015hidden, mayer2020scalable, sculley2015hidden}.
@@
 as first class citizens.
-Thus, VDMS' API revolves around visual data operations and retrieval.
+Thus, VDMS' API revolves around visual data operations and retrieval,
+but at the same time, enables applications to store any other
+application-specific metadata.
 VDMS API is easy to use and explicitly pre-defines certain
@@
 
-Listing~\ref{addimageandautotag} shows a sample query for inserting an image
-and its associated metadata to VDMS.
+Listing~\ref{addimageandautotag} shows a sample query for inserting
+an image, properties associated with the image, and metadata to VDMS.
+The metadata, in this case, is the information about the "autotag",
+which is provided by the dataset in this specific
+\textit{image-search} application.
 In this example, the transaction inserts an application-defined "Entity" of
@@
 containing an object of type  \textit{alligator}) with a specific value.
+This query is performed \textit{transactionally}: either all the image, metadata
+and connection are inserted, or none of them are and an error is returned.
 Note that no schema needs to be defined in advance.
@@
 (changing output formats and encodings).
-The API allows interaction with metadata, images, videos,
-and feature vectors in a similar fashion, and it is fully
-documented on the project's Github
+The API allows interaction with metadata, images, videos, bounding boxes,
+frames, feature vectors, and more in a similar fashion,
+and it is fully documented on the project's Github
 wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.
@@
     "format": "png",
-    "link": {
-        "ref":1,
-        "constraints": {
-            "prob": [">=", 0.66]
-        }
-    },
     "constraints": {
@@
             "angle": 45.34
-    }]
+    }],
+    "link": {
+        "ref":1,
+        "constraints": {
+            "prob": [">=", 0.66]
+        }
+    }
 }
@@
 PMGD is also fully open-sourced \footnote{https://github.com/IntelLabs/pmgd}.
-PMGD implements an in-persistent-memory graph database, PMGD, optimized
+PMGD implements an in-persistent-memory graph database optimized
 to run on a platform equipped with persistent memory, but
@@
 (ACID) properties expected from databases.
-PMGD comes as a C++ library that VDMS uses for its internal query engine.
+PMGD comes as a C++ library that VDMS uses for implementing its query engine.
 Because PMGD already supports most of the graph abstractions we wanted
@@
 an internal abstraction layer for interacting with visual data.
-It enables simple visual data handling and processing
-(i.e., basic building block operations like crop, resize, etc).
+It enables the query engines to coordinate and perform
+visual data handling and processing
+(i.e., basic building block operations like crop, resize, etc for images/videos
+and k-nearest neighbor search for feature vectors),
+shown in Figure \ref{fig:arch}.
 For traditional formats (jpg, png, tiff, mp4, etc.),
@@
 The performance comparison between the array-based lossless image format
-we develop and other similar formats (like PNG or TIFF) are outside of the
+we develop and other similar formats (like png or tiff) are outside of the
 scope of this evaluation.
@@
 This is supported through the use of either OpenCV~\cite{opencv}
-or \textit{libffmpeg}\cite{ffmpeg}, or both.
+or \textit{libffmpeg}\cite{ffmpeg}, or both,
+plus additional implementation to support fast random access to video frames.
 All operations supported for images in VDMS are also supported at the
@@
 enabling users to retrieve clips at different
-frames-per-second (FPS) versions of the video.
+frames-per-second (fps) versions of the video.
 
@@
 Facebook's Faiss~\cite{faiss}.
+Users can, through our API, use different indexing techniques
+for feature vectors, depending on their application's need.
 
@@
 similar to most databases\cite{memsql, mysql, postgresql}.
-Users can connect to VDMS and implement queries using VDMS' API
-by defining JSON commands conforming to the query protocol we have defined.
+Users can connect to VDMS and run queries using VDMS' API
+by defining a transaction using JSON objects.
 The client library provides a simple method that
@@
 and returns it to the client.
-The responses require JSON parsing on the client
-side for the metadata string that indicates how to interpret the blobs field.
 Currently, client libraries are implemented for Python and C++ client.

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Minor fixes

@@
 used in the industry, and following what we have done in the past
-in order to solve the data search problem,
-which we describe in Section \ref{images}.
+in order to build an image search engine, which we describe in Section \ref{images}.
 % We have also included performance evaluation on the Video and Feature Vectors
@@
 probability of that \textit{autotags} being present in that image/frame.
-We have also used feature vectors generated for every image and first frame
-of every video \cite{features} to implement a similarity search.
+% We have also used feature vectors generated for every image and first frame
+% of every video \cite{features} to implement a similarity search.
 Given that there is no standard benchmark oriented towards visual data queries,
@@
 and metadata behind a single interface), we have implemented two equivalent
-visual data management system as baselines, comprised of a combination of
-widely available, off-the-shelf components.
+visual data management system as baselines,
+desing specifically for the image search use case,
+and comprised of a combination of widely available, off-the-shelf components.
 The first baseline uses MySQL Server 5.7 (for storing metadata),
@@
 OpenCV 3.3 (to provide pre-processing operations on images).
+We chose Apache Web Server to work as a file server (serving the image files),
+because it provides the lowest possible overhead
+(when compared to other object store system), at the expenses of providing less
+functionality (authentication, data integrity checking, etc).
 The other baseline system replaces the MySQL Server 5.7 with a
@@
     and durability (ACID) while non-relational may compromise some ACID properties.
-    \item The YFCC100m data is clearly structured.
-    \item We need to efficiently collate and return metadata records.
+    \item The YFCC100m data is clearly structured, and can be modeled with relatively ease with a relational database.
+    \item We need to efficiently collate and return metadata records,
+    and SQL engines are very mature and optimized for that task.
 \end{itemize}
@@
 The baseline implementations are based upon internal tools used for ML-based
-pipelines for media, the common practice in the industry \cite{haystack, tao}.
+pipelines for visual data, which is common practice
+in the industry \cite{haystack, tao}.
 We have implemented a set of client-side applications that take care
@@
 To build both MySQL and PostgreSQL on the same machine,
-we stored each relational database on separate SSD drives.
+we stored each database on separate SSD drives.
 Other than the difference in DRAM space and storage needed for two baselines,
@@
 Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
-it can deliver good performance when using SSDs directly, while still
-providing ACID-compliant transactions.
+it can deliver good performance when using SSDs directly.
 
-For the metadata, we built VDMS, MySQL, and PostgreSQL databases
-using the YFCC100M dataset with incremental database sizes.
+For the metadata, we built VDMS, MySQL, and PostgreSQL databases using
+the YFCC100M dataset with incremental database sizes to study the effects at scale.
 For simplicity, we named the database based on the approximate number of images
 it contains, as follows: 1M, 5M, 10M, 50M, 100M.
-The databases have comparable number of elements.
+The baseline systems have comparable number of elements.
 The exact number of images/elements in each database are shown in
@@
 \subsubsection{Data Representation}
+We detail the data representation for each of the systems we implemented:
 
@@
 
-\textbf{MySQL/Baseline 1:}
+\textbf{MySQL-Based System:}
 Each MySQL database is created in a similar manner as VDMS
@@
 
-\textbf{PostgreSQL (Baseline 2):}
+\textbf{PostgreSQL-Based System:}
 Each PostgreSQL database is created in the exact same manner as MySQL

### 2021-02-26 yfcc100m/paper/vldb2021/intro.tex
Minor fixes

@@
 results in resource utilization inefficiencies\cite{rasdaman}.
-For example, consider a ML developer constructing a pipeline for extracting
+The combination of systems is required to manage both \textit{metadata} and
+\textit{visual data}.
+We use \textit{visual data} to refer to any pixel-data (images, videos,
+frames in a video, etc), and feature vectors (a.k.a. descriptors, or embeddings),
+which are representations of the pixel data.
+We use \textit{metadata} to refer to any data that is important within
+the context of specific applications.
+Information about a patient (name, last name, unique identifiers), or
+about a clinic (name, location, etc) are examples of what we refer to
+as \textit{metadata} in the context of a health-care IT system.
+One can think of \textit{metadata} as the data an application would store
+in rows and tables following the relational model.
+
+To illustrate the point on the need for multiple system to build the data
+infrastructure for applications,
+consider a ML developer constructing a pipeline for extracting
 brain tumor information from existing brain images in a classic medical
@@
 This requires assigning consistent identifiers for the scans and adding their
-metadata in a relational or key-value database\cite{kumar2017data}.
+\textit{metadata} in a relational or key-value database\cite{kumar2017data}.
+If the queries require a search over patient information,
+then patients are associated with their brain scans.
 If the queries require a search over patient information,
@@
 than the original, there is additional compute
-diverted towards pre-processing the original images which are typically larger.
+diverted towards pre-processing the original images.
 All these steps require understanding different software
 solutions that provide various functionalities that can then be stitched
-together with a script for this specific use case.
+together with many scripts for this specific use case.
 Moreover, if the pipeline identifies new metadata to be added for
@@
 subset of the data before running any processing, and a large effort
-is devoted to cleaning and pre-processing the data.
+is devoted to filtering, curating, and pre-processing the data.
 Selecting subsets of data is by itself a time consuming task,
@@
 usually end up building an ad-hoc solution that results in a
-combination of databases and file systems to store
+combination of databases and file servers to store
 metadata and visual data (images, videos), respectively \cite{sculley2015hidden}.
@@
 we use the YFCC100M dataset\cite{Thomee_2016}.
-The YFCC100M is the largest public multimedia collection.
+The YFCC100M dataset is the largest public multimedia collection.
 It contains the metadata of around 99.2 million photos
 and 0.8 million videos from Flickr,
-plus expansion packs that include a variety of multidimensional data,
-all of which were shared under one of the various Creative Commons licenses.
+plus expansion packs \cite{features} that include a variety of multidimensional
+data, all of which were shared under one of the various Creative Commons licenses.
 We have used this dataset for multiple proof of concepts and

### 2021-02-26 yfcc100m/paper/vldb2021/appendix-eval-plots.tex

@@
 In the case of \textit{q4} and \textit{q6}, the throughput for 50M (and 100M for \textit{q4})
-increase by a large margin which outperforms both baselines while in figure~\ref{fig:q_throughput_56}
+increase by a large margin which outperforms both baselines while, in figure~\ref{fig:q_throughput_56},
 PostgreSQL slightly performed better than VDMS. 
@@
 We also see how \textit{q3} and \textit{q5} have poor performances and scalability
-as the database size grows same as shown in Figure~\ref{fig:summary_postgresql}.
+as the database size grows, same as shown in Figure~\ref{fig:summary_postgresql}.
 Considering all other queries (\textit{q1}, \textit{q2}, \textit{q4}, and \textit{q6}), 

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex

@@
 each image/video identifier, and the list of 1,570 \textit{autotags}.
-Internally, that information is represented as a property graph,
+Internally, that information is represented as a graph,
 where we have one node for each image, one node for each tag
@@
 However, the time to build a database increased.
-We later changed  the buffer pool size to a multiple of the default value, i.e. 16x,
-which produced the best results for loading time.
+We later changed  the buffer pool size to be 16x the default value,
+where we saw the best performance.
 
@@
 for the tablespace of all databases.
+This was required, specially for 100M database, in order to make the queries
+complete execution without failures, and represented a big disadvantage
+in terms of the disk space footprint during query execution.
 
@@
 using the default parameters provided by the implementation.
+
+\subsubsection{Indexes:}
 For all systems, we created indexes over the
@@
 
-% \subsubsection{Database Building Time}
-
-% One of the first things we noticed is the difference in the time needed to
-% build each database, where VDMS outperforms MySQL by a large margin
-% but PostgreSQL outperforms VDMS by a large margin as well.
-% This analysis includes only the metadata, as the images are stored
-% in a shared filesystem. Figure~\ref{fig:db_time_size} illustrates the
-% comparison in time to build the databases, and how the speedup
-% is sustained as the database size grows.
-% Key difference in the build times are attributed to the low-level
-% implementation of how MySQL reads and stores data from the files, the
-% optimizations (increased InnoDB pool size, etc.) needed
-% to handle large datasets such as YFCC100M, and the efficiency of PMGD.
-% PostgreSQL Python API can efficiently copy data from files to the database
-% which provides an advantage over VDMS which must insert
-% multiple nodes/connections per image.
-% On average, it took MySQL around $\sim$4x longer to build each database than VDMS,
-% and PostgreSQL was around $\sim$21x faster than VDMS.
-
 \subsubsection{Database Storage Footprint}
 
-% Another important aspect to note is the storage footprint of the databases.
+The Graph Database used by VDMS (PMGD) was designed for performance,
+especially in environments where persistent memory is present.
+This design decision comes as a trade-off for storage footprint, which is
+noticeable in other PMGD evaluation we have run in the past.
+Because PMGD was not optimized for low storage footprint, authors feared that
+any improvement would come at the expenses of a significant increase on
+disk occupancy.
+We measured metadata disk footprint for that reason, and show the results
+in Figure~\ref{fig:db_sizes}.
 PostgreSQL and VDMS require comparable amount of storage for metadata which
@@
 each node/connection.
-The Graph Database internal to VDMS (called PMGD) was designed for performance,
-especially in environments where persistent memory is present.
-This design decision comes as a trade-off for storage footprint, which is
-noticeable in our results.
 VDMS required 37-41\% more storage than MySQL for storing the same amount

### 2021-02-26 yfcc100m/paper/vldb2021/conclusion.tex
Minor changes on eval and conclusions

@@
 together with the expansions packs that include
-machine-generated labels and feature vectors.
+machine-generated labels.
 We show how VDMS compares against a combination of
@@
 We see improvements up to 364x in certain queries,
-and an average improvement of about 85x when compared to PostgreSQL.
-When compared to MySQL, we see up to 96x speedup
+and an average improvement of about 85x when compared to PostgreSQL+Apache.
+When compared to MySQL+Apache, we see up to 96x speedup
 and and average improvement of 31x.
-% Finally, we analyze the different trade-offs of VDMS' 
-% descriptors indexes, a functionality 
-% that fully integrates with the rest of VDMS interface, 
-% providing a powerful and comprehensive API.
 The design of VDMS, which was conceived as a

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Minor changes on eval and conclusions

@@
 
-\textbf{MySQL-Based System:}
+\textbf{MySQL-Based System (\textit{mysql}):}
 Each MySQL database is created in a similar manner as VDMS
@@
 
-\textbf{PostgreSQL-Based System:}
+\textbf{PostgreSQL-Based System (\textit{postgresql}):}
 Each PostgreSQL database is created in the exact same manner as MySQL
@@
 
-In order to evaluate VDMS and the baseline systems on our use-case queries,
+In order to evaluate VDMS and the baseline systems for our use-case queries,
 we implemented 6 queries that filter and retrieve a specific set of images.
-We chose these queries because they represent typical use-cases where a
-cohort of images is to be retrieved and processed from a large corpus of data.
+We chose these queries because they are the same that we use when filtering a
+cohort of images to be retrieved and processed from a large corpus of data.
 As we mentioned before, we took this approach mainly because we wanted to
@@
 benchmarks that are oriented towards visual data retrieval.
+
+The implementation of this evaluation, as well as all the queries, are available
+open-source for reproducibility
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m},
+together with all the results of our evaluation
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}.
 
@@
 To evaluate the access to metadata and images,
-we use the following queries, modeled after our internal use cases:
+we use the following queries, based on our internal use cases:
 
@@
 of the \textit{connection} between the image and that \textit{autotag}),
-as well as in MySQL and PostgreSQL (in the form of a column in the
+as well as in \textit{mysql} and \textit{postgresql} (in the form of a column in the
 \textit{autotags} table that links images with tags).
@@
 the query involves JOIN operations between the 3 tables.
-The implementation of this evaluation, as well as all the queries, are available
-under the benchmarks branch of the VDMS project
-\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m},
-and some examples of the queries can be found in the appendix of this paper.
 
@@
 This way, we remove bottleneck introduced by network bandwidth that would
-otherwise over-complicate the understanding of the results.
+otherwise over-complicate the understanding of the results,
+and play in disadvantage for the baseline implementations.
 
@@
 applicability of visual data management systems like VDMS.
-Because of this, we put emphasis on the analysis of concurrency and throughput,
-rather than latency.
+Because of this, \textit{we put emphasis on the analysis of
+\textbf{concurrency and throughput}, rather than latency}.
 
@@
 described in the Experimental Setup Section.
-We show each database in different figures for readability reasons.
 The figures show the throughput (images per second) when retrieving
@@
 
-Figure~\ref{fig:concurrency_comparison_q2} illustrates a concurrency analysis for
-\textit{q2} (\textit{1tag\_geo\_resize}), described above,
+To analyze these results, one needs to compare the full-line (VDMS) versus the
+dotted-lines (two baselines).
+We start with a simple query that involves a simple
+metadata filtering over 1 tag and also geo-location,
+plus pre-processing operations.
+Figure~\ref{fig:concurrency_comparison_q2} illustrates a concurrency analysis
+for \textit{q2} (\textit{1tag\_geo\_resize}), described above,
 using VDMS and the baselines.
-Here we evaluate the scalability of all systems, as the number of concurrent
+Here we evaluate the concurrency of all systems, as the number of concurrent
 clients grows (x-axis) and as the size of the databases grow.
@@
 throughput after 16 threads except for the 100M case.
-In this case, PostgreSQL has a performance spike at 32 clients
+In this case, \textit{postgresql} has a performance spike at 32 clients
 and the performance for both baselines is less stable when compared
@@
 \caption{Concurrency Analysis for all the queries using the 10M database).
-We show queries in different figures for readability reasons.
 Each figure shows the throughput (images per second) when retrieving
@@
 concurrency for each of the six queries.
-It is obvious that VDMS consistently outperforms each of the baseline
+VDMS consistently outperforms each of the baseline
 systems for queries \textit{q1}, \textit{q2}, \textit{q4},
@@
 when it comes to concurrency.  The
-VDMS throughput continues to increase up to 56 concurrent clients, which is
-the hardware concurrency of the system, for all queries except \textit{q3}.
+VDMS throughput continues to significantly increase up to 56 concurrent clients,
+which is the hardware concurrency of the system, for all queries except \textit{q3}.
 In this query, the throughput seems to decrease after 32 clients
 and both baselines outperform VDMS after 8 concurrent clients.
-Prior to 8 clients, VDMS has a slight advantage over PostgreSQL which has
+Prior to 8 clients, VDMS has a slight advantage over \textit{postgresql} which has
 a major performance improvement at 16 concurrent clients. On the other hand,
-MySQL maintains its improvement in throughput over PostgreSQL and VDMS over all
-concurrent clients. In the case of \textit{q5}, the throughput of the baselines
+\textit{mysql} maintains its improvement in throughput over \textit{postgresql}
+and VDMS over all concurrent clients.
+In the case of \textit{q5}, the throughput of the baselines
 are less aggregated than those of VDMS up to 16 concurrent clients. When it comes
 to low concurrency, the baselines do as good and even better than VDMS with 2 or
-4 concurrent clients.  At 16 concurrent client, the performance of the MySQL baseline
-begin to stabilize while the throughput of PostgreSQL begins to degrade. However, as
-concurrency increases beyond 16 concurrent clients, the difference
+4 concurrent clients.
+At 16 concurrent client, the performance of the \textit{mysql} baseline
+begin to stabilize while the throughput of \textit{postgresql} begins to degrade.
+However, as concurrency increases beyond 16 concurrent clients, the difference
 in throughput becomes clear, with VDMS reaching its peak performance at
@@
 In the case of \textit{q4} and \textit{q6}, it is worth noting
-that for the OR operation, there is no need for 2 retrievals.
-Rather, a single retrieval is performed and the result filtered on the client.
-Future release will add more of such operators (AND, OR, etc) in order to prevent
-unnecessary retrievals and extra filtering on the client side.
+that for the OR operation, there is no need for 2 retrievals (2 transactions).
+Rather, a single transaction is performed and the result filtered on the client.
+Future VDMS releases will add more of such operators (AND, OR, etc)
+in order to prevent unnecessary retrievals and extra filtering on the client side.
 
@@
 described in the Experimental Setup Section.
-We show different number of concurrent clients in different figures for readability reasons.
 The experiments show the throughput as images per second for
@@
 
-In the case of \textit{q4} and \textit{q6}, we see the performance for MySQL
+In the case of \textit{q4} and \textit{q6}, we see the performance for \textit{mysql}
 degrades up to 64 concurrent clients, then there's an improvement with 112 concurrent clients.
@@
 As the number of clients and the size of database grows, the performance for this query
-drops and the performance of PostgreSQL becomes more comparable to VDMS. Lets look at the
+drops and the performance of \textit{postgresql} becomes more comparable to VDMS. Lets look at the
 case with 32 concurrent clients.
 In this case, the throughput of VDMS drops drastically with
-the 100M database and PostgreSQL outperforms by a small margin.
+the 100M database and \textit{postgresql} outperforms by a small margin.
 As the clients increase, we see this trade-off occurs with the 50M database
 instead of 100M while in the 100M database, VDMS outperforms both baselines.
-In this case, the throughput of PostgreSQL drastically degrades
+In this case, the throughput of \textit{postgresql} drastically degrades
 for the 100M database and the performance of the baseline systems are more comparable.
@@
 Many of the inefficiencies that come with combining tools that were designed
-for other use cases simply disappear when building a tool that treats
+without visual data in mind simply disappear when building a tool that treats
 visual entities as first class citizens, as it is the case of VDMS.
@@
 
-\subsubsection{Query Execution Analysis}
+\subsubsection{Scalability Analysis}
 
@@
 database size increases.
-These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged out of 5 runs.}
+These queries were run using 56 simultaneous clients,
+and averaged out of 10 runs.}
 \label{fig:q_throughput_56}
@@
 
-The next step in our analysis involved running a set of queries
-(described above in this section),
-to better understand the performance of the systems under different query conditions.
+The next step in our analysis involves looking more deeply at how
+the performance scales with the number of elements in the database.
 Figure~\ref{fig:q_throughput_56} shows the evaluation of the
@@
 database size increases in terms of number of images.
-These queries were run using 56 simultaneous clients (nthr = 56),
-and averaged over 5 runs (niter = 5).
+These queries were run using 56 simultaneous clients,
+and averaged over 10 runs.
 To analyze these plots, one needs to compare the full-line (VDMS) versus the
@@
 the performance degrades as well.
-For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS
+For this query, \textit{mysql} (and in some cases \textit{postgresql}) outperforms VDMS
 for databases larger than 1M.
@@
 (\textit{q5}) decreases the performance as the database sizes scales.
-The behaviour is different in the case of \textit{q6},
+The behavior is different in the case of \textit{q6},
 which also filters by geo-location.
@@
 In comparison to figure~\ref{fig:q_throughput_56},
-when using fewer concurrent clients (nthr = 16),
-the throughput for VDMS for queries \textit{q4} (\textit{2tag\_or\_resize}),
-\textit{q5} (\textit{2tag\_resize\_geo\_and}), and
-\textit{q6} (\textit{2tag\_resize\_geo\_or}) improved by a large margin
-for the 50M and 100M databases.
+when using fewer concurrent clients (16),
+the throughput for VDMS for queries \textit{q4}, \textit{q5}, and \textit{q6}
+improved by a large margin for the 50M and 100M databases.
 In the case of \textit{q4} and \textit{q6},
@@
 increase by a large margin which outperforms both baselines while in
-Figure~\ref{fig:q_throughput_56} PostgreSQL slightly performed better than VDMS.
+Figure~\ref{fig:q_throughput_56} \textit{postgresql} slightly performed better than VDMS.
 In the case of \textit{q5}, all systems perform a little differently for databases
@@
 Finally, Figure~\ref{fig:summary_postgresql} summarizes the results
-comparing VDMS and PostgreSQL. A comparison with MySQL is
-provided in Figure~\ref{fig:summary_mysql} of the appendix.
-In comparison to PostgreSQL, VDMS provides up to 364x speedup
+comparing VDMS and \textit{postgresql}.
+In comparison to \textit{postgresql}, VDMS provides up to 364x speedup
 (for the case of \textit{q1}),
-and an average improvement in throughput of about 85x. This is
-an impressive improvement over the two baseline systems.
+and an average improvement in throughput of about 85x.
+This is an impressive improvement over the two baseline systems.
 More importantly, we see that the speedup increases as the database size grows,
@@
 We also see how \textit{q3} and \textit{q5} have poor performances and scalability
-when compared to the baselines as the database size grows.
+when compared to the baselines as the database size grows, except in the
+case of the largest database (100M), where the baselines performed very poorly.
 This evaluation served the
@@
 re-purposing systems that were designed to handle types of date other than visual.
-VDMS, by design, eliminates all the inefficiencies that result
+VDMS, by design, eliminates most of the inefficiencies that result
 from a forced integration of components designed for a
@@
 Because of space constraints, we are only showing a subset of our evaluation.
-More details about our comprehensive evaluation are available.
-\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}
+More details about our comprehensive evaluation are available
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}.
 % TODO. Upload the final version of the plots once we are sure we are not making

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Minor update on eval
Minor fixes

@@
 It is worth noting that the images are stored in a shared repository
-(ext4 filesystem on a RAID 6 configuration of 16TB) that both
+(ext4 filesystem on a RAID 5 configuration of 16TB) that both
 Apache WebServer and VDMS have direct access.
@@
 More importantly, we see that the speedup grow as the database size increases,
-showing that VDMS scales better than the PostGreSQL baseline.}
+showing that VDMS scales better than the PostgreSQL+Apache baseline.}
 \label{fig:summary_postgresql}

### 2021-03-03 yfcc100m/python/eval/DBEvalFramework/Plotting.py
Fix issue with plotting

@@ -121,7 +121,7 @@ class Plotting(object):
         plt.xlabel(xlabel, fontsize=12)
         plt.ylabel(ylabel, fontsize=12)
 
-        if len(engines) > 2 or len(n_queries > 2):
+        if len(engines) > 2 or n_queries > 2:
             plt.legend(bbox_to_anchor=(0, -0.15),
                        loc="upper left",
                        # mode="expand",

### 2021-03-03 yfcc100m/python/eval/run_performance.sh
Fix run_performance

@@ -1,6 +1,7 @@
 #!/bin/bash
-NITER=5
-experiment_name=perf_all_threads_images
+
+NITER=10
+experiment_name=perf_results
 
 running_folder=running_experiment
 info=${running_folder}/info.txt
@@ -15,23 +16,21 @@ echo $experiment_name "started..." >> $info
 date >> $info
 
 for th in 2 4 8 16 32 56 64 112
-# for th in 4 8 16
 do
-    # for size in 1M 5M 10M 50M
     for size in 1M 5M 10M 50M 100M
     do
-        for db in vdms postgresql
+        for db in vdms mysql postgresql
         do
-            # Run VDMS Queries
             echo "Running $th $db ${size}..." >> $info
+
             python3 performance.py \
                     -db_type=$db \
                     -db_name=$size \
                     -numthreads=$th \
                     -numiters=$NITER \
                     -out_folder=$running_folder \
-                    >> $running_folder/log.log
-                    2>> $running_folder/log_error.log
+                    >>  $running_folder/log.log \
+                    2>> $running_folder/log.log
         done
     done
 done

### 2021-03-03 yfcc100m/paper/vldb2021/design.tex

@@
 "properties" of the node in the graph.
-Users can query, filter, and retrieve this objects based on its properties.
+Users can query, filter, and retrieve these objects based on its properties.
 VDMS does not simply treat these objects as binary blobs of data, but rather
@@
 or retrieve only a clip from a large video.
-This is one of the main differentiation aspect when compared to other
+This is one of the main differentiating aspects when compared to other
 database systems, including relational databases.

### 2021-04-22 yfcc100m/paper/vldb2021/reviewer_response.txt
added file for reviewer responses and fixed typos from reviewer 5

Reviews: Point addressed
========================================
Reviewer #1

Reviewer #4

Reviewer #5
    D4: Fixed all typos

### 2021-04-22 yfcc100m/paper/vldb2021/eval.tex
added file for reviewer responses and fixed typos from reviewer 5

@@
 visual data management system as baselines,
-desing specifically for the image search use case,
+design specifically for the image search use case,
 and comprised of a combination of widely available, off-the-shelf components.
@@
 where data is represented as three tables:
-\textit{images}, \textit{taglist}, and \textit{taglist}.
+\textit{images}, \textit{taglist}, and \textit{autotags}.
 PostgreSQL works well out-of-box and optimizations were not necessary,
@@
 \includegraphics[width=\columnwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
-\caption{Concurrency Analysis for all the queries using the 10M database).
+\caption{Concurrency Analysis for all the queries using the 10M database.
 Each figure shows the throughput (images per second) when retrieving

### 2021-04-22 yfcc100m/paper/vldb2021/intro.tex
minor edits to intro
minor edit to intro
added file for reviewer responses and fixed typos from reviewer 5

@@
 
-To illustrate the point on the need for multiple system to build the data
+To illustrate the point on the need for multiple systems to build the data
 infrastructure for applications,
@@
 \textit{metadata} in a relational or key-value database\cite{kumar2017data}.
-If the queries require a search over patient information,
-then patients are associated with their brain scans.
 If the queries require a search over patient information,

### 2021-04-28 yfcc100m/python/eval/DBEvalFramework/EvalFramework.py
Remove DBEvalFramework, not it is a separate package

@@ -1,594 +1,0 @@
-import os
-
-import numpy as np
-from numpy import inf
-from numpy import nan
-
-import pandas as pd
-
-from DBEvalFramework import Plotting
-
-class EvalFramework(object):
-
-    def __init__(self, experiment_name):
-
-        self.experiment_name = experiment_name
-
-        try:
-            self.data = pd.read_csv(experiment_name + ".csv", index_col=0)
-        except:
-            print("File not found, creating a new onself...")
-            self.data = pd.DataFrame(columns=[
-                    "query", "engine", "db_size",
-                    "n_threads",
-                    "n_samples",
-                    "query_time_avg", "query_time_std",
-                    "n_results", "n_results_std",
-                    ])
-
-        self.export_to_csv()
-
-    def add_row(self, query, engine, db_size,
-            n_threads,
-            n_samples,
-            query_time_avg, query_time_std,
-            n_results, n_results_std):
-
-        self.data.loc[len(self.data)] = [
-                query,
-                engine,
-                db_size,
-                n_threads,
-                n_samples,
-                query_time_avg,
-                query_time_std,
-                n_results,
-                n_results_std,
-            ]
-
-    def clear(self):
-
-        self.data = pd.DataFrame(columns=[
-                    "query", "engine", "db_size",
-                    "n_threads",
-                    "n_samples",
-                    "query_time_avg", "query_time_std",
-                    "n_results", "n_results_std",
-                    ])
-
-        self.export_to_csv()
-
-    def export_to_csv(self):
-
-        self.data.to_csv(self.experiment_name + ".csv")
-
-    def get_unique(self, column):
-
-        arr = []
-        vals = self.data.loc[:,column]
-
-        # arr = [v for v in vals if v not in arr]
-        for val in vals:
-            if val not in arr:
-                arr.append(val)
-
-        return arr
-
-    def get_arr_for_eng_q_dbsize(self, engine, query, db_size, key):
-
-        arr = self.data[self.data["query"] == query]
-        arr = arr[self.data["engine"]      == engine]
-        arr = arr[self.data["db_size"]     == db_size]
-        arr = arr[key].to_numpy()
-
-        return arr
-
-    def get_arr_for_eng_q_threads(self, engine, query, threads, key):
-
-        arr = self.data[self.data["query"] == query]
-        arr = arr[self.data["engine"]      == engine]
-        arr = arr[self.data["n_threads"]   == threads]
[…]

### 2021-04-28 yfcc100m/python/eval/DBEvalFramework/Plotting.py

@@ -1,291 +1,0 @@
-import csv
-import os
-import re
-import argparse
-from pathlib import Path
-
-# from cycler import cycler
-import numpy as np
-import matplotlib
-
-matplotlib.use('pdf')
-import matplotlib.pyplot as plt
-from matplotlib.backends.backend_pdf import PdfPages
-
-
-color      = ['red', 'blue', 'orange', 'black', 'pink', 'brown', 'violet', 'green']
-linestyles = ['-', '--', '-.']
-markers    = ['o', '*', 'd']
-patterns = [ "" ,"" , "", "", "\\", ".", "*",  "\\" , "|" , "-" ]
-
-class Plotting(object):
-
-    def __init__(self):
-
-        return
-
-    def isfloat(self, value):
-        try:
-            float(value)
-            return True
-        except ValueError:
-            return False
-
-    def instr2bool(self, in_value):
-        if in_value.lower() in ['true', 't']:
-            return True
-        else:
-            return False
-
-    def value_to_float(self, x):
-        if type(x) == float or type(x) == int or type(x) == np.int64:
-            return x
-
-        if 'k' in x:
-            if len(x) > 1:
-                return float(x.replace('k', '')) * 1000
-            return 1000.0
-        if 'M' in x:
-            if len(x) > 1:
-                return float(x.replace('M', '')) * 1000000
-            return 1000000.0
-        return 0.0
-
-
-    # This will plot query time (in seconds) as a function of
-    # database size for all the queries.
-    # This will name que queries using engine_query for all the queries.
-    # values is in the format:
-    #   query0_dbsize0_time,query0_dbsize0_std, ... , query0_dbsizeN_time,query0_dbsizeN_std
-    def plot_lines_all(self, queries, db_sizes, engines, values,
-                        log="y",
-                        title="Query Time",
-                        filename="plot_unnamed.pdf",
-                        xlabel="Database Size",
-                        ylabel="None"):
-
-        # print(len(values))
-        # print(len(values[1:]))
-        # print(values)
-        if (len(db_sizes) * 2) != len(values[1,:]):
-            print("Error input size")
-            print("db_sizes:", db_sizes)
-            print("values:", values)
-            return
-
-        if len(queries) * len(engines) != len(values[:,1]):
-            print("Error input size")
-            print("queries:", queries)
-            print("values:", values)
-            return
-
-        n_queries = len(queries)
-
-        # local_markers = []
-        # local_linesty = []
[…]

### 2021-04-28 yfcc100m/paper/vldb2021/futurework.tex
Add future work section

\section{Future Work}

Add future work here, responding reviewrs questions.

### 2021-04-28 yfcc100m/paper/vldb2021/design.tex
Change font of API snippets

@@
               xleftmargin=21pt,
+              fontsize=\footnotesize,
               tabsize=4]{js}
@@
     "format": "jpg"
-}
+},
 "AddConnection": {
@@
               xleftmargin=21pt,
+              fontsize=\footnotesize,
               tabsize=4]{js}
 "FindEntity"{
+    "_ref" : 1,
     "class": "autotag",
@@
     }
-    "_ref" : 1
 },

### 2021-04-28 yfcc100m/paper/vldb2021/eval.tex
Add resize comparison plot to eval
Move figures around to make them fit better
Change font of API snippets

@@
 
+\begin{figure}[ht!]
+\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
+\caption{Throughput Analysis using all queries from our use-case
+described in the Experimental Setup Section.
+We show queries in different figures for readability reasons.
+The experiments show the performance of all systems (VDMS and the two baselines) as the
+database size increases.
+These queries were run using 16 simultaneous clients, and averaged out of 10 runs.}
+\label{fig:q_throughput_16}
+\end{figure}
+
+%% ========= Resize Comparison
+
+\begin{figure}[h!]
+% WAS: begin{figure*}[ht]; textwidth
+\includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput_resize}
+\caption{Throughput Analysis using a subset of the queries, with and without resize.}
+\label{fig:q_throughput_56_resize}
+\end{figure}
+
+\begin{figure*}[ht!]
+\centering
+\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_postgresql}
+\caption{Summary of performance gains for all queries.
+We see up to 364x speedup (\textit{q1}), and an average of about 85x.
+More importantly, we see that the speedup grow as the database size increases,
+showing that VDMS scales better than the PostgreSQL+Apache baseline.}
+\label{fig:summary_postgresql}
+\end{figure*}
+
 From the first 2 queries, as well as \textit{q4} and
@@
 
-\begin{figure}[ht!]
-\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
-\caption{Throughput Analysis using all queries from our use-case
-described in the Experimental Setup Section.
-We show queries in different figures for readability reasons.
-The experiments show the performance of all systems (VDMS and the two baselines) as the
-database size increases.
-These queries were run using 16 simultaneous clients (nthr = 16),
-and averaged out of 5 runs.}
-\label{fig:q_throughput_16}
-\end{figure}
-
 %% ========= Queries Analysis - END
@@
 %% ========= Summary Analysis
-
-\begin{figure*}[ht!]
-\centering
-\includegraphics[width=\textwidth]{figures/plot_th_56_query_times_speedup_postgresql}
-\caption{Summary of performance gains for all queries.
-We see up to 364x speedup (\textit{q1}), and an average of about 85x.
-More importantly, we see that the speedup grow as the database size increases,
-showing that VDMS scales better than the PostgreSQL+Apache baseline.}
-\label{fig:summary_postgresql}
-\end{figure*}
 

### 2021-04-28 yfcc100m/paper/vldb2021/reviewer_response.txt
Updates for second reviewer

@@
 Reviewer #4
+    D1. The proposed solution seems to use the functionalities of the existing systems as it is. Authors should provide
+    more original ideas or more innovative designs. For example, providing optimizations that specifically target visual
+    data or query patterns associated could have been more meaningful to the research communities and convince
+    users to replace their customized solution to VDMS.
+        - Partially addressed in D3
+
+    D2. Both baselines are similar in architecture, and the evaluation has not effectively highlighted the performance
+    gain of other design choices authors made in VDMS, except that the VDMS server process both metadata operation
+    and image operation within only one round of communication with clients. For example, benchmarking systems can
+    be built to test in isolation the improvement on metadata handling.
+        - Should we include metadata only results?
+
+    D3. The paper should detail the types of queries, and image preprocessing it supports, as this being “one of the
+    most differentiating aspects” as mentioned by them. And for the selection of operations supported, authors need to
+    argue for their decision in the application contexts.
+        - In Section 3: "VDMS API also allows users to insert application-defined "Entities", ... In some cases,
+            users may need to apply different processing operations to the visual data
+            for their application. We chose to support
+            specific operations due to their frequent use in ML applications using
+            visual data. Common operations supported for both images and videos are
+            thresholding, cropping, and resizing. Additional operations such as
+            flipping and rotation are supported for images and extracting
+            frames by interval is supported for videos.
+            In most ML applications, a subset of these operations are included in
+            the preprocessing stage of model training and/or inferencing which supports
+            majority of the research communities who analyze visual data and may benefit
+            from replacing customized solutions with VDMS."
+
+    D4. Data scientists commonly perform data exploration on raw data and interactively curate their datasets for
+    downstream tasks like model training. VDMS could design new mechanisms to support interactive analysis, which
+    would be beneficial.
+        - Out of scope of paper, would require new API to allow interactive analysis
+
+    D5. In terms of scalability, the paper only shows that VDMS can handle large datasets, yet it is not known how
+    VDMS can scale horizontally with more computation resources. For more complex logic of image preprocessing,
+    VDMS will suffer from its single server capacity.
+    - Sounds more like distributed VDMS to me
+    - Any experiements comparing more complex image processing operations?
 
 Reviewer #5
+    D3: The experiments section seems unnecessarily long, yet missing some important details.
+All the experiments are performed on images and do not handle videos or feature values at all.
+For example, video operations may include interval operations or video clip retrieval.
+Some parts seem redundant, e.g., the description of Figure 9 seems near-identical to that of Figure 8.
+        - POssibly remove figure 8 or 9 and use space for either PMGD (D2) or something else
+
     D4: Fixed all typos
+        -Page 1: “need for multiple system” -> “need for multiple systems”
+        -Page 5: “desing specifically” -> “designed specifically”
+        -Page 7: “images, taglist, and taglist” -> “images, taglist, and autotags”
+        -Page 10, Figure 6: “10M database).” -> “10M database”

### 2021-04-28 yfcc100m/paper/vldb2021/design.tex
Updates for second reviewer

@@
 providing the ability to run processing on-the-fly, both at insertion and
-retrieval time.
-Users can, for instance, retrieve a resized version of an image,
-or retrieve only a clip from a large video.
-This is one of the main differentiating aspects when compared to other
+retrieval time. This is one of the main differentiating aspects when compared to other
 database systems, including relational databases.
@@
 Later, the user can retrieve the Entity object corresponding to a person,
-together with all the images \textit{connected} to it.
+together with all the images \textit{connected} to it. In some cases,
+users may need to apply different processing operations to the visual data
+for their application. We chose to support
+specific operations due to their frequent use in ML applications using
+visual data. Common operations supported for both images and videos are
+thresholding, cropping, and resizing. Additional operations such as
+flipping and rotation are supported for images and extracting
+frames by interval is supported for videos.
+In most ML applications, a subset of these operations are included in
+the preprocessing stage of model training and/or inferencing which supports
+majority of the research communities who analyze visual data and may benefit
+from replacing customized solutions with VDMS.
 

### 2021-04-28 yfcc100m/paper/vldb2021/reviewer_response.txt
Add crash answer for pmgd in section 3.2 and expanded benefit of no schema in section 3.1

@@
         - Partially addressed in D3
+        - Section 3.1: "...schema of objects(and its properties) before hand. This is a benefit over relational
+            databases which require the user to identify how data is divided into
+            different tables and determine the relationship between tables."
 
@@
         - Out of scope of paper, would require new API to allow interactive analysis
+        - Future work?
 
@@
 Reviewer #5
+    D1: VDMS is a valuable contribution to Database systems where it takes some first important steps to handle visual data as a first-class citizen instead of considering it a blob of information.
+However, the description of VDMS is rather short (only 4 pages), and some critical technical details are missing.
+In particular, what are the technical challenges for supporting ACID properties using the graph model abstraction?
+Are the solutions analogous to RDBMS’s?
+Does VDMS also support crash recovery like an RDBMS?
+What are the key challenges for implementing the new visual object computations?
+        - Section 3.2: "...(ACID) properties expected from databases.
+            In cases where the system may crash or fail, PMGD has the ability
+            to recover to a consistent state without loading its content into memory.
+            Unlike existing graph databases that are disk-based, our graph engine
+            avoids converting data from a disk-friendly format to an usable format for an application."
+
     D3: The experiments section seems unnecessarily long, yet missing some important details.

### 2021-04-28 yfcc100m/paper/vldb2021/design.tex
Add crash answer for pmgd in section 3.2 and expanded benefit of no schema in section 3.1

@@
 at insertion time, without any need to define a schema of objects
-(and its properties) before hand.
+(and its properties) before hand. This is a benefit over relational
+databases which require the user to identify how data is divided into
+different tables and determine the relationship between tables.
 
@@
 (ACID) properties expected from databases.
+In cases where the system may crash or fail, PMGD has the ability
+to recover to a consistent state without loading its content into memory.
+Unlike existing graph databases that are disk-based, our graph engine
+avoids converting data from a disk-friendly format to an usable format for an application.
 PMGD comes as a C++ library that VDMS uses for implementing its query engine.

### 2021-04-28 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex

\documentclass[11pt]{proposalnsf}
\usepackage{times}
\usepackage{mathptmx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{array}
\usepackage[noend,nofillcomment]{algorithm2e}
\usepackage{verbatim}
%\usepackage{algorithmic}
%\usepackage{amsfonts}
%\usepackage{lineno}
\usepackage{ifthen}
\usepackage{booktabs}

\usepackage{color}

\usepackage{multirow}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

\makeatletter
\def\tightitemize{\ifnum \@itemdepth >3 \@toodeep\else \advance\@itemdepth \@ne
\edef\@itemitem{labelitem\romannumeral\the\@itemdepth}%
\list{\csname\@itemitem\endcsname}{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\def\makelabel##1{\hss\llap{##1}}}\fi}
\let\endtightitemize =\endlist
\def\tightenumerate{%
\ifnum \@enumdepth >\thr@@\@toodeep\else
\advance\@enumdepth\@ne
\edef\@enumctr{enum\romannumeral\the\@enumdepth}%
\expandafter
\list
\csname label\@enumctr\endcsname
{\setlength{\topsep}{-\parskip}\setlength{\parsep}{0in}\setlength{\itemsep}{0in}\setlength{\parskip}{0in}\usecounter\@enumctr\def\makelabel##1{\hss\llap{##1}}}%
\fi}
\let\endtightenumerate =\endlist
\makeatother
\renewcommand{\refname}{\centerline{References cited}}

% this handles hanging indents for publications
\def\rrr#1\\{\par
\medskip\hbox{\vbox{\parindent=2em\hsize=6.12in
\hangindent=4em\hangafter=1#1}}}

\def\baselinestretch{1}

\begin{document}

\begin{center}
{\Large{\bf VLDB 2021 - Notes for Reviewers}}

\end{center}

Here we address rebviews comments 1 by one.

\newpage
\pagenumbering{arabic}

\end{document}

### 2021-04-28 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Update response doc

@@
 \begin{center}
-{\Large{\bf VLDB 2021 - Notes for Reviewers}}
+{\Large{\bf VLDB 2021 - Response to Reviews: \\
+Using VDMS to Index and Search 100M Images}}
 
@@
 
-Here we address rebviews comments 1 by one.
+\section{Reviewer \#1}
+
+\textbf{
+D1 (W1): Although VDMS mentions that it supports Visual Feature Vectors as
+first class citizens in Section 3.3, the queries used in evaluation
+do not include an example.
+}\bigskip
+
+Not for this work, we will add to future work.
+Also, building a comparison baseline that integrates all that functionality
+is very ad-hoc, and would complicate the understanding of the origins
+of the benefits of VDMS.
+
+We can provide appendix.
+
+\bigskip
+\textbf{
+D2 (W1): All queries involve “resize” operation, which is expensive for the
+baselines since it is happening on the client side.
+A simplistic query without resizing can show how the graph model performs
+against traditional RDS DBs (i.e., will show that VDMS has comparable
+performance even for such use-case without "visual manipulation").
+}\bigskip
+
+This is a great point, we appreciate the suggestion.
+We have added
+
+\bigskip
+\textbf{
+D3 (W2): It is not clear which queries are supported or not.
+The queries used for evaluation are based on tags and geo search.
+What about visual-based queries? For example, is image similarity search based
+on Visual Feature Vectors within a geographical region supported?
+}\bigskip
+
+VDMS do support the suggested query, i.e., perform a similarity search based
+on feature vectors and also filter by geographical location.
+This can be achieved using the very similar queries as the one used in
+the present evaluation, plus the use of the feature vector queries,
+that are natively supported using the same set of APIs.
+Because of space constraints, we were not able to accommodate initial evaluation
+we have been working on with Feature Vector search specifically,
+and we plan to perform a more comprehensive evaluation on this functionality
+of this feature because of its relevance when processing visual data, as well
+as our own needs in our use cases.
+
+\bigskip
+\textbf{
+D4 (W2): Section 4.3.1, paragraph 5, mentions that INTERSECTION operation is not
+yet implemented at the server side.
+A discussion on current limitations/future work for such queries or queries
+such as in D3, will be better to be clarified in paper.
+}\bigskip
+
+Thanks for this comment, we have added a "Future Work" section where we cover
+this and other question, expanding further on the most immediate lines of
+improvement and research for the VDMS project.
+
+\bigskip
+\textbf{
+D5 (W3): Section 4.2 paragraph 1, desing
+}\bigskip
+
+Fixed in final version.
+
+\bigskip
 
 \newpage
+\section{Reviewer \#4}
+
+\newpage
+\section{Reviewer \#5}
+
 \pagenumbering{arabic}

### 2021-04-28 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Add mention to union/intersetction operations in future work

@@
 
-This is a great point, we appreciate the suggestion.
-We have added
+This is a great point, we very much appreciate the suggestion.
+We have run a new experiment and added a new figure to the paper (Figure 10),
+where we compare 3 of the queries with and without the resize operation
+to compare the results.
+We have also added an explanation on the differences in performance
+in the final version.
 
@@
 
-VDMS do support the suggested query, i.e., perform a similarity search based
+VDMS does support the suggested query, i.e., perform a similarity search based
 on feature vectors and also filter by geographical location.
@@
 Thanks for this comment, we have added a "Future Work" section where we cover
-this and other question, expanding further on the most immediate lines of
+this and D3, expanding further on the most immediate lines of
 improvement and research for the VDMS project.

### 2021-04-28 yfcc100m/paper/vldb2021/futurework.tex
Add mention to union/intersetction operations in future work

@@
 
-Add future work here, responding reviewrs questions.
+Based on the results of this evaluation and our experience with VDMS
+for this and other internal use cases, we have identified certain features
+that are needed to fully exploit the potential of the unified back-end
+for visual data infrastructure.
+In particular, we have identified the urgent need for a set of UNION/INTERSECTION
+interfaces to enable richer queries and improve performance,
+given the limitations discussed on \ref{concurrency_analysis}.

### 2021-04-30 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Update review response files

@@
 
+\bigskip
+\textbf{
+D1.
+The proposed solution seems to use the functionalities of the existing systems as it is.
+Authors should provide more original ideas or more innovative designs.
+For example, providing optimizations that specifically target visual data or query patterns associated could have been more meaningful to the research communities and convince users to replace their customized solution to VDMS.
+}\bigskip
+
+We partially address this in D3 and also elaborated more on the benefits of using VDMS over relational databases in Section 3.1.
+
+\bigskip
+\textbf{
+D2.
+Both baselines are similar in architecture, and the evaluation has not effectively highlighted the performance gain of other design choices authors made in VDMS, except that the VDMS server process both metadata operation and image operation within only one round of communication with clients.
+For example, benchmarking systems can be built to test in isolation the improvement on metadata handling.
+}\bigskip
+
+Thank you for the input.
+To demonstrate the benefits of VDMS more clearly, we have added Figure 10 which compares queries with and without the resize operation.
+Without the resize operation, the evaluation demonstrates the improvement on metadata handling better.
+
+\bigskip
+\textbf{
+D3.
+The paper should detail the types of queries, and image preprocessing it supports, as this being “one of the most differentiating aspects” as mentioned by them.
+And for the selection of operations supported, authors need to argue for their decision in the application contexts.
+}\bigskip
+
+Thank you for this recommendation.  We have provided additional details on which operations we support and our decision for chosing the operations in section 3.
+
+\bigskip
+\textbf{
+D4.
+Data scientists commonly perform data exploration on raw data and interactively curate their datasets for downstream tasks like model training.
+VDMS could design new mechanisms to support interactive analysis, which would be beneficial.
+}\bigskip
+
+Interactive analysis is out of scope for this paper but this can easily be done using the current implementation of VDMS.
+To support interactive analysis, a new API/frontend could be implemented on top of VDMS as a webservice without adding any additional abstractions.
+An example of such frontend can be found <insert URL>.
+
+\bigskip
+\textbf{
+D5.
+In terms of scalability, the paper only shows that VDMS can handle large datasets, yet it is not known how VDMS can scale horizontally with more computation resources.
+For more complex logic of image preprocessing, VDMS will suffer from its single server capacity.
+}\bigskip
+
+Thanks for this comment, we have added a "Future Work" section where the need for this capability in a distributed version of VDMS.
+
 \newpage
 \section{Reviewer \#5}
+
+\bigskip
+\textbf{
+D1: VDMS is a valuable contribution to Database systems where it takes
+some first important steps to handle visual data as a first-class citizen
+instead of considering it a blob of information. However, the description
+of VDMS is rather short (only 4 pages), and some critical technical
+details are missing. In particular, what are the technical challenges
+for supporting ACID properties using the graph model abstraction? Are
+the solutions analogous to RDBMS’s? Does VDMS also support crash
+recovery like an RDBMS? What are the key challenges for implementing
+the new visual object computations?
+}\bigskip
+
+Thank you for the comment.  In Section 3.2, we have expanded on the crash recovery of PMGD in comparison to other graph databases.
+
+\bigskip
+\textbf{
+D2: The coverage of persistent memory is underwhelming because persistent memory is never used in the experiments.
+What is the key advantage of using persistent memory and how is it more suitable for visual data?
+Why not add experiments comparing VDMS with and without using PMGD?
+}\bigskip
+
+Due to page limitations and the use of SSDs instead of persistent memory in the evaluation, we did not elaborate on advantages of PMGD and related experiments.
+
+\bigskip
+\textbf{
+D3: The experiments section seems unnecessarily long, yet missing some important details.
+All the experiments are performed on images and do not handle videos or feature values at all.
+For example, video operations may include interval operations or video clip retrieval.
+Some parts seem redundant, e.g., the description of Figure 9 seems near-identical to that of Figure 8.
+}\bigskip
+
+The paper is focused on an image search application so we did not include any experiments on videos or feature vectors.
+We have experiments that we can provide in appendix but due to page limitations, it did not fit the scope of this paper.
+
+Figure 8 and 9 are quite similar but provide the performance difference with different numer of concurrent clients.
+We have removed figure 9 and replaced with a figure demonstrating the difference between queries with and without the resize operation.
+
+\bigskip
+\textbf{
+D4: Fixed all typos
+Page 1: “need for multiple system” -> “need for multiple systems”
+Page 5: “desing specifically” -> “designed specifically”
+Page 7: “images, taglist, and taglist” -> “images, taglist, and autotags”
+Page 10, Figure 6: “10M database).” -> “10M database”
+}\bigskip
+
+All typos are fixed in the final version.
 

### 2021-05-03 yfcc100m/paper/vldb2021/eval.tex
Add description for resize different, remove figure 9 (16th analysis)

@@
 
-\begin{figure}[ht]
-% WAS: begin{figure*}[ht]; textwidth
-\includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput}
-\caption{Throughput Analysis using all queries from our use-case
-described in the Experimental Setup Section.
-We show queries in different figures for readability reasons.
-The experiments show the performance of all systems (VDMS and the two baselines) as the
-database size increases.
-These queries were run using 56 simultaneous clients,
-and averaged out of 10 runs.}
-\label{fig:q_throughput_56}
-\end{figure}
+% \begin{figure}[ht]
+% % WAS: begin{figure*}[ht]; textwidth
+% \includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput}
+% \caption{Throughput Analysis using all queries from our use-case
+% described in the Experimental Setup Section.
+% We show queries in different figures for readability reasons.
+% The experiments show the performance of all systems (VDMS and the two baselines) as the
+% database size increases.
+% These queries were run using 56 simultaneous clients,
+% and averaged out of 10 runs.}
+% \label{fig:q_throughput_56}
+% \end{figure}
 
@@
 
-Figure~\ref{fig:q_throughput_16} shows the measured throughput,
-as images per second, delivered by each system for all queries and databases.
-In comparison to figure~\ref{fig:q_throughput_56},
-when using fewer concurrent clients (16),
-the throughput for VDMS for queries \textit{q4}, \textit{q5}, and \textit{q6}
-improved by a large margin for the 50M and 100M databases.
-In the case of \textit{q4} and \textit{q6},
-the throughput for 50M (and 100M for \textit{q4})
-increase by a large margin which outperforms both baselines while in
-Figure~\ref{fig:q_throughput_56} \textit{postgresql} slightly performed better than VDMS.
-In the case of \textit{q5}, all systems perform a little differently for databases
-larger than 5M.
-The gaps between the throughput of each system narrows at 10M.
-At a database size of 50M, the performance of the baselines drastically drop
-while the throughput of VDMS has a large improvement in comparison.
+% Figure~\ref{fig:q_throughput_16} shows the measured throughput,
+% as images per second, delivered by each system for all queries and databases.
+% In comparison to figure~\ref{fig:q_throughput_56},
+% when using fewer concurrent clients (16),
+% the throughput for VDMS for queries \textit{q4}, \textit{q5}, and \textit{q6}
+% improved by a large margin for the 50M and 100M databases.
+% In the case of \textit{q4} and \textit{q6},
+% the throughput for 50M (and 100M for \textit{q4})
+% increase by a large margin which outperforms both baselines while in
+% Figure~\ref{fig:q_throughput_56} \textit{postgresql} slightly performed better than VDMS.
+% In the case of \textit{q5}, all systems perform a little differently for databases
+% larger than 5M.
+% The gaps between the throughput of each system narrows at 10M.
+% At a database size of 50M, the performance of the baselines drastically drop
+% while the throughput of VDMS has a large improvement in comparison.
+
+We designed VDMS with the idea in mind that pre-processing operations will be
+commonly performed, as visual data is generally pre-processed before analytics
+computation are performed on them.
+Figure \ref{fig:q_throughput_56_resize} shows how the query throughput varies
+when a pre-processing operations is applied (in this case, a resize operation).
+The first thing to notice is that a resize operation is a significant part of
+the overall operation (retrieval + preprocessing).
+Doing a resize operation as part of the retrieval process
+brings the throughput orders of magnitude down.
+This is the case for both VDMS and the baseline.
+Nonetheless, VDMS is orders of magnitude faster when retrieving images
+both with and without resize operations.
+An important thing to notice is that, even if a resized image is smaller,
+and thus transfered faster between the server and the client,
+transferring the image as is (without resize) is faster in this case.
+This is a characteristic of the dataset, which has relatively small images
+(the average image is about 150KB in size).
+We have seen on other applications and datasets with higher resolution images
+that transferring resized (or cropped) images is significantly faster than
+transferring images as they are stored.
+An interesting aspect to notice is that, when removing the resize operations,
+differences between the 2 baseline systems become more noticeable.
+For instance, for an image retrieval based on a single tag, we see how
+\textit{postgresql} outperforms \textit{mysql}, but when adding geo-location
+filtering, we see \textit{mysql} showing a small advantage.
+When the resize operations is performed, both baselines do very similar
+on 1tag based queries, and orders of magnitude worst than VDMS.
+Note that in the baseline systems, the resize happens on the client side,
+rather than on the server side.
+This is the common case in most analytics applications, where pre-processing
+operations happen on the client side (or when the analytics computation happens),
+rather than near where the data is located, as it is the case with VDMS.
+We kept both the server and client systems the same to avoid the computation
+capabilities difference complicating our understanding of the results.
 

### 2021-05-03 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Update response to reviewrs

@@
 This is a great point, we very much appreciate the suggestion.
-We have run a new experiment and added a new figure to the paper (Figure 10),
+We have run a new experiment and added a new figure to the paper (Figure 9),
 where we compare 3 of the queries with and without the resize operation
@@
 \textbf{
-D1.
-The proposed solution seems to use the functionalities of the existing systems as it is.
-Authors should provide more original ideas or more innovative designs.
-For example, providing optimizations that specifically target visual data or query patterns associated could have been more meaningful to the research communities and convince users to replace their customized solution to VDMS.
+D1. The proposed solution seems to use the functionalities of the existing systems
+as it is. Authors should provide more original ideas or more innovative designs.
+For example, providing optimizations that specifically target visual data or query
+patterns associated could have been more meaningful to the research communities and
+convince users to replace their customized solution to VDMS.
 }\bigskip
 
-We partially address this in D3 and also elaborated more on the benefits of using VDMS over relational databases in Section 3.1.
+We partially address this in D3 and also elaborated more on the benefits of
+using VDMS over relational databases in Section 3.1.
 
@@
 \textbf{
-D2.
-Both baselines are similar in architecture, and the evaluation has not effectively highlighted the performance gain of other design choices authors made in VDMS, except that the VDMS server process both metadata operation and image operation within only one round of communication with clients.
-For example, benchmarking systems can be built to test in isolation the improvement on metadata handling.
+D2. Both baselines are similar in architecture, and the evaluation has not effectively
+highlighted the performance gain of other design choices authors made in VDMS,
+except that the VDMS server process both metadata operation and image operation
+within only one round of communication with clients.
+For example, benchmarking systems can be built to test in isolation the
+improvement on metadata handling.
 }\bigskip
 
-Thank you for the input.
-To demonstrate the benefits of VDMS more clearly, we have added Figure 10 which compares queries with and without the resize operation.
-Without the resize operation, the evaluation demonstrates the improvement on metadata handling better.
+Thank you for the input. To demonstrate the benefits of VDMS more clearly,
+we have added Figure 9 which compares queries with and without the resize operation.
+Without the resize operation, the evaluation demonstrates the
+improvement on metadata handling better.
 
@@
 \textbf{
-D3.
-The paper should detail the types of queries, and image preprocessing it supports, as this being “one of the most differentiating aspects” as mentioned by them.
-And for the selection of operations supported, authors need to argue for their decision in the application contexts.
+D3. The paper should detail the types of queries, and image preprocessing it supports,
+as this being “one of the most differentiating aspects” as mentioned by them.
+And for the selection of operations supported, authors need to argue for their
+decision in the application contexts.
 }\bigskip
 
-Thank you for this recommendation.  We have provided additional details on which operations we support and our decision for chosing the operations in section 3.
+Thank you for this recommendation.
+We have provided additional details on which operations we support and our
+decision for chosing the operations in section 3.
 
@@
 \textbf{
-D4.
-Data scientists commonly perform data exploration on raw data and interactively curate their datasets for downstream tasks like model training.
-VDMS could design new mechanisms to support interactive analysis, which would be beneficial.
+D4. Data scientists commonly perform data exploration on raw data and interactively
+curate their datasets for downstream tasks like model training.
+VDMS could design new mechanisms to support interactive analysis,
+which would be beneficial.
 }\bigskip
 
-Interactive analysis is out of scope for this paper but this can easily be done using the current implementation of VDMS.
-To support interactive analysis, a new API/frontend could be implemented on top of VDMS as a webservice without adding any additional abstractions.
-An example of such frontend can be found <insert URL>.
+Interactive analysis is out of scope for this paper but this can easily be done
+using the current implementation of VDMS.
+To support interactive analysis, a new API/frontend could be implemented on top of
+VDMS as a webservice without adding any additional abstractions.
+An example of such frontend can be found <http://coco.datasets.aperturedata.io>.
 
@@
 \textbf{
-D5.
-In terms of scalability, the paper only shows that VDMS can handle large datasets, yet it is not known how VDMS can scale horizontally with more computation resources.
-For more complex logic of image preprocessing, VDMS will suffer from its single server capacity.
+D5. In terms of scalability, the paper only shows that VDMS can handle large
+datasets, yet it is not known how VDMS can scale horizontally with more
+computation resources.
+For more complex logic of image preprocessing, VDMS will suffer from its
+single server capacity.
 }\bigskip
 
-Thanks for this comment, we have added a "Future Work" section where the need for this capability in a distributed version of VDMS.
+Thanks for this comment, we have added a "Future Work" section where the need
+for this capability in a distributed version of VDMS.
 
@@
 
-Thank you for the comment.  In Section 3.2, we have expanded on the crash recovery of PMGD in comparison to other graph databases.
+Thank you for the comment.
+In Section 3.2, we have expanded on the crash recovery of PMGD in
+comparison to other graph databases.
 
@@
 \textbf{
-D2: The coverage of persistent memory is underwhelming because persistent memory is never used in the experiments.
-What is the key advantage of using persistent memory and how is it more suitable for visual data?
-Why not add experiments comparing VDMS with and without using PMGD?
+D2: The coverage of persistent memory is underwhelming because persistent memory
+is never used in the experiments.
+What is the key advantage of using persistent memory and how is it
+more suitable for visual data? Why not add experiments comparing VDMS with
+and without using PMGD?
 }\bigskip
 
-Due to page limitations and the use of SSDs instead of persistent memory in the evaluation, we did not elaborate on advantages of PMGD and related experiments.
+Due to page limitations and the use of SSDs instead of persistent memory in the
+evaluation, we did not elaborate on advantages of PMGD and related experiments.
 
@@
 \textbf{
-D3: The experiments section seems unnecessarily long, yet missing some important details.
-All the experiments are performed on images and do not handle videos or feature values at all.
-For example, video operations may include interval operations or video clip retrieval.
-Some parts seem redundant, e.g., the description of Figure 9 seems near-identical to that of Figure 8.
+D3: The experiments section seems unnecessarily long, yet missing some important
+details. All the experiments are performed on images and do not handle videos or feature
+values at all. For example, video operations may include interval operations or video
+clip retrieval. Some parts seem redundant, e.g., the description of Figure 9 seems
+near-identical to that of Figure 8.
 }\bigskip
 
-The paper is focused on an image search application so we did not include any experiments on videos or feature vectors.
-We have experiments that we can provide in appendix but due to page limitations, it did not fit the scope of this paper.
+The paper is focused on an image search application so we did not include any
+experiments on videos or feature vectors. We have experiments that we can provide in
+appendix but due to page limitations, it did not fit the scope of this paper.
 
-Figure 8 and 9 are quite similar but provide the performance difference with different numer of concurrent clients.
-We have removed figure 9 and replaced with a figure demonstrating the difference between queries with and without the resize operation.
+Figure 8 and 9 are quite similar but provide the performance difference with different
+numer of concurrent clients. We have removed figure 9 and replaced with a figure
+demonstrating the difference between queries with and without the resize operation.
 

### 2021-05-03 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Minor updates on reviewers response

--- 2021-04-30 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
@@
 This is a great point, we very much appreciate the suggestion.
-We have run a new experiment and added a new figure to the paper (Figure 10),
+We have run a new experiment and added a new figure to the paper (Figure 9),
 where we compare 3 of the queries with and without the resize operation
-to compare the results.
-We have also added an explanation on the differences in performance
+to understand the difference in performance better.
+We have also added an explanation on these differences in performance
 in the final version.
+Note that, even if the resize happens on the client side in the case of both
+baselines we have built, the computing capabilities of the server and the
+client are identical. This is on purpose, to avoid the difference in
+computing power for the resize operation to affect the understanding
+of the results.
 
@@
 \textbf{
-D2.
-Both baselines are similar in architecture, and the evaluation has not effectively highlighted the performance gain of other design choices authors made in VDMS, except that the VDMS server process both metadata operation and image operation within only one round of communication with clients.
-For example, benchmarking systems can be built to test in isolation the improvement on metadata handling.
+D2. Both baselines are similar in architecture, and the evaluation has not effectively
+highlighted the performance gain of other design choices authors made in VDMS,
+except that the VDMS server process both metadata operation and image operation
+within only one round of communication with clients.
+For example, benchmarking systems can be built to test in isolation the
+improvement on metadata handling.
 }\bigskip
 
-Thank you for the input.
-To demonstrate the benefits of VDMS more clearly, we have added Figure 10 which compares queries with and without the resize operation.
-Without the resize operation, the evaluation demonstrates the improvement on metadata handling better.
+Thank you for the input. To demonstrate the benefits of VDMS more clearly,
+we have added Figure 9, which compares queries with and without the resize operation.
+Without the resize operation, the evaluation demonstrates the
+improvement on metadata handling better, plus the effect of a single round
+of communication between client and the server (as it is the case with VDMS).
 
@@
 \textbf{
-D3.
-The paper should detail the types of queries, and image preprocessing it supports, as this being “one of the most differentiating aspects” as mentioned by them.
-And for the selection of operations supported, authors need to argue for their decision in the application contexts.
+D3. The paper should detail the types of queries, and image preprocessing it supports,
+as this being “one of the most differentiating aspects” as mentioned by them.
+And for the selection of operations supported, authors need to argue for their
+decision in the application contexts.
 }\bigskip
 
-Thank you for this recommendation.  We have provided additional details on which operations we support and our decision for chosing the operations in section 3.
+Thank you for this recommendation.
+We have provided additional details on which operations we support and our
+decision for choosing the operations in section 3.
 
@@
 \textbf{
-D4.
-Data scientists commonly perform data exploration on raw data and interactively curate their datasets for downstream tasks like model training.
-VDMS could design new mechanisms to support interactive analysis, which would be beneficial.
+D4. Data scientists commonly perform data exploration on raw data and interactively
+curate their datasets for downstream tasks like model training.
+VDMS could design new mechanisms to support interactive analysis,
+which would be beneficial.
 }\bigskip
 
-Interactive analysis is out of scope for this paper but this can easily be done using the current implementation of VDMS.
-To support interactive analysis, a new API/frontend could be implemented on top of VDMS as a webservice without adding any additional abstractions.
-An example of such frontend can be found <insert URL>.
+Interactive analysis is out of scope for this paper but this can easily be done
+using the current implementation of VDMS.
+To support interactive analysis, a new API/frontend could be implemented on top of
+VDMS as a web app without adding any additional abstractions.
+An example of such front-end can be found <http://coco.datasets.aperturedata.io>.
+The front-end is very thin, as most of the heavy-load is done by VDMS through
+the interfaces mentioned in this work
+(plus a thin REST abstraction we have added recently).
 

### 2021-05-04 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
Minor updates to reviews response

@@
 
+\noindent %
 \textbf{
@@
 \bigskip
+\noindent %
 \textbf{
@@
 
-\bigskip
+\noindent %
 \textbf{
@@
 using the current implementation of VDMS.
-To support interactive analysis, a new API/frontend could be implemented on top of
+To support interactive analysis, a new API/front-end could be implemented on top of
 VDMS as a web app without adding any additional abstractions.
-An example of such front-end can be found <http://coco.datasets.aperturedata.io>.
+An example of such front-end can be found \url{http://coco.datasets.aperturedata.io}.
 The front-end is very thin, as most of the heavy-load is done by VDMS through
@@
 Thank you for the comment.
+We have expanded portions of the VDMS description where necessary.
 In Section 3.2, we have expanded on the crash recovery of PMGD in
@@
 Figure 8 and 9 are quite similar but provide the performance difference with different
-numer of concurrent clients. We have removed figure 9 and replaced with a figure
+number of concurrent clients. We have removed figure 9 and replaced with a figure
 demonstrating the difference between queries with and without the resize operation.
@@
 \bigskip
+\noindent %
 \textbf{
-D4: Fixed all typos
-Page 1: “need for multiple system” -> “need for multiple systems”
-Page 5: “desing specifically” -> “designed specifically”
-Page 7: “images, taglist, and taglist” -> “images, taglist, and autotags”
-Page 10, Figure 6: “10M database).” -> “10M database”
+D4: There are some minor typos.\\
+Page 1: “need for multiple system”\textrightarrow “need for multiple systems”\\
+Page 5: “desing specifically”\textrightarrow “designed specifically”\\
+Page 7: “images, taglist, and taglist”\textrightarrow “images, taglist, and autotags”\\
+Page 10, Figure 6: “10M database).”\textrightarrow “10M database”
 }\bigskip

### 2021-05-04 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Corrected spelling and minor updates to design

--- 2021-05-04 yfcc100m/paper/vldb2021/submision/review_response/reviews_response.tex
@@
 D4: There are some minor typos.\\
-Page 1: “need for multiple system”\textrightarrow “need for multiple systems”\\
-Page 5: “desing specifically”\textrightarrow “designed specifically”\\
-Page 7: “images, taglist, and taglist”\textrightarrow “images, taglist, and autotags”\\
-Page 10, Figure 6: “10M database).”\textrightarrow “10M database”
+Page 1: “need for multiple system” $\rightarrow$  “need for multiple systems”\\
+Page 5: “desing specifically” $\rightarrow$ “designed specifically”\\
+Page 7: “images, taglist, and taglist” $\rightarrow$ “images, taglist, and autotags”\\
+Page 10, Figure 6: “10M database).” $\rightarrow$ “10M database”
 }\bigskip

### 2021-05-04 yfcc100m/paper/vldb2021/design.tex
Corrected spelling and minor updates to design

@@
 (ACID) properties expected from databases.
-This is, users interact with their objects (metadata, images, videos, etc),
+This is, users interact with their objects (metadata, images, videos, etc.),
 as if these objects were in a connected graph.
@@
 Each object is a node in the graph.
-The information associated to each visual object (image, video, etc) is modeled as
+The information associated to each visual object (image, video, etc.) is modeled as
 "properties" of the node in the graph.
@@
 provides the ability to \textit{compute} on these visual objects
-(image, videos, etc).
+(image, videos, etc.).
 
@@
 Figure \ref{fig:arch} depicts the high-level architecture of VDMS, which
-is composed by several subcomponents, and a Query Engine that implements
+is composed by several sub-components, and a Query Engine that implements
 the unified API and hides all the complexity underneath.
-We first describe the VDMS API, and then describe its subcomponents
+We first describe the VDMS API, and then describe its sub-components
 and design decisions.
@@
 One of the most important differentiating aspects of VDMS is its API.
-VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
+VDMS is unique in recognizing visual entities (i.e., images, videos, etc.)
 as first class citizens.
@@
 (ACID) properties expected from databases.
+Specifically, the database remains consistent, each transaction either
+completes entirely or not at all, and the effects of a completed transaction
+survives any failure.
 In cases where the system may crash or fail, PMGD has the ability
@@
 than other graph databases, including Neo4j\cite{miller2013graph}.
-PMGD performance evaluation will be published separately.
+In this paper, the evaluation uses SSDs instead of persistent memory.
+Hence, PMGD performance evaluation will be published separately to highlight
+the advantages of using persistent memory for visual data.
 
@@
 visual data handling and processing
-(i.e., basic building block operations like crop, resize, etc for images/videos
+(i.e., basic building block operations like crop, resize, etc. for images/videos
 and k-nearest neighbor search for feature vectors),
@@
 The performance comparison between the array-based lossless image format
-we develop and other similar formats (like png or tiff) are outside of the
+we developed and other similar formats (like png or tiff) are outside of the
 scope of this evaluation.

### 2021-05-04 yfcc100m/paper/vldb2021/eval.tex
Update query names to match figures

@@
 
-\item \textit{q3} - {\bf {\em 2tag\_and\_resize}}: Retrieve images with two specific autotags (i.e. alligator AND lake), and resize to 224x224.
+\item \textit{q3} - {\bf {\em 2tag\_resize\_and}}: Retrieve images with two specific autotags (i.e. alligator AND lake), and resize to 224x224.
 
-\item \textit{q4} - {\bf {\em 2tag\_or\_resize}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), and resize to 224x224.
+\item \textit{q4} - {\bf {\em 2tag\_resize\_or}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), and resize to 224x224.
 
-\item \textit{q5} - {\bf {\em 2tag\_resize\_geo\_and}}: Retrieve images with two specific autotags (i.e. alligator AND lake), in a particular geo-location, and resize to 224x224.
+\item \textit{q5} - {\bf {\em 2tag\_geo\_resize\_and}}: Retrieve images with two specific autotags (i.e. alligator AND lake), in a particular geo-location, and resize to 224x224.
 
-\item \textit{q6} - {\bf {\em 2tag\_resize\_geo\_or}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), in a particular geo-location, and resize to 224x224.
+\item \textit{q6} - {\bf {\em 2tag\_geo\_resize\_or}}: Retrieve images with either of two specific autotags (i.e. alligator OR lake), in a particular geo-location, and resize to 224x224.
 
@@
 \includegraphics[width=\columnwidth]{figures/plot_q_2tag_resize_or_mosaic_results_throughput_threads}
-\caption{Concurrency Analysis on \textit{q4} (\textit{2tag\_or\_resize}) from our use-case
+\caption{Concurrency Analysis on \textit{q4} (\textit{2tag\_resize\_or}) from our use-case
 described in the Experimental Setup Section.

### 2021-05-04 yfcc100m/paper/vldb2021/eval.tex
Corrected figure 8 to use correct figure based on text, fixed minor errors

@@
 
-\subsection{Images Search}
+\subsection{Image Search}
 \label{images}
@@
 Note that one could argue that the opposite will happen
-when the resize operation retrieves a an up-sampled (larger) version of the image
+when the resize operation retrieves an up-sampled (larger) version of the image
 instead of a down-sampled (smaller) one.
@@
 
-% \begin{figure}[ht]
-% % WAS: begin{figure*}[ht]; textwidth
-% \includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput}
-% \caption{Throughput Analysis using all queries from our use-case
-% described in the Experimental Setup Section.
-% We show queries in different figures for readability reasons.
-% The experiments show the performance of all systems (VDMS and the two baselines) as the
-% database size increases.
-% These queries were run using 56 simultaneous clients,
-% and averaged out of 10 runs.}
-% \label{fig:q_throughput_56}
-% \end{figure}
+\begin{figure}[ht]
+% WAS: begin{figure*}[ht]; textwidth
+\includegraphics[width=\columnwidth]{figures/plot_th_56_mosaic_results_throughput}
+\caption{Throughput Analysis using all queries from our use-case
+described in the Experimental Setup Section.
+We show queries in different figures for readability reasons.
+The experiments show the performance of all systems (VDMS and the two baselines) as the
+database size increases.
+These queries were run using 56 simultaneous clients,
+and averaged out of 10 runs.}
+\label{fig:q_throughput_56}
+\end{figure}
 
@@
 
-\begin{figure}[ht!]
-\includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
-\caption{Throughput Analysis using all queries from our use-case
-described in the Experimental Setup Section.
-We show queries in different figures for readability reasons.
-The experiments show the performance of all systems (VDMS and the two baselines) as the
-database size increases.
-These queries were run using 16 simultaneous clients, and averaged out of 10 runs.}
-\label{fig:q_throughput_16}
-\end{figure}
+% \begin{figure}[ht!]
+% \includegraphics[width=\columnwidth]{figures/plot_th_16_mosaic_results_throughput}
+% \caption{Throughput Analysis using all queries from our use-case
+% described in the Experimental Setup Section.
+% We show queries in different figures for readability reasons.
+% The experiments show the performance of all systems (VDMS and the two baselines) as the
+% database size increases.
+% These queries were run using 16 simultaneous clients, and averaged out of 10 runs.}
+% \label{fig:q_throughput_16}
+% \end{figure}
 
@@
 Doing a resize operation as part of the retrieval process
-brings the throughput orders of magnitude down.
-This is the case for both VDMS and the baseline.
+brings the throughput down by orders of magnitude.
+This is the case for both VDMS and the baselines.
 Nonetheless, VDMS is orders of magnitude faster when retrieving images

### 2021-05-04 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Cleaned some responses to reviewer

@@
 
-We partially address this in D3 and also elaborated more on the benefits of
+We have provided additional details on which operations we support and our
+decision for choosing the operations in section 3.
+We also elaborated more on the benefits of
 using VDMS over relational databases in Section 3.1.
@@
 (plus a thin REST abstraction we have added recently).
+In section 3 of final version, we briefly mention this ability.
 
@@
 
-Thanks for this comment, we have added a "Future Work" section where the need
-for this capability in a distributed version of VDMS.
+Thanks for this comment.
+We have included this capability in the "Future Work" section
+added to the final version.
 
@@
 Thank you for the comment.
-We have expanded portions of the VDMS description where necessary.
 In Section 3.2, we have expanded on the crash recovery of PMGD in
@@
 
-Due to page limitations and the use of SSDs instead of persistent memory in the
-evaluation, we did not elaborate on advantages of PMGD and related experiments.
+We did not elaborate on advantages of using persistent memory or include
+any PMGD experiments due to page limitations and the use of SSDs
+instead of persistent memory in the evaluation.
+PMGD performance evaluation will be published separately to highlight
+the advantages of using persistent memory for visual data.
 
@@
 
+Thank you for this comment.
 The paper is focused on an image search application so we did not include any
@@
 
-All typos are fixed in the final version.
+Thank you for highlighting this. All typos are fixed in the final version.
 

### 2021-05-04 yfcc100m/paper/vldb2021/design.tex
Cleaned some responses to reviewer
Update design section based on reviewer's comment

@@
 all the data the application need behind a single, unified API.
+This means users can access all data (metadata, images, videos, etc.)
+from the VDMS API and even implement their own API/front-end on top
+of VDMS without adding any additional abstractions.
 This is in contrast with current applications that rely on a combination
@@
 (changing output formats and encodings).
+Moreover, because VDMS is in control of the metadata and the image data,
+re-ordering or pre-processing operations can be done transparently,
+increasing the performance of the retrieval process.
 The API allows interaction with metadata, images, videos, bounding boxes,
@@
 wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.
+The visual data pre-processing operations supported in VDMS are generic and
+application independent.
+By design, we only include pre-processing operations that are not specific
+to one domain, but rather general for visual analytics.
 

### 2021-05-04 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Update reviews response

@@
 \usepackage{graphicx}
+\usepackage{pdfpages}
 \usepackage{url}
@@
 
-Not for this work, we will add to future work.
-Also, building a comparison baseline that integrates all that functionality
-is very ad-hoc, and would complicate the understanding of the origins
-of the benefits of VDMS.
-
-We can provide appendix.
+Thanks for your comment.
+We have decided to avoid going in detail on the Feature Vector support
+and interfaces primarily due to space constraints.
+We should note that this functionality is fully implemented,
+and you can find the more details on the github wiki of the project
+\footnote{https://github.com/IntelLabs/vdms/wiki/Descriptors-Tutorial,
+https://github.com/IntelLabs/vdms/wiki/FindDescriptor}.
+Based on feedback from last year submission at this same venue, we decided
+to skip the feature vector functionality evaluation and focus on image search
+based on metadata to better be able to compare against well-known baselines.
+Including an analysis of similarity search in this same work would also
+complicate the implementation of the baseline, making it harder to understand
+and isolate the main sources of performance gains.
+Implementing a baseline that also includes similarity search, even if doable,
+would mean combining more than 4 systems (SQL database, web server, client side
+pre-processing with OpenCV, some quick-and-dirty implementation
+to provide client-server architecture to Faiss/Annoy, given that these
+are just C++ libraries).
+This would increase the complexity of the baseline systems, adding
+more parameters to tune, and complicating the understanding of the results.
+We have decided to focus this work on image search based on metadata tags
+and the addition of needed pre-processing operations, and
+work on a similarity-search specific evaluation as a next step.
+If reviewers are interested in learning more about our similarity search functionality,
+we have attached as an appendix part of the evaluation we submitted last year.
+It also includes initial evaluation on the Video API, which works in a very
+similar way as it does for images.
+We will continue to improve this evaluation, and submit to arxiv directly
+as an extension of this work.
 
@@
 baselines we have built, the computing capabilities of the server and the
-client are identical. This is on purpose, to avoid the difference in
-computing power for the resize operation to affect the understanding
-of the results.
+client are identical.
+This is on purpose, to avoid the difference in computing power for the
+resize operation to affect the understanding of the results.
 
@@
 the present evaluation, plus the use of the feature vector queries,
-that are natively supported using the same set of APIs.
+that are natively supported using the same set of APIs (as we mention in D1).
 Because of space constraints, we were not able to accommodate initial evaluation
@@
 
-Thanks for this comment, we have added a "Future Work" section where we cover
-this and D3, expanding further on the most immediate lines of
-improvement and research for the VDMS project.
+Thanks for this comment, great suggestion.
+We have added a "Future Work" section where we cover this and D3,
+expanding further on the most immediate lines of improvement and
+research for the VDMS project.
 
@@
 
-We have provided additional details on which operations we support and our
-decision for choosing the operations in section 3.
+Thanks for this comment.
+We avoided expanding on supported operations for images due to space constraints,
+and because these are fully documented on the projects wiki
+\footnote{https://github.com/IntelLabs/vdms/wiki}.
+The wiki also contains the full documentation on APIs for video and similarity
+search, among other things, which work together with the API we used for
+image search for this work, providing a unified API for all the data applications need.
+We have expanded on additional details on which operations we support and our
+decision for choosing the operations in Section 3.
 We also elaborated more on the benefits of
@@
 
-Thank you for the input. To demonstrate the benefits of VDMS more clearly,
-we have added Figure 9, which compares queries with and without the resize operation.
+Thank you for the input.
+To demonstrate the benefits of VDMS more clearly, and to address another
+reviewers' comment,  we have added Figure 9, which compares queries
+with and without the resize operation.
 Without the resize operation, the evaluation demonstrates the
@@
 
-Thank you for this recommendation.
+Thank you for this recommendation, it is an important clarification we have added.
+In a nutshell, we only add pre-processing operations to VDMS API only if such
+operation are not application-dependent.
+Operations like resize, crop, zoom, transcoding, convert to black and white,
+re-sample video, and access to random frames in a video, among others, are
+example of operations that are needed in many different applications,
+and that it is performed in the same way in all uses.
 We have provided additional details on which operations we support and our
-decision for choosing the operations in section 3.
+decision for choosing the operations in Section 3.
 
@@
 
-Interactive analysis is out of scope for this paper but this can easily be done
-using the current implementation of VDMS.
+Interactive analysis of visual data is out of scope for this paper
+but this can easily be done using the current implementation of VDMS.
 To support interactive analysis, a new API/front-end could be implemented on top of
-VDMS as a web app without adding any additional abstractions.
+VDMS API as a web app without adding any additional abstractions.
 An example of such front-end can be found \url{http://coco.datasets.aperturedata.io}.
@@
 added to the final version.
+We are currently working on a distributed version of the system.
 
@@
 Thank you for the comment.
-In Section 3.2, we have expanded on the crash recovery of PMGD in
-comparison to other graph databases.
+VDMS does support crash recovery through the use of the ACID properties
+of the Graph Engine (PMGD).
+The design and implementation of PMGD, and the mechanism by which ACID properties
+are guaranteed, are currently under submission, and will be published separately.
+In Section 3.2, we have added a clarification regarding this detail.
 
@@
 instead of persistent memory in the evaluation.
+In our evaluation, we use SSDs (and not persistent memory) to avoid
+being unfair with the baseline systems, and have a better understanding of
+the results.
 PMGD performance evaluation will be published separately to highlight
@@
 
-Thank you for this comment.
+Thank you for this comment, and it is a great suggestion.
+Figure 8 and 9 are quite similar but provide the performance difference with different
+number of concurrent clients we found interesting to be shared.
+We have removed figure 9 and replaced with a figure
+demonstrating the difference between queries with and without the resize operation.
+
 The paper is focused on an image search application so we did not include any
-experiments on videos or feature vectors. We have experiments that we can provide in
-appendix but due to page limitations, it did not fit the scope of this paper.
-
-Figure 8 and 9 are quite similar but provide the performance difference with different
-number of concurrent clients. We have removed figure 9 and replaced with a figure
-demonstrating the difference between queries with and without the resize operation.
+experiments on videos or feature vectors, but we have attached as an appendix
+part of the evaluation we submitted last year.
+It also includes initial evaluation on the Video API, which works in a very
+similar way as it does for images.
 

### 2021-05-05 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Update review response

@@
 
+%===========================================================
+\newpage
 \section{Reviewer \#1}
+%===========================================================
 
@@
 We should note that this functionality is fully implemented,
-and you can find the more details on the github wiki of the project
-\footnote{https://github.com/IntelLabs/vdms/wiki/Descriptors-Tutorial,
-https://github.com/IntelLabs/vdms/wiki/FindDescriptor}.
+and you can find the more details on the project's wiki
+\footnote{https://github.com/IntelLabs/vdms/wiki/Descriptors-Tutorial}
+\footnote{https://github.com/IntelLabs/vdms/wiki/FindDescriptor}.
 Based on feedback from last year submission at this same venue, we decided
@@
 
+%===========================================================
 \newpage
 \section{Reviewer \#4}
+%===========================================================
 
@@
 
-Thanks for this comment.
+Thanks for your comment, these are important clarifications to add.
+We have expanded on additional details on which operations we support and our
+decision for choosing the operations in Section 3.
+We also elaborated more on the benefits of
+using VDMS over relational databases in Section 3.1.
 We avoided expanding on supported operations for images due to space constraints,
-and because these are fully documented on the projects wiki
+and because these are fully documented on the project's wiki
 \footnote{https://github.com/IntelLabs/vdms/wiki}.
@@
 image search for this work, providing a unified API for all the data applications need.
-We have expanded on additional details on which operations we support and our
-decision for choosing the operations in Section 3.
-We also elaborated more on the benefits of
-using VDMS over relational databases in Section 3.1.
 
@@
 with and without the resize operation.
-Without the resize operation, the evaluation demonstrates the
+Without the resize operation, the evaluation shows the
 improvement on metadata handling better, plus the effect of a single round
@@
 but this can easily be done using the current implementation of VDMS.
-To support interactive analysis, a new API/front-end could be implemented on top of
-VDMS API as a web app without adding any additional abstractions.
-An example of such front-end can be found \url{http://coco.datasets.aperturedata.io}.
-The front-end is very thin, as most of the heavy-load is done by VDMS through
-the interfaces mentioned in this work
+To support interactive analysis, a front-end could be implemented on top of
+VDMS API as a web app without needing any additional abstractions.
+An example of such front-end can be found here: \url{http://coco.datasets.aperturedata.io}.
+This front-end is very thin, as most of the heavy-load is done by VDMS through
+the interfaces described in this work
 (plus a thin REST abstraction we have added recently).
-In section 3 of final version, we briefly mention this ability.
+In section 3 of final version, we have added a brief mention to this mechanism.
 
@@
 
+%===========================================================
 \newpage
 \section{Reviewer \#5}
+%===========================================================
 

### 2021-05-05 yfcc100m/paper/vldb2021/futurework.tex
Update future work

@@
 given the limitations discussed on \ref{concurrency_analysis}.
+Even if this evaluation focuses on image search using metadata properties,
+VDMS has the capability to also perform similarity search based on
+feature vectors.
+The team will work on a evaluation of this feature as future work,
+as more work is needed to arrive at a comparable and fair baseline.
+Last, but not least, we recognize the need for a distributed
+implementation of VDMS.
+Most of the subcomponents and abstractions were design with a distributed
+architecture in mind, but also most of the applications we encountered in our
+internal use cases would fit on a single server plus a distributed file system.
+With a validation of the benefits of the abstractions provided by VDMS,
+the natural next step is working on scaling the system out.

### 2021-05-09 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Minor fix on reviews response
Update reviews response

@@
 %===========================================================
-\newpage
+\section{General Description of Changes}
+%===========================================================
+
+The authors would like to thank the program committee and the reviewers
+for the comments and suggestions,
+instrumental to increase the quality of our work and its presentation.
+
+We have addressed comments from every reviewer separately, but the following
+list summarize the main updates:
+
+\begin{itemize}
+  \item We have added and run a new experiment that compares VDMS and baseline
+        for queries that do not run a pre-processing operations, following
+        reviewer's suggestions.
+        We expand the analysis of the results including these findings.
+  \item We have expanded on the type of supported pre-processing
+        operations, and more details on VDMS API design in general.
+  \item We have expanded on the description of the ACID properties offered
+        by VDMS and how that mechanism work.
+  \item We have expanded on the type of queries supported by VDMS,
+        including similarity-search based on feature vectors,
+        which is a feature currently supported by the system.
+  \item We have added a "Future Work" sections where we describe some
+        of the limitations found during this experiments, as well as the
+        near and long term line of work after this paper.
+\end{itemize}
+
+%===========================================================
 \section{Reviewer \#1}
@@
 Thanks for your comment.
-We have decided to avoid going in detail on the Feature Vector support
-and interfaces primarily due to space constraints.
-We should note that this functionality is fully implemented,
-and you can find the more details on the project's wiki
+Following feedback and suggestions from last year submission at this same venue,
+we decided to skip the feature vector functionality evaluation and
+focus on image search based on metadata to better be able to compare
+against well-known baselines.
+We should note that similarity search based on feature vectors
+is fully implemented, and you can find the more details on the project's wiki
 \footnote{https://github.com/IntelLabs/vdms/wiki/Descriptors-Tutorial}
 \footnote{https://github.com/IntelLabs/vdms/wiki/FindDescriptor}.
-Based on feedback from last year submission at this same venue, we decided
-to skip the feature vector functionality evaluation and focus on image search
-based on metadata to better be able to compare against well-known baselines.
+
 Including an analysis of similarity search in this same work would also
@@
 and isolate the main sources of performance gains.
-Implementing a baseline that also includes similarity search, even if doable,
-would mean combining more than 4 systems (SQL database, web server, client side
+We have a preliminary implementation of baselines for our internal use cases,
+and it meant combining more than 4 systems (SQL database, web server, client side
 pre-processing with OpenCV, some quick-and-dirty implementation
@@
 are just C++ libraries).
-This would increase the complexity of the baseline systems, adding
+This increases the complexity of the baseline systems, adding
 more parameters to tune, and complicating the understanding of the results.
@@
 We have also added an explanation on these differences in performance
-in the final version.
+in the final version, and made these results publicly available
+\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results/resize\_comparison}
+(as the rest of the evaluation in this work).
 Note that, even if the resize happens on the client side in the case of both
@@
 Because of space constraints, we were not able to accommodate initial evaluation
-we have been working on with Feature Vector search specifically,
+we have been working on for Feature Vector-based search specifically,
 and we plan to perform a more comprehensive evaluation on this functionality
@@
 We have expanded on additional details on which operations we support and our
-decision for choosing the operations in Section 3.
+decision for choosing the operations in Section 3, and how supporting
+these operations natively opens optimization possibilities that we exploit.
 We also elaborated more on the benefits of
-using VDMS over relational databases in Section 3.1.
+using VDMS over ad-hoc solutions in Section 3.1.
 We avoided expanding on supported operations for images due to space constraints,
@@
 Thank you for the input.
-To demonstrate the benefits of VDMS more clearly, and to address another
-reviewers' comment,  we have added Figure 9, which compares queries
-with and without the resize operation.
+To demonstrate the benefits of VDMS more clearly, we have added Figure 9,
+which compares queries with and without the resize operation.
 Without the resize operation, the evaluation shows the
@@
 of communication between client and the server (as it is the case with VDMS).
+Because we design VDMS for visual data as the first class elements,
+we focused our evaluation on the retrieval of visual objects rather
+than metadata only.
+When it comes to building solutions that primarily depend on metadata,
+rather than actual visual objects, the authors believe that VDMS is not
+the right solution, given that there are many other commercially available
+solutions (relational databases, graph databases, document-oriented databases, etc)
+that have been optimized over many decades and provide a level of
+functionality and maturity that is sufficient for the majority of the use cases.
 
@@
 Thank you for this recommendation, it is an important clarification we have added.
-In a nutshell, we only add pre-processing operations to VDMS API only if such
-operation are not application-dependent.
-Operations like resize, crop, zoom, transcoding, convert to black and white,
+In a nutshell, we add pre-processing operations to VDMS API only if such
+operations are application-independent.
+Operations like resize, crop, zoom, trans-coding, convert to black and white,
 re-sample video, and access to random frames in a video, among others, are
@@
 
-Interactive analysis of visual data is out of scope for this paper
-but this can easily be done using the current implementation of VDMS.
+This is a great point, and a primary motivation for building a system like VDMS.
+VDMS, by design, supports functionality that makes it easy to build
+iterative analysis using its API.
 To support interactive analysis, a front-end could be implemented on top of
 VDMS API as a web app without needing any additional abstractions.
-An example of such front-end can be found here: \url{http://coco.datasets.aperturedata.io}.
+An example of such front-end can be found here:
+\url{http://coco.datasets.aperturedata.io}.
 This front-end is very thin, as most of the heavy-load is done by VDMS through
@@
 We did not elaborate on advantages of using persistent memory or include
-any PMGD experiments due to page limitations and the use of SSDs
-instead of persistent memory in the evaluation.
+any PMGD experiments primarily due to space constraints, and to avoid
+complicating the understand of the experimental evaluation.
 In our evaluation, we use SSDs (and not persistent memory) to avoid
-being unfair with the baseline systems, and have a better understanding of
+being unfair with the baseline systems (which do not have explicit support for
+persistent memory abstractions), and have a better understanding of
 the results.
 PMGD performance evaluation will be published separately to highlight
-the advantages of using persistent memory for visual data.
+the advantages of using persistent memory in general.
 
@@
 number of concurrent clients we found interesting to be shared.
-We have removed figure 9 and replaced with a figure
+We have removed figure 9 and replaced it with a figure
 demonstrating the difference between queries with and without the resize operation.
@@
 experiments on videos or feature vectors, but we have attached as an appendix
-part of the evaluation we submitted last year.
+part of our preliminary evaluation using similarity search.
 It also includes initial evaluation on the Video API, which works in a very

### 2021-05-09 yfcc100m/paper/vldb2021/design.tex
Minor updates based on reviewers feedback

@@
 to recover to a consistent state without loading its content into memory.
-Unlike existing graph databases that are disk-based, our graph engine
-avoids converting data from a disk-friendly format to an usable format for an application.
+This is done through an undo log, chosen because it is optimistic and
+it handles reads after writes naturally.
+Upon failures, PMGD consults the log to restore the database
+to its state just past the last committed transaction.
 PMGD comes as a C++ library that VDMS uses for implementing its query engine.

### 2021-05-09 yfcc100m/paper/vldb2021/eval.tex
Reduce systems figure to single column
Minor updates based on reviewers feedback

@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=\textwidth]{figures/comparison_system}
+\includegraphics[width=\columnwidth]{figures/comparison_system}
 \caption{Comparison Systems: Logical view of the interaction between the client application with VDMS (left) and the baseline systems based on Apache Web Server and PostgreSQL (middle), or MySQL (right).
@@
 \label{fig:systems}
-\end{figure*}
+\end{figure}
 
@@
 commonly performed, as visual data is generally pre-processed before analytics
-computation are performed on them.
+computation are performed on them (i.e., a resize plus
+normalization done on images before passing them through a neural network).
 Figure \ref{fig:q_throughput_56_resize} shows how the query throughput varies
@@
 This is the common case in most analytics applications, where pre-processing
-operations happen on the client side (or when the analytics computation happens),
+operations happen on the client side (or wherever the analytics computation happens),
 rather than near where the data is located, as it is the case with VDMS.

### 2021-05-09 yfcc100m/python/build_databases/vdms_build/build_all.sh
wip descriptors stuff

@@ -3,19 +3,29 @@
 rm -rf logs
 mkdir logs
 
-input_root="/mnt/yfcc100m/metadata/processed/splitted/"
-output_root="/mnt/yfcc100m/golden_building/"
+echo "$(date) - Build start..."
+
+input_root="/mnt/largedata/yfcc100m/metadata/processed/splitted/"
+output_root="/mnt/data/yfcc100m/"
 
 python3 build_yfcc_db.py \
           -data_file=${input_root}/yfcc100m_dataset_000 \
           -tag_file=${input_root}/yfcc100m_dataset_autotags_000 \
           -tag_list=${input_root}/../autotag_list.txt \
           -add_tags > logs/screen_000.log 2> logs/error_000.log
+sync
 
-# sync
-# (cd ${output_root}/vdms/ && \
-#  bash check_graph.sh && \
-#  bsdtar cvfz ${output_root}/vdms_archive_000.tar.gz db )
+echo "$(date) - Done building 1M."
+
+echo "$(date) - Compressing..."
+
+(cd ${output_root} && \
+ bsdtar cfz ${output_root}/vdms_archive_000.tar.gz db )
+
+echo "$(date) - Done compressing."
+# bash check_graph.sh && \
+
+exit 0
 
 for i in $(seq -f "%03g" 1 99)
 do

### 2021-05-09 yfcc100m/python/build_databases/vdms_build/build_yfcc_db.py

@@ -15,15 +15,15 @@ import sys
 # process_first_n = 500000
 
 log_name = sys.argv[0].split('/')[-1].replace('.py','.log')
-connection_batch_limit = 100  #10
+connection_batch_limit = 200  #10
 
 def get_args():
     parserobj = argparse.ArgumentParser()
     # parserobj.add_argument('-process_first_n', type=int, default=None,
                            # help='Process only the first N lines of image/connection data [default: None-> all data]')
-    parserobj.add_argument('-num_threads', type=int, default=100,
-                           help='Number of threads to use [default: 100]')
-    parserobj.add_argument('-batch_size', type=int, default=100,
+    parserobj.add_argument('-num_threads', type=int, default=16,
+                           help='Number of threads to use [default: 32]')
+    parserobj.add_argument('-batch_size', type=int, default=150,
                            help='Number of entries per thread for autotags and images [default: 100; connections: max {}]'.format(connection_batch_limit))
     parserobj.add_argument('-data_file', type=str,
                            default='yfcc100m_dataset_1M',
@@ -54,7 +54,7 @@ def get_db_list(num):
     dbs = []
     for i in range(0, num):
         db = vdms.vdms()
-        db.connect("sky4.local")
+        db.connect("localhost")
         dbs.append(db)
     return dbs
 
@@ -116,7 +116,6 @@ def process_tag_entities(params, dbs, all_data):
             idx += batch
 
     return error_counter
-
 
 def process_image_entities(params, dbs, all_data):
     batch, num_lines, blocks, results = get_parameters(params, all_data)

### 2021-05-09 yfcc100m/python/build_databases/vdms_build/util.py

@@ -180,8 +180,8 @@ def add_autotag_connection_batch(index, database, row_data, start, end, results)
         if not pd.isna(row['autotags']):
             ind = [None, None]
             # Find Image
-            parentref = (100 * rix) % 19000
-            parentref +=1  #Max ref 20000
+            parentref = (100 * rix) % 9000
+            parentref +=1  #Max ref 10000
             query = {}
             findImage = {}
             findImage["_ref"] = parentref
@@ -242,10 +242,17 @@ def add_autotag_connection_batch(index, database, row_data, start, end, results)
 """ Using single entry per thread """
 
 def add_autotags_entity(index, database, tag, results):
-    query = {}
-    add_entity = {"class": 'autotags', "properties": {"name": tag}}
-    query["AddEntity"] = add_entity
-    res = database.query([query])
+
+    query = [ {
+        "AddEntity": {
+            "class": "autotags",
+            "properties": {
+                "name": tag
+            }
+        }
+    }]
+
+    res = database.query(query)
     # print('res[0][0]', res[0][0])
     results[index] = res[0][0]['AddEntity']["status"]
 

### 2021-05-09 yfcc100m/python/load_descriptors/descriptors_io.py

@@ -4,17 +4,21 @@ import numpy
 
 class descriptors_reader(object):
 
-    def __init__(self, prefix):
+    def __init__(self, prefix="", just_one=False, just_one_path=""):
 
         self.file = None
         self.prefix = prefix
         self.file_counter = 0
+        self.just_one = just_one
 
         self.dimensions = 4096
 
         self.desc_counter = 0
 
-        self.read_next_file()
+        if just_one:
+            self.file = open(just_one_path, 'rb')
+        else:
+            self.read_next_file()
 
     def get_next_n(self, n):
 
@@ -27,8 +31,10 @@ class descriptors_reader(object):
 
             id_bin = self.file.read(8)
             if len(id_bin) != 8 :
-                self.read_next_file()
-                continue
+                if self.read_next_file():
+                    continue
+                else:
+                    break
 
             id_long = struct.unpack('@l', id_bin)[0]
 
@@ -43,12 +49,17 @@ class descriptors_reader(object):
 
     def read_next_file(self):
 
+        if self.just_one:
+            return False
+
         filename = self.prefix + str(self.file_counter) + ".bin"
 
         print("Reading", filename, "...")
 
         self.file = open(filename, 'rb')
         self.file_counter += 1
+
+        return True
 
 class descriptors_writer(object):
 

### 2021-05-09 yfcc100m/tools/downloader/videos/yfcc100m_videos_download.py

@@ -13,6 +13,7 @@ import logging
 
 FROOT = "input_files"
 LICENSETYPE = "Attribution License"
+YFCC_VIDEOS = "/mnt/largedata/yfcc100m/metadata/yfcc100m_dataset_videos"
 
 '''
 License Options:
@@ -28,7 +29,7 @@ BASEURL = 'https://multimedia-commons.s3-us-west-2.amazonaws.com/data/videos/mp4
 SEARCH1 = "backpack"
 
 def main():
-    print("Starting dataset processing")
+    # print("Starting dataset processing")
     mdpdf = getMetaData()
 
     ''' Get a subset that has appropriate license type '''
@@ -41,13 +42,13 @@ def main():
     ''' Fetch those videos '''
     getVideos(searchpdf)
 
-    print("Finished downloading")
+    # print("Finished downloading")
     sys.exit(0)
 
 def getMetaData():
 
     ''' PDF for yf100m metadata '''
-    mdfile = getPath("yfcc100m_videos_metadata.tsv")
+    mdfile = getPath(YFCC_VIDEOS)
     mdheaderfile = getPath("yfcc100m_videos_metadata_headers.csv")
     mdheader = [x.strip() for x in open(mdheaderfile)]
     mdpdf = pd.read_csv(mdfile, "\t", header = None)
@@ -86,7 +87,7 @@ def getVideo(rowd,baseurl):
     hashf = hashd + ".mp4"
     fullurl = os.path.join(baseurl,*[sub1,sub2,hashf])
     print(fullurl)
-    # return
+    return
     try:
         urllib.request.urlretrieve(fullurl,idf)
     except Exception as e:

### 2021-05-10 yfcc100m/paper/vldb2021/submission/review_response/reviews_response.tex
Minor updates to reviews_response

@@
   \item We have added a "Future Work" sections where we describe some
-        of the limitations found during this experiments, as well as the
+        of the limitations found during these experiments, as well as the
         near and long term line of work after this paper.
@@
 Thanks for your comment.
-Following feedback and suggestions from last year submission at this same venue,
+Following feedback and suggestions from last year's submission at this same venue,
 we decided to skip the feature vector functionality evaluation and
-focus on image search based on metadata to better be able to compare
-against well-known baselines.
+focus on image search based on metadata to compare
+against well-known baselines better.
 We should note that similarity search based on feature vectors
@@
 Including an analysis of similarity search in this same work would also
-complicate the implementation of the baseline, making it harder to understand
+complicate the implementation of the baselines, making it harder to understand
 and isolate the main sources of performance gains.
@@
 If reviewers are interested in learning more about our similarity search functionality,
-we have attached as an appendix part of the evaluation we submitted last year.
+we have attached, as an appendix, part of the evaluation we submitted last year.
 It also includes initial evaluation on the Video API, which works in a very
@@
 on feature vectors and also filter by geographical location.
-This can be achieved using the very similar queries as the one used in
+This can be achieved using very similar queries as the one used in
 the present evaluation, plus the use of the feature vector queries,
@@
 the right solution, given that there are many other commercially available
-solutions (relational databases, graph databases, document-oriented databases, etc)
+solutions (relational databases, graph databases, document-oriented databases, etc.)
 that have been optimized over many decades and provide a level of

### 2021-05-10 yfcc100m/paper/vldb2021/intro.tex
Minor changes to intro

@@
 We use \textit{visual data} to refer to any pixel-data (images, videos,
-frames in a video, etc), and feature vectors (a.k.a. descriptors, or embeddings),
+frames in a video, etc.), and feature vectors (a.k.a. descriptors, or embeddings),
 which are representations of the pixel data.
@@
 Information about a patient (name, last name, unique identifiers), or
-about a clinic (name, location, etc) are examples of what we refer to
+about a clinic (name, location, etc.) are examples of what we refer to
 as \textit{metadata} in the context of a health-care IT system.
@@
 as it involves loading all metadata into a solution that enables searching
-based on tags (relational database, graph database, csv files, etc), and
+based on tags (relational database, graph database, csv files, etc.), and
 building the necessary pipelines for querying and retrieving the right data.

### 2021-05-10 yfcc100m/paper/vldb2021/design.tex
Minor update to design

@@
 it goes beyond the typical supported data types (string, integers, floats,
-blobs, JSON-documents, etc), recognizing visual entities (image, videos,
-feature vectors, etc) as first class citizens.
+blobs, JSON-documents, etc.), recognizing visual entities (image, videos,
+feature vectors, etc.) as first class citizens.
 VDMS API enable users to insert, index, process, and query visual data,

### 2021-05-10 yfcc100m/paper/vldb2021/eval.tex
Minor updates to eval

@@
 (when compared to other object store system), at the expenses of providing less
-functionality (authentication, data integrity checking, etc).
+functionality (authentication, data integrity checking, etc.).
 The other baseline system replaces the MySQL Server 5.7 with a
@@
 Rather, a single transaction is performed and the result filtered on the client.
-Future VDMS releases will add more of such operators (AND, OR, etc)
+Future VDMS releases will add more of such operators (AND, OR, etc.)
 in order to prevent unnecessary retrievals and extra filtering on the client side.
@@
 filtering, we see \textit{mysql} showing a small advantage.
-When the resize operations is performed, both baselines do very similar
+When the resize operation is performed, both baselines do very similar
 on 1tag based queries, and orders of magnitude worst than VDMS.
@@
 principles of VDMS, which aims to eliminate the need of combining and
-re-purposing systems that were designed to handle types of date other than visual.
+re-purposing systems that were designed to handle types of data other than visual.
 VDMS, by design, eliminates most of the inefficiencies that result

### 2021-07-22 yfcc100m/paper/vldb2021/intro.tex
Added accessed month/yr to URLs based on date added to .bib. Removed duplicate references

@@
 and systems that manipulate a specific category of
-visual data~\cite{scidb, rasdaman},
+visual data~\cite{brown2010overview, rasdaman},
 VDMS can be distinguished from them on the following aspects:
@@
 the Facebook architecture for photos that combine Tao~\cite{tao},
-Haystack~\cite{beaver2010finding}, and f4~\cite{muralidhar2014f4} for metadata,
+Haystack~\cite{haystack}, and f4~\cite{muralidhar2014f4} for metadata,
 hot/recent data, and warm data respectively.

### 2021-08-03 yfcc100m/paper/vldb2021/eval.tex
Update tex files for camera ready

@@
 We use the \textit{autotags} (as they contain information about the content
-of the image), and geo-location information (latitude/longitude)
+of the image), and geo-location information (latitude and longitude)
 of the images for search and filtering.
@@
 Figure~\ref{fig:concurrency_comparison_q2} illustrates a concurrency analysis
-for \textit{q2} (\textit{1tag\_geo\_resize}), described above,
+for \textit{q2} described above (\textit{1tag\_geo\_resize}),
 using VDMS and the baselines.

### 2021-08-03 yfcc100m/paper/vldb2021/acks.tex
Update tex files for camera ready

--- 2020-02-27 yfcc100m/paper/vldb2020/acks.tex
@@
-\section{Acknowledgments}
-
+\begin{acks}
 We would like to thank the many people that made this project possible and
@@
 the servers and infrastructure needed to conduct our experiments.
+\end{acks}

### 2021-08-03 yfcc100m/paper/vldb2021/legal.tex
Update tex files for camera ready

--- 2020-02-27 yfcc100m/paper/vldb2020/legal.tex
@@
+% \balance
+\begin{framed}
 \scriptsize
-\begin{framed}
+
 \noindent \textbf{Notices \& Disclaimers}
+\\
 
-\copyright  Intel Corporation.  Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.  Other names and brands may be claimed as the property of others.
+\noindent Intel technologies may require enabled hardware, software or service activation.
+\\
 
-Performance results are based on testing as of dates shown in configurations and may not reflect all publicly available security updates.
+\noindent No product or component can be absolutely secure.
+\\
+
+\noindent Your costs and results may vary.
+\\
+
+\noindent \copyright Intel Corporation.
+Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.
+Other names and brands may be claimed as the property of others.
+\\
+
+\noindent Intel does not control or audit third-party data.  You should consult other sources to evaluate accuracy.
+\\
+
+\noindent Intel disclaims all express and implied warranties, including without limitation, the implied warranties of merchantability, fitness for a particular purpose, and non-infringement, as well as any warranty arising from course of performance, course of dealing, or usage in trade.
+
+% \copyright  Intel Corporation.
+Intel, the Intel logo, and other Intel marks are trademarks of Intel Corporation or its subsidiaries.
+Other names and brands may be claimed as the property of others.
+
+% Performance results are based on testing as of dates shown in configurations and may not reflect all publicly available security updates.
 See backup for configuration details.
@@
 
-Your costs and results may vary.
+% Your costs and results may vary.
 
-Intel technologies may require enabled hardware, software or service activation.
+% Intel technologies may require enabled hardware, software or service activation.
 
-Software and workloads used in performance tests may have been optimized for performance only on Intel microprocessors.
+\noindent Software and workloads used in performance tests may have been optimized for performance only on Intel microprocessors.
 Performance tests, such as SYSmark and MobileMark, are measured using specific computer systems, components, software, operations and functions.
@@
 For more complete information visit www.intel.com/benchmarks.
+\\
 
-Intel's compilers may or may not optimize to the same degree for non-Intel microprocessors for optimizations that are not unique to Intel microprocessors.
+\noindent Intel's compilers may or may not optimize to the same degree for non-Intel microprocessors for optimizations that are not unique to Intel microprocessors.
 These optimizations include SSE2, SSE3, and SSSE3 instruction sets and other optimizations.
